### 2021-03-23 paper.Rmd

---
title: "COVID-19 Vaccine Acceptance and Hesitancy in Low and Middle Income Countries, and Implications for Messaging"
author:
- name: Julio S. Solís Arce, BA^[First author]
  affiliation: WZB Berlin Social Science Center
  number: 1
- name: Shana S. Warren, PhD\(^*\)
  affiliation: Innovations for Poverty Action (IPA)
  number: 2
- name: Niccoló F. Meriggi, MSc\(^*\)
  affiliation: International Growth Centre (IGC)
  number: 3
- name: Alexandra Scacco, PhD\(^*\)
  number: 1
- name: Nina McMurry, PhD\(^*\)
  number: 1
- name: Maarten Voors, PhD\(^*\)
  affiliation: Wageningen University & Research
  number: 4
- name: Georgiy Syunyaev, MPhil\(^*\)
  affiliation: International Center for the Study of Institutions and Development (HSE University, Moscow, Russia)
  number: [1, ", ", 5, ", ",  24]
- name: Amyn Abdul Malik, PhD\(^*\)
  affiliation: Yale Institute for Global Health
  number: 6
- name: Samya Aboutajdine, MA
  number: 3
- name: Alex Armand, PhD
  affiliation: Nova School of Business and Economics
  number: 7
- name: Saher Asad, PhD
  affiliation: Lahore University of Management Sciences
  number: 8
- name: Britta Augsburg, PhD
  affiliation: The Institute for Fiscal Studies
  number: 9
- name: Antonella Bancalari, PhD
  affiliation: University of St. Andrews & The Institute for Fiscal Studies
  number: 10
- name: Martina Björkman Nyqvist, PhD
  affiliation: Stockholm School of Economics and Misum
  number: 11
- name: Ekaterina Borisova, PhD
  affiliation: Economics Department of Ghent University
  number: [5, ", ", 12]
- name: Constantin Manuel Bosancianu, PhD
  number: 1
- name: Ali Cheema, PhD
  affiliation: Institute of Development and Economic Alternatives
  number: [8, ", ", 13]
- name: Elliott Collins, PhD
  number: 2
- name: Ahsan Zia Farooqi, MA
  number: 13
- name: Mattia Fracchia, MA
  number: 7
- name: Andrea Guariso, PhD
  affiliation: Trinity College Dublin
  number: 14
- name: Ali Hasanain, PhD
  number: 8
- name: Anthony Kamwesigye, BSc
  number: 2
- name: Sarah Kreps, PhD
  affiliation: Cornell University
  number: 15
- name: Madison Levine, MSc
  number: 4 
- name: Rebecca Littman, PhD
  affiliation: University of Illinois Chicago
  number: 16
- name: Melina Platas, PhD
  affiliation: NYU Abu Dhabi
  number: 17
- name: Vasudha Ramakrishna, MSc
  affiliation: Yale Research Initiative on Innovation and Scale (Y-RISE)
  number: 18
- name: Jacob N. Shapiro, PhD
  affiliation: Princeton University
  number: 19
- name: Jakob Svensson, PhD
  affiliation: Institute for International Economic Studies (IIES), Stockholm University
  number: 20
- name: Corey Vernot, BS
  number: 18
- name: Pedro C. Vicente, PhD
  number: 7
- name: Laurin B Weissinger, DPhil
  affiliation: Tufts University
  number: 21
- name: Baobao Zhang, PhD
  number: 15
- name: Dean Karlan, PhD^[Last author]
  affiliation: Kellogg School of Management at Northwestern University
  number: [2, ", ", 22]
- name: Michael Callen, PhD\(^\dagger\)
  affiliation: London School of Economics
  number: [3, ", ", 23]
- name: Matthieu Teachout, PhD\(^\dagger\)
  number: 3
- name: Macartan Humphreys, PhD\(^\dagger\)
  affiliation: Columbia University
  number: [1, ", ", 24]
- name: Saad B. Omer, PhD^[Corresponding author. 1 Church Street, New Haven, CT 06510. Phone  203 432 3656.  Email  \href{mailto:saad.omer@yale.edu}{saad.omer@yale.edu}]
  number: 6
- name: Ahmed Mushfiq Mobarak, PhD^[Corresponding author. 165 Whitney Avenue, New Haven, CT 06520-8200. Phone 203 432 5787. Email  \href{mailto:ahmed.mobarak@yale.edu}{ahmed.mobarak@yale.edu}]
  affiliation: Yale University
  number: 25
output: 
  bookdown::pdf_document2:
    keep_tex: true
    pandoc_args:
      - --template=4_paper_supplementary_files/mytemplate.latex
    toc: false
    extra_dependencies:
      caption: ["font={bf,normalsize}"]
      mathptmx: null
      fontenc: ["T1"]
  bookdown::word_document2: default
bibliography: 4_paper_supplementary_files/bib_final.bib
csl: 4_paper_supplementary_files/nature.csl
header-includes:
  - \usepackage{titlesec}
  - \titleformat*{\subsection}{\normalsize\bfseries}
  - \titleformat*{\section}{\large\bfseries}
# linestretch: 1.5
urlcolor: blue
link-citations: yes
linkcolor: blue
fontsize: 12pt
---

```{r setup, include=FALSE}

library(knitr)

knitr::opts_chunk$set(echo = FALSE)

```

```{r cleaning, child = "1_cleaning.Rmd", include = FALSE, eval = FALSE}

```

```{r tables, include = FALSE, message = FALSE}

knitr::purl("2_tables.Rmd", documentation = 2)

source("2_tables.R")

file.remove("2_tables.R")

nice <- function(x) prettyNum(x, big.mark = ",", scientific = FALSE)

```

```{r load_gdoc, include = FALSE, eval = TRUE}

googledrive::drive_download(file = "NatureMedicine Covid vaccine hesitancy", 
                            path = "4_paper_supplementary_files/Intro Vaccine.txt", overwrite = TRUE)

file.rename(from = "4_paper_supplementary_files/Intro Vaccine.txt", 
            to = "4_paper_supplementary_files/temp_download.Rmd")

knitr::opts_chunk$set(echo = FALSE)

```

```{r main_text, child = "4_paper_supplementary_files/temp_download.Rmd"}

```