### 2021-03-23 paper_files/figure-latex/fig2paper-1.tex

% Created by tikzDevice version 0.12.3.1 on 2021-03-23 16:52:30
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}

\nonstopmode

\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (722.70,542.02);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.00) rectangle (722.70,542.02);
\end{scope}
\begin{scope}
\path[clip] ( 78.32, 67.14) rectangle (137.25,464.78);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 78.32, 67.14) rectangle (137.25,464.78);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 87.69, 67.14) --
	( 87.69,464.78);

\path[draw=drawColor,line width= 0.3pt,line join=round] (101.09, 67.14) --
	(101.09,464.78);

\path[draw=drawColor,line width= 0.3pt,line join=round] (114.48, 67.14) --
	(114.48,464.78);

\path[draw=drawColor,line width= 0.3pt,line join=round] (127.88, 67.14) --
	(127.88,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 78.32, 83.94) --
	(137.25, 83.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 78.32,111.94) --
	(137.25,111.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 78.32,139.95) --
	(137.25,139.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 78.32,167.95) --
	(137.25,167.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 78.32,195.95) --
	(137.25,195.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 78.32,223.96) --
	(137.25,223.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 78.32,251.96) --
	(137.25,251.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 78.32,279.96) --
	(137.25,279.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 78.32,307.96) --
	(137.25,307.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 78.32,335.97) --
	(137.25,335.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 78.32,363.97) --
	(137.25,363.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 78.32,391.97) --
	(137.25,391.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 78.32,419.98) --
	(137.25,419.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 78.32,447.98) --
	(137.25,447.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 81.00, 67.14) --
	( 81.00,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 94.39, 67.14) --
	( 94.39,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (107.79, 67.14) --
	(107.79,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (121.18, 67.14) --
	(121.18,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (134.58, 67.14) --
	(134.58,464.78);
\definecolor{drawColor}{RGB}{204,102,119}

\path[draw=drawColor,line width= 0.6pt,line join=round] (111.16,137.15) --
	(111.16,142.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (111.16,139.95) --
	( 94.55,139.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 94.55,137.15) --
	( 94.55,142.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (105.80,445.18) --
	(105.80,450.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (105.80,447.98) --
	(100.02,447.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (100.02,445.18) --
	(100.02,450.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (101.14,417.18) --
	(101.14,422.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (101.14,419.98) --
	( 94.07,419.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 94.07,417.18) --
	( 94.07,422.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 90.75,361.17) --
	( 90.75,366.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 90.75,363.97) --
	( 81.16,363.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 81.16,361.17) --
	( 81.16,366.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 94.66,333.17) --
	( 94.66,338.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 94.66,335.97) --
	( 90.37,335.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 90.37,333.17) --
	( 90.37,338.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 97.50,305.16) --
	( 97.50,310.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 97.50,307.96) --
	( 89.09,307.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 89.09,305.16) --
	( 89.09,310.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (101.57,109.14) --
	(101.57,114.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] (101.57,111.94) --
	( 99.86,111.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 99.86,109.14) --
	( 99.86,114.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] (107.95,277.16) --
	(107.95,282.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (107.95,279.96) --
	( 95.41,279.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 95.41,277.16) --
	( 95.41,282.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (113.09,249.16) --
	(113.09,254.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (113.09,251.96) --
	(106.23,251.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (106.23,249.16) --
	(106.23,254.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (116.20,221.16) --
	(116.20,226.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (116.20,223.96) --
	(107.84,223.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (107.84,221.16) --
	(107.84,226.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (129.00,193.15) --
	(129.00,198.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (129.00,195.95) --
	(124.23,195.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (124.23,193.15) --
	(124.23,198.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (109.13,165.15) --
	(109.13,170.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (109.13,167.95) --
	(103.61,167.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (103.61,165.15) --
	(103.61,170.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (126.00, 81.14) --
	(126.00, 86.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] (126.00, 83.94) --
	(120.97, 83.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (120.97, 81.14) --
	(120.97, 86.74);
\definecolor{fillColor}{RGB}{204,102,119}

\path[fill=fillColor] (102.86,139.95) circle (  3.57);

\path[fill=fillColor] (102.91,447.98) circle (  3.57);

\path[fill=fillColor] ( 97.61,419.98) circle (  3.57);

\path[fill=fillColor] ( 85.98,363.97) circle (  1.43);

\path[fill=fillColor] ( 92.52,335.97) circle (  3.57);

\path[fill=fillColor] ( 93.32,307.96) circle (  3.57);

\path[fill=fillColor] (100.71,111.94) circle (  3.57);

\path[fill=fillColor] (101.68,279.96) circle (  3.57);

\path[fill=fillColor] (109.66,251.96) circle (  3.57);

\path[fill=fillColor] (112.02,223.96) circle (  3.57);

\path[fill=fillColor] (126.59,195.95) circle (  3.57);

\path[fill=fillColor] (106.34,167.95) circle (  3.57);

\path[fill=fillColor] (123.48, 83.94) circle (  3.57);
\definecolor{drawColor}{RGB}{169,169,169}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 78.32,153.95) -- (137.25,153.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 78.32,125.95) -- (137.25,125.95);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 78.32, 67.14) rectangle (137.25,464.78);
\end{scope}
\begin{scope}
\path[clip] (142.75, 67.14) rectangle (201.69,464.78);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (142.75, 67.14) rectangle (201.69,464.78);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] (152.13, 67.14) --
	(152.13,464.78);

\path[draw=drawColor,line width= 0.3pt,line join=round] (165.53, 67.14) --
	(165.53,464.78);

\path[draw=drawColor,line width= 0.3pt,line join=round] (178.92, 67.14) --
	(178.92,464.78);

\path[draw=drawColor,line width= 0.3pt,line join=round] (192.32, 67.14) --
	(192.32,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (142.75, 83.94) --
	(201.69, 83.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (142.75,111.94) --
	(201.69,111.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (142.75,139.95) --
	(201.69,139.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (142.75,167.95) --
	(201.69,167.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (142.75,195.95) --
	(201.69,195.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (142.75,223.96) --
	(201.69,223.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (142.75,251.96) --
	(201.69,251.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (142.75,279.96) --
	(201.69,279.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (142.75,307.96) --
	(201.69,307.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (142.75,335.97) --
	(201.69,335.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (142.75,363.97) --
	(201.69,363.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (142.75,391.97) --
	(201.69,391.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (142.75,419.98) --
	(201.69,419.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (142.75,447.98) --
	(201.69,447.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (145.43, 67.14) --
	(145.43,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (158.83, 67.14) --
	(158.83,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (172.22, 67.14) --
	(172.22,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (185.62, 67.14) --
	(185.62,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (199.01, 67.14) --
	(199.01,464.78);
\definecolor{drawColor}{RGB}{221,204,119}

\path[draw=drawColor,line width= 0.6pt,line join=round] (159.63,137.15) --
	(159.63,142.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (159.63,139.95) --
	(150.10,139.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (150.10,137.15) --
	(150.10,142.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (151.33,445.18) --
	(151.33,450.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (151.33,447.98) --
	(148.11,447.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (148.11,445.18) --
	(148.11,450.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (157.97,417.18) --
	(157.97,422.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (157.97,419.98) --
	(152.24,419.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (152.24,417.18) --
	(152.24,422.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (154.17,361.17) --
	(154.17,366.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (161.72,333.17) --
	(161.72,338.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (161.72,335.97) --
	(157.11,335.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (157.11,333.17) --
	(157.11,338.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (161.19,305.16) --
	(161.19,310.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (161.19,307.96) --
	(153.10,307.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (153.10,305.16) --
	(153.10,310.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (153.52,109.14) --
	(153.52,114.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] (153.52,111.94) --
	(152.29,111.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (152.29,109.14) --
	(152.29,114.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] (154.70,277.16) --
	(154.70,282.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (154.70,279.96) --
	(146.93,279.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (146.93,277.16) --
	(146.93,282.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (169.12,249.16) --
	(169.12,254.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (169.12,251.96) --
	(162.37,251.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (162.37,249.16) --
	(162.37,254.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (153.04,165.15) --
	(153.04,170.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (153.04,167.95) --
	(149.24,167.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (149.24,165.15) --
	(149.24,170.75);
\definecolor{fillColor}{RGB}{221,204,119}

\path[fill=fillColor] (154.86,139.95) circle (  3.57);

\path[fill=fillColor] (149.72,447.98) circle (  3.57);

\path[fill=fillColor] (155.13,419.98) circle (  3.57);

\path[fill=fillColor] (149.67,363.97) circle (  1.43);

\path[fill=fillColor] (159.42,335.97) circle (  3.57);

\path[fill=fillColor] (157.17,307.96) circle (  3.57);

\path[fill=fillColor] (152.88,111.94) circle (  3.57);

\path[fill=fillColor] (150.85,279.96) circle (  3.57);

\path[fill=fillColor] (165.74,251.96) circle (  3.57);

\path[fill=fillColor] (151.17,167.95) circle (  3.57);
\definecolor{drawColor}{RGB}{169,169,169}

\path[draw=drawColor,line width= 0.6pt,line join=round] (142.75,153.95) -- (201.69,153.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (142.75,125.95) -- (201.69,125.95);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (142.75, 67.14) rectangle (201.69,464.78);
\end{scope}
\begin{scope}
\path[clip] (207.19, 67.14) rectangle (266.13,464.78);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (207.19, 67.14) rectangle (266.13,464.78);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] (216.57, 67.14) --
	(216.57,464.78);

\path[draw=drawColor,line width= 0.3pt,line join=round] (229.96, 67.14) --
	(229.96,464.78);

\path[draw=drawColor,line width= 0.3pt,line join=round] (243.36, 67.14) --
	(243.36,464.78);

\path[draw=drawColor,line width= 0.3pt,line join=round] (256.76, 67.14) --
	(256.76,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (207.19, 83.94) --
	(266.13, 83.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (207.19,111.94) --
	(266.13,111.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (207.19,139.95) --
	(266.13,139.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (207.19,167.95) --
	(266.13,167.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (207.19,195.95) --
	(266.13,195.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (207.19,223.96) --
	(266.13,223.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (207.19,251.96) --
	(266.13,251.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (207.19,279.96) --
	(266.13,279.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (207.19,307.96) --
	(266.13,307.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (207.19,335.97) --
	(266.13,335.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (207.19,363.97) --
	(266.13,363.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (207.19,391.97) --
	(266.13,391.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (207.19,419.98) --
	(266.13,419.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (207.19,447.98) --
	(266.13,447.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (209.87, 67.14) --
	(209.87,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (223.27, 67.14) --
	(223.27,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (236.66, 67.14) --
	(236.66,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (250.06, 67.14) --
	(250.06,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (263.45, 67.14) --
	(263.45,464.78);
\definecolor{drawColor}{RGB}{17,119,51}

\path[draw=drawColor,line width= 0.6pt,line join=round] (219.95,137.15) --
	(219.95,142.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (219.95,139.95) --
	(213.30,139.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (213.30,137.15) --
	(213.30,142.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (215.34,445.18) --
	(215.34,450.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (215.34,447.98) --
	(212.28,447.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (212.28,445.18) --
	(212.28,450.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (216.30,417.18) --
	(216.30,422.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (216.30,419.98) --
	(211.96,419.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (211.96,417.18) --
	(211.96,422.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (213.14,389.17) --
	(213.14,394.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (228.14,361.17) --
	(228.14,366.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (228.14,363.97) --
	(213.46,363.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (213.46,361.17) --
	(213.46,366.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (220.27,333.17) --
	(220.27,338.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (220.27,335.97) --
	(216.46,335.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (216.46,333.17) --
	(216.46,338.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (230.18,305.16) --
	(230.18,310.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (230.18,307.96) --
	(221.07,307.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (221.07,305.16) --
	(221.07,310.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (213.14,109.14) --
	(213.14,114.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] (213.14,111.94) --
	(212.34,111.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (212.34,109.14) --
	(212.34,114.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] (224.93,277.16) --
	(224.93,282.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (224.93,279.96) --
	(214.86,279.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (214.86,277.16) --
	(214.86,282.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (220.16,249.16) --
	(220.16,254.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (220.16,251.96) --
	(215.28,251.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (215.28,249.16) --
	(215.28,254.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (213.03,193.15) --
	(213.03,198.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (213.03,195.95) --
	(210.78,195.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (210.78,193.15) --
	(210.78,198.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (213.78,165.15) --
	(213.78,170.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (213.78,167.95) --
	(211.32,167.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (211.32,165.15) --
	(211.32,170.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (233.98, 81.14) --
	(233.98, 86.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] (233.98, 83.94) --
	(227.82, 83.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (227.82, 81.14) --
	(227.82, 86.74);
\definecolor{fillColor}{RGB}{17,119,51}

\path[fill=fillColor] (216.62,139.95) circle (  3.57);

\path[fill=fillColor] (213.84,447.98) circle (  3.57);

\path[fill=fillColor] (214.16,419.98) circle (  3.57);

\path[fill=fillColor] (211.32,391.97) circle (  1.43);

\path[fill=fillColor] (220.80,363.97) circle (  3.57);

\path[fill=fillColor] (218.39,335.97) circle (  3.57);

\path[fill=fillColor] (225.62,307.96) circle (  3.57);

\path[fill=fillColor] (212.77,111.94) circle (  3.57);

\path[fill=fillColor] (219.89,279.96) circle (  3.57);

\path[fill=fillColor] (217.70,251.96) circle (  3.57);

\path[fill=fillColor] (211.91,195.95) circle (  3.57);

\path[fill=fillColor] (212.55,167.95) circle (  3.57);

\path[fill=fillColor] (230.93, 83.94) circle (  3.57);
\definecolor{drawColor}{RGB}{169,169,169}

\path[draw=drawColor,line width= 0.6pt,line join=round] (207.19,153.95) -- (266.13,153.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (207.19,125.95) -- (266.13,125.95);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (207.19, 67.14) rectangle (266.13,464.78);
\end{scope}
\begin{scope}
\path[clip] (271.63, 67.14) rectangle (330.57,464.78);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (271.63, 67.14) rectangle (330.57,464.78);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] (281.01, 67.14) --
	(281.01,464.78);

\path[draw=drawColor,line width= 0.3pt,line join=round] (294.40, 67.14) --
	(294.40,464.78);

\path[draw=drawColor,line width= 0.3pt,line join=round] (307.80, 67.14) --
	(307.80,464.78);

\path[draw=drawColor,line width= 0.3pt,line join=round] (321.19, 67.14) --
	(321.19,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (271.63, 83.94) --
	(330.57, 83.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (271.63,111.94) --
	(330.57,111.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (271.63,139.95) --
	(330.57,139.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (271.63,167.95) --
	(330.57,167.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (271.63,195.95) --
	(330.57,195.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (271.63,223.96) --
	(330.57,223.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (271.63,251.96) --
	(330.57,251.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (271.63,279.96) --
	(330.57,279.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (271.63,307.96) --
	(330.57,307.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (271.63,335.97) --
	(330.57,335.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (271.63,363.97) --
	(330.57,363.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (271.63,391.97) --
	(330.57,391.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (271.63,419.98) --
	(330.57,419.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (271.63,447.98) --
	(330.57,447.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (274.31, 67.14) --
	(274.31,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (287.71, 67.14) --
	(287.71,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (301.10, 67.14) --
	(301.10,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (314.50, 67.14) --
	(314.50,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (327.89, 67.14) --
	(327.89,464.78);
\definecolor{drawColor}{RGB}{51,34,136}

\path[draw=drawColor,line width= 0.6pt,line join=round] (287.55,137.15) --
	(287.55,142.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (287.55,139.95) --
	(281.70,139.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (281.70,137.15) --
	(281.70,142.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (287.06,445.18) --
	(287.06,450.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (287.06,447.98) --
	(282.40,447.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (282.40,445.18) --
	(282.40,450.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (282.08,417.18) --
	(282.08,422.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (282.08,419.98) --
	(277.47,419.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (277.47,417.18) --
	(277.47,422.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (296.17,389.17) --
	(296.17,394.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (296.17,391.97) --
	(284.28,391.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (284.28,389.17) --
	(284.28,394.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (288.88,361.17) --
	(288.88,366.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (288.88,363.97) --
	(276.03,363.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (276.03,361.17) --
	(276.03,366.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (280.79,333.17) --
	(280.79,338.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (280.79,335.97) --
	(277.74,335.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (277.74,333.17) --
	(277.74,338.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (292.53,305.16) --
	(292.53,310.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (292.53,307.96) --
	(283.96,307.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (283.96,305.16) --
	(283.96,310.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (290.97,109.14) --
	(290.97,114.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] (290.97,111.94) --
	(289.37,111.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (289.37,109.14) --
	(289.37,114.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] (291.13,277.16) --
	(291.13,282.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (291.13,279.96) --
	(280.53,279.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (280.53,277.16) --
	(280.53,282.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (280.15,249.16) --
	(280.15,254.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (280.15,251.96) --
	(276.56,251.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (276.56,249.16) --
	(276.56,254.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (286.47,221.16) --
	(286.47,226.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (286.47,223.96) --
	(280.69,223.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (280.69,221.16) --
	(280.69,226.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (289.96,193.15) --
	(289.96,198.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (289.96,195.95) --
	(284.60,195.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (284.60,193.15) --
	(284.60,198.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (293.71,165.15) --
	(293.71,170.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (293.71,167.95) --
	(288.19,167.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (288.19,165.15) --
	(288.19,170.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (302.49, 81.14) --
	(302.49, 86.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] (302.49, 83.94) --
	(296.28, 83.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (296.28, 81.14) --
	(296.28, 86.74);
\definecolor{fillColor}{RGB}{51,34,136}

\path[fill=fillColor] (284.60,139.95) circle (  3.57);

\path[fill=fillColor] (284.76,447.98) circle (  3.57);

\path[fill=fillColor] (279.78,419.98) circle (  3.57);

\path[fill=fillColor] (290.22,391.97) circle (  3.57);

\path[fill=fillColor] (282.45,363.97) circle (  3.57);

\path[fill=fillColor] (279.29,335.97) circle (  3.57);

\path[fill=fillColor] (288.24,307.96) circle (  3.57);

\path[fill=fillColor] (290.17,111.94) circle (  3.57);

\path[fill=fillColor] (285.83,279.96) circle (  3.57);

\path[fill=fillColor] (278.33,251.96) circle (  3.57);

\path[fill=fillColor] (283.58,223.96) circle (  3.57);

\path[fill=fillColor] (287.28,195.95) circle (  3.57);

\path[fill=fillColor] (290.92,167.95) circle (  3.57);

\path[fill=fillColor] (299.39, 83.94) circle (  3.57);
\definecolor{drawColor}{RGB}{169,169,169}

\path[draw=drawColor,line width= 0.6pt,line join=round] (271.63,153.95) -- (330.57,153.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (271.63,125.95) -- (330.57,125.95);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (271.63, 67.14) rectangle (330.57,464.78);
\end{scope}
\begin{scope}
\path[clip] (336.07, 67.14) rectangle (395.01,464.78);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (336.07, 67.14) rectangle (395.01,464.78);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] (345.45, 67.14) --
	(345.45,464.78);

\path[draw=drawColor,line width= 0.3pt,line join=round] (358.84, 67.14) --
	(358.84,464.78);

\path[draw=drawColor,line width= 0.3pt,line join=round] (372.24, 67.14) --
	(372.24,464.78);

\path[draw=drawColor,line width= 0.3pt,line join=round] (385.63, 67.14) --
	(385.63,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (336.07, 83.94) --
	(395.01, 83.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (336.07,111.94) --
	(395.01,111.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (336.07,139.95) --
	(395.01,139.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (336.07,167.95) --
	(395.01,167.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (336.07,195.95) --
	(395.01,195.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (336.07,223.96) --
	(395.01,223.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (336.07,251.96) --
	(395.01,251.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (336.07,279.96) --
	(395.01,279.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (336.07,307.96) --
	(395.01,307.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (336.07,335.97) --
	(395.01,335.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (336.07,363.97) --
	(395.01,363.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (336.07,391.97) --
	(395.01,391.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (336.07,419.98) --
	(395.01,419.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (336.07,447.98) --
	(395.01,447.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (338.75, 67.14) --
	(338.75,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (352.14, 67.14) --
	(352.14,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (365.54, 67.14) --
	(365.54,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (378.93, 67.14) --
	(378.93,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (392.33, 67.14) --
	(392.33,464.78);
\definecolor{drawColor}{RGB}{170,68,153}

\path[draw=drawColor,line width= 0.6pt,line join=round] (346.73,137.15) --
	(346.73,142.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (346.73,139.95) --
	(340.03,139.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (340.03,137.15) --
	(340.03,142.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (347.96,445.18) --
	(347.96,450.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (347.96,447.98) --
	(344.00,447.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (344.00,445.18) --
	(344.00,450.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (341.05,417.18) --
	(341.05,422.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (341.05,419.98) --
	(338.91,419.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (338.91,417.18) --
	(338.91,422.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (353.38,361.17) --
	(353.38,366.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (353.38,363.97) --
	(340.89,363.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (340.89,361.17) --
	(340.89,366.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (355.52,305.16) --
	(355.52,310.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (355.52,307.96) --
	(345.61,307.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (345.61,305.16) --
	(345.61,310.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (342.66,109.14) --
	(342.66,114.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] (342.66,111.94) --
	(341.75,111.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (341.75,109.14) --
	(341.75,114.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] (344.86,277.16) --
	(344.86,282.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (344.86,279.96) --
	(338.80,279.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (338.80,277.16) --
	(338.80,282.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (342.39,249.16) --
	(342.39,254.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (342.39,251.96) --
	(339.61,251.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (339.61,249.16) --
	(339.61,254.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (340.46,193.15) --
	(340.46,198.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (340.46,195.95) --
	(338.86,195.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (338.86,193.15) --
	(338.86,198.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (342.07,165.15) --
	(342.07,170.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (342.07,167.95) --
	(339.77,167.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (339.77,165.15) --
	(339.77,170.75);
\definecolor{fillColor}{RGB}{170,68,153}

\path[fill=fillColor] (343.41,139.95) circle (  3.57);

\path[fill=fillColor] (345.98,447.98) circle (  3.57);

\path[fill=fillColor] (339.98,419.98) circle (  1.43);

\path[fill=fillColor] (347.16,363.97) circle (  3.57);

\path[fill=fillColor] (350.59,307.96) circle (  3.57);

\path[fill=fillColor] (342.18,111.94) circle (  3.57);

\path[fill=fillColor] (341.86,279.96) circle (  1.43);

\path[fill=fillColor] (341.00,251.96) circle (  3.57);

\path[fill=fillColor] (339.66,195.95) circle (  1.43);

\path[fill=fillColor] (340.95,167.95) circle (  3.57);
\definecolor{drawColor}{RGB}{169,169,169}

\path[draw=drawColor,line width= 0.6pt,line join=round] (336.07,153.95) -- (395.01,153.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (336.07,125.95) -- (395.01,125.95);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (336.07, 67.14) rectangle (395.01,464.78);
\end{scope}
\begin{scope}
\path[clip] (400.51, 67.14) rectangle (459.45,464.78);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (400.51, 67.14) rectangle (459.45,464.78);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] (409.88, 67.14) --
	(409.88,464.78);

\path[draw=drawColor,line width= 0.3pt,line join=round] (423.28, 67.14) --
	(423.28,464.78);

\path[draw=drawColor,line width= 0.3pt,line join=round] (436.68, 67.14) --
	(436.68,464.78);

\path[draw=drawColor,line width= 0.3pt,line join=round] (450.07, 67.14) --
	(450.07,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (400.51, 83.94) --
	(459.45, 83.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (400.51,111.94) --
	(459.45,111.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (400.51,139.95) --
	(459.45,139.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (400.51,167.95) --
	(459.45,167.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (400.51,195.95) --
	(459.45,195.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (400.51,223.96) --
	(459.45,223.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (400.51,251.96) --
	(459.45,251.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (400.51,279.96) --
	(459.45,279.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (400.51,307.96) --
	(459.45,307.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (400.51,335.97) --
	(459.45,335.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (400.51,363.97) --
	(459.45,363.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (400.51,391.97) --
	(459.45,391.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (400.51,419.98) --
	(459.45,419.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (400.51,447.98) --
	(459.45,447.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (403.19, 67.14) --
	(403.19,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (416.58, 67.14) --
	(416.58,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (429.98, 67.14) --
	(429.98,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (443.37, 67.14) --
	(443.37,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (456.77, 67.14) --
	(456.77,464.78);
\definecolor{drawColor}{RGB}{68,170,153}

\path[draw=drawColor,line width= 0.6pt,line join=round] (406.83,137.15) --
	(406.83,142.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (406.83,139.95) --
	(404.10,139.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (404.10,137.15) --
	(404.10,142.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (406.19,445.18) --
	(406.19,450.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (406.19,447.98) --
	(403.99,447.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (403.99,445.18) --
	(403.99,450.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (404.15,417.18) --
	(404.15,422.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (408.87,361.17) --
	(408.87,366.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (412.51,305.16) --
	(412.51,310.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (412.51,307.96) --
	(406.13,307.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (406.13,305.16) --
	(406.13,310.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (405.49,109.14) --
	(405.49,114.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] (405.49,111.94) --
	(404.85,111.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (404.85,109.14) --
	(404.85,114.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] (410.26,277.16) --
	(410.26,282.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (410.26,279.96) --
	(403.67,279.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (403.67,277.16) --
	(403.67,282.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (405.97,249.16) --
	(405.97,254.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (405.97,251.96) --
	(403.62,251.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (403.62,249.16) --
	(403.62,254.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (407.37,221.16) --
	(407.37,226.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (407.37,223.96) --
	(404.53,223.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (404.53,221.16) --
	(404.53,226.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (404.90,193.15) --
	(404.90,198.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (404.90,195.95) --
	(403.29,195.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (403.29,193.15) --
	(403.29,198.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (404.74,165.15) --
	(404.74,170.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (404.74,167.95) --
	(403.29,167.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (403.29,165.15) --
	(403.29,170.75);
\definecolor{fillColor}{RGB}{68,170,153}

\path[fill=fillColor] (405.49,139.95) circle (  3.57);

\path[fill=fillColor] (405.06,447.98) circle (  3.57);

\path[fill=fillColor] (403.51,419.98) circle (  1.43);

\path[fill=fillColor] (405.54,363.97) circle (  1.43);

\path[fill=fillColor] (409.35,307.96) circle (  3.57);

\path[fill=fillColor] (405.17,111.94) circle (  3.57);

\path[fill=fillColor] (406.94,279.96) circle (  1.43);

\path[fill=fillColor] (404.79,251.96) circle (  3.57);

\path[fill=fillColor] (405.92,223.96) circle (  3.57);

\path[fill=fillColor] (404.10,195.95) circle (  1.43);

\path[fill=fillColor] (404.04,167.95) circle (  3.57);
\definecolor{drawColor}{RGB}{169,169,169}

\path[draw=drawColor,line width= 0.6pt,line join=round] (400.51,153.95) -- (459.45,153.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (400.51,125.95) -- (459.45,125.95);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (400.51, 67.14) rectangle (459.45,464.78);
\end{scope}
\begin{scope}
\path[clip] (464.95, 67.14) rectangle (523.88,464.78);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (464.95, 67.14) rectangle (523.88,464.78);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] (474.32, 67.14) --
	(474.32,464.78);

\path[draw=drawColor,line width= 0.3pt,line join=round] (487.72, 67.14) --
	(487.72,464.78);

\path[draw=drawColor,line width= 0.3pt,line join=round] (501.11, 67.14) --
	(501.11,464.78);

\path[draw=drawColor,line width= 0.3pt,line join=round] (514.51, 67.14) --
	(514.51,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (464.95, 83.94) --
	(523.88, 83.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (464.95,111.94) --
	(523.88,111.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (464.95,139.95) --
	(523.88,139.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (464.95,167.95) --
	(523.88,167.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (464.95,195.95) --
	(523.88,195.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (464.95,223.96) --
	(523.88,223.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (464.95,251.96) --
	(523.88,251.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (464.95,279.96) --
	(523.88,279.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (464.95,307.96) --
	(523.88,307.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (464.95,335.97) --
	(523.88,335.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (464.95,363.97) --
	(523.88,363.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (464.95,391.97) --
	(523.88,391.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (464.95,419.98) --
	(523.88,419.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (464.95,447.98) --
	(523.88,447.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (467.63, 67.14) --
	(467.63,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (481.02, 67.14) --
	(481.02,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (494.42, 67.14) --
	(494.42,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (507.81, 67.14) --
	(507.81,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (521.21, 67.14) --
	(521.21,464.78);
\definecolor{drawColor}{RGB}{153,153,51}

\path[draw=drawColor,line width= 0.6pt,line join=round] (469.39,137.15) --
	(469.39,142.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (469.13,445.18) --
	(469.13,450.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (469.13,447.98) --
	(467.73,447.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (467.73,445.18) --
	(467.73,450.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (468.32,417.18) --
	(468.32,422.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (470.57,361.17) --
	(470.57,366.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (468.00,333.17) --
	(468.00,338.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (473.63,109.14) --
	(473.63,114.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] (473.63,111.94) --
	(472.56,111.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (472.56,109.14) --
	(472.56,114.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] (473.57,277.16) --
	(473.57,282.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (473.57,279.96) --
	(467.68,279.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (467.68,277.16) --
	(467.68,282.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (468.80,249.16) --
	(468.80,254.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (468.11,165.15) --
	(468.11,170.75);
\definecolor{fillColor}{RGB}{153,153,51}

\path[fill=fillColor] (468.43,139.95) circle (  3.57);

\path[fill=fillColor] (468.43,447.98) circle (  1.43);

\path[fill=fillColor] (467.84,419.98) circle (  1.43);

\path[fill=fillColor] (468.59,363.97) circle (  1.43);

\path[fill=fillColor] (467.73,335.97) circle (  1.43);

\path[fill=fillColor] (473.09,111.94) circle (  3.57);

\path[fill=fillColor] (470.63,279.96) circle (  1.43);

\path[fill=fillColor] (468.11,251.96) circle (  1.43);

\path[fill=fillColor] (467.79,167.95) circle (  1.43);
\definecolor{drawColor}{RGB}{169,169,169}

\path[draw=drawColor,line width= 0.6pt,line join=round] (464.95,153.95) -- (523.88,153.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (464.95,125.95) -- (523.88,125.95);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (464.95, 67.14) rectangle (523.88,464.78);
\end{scope}
\begin{scope}
\path[clip] (529.38, 67.14) rectangle (588.32,464.78);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (529.38, 67.14) rectangle (588.32,464.78);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] (538.76, 67.14) --
	(538.76,464.78);

\path[draw=drawColor,line width= 0.3pt,line join=round] (552.16, 67.14) --
	(552.16,464.78);

\path[draw=drawColor,line width= 0.3pt,line join=round] (565.55, 67.14) --
	(565.55,464.78);

\path[draw=drawColor,line width= 0.3pt,line join=round] (578.95, 67.14) --
	(578.95,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (529.38, 83.94) --
	(588.32, 83.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (529.38,111.94) --
	(588.32,111.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (529.38,139.95) --
	(588.32,139.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (529.38,167.95) --
	(588.32,167.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (529.38,195.95) --
	(588.32,195.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (529.38,223.96) --
	(588.32,223.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (529.38,251.96) --
	(588.32,251.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (529.38,279.96) --
	(588.32,279.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (529.38,307.96) --
	(588.32,307.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (529.38,335.97) --
	(588.32,335.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (529.38,363.97) --
	(588.32,363.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (529.38,391.97) --
	(588.32,391.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (529.38,419.98) --
	(588.32,419.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (529.38,447.98) --
	(588.32,447.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (532.06, 67.14) --
	(532.06,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (545.46, 67.14) --
	(545.46,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (558.85, 67.14) --
	(558.85,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (572.25, 67.14) --
	(572.25,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (585.64, 67.14) --
	(585.64,464.78);
\definecolor{drawColor}{RGB}{136,34,85}

\path[draw=drawColor,line width= 0.6pt,line join=round] (533.78,137.15) --
	(533.78,142.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (533.08,445.18) --
	(533.08,450.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (532.87,417.18) --
	(532.87,422.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (532.81,109.14) --
	(532.81,114.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] (532.81,111.94) --
	(532.44,111.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (532.44,109.14) --
	(532.44,114.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] (535.55,249.16) --
	(535.55,254.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (535.55,251.96) --
	(532.81,251.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (532.81,249.16) --
	(532.81,254.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (533.24,193.15) --
	(533.24,198.75);
\definecolor{fillColor}{RGB}{136,34,85}

\path[fill=fillColor] (532.76,139.95) circle (  3.57);

\path[fill=fillColor] (532.55,447.98) circle (  1.43);

\path[fill=fillColor] (532.33,419.98) circle (  1.43);

\path[fill=fillColor] (532.60,111.94) circle (  3.57);

\path[fill=fillColor] (534.21,251.96) circle (  3.57);

\path[fill=fillColor] (532.60,195.95) circle (  1.43);
\definecolor{drawColor}{RGB}{169,169,169}

\path[draw=drawColor,line width= 0.6pt,line join=round] (529.38,153.95) -- (588.32,153.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (529.38,125.95) -- (588.32,125.95);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (529.38, 67.14) rectangle (588.32,464.78);
\end{scope}
\begin{scope}
\path[clip] (593.82, 67.14) rectangle (652.76,464.78);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (593.82, 67.14) rectangle (652.76,464.78);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] (603.20, 67.14) --
	(603.20,464.78);

\path[draw=drawColor,line width= 0.3pt,line join=round] (616.59, 67.14) --
	(616.59,464.78);

\path[draw=drawColor,line width= 0.3pt,line join=round] (629.99, 67.14) --
	(629.99,464.78);

\path[draw=drawColor,line width= 0.3pt,line join=round] (643.39, 67.14) --
	(643.39,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (593.82, 83.94) --
	(652.76, 83.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (593.82,111.94) --
	(652.76,111.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (593.82,139.95) --
	(652.76,139.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (593.82,167.95) --
	(652.76,167.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (593.82,195.95) --
	(652.76,195.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (593.82,223.96) --
	(652.76,223.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (593.82,251.96) --
	(652.76,251.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (593.82,279.96) --
	(652.76,279.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (593.82,307.96) --
	(652.76,307.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (593.82,335.97) --
	(652.76,335.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (593.82,363.97) --
	(652.76,363.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (593.82,391.97) --
	(652.76,391.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (593.82,419.98) --
	(652.76,419.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (593.82,447.98) --
	(652.76,447.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (596.50, 67.14) --
	(596.50,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (609.90, 67.14) --
	(609.90,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (623.29, 67.14) --
	(623.29,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (636.69, 67.14) --
	(636.69,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (650.08, 67.14) --
	(650.08,464.78);
\definecolor{drawColor}{RGB}{102,17,0}

\path[draw=drawColor,line width= 0.6pt,line join=round] (605.61,137.15) --
	(605.61,142.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (605.61,139.95) --
	(599.77,139.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (599.77,137.15) --
	(599.77,142.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (608.34,445.18) --
	(608.34,450.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (608.34,447.98) --
	(603.84,447.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (603.84,445.18) --
	(603.84,450.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (604.11,417.18) --
	(604.11,422.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (604.11,419.98) --
	(599.61,419.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (599.61,417.18) --
	(599.61,422.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (600.36,361.17) --
	(600.36,366.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (600.25,333.17) --
	(600.25,338.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (600.25,335.97) --
	(598.00,335.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (598.00,333.17) --
	(598.00,338.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (606.90,305.16) --
	(606.90,310.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (606.90,307.96) --
	(600.31,307.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (600.31,305.16) --
	(600.31,310.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (608.72,109.14) --
	(608.72,114.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] (608.72,111.94) --
	(607.27,111.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (607.27,109.14) --
	(607.27,114.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] (613.17,277.16) --
	(613.17,282.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (613.17,279.96) --
	(602.66,279.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (602.66,277.16) --
	(602.66,282.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (610.17,249.16) --
	(610.17,254.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (610.17,251.96) --
	(604.59,251.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (604.59,249.16) --
	(604.59,254.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (599.56,221.16) --
	(599.56,226.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (599.56,223.96) --
	(597.20,223.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (597.20,221.16) --
	(597.20,226.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (603.84,165.15) --
	(603.84,170.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (603.84,167.95) --
	(600.25,167.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (600.25,165.15) --
	(600.25,170.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (601.16, 81.14) --
	(601.16, 86.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] (601.16, 83.94) --
	(598.32, 83.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (598.32, 81.14) --
	(598.32, 86.74);
\definecolor{fillColor}{RGB}{102,17,0}

\path[fill=fillColor] (602.72,139.95) circle (  3.57);

\path[fill=fillColor] (606.09,447.98) circle (  3.57);

\path[fill=fillColor] (601.86,419.98) circle (  3.57);

\path[fill=fillColor] (598.00,363.97) circle (  1.43);

\path[fill=fillColor] (599.13,335.97) circle (  3.57);

\path[fill=fillColor] (603.57,307.96) circle (  3.57);

\path[fill=fillColor] (607.97,111.94) circle (  3.57);

\path[fill=fillColor] (607.91,279.96) circle (  3.57);

\path[fill=fillColor] (607.38,251.96) circle (  3.57);

\path[fill=fillColor] (598.38,223.96) circle (  3.57);

\path[fill=fillColor] (602.02,167.95) circle (  3.57);

\path[fill=fillColor] (599.72, 83.94) circle (  3.57);
\definecolor{drawColor}{RGB}{169,169,169}

\path[draw=drawColor,line width= 0.6pt,line join=round] (593.82,153.95) -- (652.76,153.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (593.82,125.95) -- (652.76,125.95);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (593.82, 67.14) rectangle (652.76,464.78);
\end{scope}
\begin{scope}
\path[clip] (658.26, 67.14) rectangle (717.20,464.78);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (658.26, 67.14) rectangle (717.20,464.78);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] (667.64, 67.14) --
	(667.64,464.78);

\path[draw=drawColor,line width= 0.3pt,line join=round] (681.03, 67.14) --
	(681.03,464.78);

\path[draw=drawColor,line width= 0.3pt,line join=round] (694.43, 67.14) --
	(694.43,464.78);

\path[draw=drawColor,line width= 0.3pt,line join=round] (707.82, 67.14) --
	(707.82,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (658.26, 83.94) --
	(717.20, 83.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (658.26,111.94) --
	(717.20,111.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (658.26,139.95) --
	(717.20,139.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (658.26,167.95) --
	(717.20,167.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (658.26,195.95) --
	(717.20,195.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (658.26,223.96) --
	(717.20,223.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (658.26,251.96) --
	(717.20,251.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (658.26,279.96) --
	(717.20,279.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (658.26,307.96) --
	(717.20,307.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (658.26,335.97) --
	(717.20,335.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (658.26,363.97) --
	(717.20,363.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (658.26,391.97) --
	(717.20,391.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (658.26,419.98) --
	(717.20,419.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (658.26,447.98) --
	(717.20,447.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (660.94, 67.14) --
	(660.94,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (674.34, 67.14) --
	(674.34,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (687.73, 67.14) --
	(687.73,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (701.13, 67.14) --
	(701.13,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (714.52, 67.14) --
	(714.52,464.78);
\definecolor{drawColor}{RGB}{102,153,204}

\path[draw=drawColor,line width= 0.6pt,line join=round] (673.64,137.15) --
	(673.64,142.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (673.64,139.95) --
	(666.83,139.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (666.83,137.15) --
	(666.83,142.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (667.26,445.18) --
	(667.26,450.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (667.26,447.98) --
	(663.94,447.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (663.94,445.18) --
	(663.94,450.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (681.41,417.18) --
	(681.41,422.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (681.41,419.98) --
	(674.39,419.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (674.39,417.18) --
	(674.39,422.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (677.55,389.17) --
	(677.55,394.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (677.55,391.97) --
	(667.48,391.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (667.48,389.17) --
	(667.48,394.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (673.53,361.17) --
	(673.53,366.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (673.53,363.97) --
	(661.37,363.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (661.37,361.17) --
	(661.37,366.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (677.60,333.17) --
	(677.60,338.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (677.60,335.97) --
	(673.00,335.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (673.00,333.17) --
	(673.00,338.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (676.37,305.16) --
	(676.37,310.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (676.37,307.96) --
	(666.51,307.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (666.51,305.16) --
	(666.51,310.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (664.05,109.14) --
	(664.05,114.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] (664.05,111.94) --
	(663.30,111.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (663.30,109.14) --
	(663.30,114.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] (680.39,277.16) --
	(680.39,282.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (680.39,279.96) --
	(669.14,279.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (669.14,277.16) --
	(669.14,282.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (665.60,249.16) --
	(665.60,254.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (665.60,251.96) --
	(662.44,251.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (662.44,249.16) --
	(662.44,254.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (677.18,221.16) --
	(677.18,226.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (677.18,223.96) --
	(671.28,223.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (671.28,221.16) --
	(671.28,226.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (666.83,193.15) --
	(666.83,198.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (666.83,195.95) --
	(663.57,195.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (663.57,193.15) --
	(663.57,198.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (665.50,165.15) --
	(665.50,170.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (665.50,167.95) --
	(662.76,167.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (662.76,165.15) --
	(662.76,170.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (690.36, 81.14) --
	(690.36, 86.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] (690.36, 83.94) --
	(684.14, 83.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (684.14, 81.14) --
	(684.14, 86.74);
\definecolor{fillColor}{RGB}{102,153,204}

\path[fill=fillColor] (670.21,139.95) circle (  3.57);

\path[fill=fillColor] (665.60,447.98) circle (  3.57);

\path[fill=fillColor] (677.87,419.98) circle (  3.57);

\path[fill=fillColor] (672.51,391.97) circle (  3.57);

\path[fill=fillColor] (667.42,363.97) circle (  3.57);

\path[fill=fillColor] (675.30,335.97) circle (  3.57);

\path[fill=fillColor] (671.44,307.96) circle (  3.57);

\path[fill=fillColor] (663.67,111.94) circle (  3.57);

\path[fill=fillColor] (674.76,279.96) circle (  3.57);

\path[fill=fillColor] (663.99,251.96) circle (  3.57);

\path[fill=fillColor] (674.23,223.96) circle (  3.57);

\path[fill=fillColor] (665.23,195.95) circle (  3.57);

\path[fill=fillColor] (664.16,167.95) circle (  3.57);

\path[fill=fillColor] (687.25, 83.94) circle (  3.57);
\definecolor{drawColor}{RGB}{169,169,169}

\path[draw=drawColor,line width= 0.6pt,line join=round] (658.26,153.95) -- (717.20,153.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (658.26,125.95) -- (717.20,125.95);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (658.26, 67.14) rectangle (717.20,464.78);
\end{scope}
\begin{scope}
\path[clip] ( 78.32,464.78) rectangle (137.25,519.37);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 78.32,464.78) rectangle (137.25,519.37);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (107.79,498.55) {Concerned};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (107.79,489.04) {about side};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (107.79,479.54) {effects};
\end{scope}
\begin{scope}
\path[clip] (142.75,464.78) rectangle (201.69,519.37);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (142.75,464.78) rectangle (201.69,519.37);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (172.22,508.05) {Concerned};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (172.22,498.55) {about getting};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (172.22,489.04) {coronavirus};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (172.22,479.54) {from the};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (172.22,470.04) {vaccine};
\end{scope}
\begin{scope}
\path[clip] (207.19,464.78) rectangle (266.13,519.37);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (207.19,464.78) rectangle (266.13,519.37);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (236.66,498.55) {Not concerned};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (236.66,489.04) {about getting};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (236.66,479.54) {seriously ill};
\end{scope}
\begin{scope}
\path[clip] (271.63,464.78) rectangle (330.57,519.37);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (271.63,464.78) rectangle (330.57,519.37);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (301.10,498.55) {Doesn't think};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (301.10,489.04) {vaccines are};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (301.10,479.54) {effective};
\end{scope}
\begin{scope}
\path[clip] (336.07,464.78) rectangle (395.01,519.37);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (336.07,464.78) rectangle (395.01,519.37);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (365.54,508.05) {Doesn't think};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (365.54,498.55) {Coronavirus};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (365.54,489.04) {outbreak is as};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (365.54,479.54) {serious as};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (365.54,470.04) {people say};
\end{scope}
\begin{scope}
\path[clip] (400.51,464.78) rectangle (459.45,519.37);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (400.51,464.78) rectangle (459.45,519.37);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (429.98,493.80) {Doesn't like};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (429.98,484.29) {needles};
\end{scope}
\begin{scope}
\path[clip] (464.95,464.78) rectangle (523.88,519.37);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (464.95,464.78) rectangle (523.88,519.37);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (494.42,493.80) {Allergic to};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (494.42,484.29) {vaccines};
\end{scope}
\begin{scope}
\path[clip] (529.38,464.78) rectangle (588.32,519.37);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (529.38,464.78) rectangle (588.32,519.37);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (558.85,498.55) {Won't have};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (558.85,489.04) {time to get};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (558.85,479.54) {vaccinated};
\end{scope}
\begin{scope}
\path[clip] (593.82,464.78) rectangle (652.76,519.37);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (593.82,464.78) rectangle (652.76,519.37);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (623.29,498.55) {Mentions a};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (623.29,489.04) {conspiracy};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (623.29,479.54) {theory};
\end{scope}
\begin{scope}
\path[clip] (658.26,464.78) rectangle (717.20,519.37);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (658.26,464.78) rectangle (717.20,519.37);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (687.73,489.04) {Other reasons};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 81.00, 64.39) --
	( 81.00, 67.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 94.39, 64.39) --
	( 94.39, 67.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (107.79, 64.39) --
	(107.79, 67.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (121.18, 64.39) --
	(121.18, 67.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (134.58, 64.39) --
	(134.58, 67.14);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 81.00, 56.13) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 94.39, 56.13) {25};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (107.79, 56.13) {50};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (121.18, 56.13) {75};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (134.58, 56.13) {100};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (145.43, 64.39) --
	(145.43, 67.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (158.83, 64.39) --
	(158.83, 67.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (172.22, 64.39) --
	(172.22, 67.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (185.62, 64.39) --
	(185.62, 67.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (199.01, 64.39) --
	(199.01, 67.14);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (145.43, 56.13) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (158.83, 56.13) {25};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (172.22, 56.13) {50};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (185.62, 56.13) {75};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (199.01, 56.13) {100};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (209.87, 64.39) --
	(209.87, 67.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (223.27, 64.39) --
	(223.27, 67.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (236.66, 64.39) --
	(236.66, 67.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (250.06, 64.39) --
	(250.06, 67.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (263.45, 64.39) --
	(263.45, 67.14);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (209.87, 56.13) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (223.27, 56.13) {25};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (236.66, 56.13) {50};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (250.06, 56.13) {75};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (263.45, 56.13) {100};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (274.31, 64.39) --
	(274.31, 67.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (287.71, 64.39) --
	(287.71, 67.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (301.10, 64.39) --
	(301.10, 67.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (314.50, 64.39) --
	(314.50, 67.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (327.89, 64.39) --
	(327.89, 67.14);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (274.31, 56.13) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (287.71, 56.13) {25};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (301.10, 56.13) {50};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (314.50, 56.13) {75};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (327.89, 56.13) {100};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (338.75, 64.39) --
	(338.75, 67.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (352.14, 64.39) --
	(352.14, 67.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (365.54, 64.39) --
	(365.54, 67.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (378.93, 64.39) --
	(378.93, 67.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (392.33, 64.39) --
	(392.33, 67.14);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (338.75, 56.13) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (352.14, 56.13) {25};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (365.54, 56.13) {50};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (378.93, 56.13) {75};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (392.33, 56.13) {100};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (403.19, 64.39) --
	(403.19, 67.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (416.58, 64.39) --
	(416.58, 67.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (429.98, 64.39) --
	(429.98, 67.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (443.37, 64.39) --
	(443.37, 67.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (456.77, 64.39) --
	(456.77, 67.14);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (403.19, 56.13) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (416.58, 56.13) {25};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (429.98, 56.13) {50};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (443.37, 56.13) {75};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (456.77, 56.13) {100};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (467.63, 64.39) --
	(467.63, 67.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (481.02, 64.39) --
	(481.02, 67.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (494.42, 64.39) --
	(494.42, 67.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (507.81, 64.39) --
	(507.81, 67.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (521.21, 64.39) --
	(521.21, 67.14);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (467.63, 56.13) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (481.02, 56.13) {25};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (494.42, 56.13) {50};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (507.81, 56.13) {75};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (521.21, 56.13) {100};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (532.06, 64.39) --
	(532.06, 67.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (545.46, 64.39) --
	(545.46, 67.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (558.85, 64.39) --
	(558.85, 67.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (572.25, 64.39) --
	(572.25, 67.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (585.64, 64.39) --
	(585.64, 67.14);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (532.06, 56.13) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (545.46, 56.13) {25};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (558.85, 56.13) {50};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (572.25, 56.13) {75};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (585.64, 56.13) {100};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (596.50, 64.39) --
	(596.50, 67.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (609.90, 64.39) --
	(609.90, 67.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (623.29, 64.39) --
	(623.29, 67.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (636.69, 64.39) --
	(636.69, 67.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (650.08, 64.39) --
	(650.08, 67.14);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (596.50, 56.13) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (609.90, 56.13) {25};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (623.29, 56.13) {50};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (636.69, 56.13) {75};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (650.08, 56.13) {100};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (660.94, 64.39) --
	(660.94, 67.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (674.34, 64.39) --
	(674.34, 67.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (687.73, 64.39) --
	(687.73, 67.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (701.13, 64.39) --
	(701.13, 67.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (714.52, 64.39) --
	(714.52, 67.14);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (660.94, 56.13) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (674.34, 56.13) {25};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (687.73, 56.13) {50};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (701.13, 56.13) {75};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (714.52, 56.13) {100};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96, 80.91) {USA};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,108.91) {Russia};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,136.92) {All LMICs};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,164.92) {Uganda 2};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,192.92) {Uganda 1};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,220.93) {Sierra Leone 2};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,248.93) {Sierra Leone 1};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,276.93) {Rwanda};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,304.93) {Pakistan 1};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,332.94) {Nigeria};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,360.94) {Nepal};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,388.94) {Mozambique};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,416.95) {Colombia};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,444.95) {Burkina Faso};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 75.57, 83.94) --
	( 78.32, 83.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 75.57,111.94) --
	( 78.32,111.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 75.57,139.95) --
	( 78.32,139.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 75.57,167.95) --
	( 78.32,167.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 75.57,195.95) --
	( 78.32,195.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 75.57,223.96) --
	( 78.32,223.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 75.57,251.96) --
	( 78.32,251.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 75.57,279.96) --
	( 78.32,279.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 75.57,307.96) --
	( 78.32,307.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 75.57,335.97) --
	( 78.32,335.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 75.57,363.97) --
	( 78.32,363.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 75.57,391.97) --
	( 78.32,391.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 75.57,419.98) --
	( 78.32,419.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 75.57,447.98) --
	( 78.32,447.98);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (397.76, 44.09) {estimate};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (284.57,  5.50) rectangle (510.95, 30.95);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.10] at (290.07, 14.44) {Number of observations};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (409.70, 11.00) rectangle (424.15, 25.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] (416.92, 18.23) circle (  1.43);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (465.45, 11.00) rectangle (479.91, 25.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] (472.68, 18.23) circle (  3.57);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (429.65, 15.20) {[50,500)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (485.41, 15.20) {500+};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.32] at (  5.50,527.43) {Why would you not take the COVID-19 vaccine?};
\end{scope}
\end{tikzpicture}

\end{document}