### 2021-03-23 paper_files/figure-latex/fig3paper-1.tex

% Created by tikzDevice version 0.12.3.1 on 2021-03-23 16:52:37
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}

\nonstopmode

\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (722.70,650.43);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.00) rectangle (722.70,650.43);
\end{scope}
\begin{scope}
\path[clip] ( 48.28,547.64) rectangle (371.71,627.77);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 48.28,547.64) rectangle (371.71,627.77);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 48.28,560.96) --
	(371.71,560.96);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 48.28,580.31) --
	(371.71,580.31);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 48.28,599.67) --
	(371.71,599.67);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 48.28,619.02) --
	(371.71,619.02);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 48.28,551.28) --
	(371.71,551.28);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 48.28,570.64) --
	(371.71,570.64);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 48.28,589.99) --
	(371.71,589.99);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 48.28,609.34) --
	(371.71,609.34);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 75.24,547.64) --
	( 75.24,627.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (120.16,547.64) --
	(120.16,627.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (165.08,547.64) --
	(165.08,627.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (209.99,547.64) --
	(209.99,627.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (254.91,547.64) --
	(254.91,627.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (299.83,547.64) --
	(299.83,627.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (344.75,547.64) --
	(344.75,627.77);
\definecolor{fillColor}{RGB}{221,204,119}

\path[fill=fillColor] (171.81,551.28) rectangle (185.29,565.53);

\path[fill=fillColor] (261.65,551.28) rectangle (275.13,552.76);

\path[fill=fillColor] (216.73,551.28) rectangle (230.21,552.68);

\path[fill=fillColor] (306.57,551.28) rectangle (320.05,560.26);

\path[fill=fillColor] (351.49,551.28) rectangle (364.97,552.60);

\path[fill=fillColor] ( 81.97,551.28) rectangle ( 95.45,591.23);

\path[fill=fillColor] (126.89,551.28) rectangle (140.37,561.27);
\definecolor{fillColor}{RGB}{17,119,51}

\path[fill=fillColor] (158.34,551.28) rectangle (171.81,566.46);

\path[fill=fillColor] (248.18,551.28) rectangle (261.65,552.83);

\path[fill=fillColor] (203.26,551.28) rectangle (216.73,551.98);

\path[fill=fillColor] (293.10,551.28) rectangle (306.57,555.00);

\path[fill=fillColor] (338.02,551.28) rectangle (351.49,551.59);

\path[fill=fillColor] ( 68.50,551.28) rectangle ( 81.97,595.49);

\path[fill=fillColor] (113.42,551.28) rectangle (126.89,562.97);
\definecolor{fillColor}{RGB}{204,102,119}

\path[fill=fillColor] (144.86,551.28) rectangle (158.34,563.83);

\path[fill=fillColor] (234.70,551.28) rectangle (248.18,552.52);

\path[fill=fillColor] (189.78,551.28) rectangle (203.26,554.15);

\path[fill=fillColor] (279.62,551.28) rectangle (293.10,570.72);

\path[fill=fillColor] (324.54,551.28) rectangle (338.02,554.54);

\path[fill=fillColor] ( 55.02,551.28) rectangle ( 68.50,582.79);

\path[fill=fillColor] ( 99.94,551.28) rectangle (113.42,557.86);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 48.28,547.64) rectangle (371.71,627.77);
\end{scope}
\begin{scope}
\path[clip] ( 48.28,462.01) rectangle (371.71,542.14);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 48.28,462.01) rectangle (371.71,542.14);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 48.28,475.33) --
	(371.71,475.33);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 48.28,494.68) --
	(371.71,494.68);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 48.28,514.04) --
	(371.71,514.04);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 48.28,533.39) --
	(371.71,533.39);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 48.28,465.65) --
	(371.71,465.65);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 48.28,485.01) --
	(371.71,485.01);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 48.28,504.36) --
	(371.71,504.36);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 48.28,523.71) --
	(371.71,523.71);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 75.24,462.01) --
	( 75.24,542.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (120.16,462.01) --
	(120.16,542.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (165.08,462.01) --
	(165.08,542.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (209.99,462.01) --
	(209.99,542.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (254.91,462.01) --
	(254.91,542.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (299.83,462.01) --
	(299.83,542.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (344.75,462.01) --
	(344.75,542.14);
\definecolor{fillColor}{RGB}{221,204,119}

\path[fill=fillColor] (171.81,465.65) rectangle (185.29,493.21);

\path[fill=fillColor] (254.91,465.65) rectangle (275.13,465.96);

\path[fill=fillColor] (216.73,465.65) rectangle (230.21,478.12);

\path[fill=fillColor] (306.57,465.65) rectangle (320.05,466.04);

\path[fill=fillColor] (351.49,465.65) rectangle (364.97,467.51);

\path[fill=fillColor] ( 81.97,465.65) rectangle ( 95.45,499.87);

\path[fill=fillColor] (126.89,465.65) rectangle (140.37,466.27);
\definecolor{fillColor}{RGB}{17,119,51}

\path[fill=fillColor] (158.34,465.65) rectangle (171.81,493.68);

\path[fill=fillColor] (234.70,465.65) rectangle (254.91,465.96);

\path[fill=fillColor] (203.26,465.65) rectangle (216.73,478.12);

\path[fill=fillColor] (293.10,465.65) rectangle (306.57,466.04);

\path[fill=fillColor] (338.02,465.65) rectangle (351.49,466.66);

\path[fill=fillColor] ( 68.50,465.65) rectangle ( 81.97,500.26);

\path[fill=fillColor] (113.42,465.65) rectangle (126.89,466.20);
\definecolor{fillColor}{RGB}{204,102,119}

\path[fill=fillColor] (144.86,465.65) rectangle (158.34,480.13);

\path[fill=fillColor] (189.78,465.65) rectangle (203.26,478.66);

\path[fill=fillColor] (279.62,465.65) rectangle (293.10,466.43);

\path[fill=fillColor] (324.54,465.65) rectangle (338.02,489.81);

\path[fill=fillColor] ( 55.02,465.65) rectangle ( 68.50,489.03);

\path[fill=fillColor] ( 99.94,465.65) rectangle (113.42,467.28);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 48.28,462.01) rectangle (371.71,542.14);
\end{scope}
\begin{scope}
\path[clip] ( 48.28,376.38) rectangle (371.71,456.51);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 48.28,376.38) rectangle (371.71,456.51);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 48.28,389.70) --
	(371.71,389.70);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 48.28,409.05) --
	(371.71,409.05);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 48.28,428.41) --
	(371.71,428.41);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 48.28,447.76) --
	(371.71,447.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 48.28,380.02) --
	(371.71,380.02);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 48.28,399.38) --
	(371.71,399.38);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 48.28,418.73) --
	(371.71,418.73);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 48.28,438.08) --
	(371.71,438.08);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 75.24,376.38) --
	( 75.24,456.51);

\path[draw=drawColor,line width= 0.6pt,line join=round] (120.16,376.38) --
	(120.16,456.51);

\path[draw=drawColor,line width= 0.6pt,line join=round] (165.08,376.38) --
	(165.08,456.51);

\path[draw=drawColor,line width= 0.6pt,line join=round] (209.99,376.38) --
	(209.99,456.51);

\path[draw=drawColor,line width= 0.6pt,line join=round] (254.91,376.38) --
	(254.91,456.51);

\path[draw=drawColor,line width= 0.6pt,line join=round] (299.83,376.38) --
	(299.83,456.51);

\path[draw=drawColor,line width= 0.6pt,line join=round] (344.75,376.38) --
	(344.75,456.51);
\definecolor{fillColor}{RGB}{221,204,119}

\path[fill=fillColor] (171.81,380.02) rectangle (185.29,391.63);

\path[fill=fillColor] (254.91,380.02) rectangle (275.13,380.49);

\path[fill=fillColor] (216.73,380.02) rectangle (230.21,380.87);

\path[fill=fillColor] (306.57,380.02) rectangle (320.05,405.72);

\path[fill=fillColor] (351.49,380.02) rectangle (364.97,380.33);

\path[fill=fillColor] ( 81.97,380.02) rectangle ( 95.45,397.83);

\path[fill=fillColor] (126.89,380.02) rectangle (140.37,400.69);
\definecolor{fillColor}{RGB}{17,119,51}

\path[fill=fillColor] (158.34,380.02) rectangle (171.81,391.71);

\path[fill=fillColor] (234.70,380.02) rectangle (254.91,380.56);

\path[fill=fillColor] (203.26,380.02) rectangle (216.73,380.80);

\path[fill=fillColor] (293.10,380.02) rectangle (306.57,404.79);

\path[fill=fillColor] (338.02,380.02) rectangle (351.49,380.10);

\path[fill=fillColor] ( 68.50,380.02) rectangle ( 81.97,398.45);

\path[fill=fillColor] (113.42,380.02) rectangle (126.89,401.23);
\definecolor{fillColor}{RGB}{204,102,119}

\path[fill=fillColor] (144.86,380.02) rectangle (158.34,389.93);

\path[fill=fillColor] (189.78,380.02) rectangle (203.26,382.27);

\path[fill=fillColor] (279.62,380.02) rectangle (293.10,421.21);

\path[fill=fillColor] (324.54,380.02) rectangle (338.02,384.28);

\path[fill=fillColor] ( 55.02,380.02) rectangle ( 68.50,387.84);

\path[fill=fillColor] ( 99.94,380.02) rectangle (113.42,392.10);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 48.28,376.38) rectangle (371.71,456.51);
\end{scope}
\begin{scope}
\path[clip] ( 48.28,290.75) rectangle (371.71,370.88);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 48.28,290.75) rectangle (371.71,370.88);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 48.28,304.07) --
	(371.71,304.07);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 48.28,323.42) --
	(371.71,323.42);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 48.28,342.78) --
	(371.71,342.78);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 48.28,362.13) --
	(371.71,362.13);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 48.28,294.39) --
	(371.71,294.39);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 48.28,313.74) --
	(371.71,313.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 48.28,333.10) --
	(371.71,333.10);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 48.28,352.45) --
	(371.71,352.45);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 75.24,290.75) --
	( 75.24,370.88);

\path[draw=drawColor,line width= 0.6pt,line join=round] (120.16,290.75) --
	(120.16,370.88);

\path[draw=drawColor,line width= 0.6pt,line join=round] (165.08,290.75) --
	(165.08,370.88);

\path[draw=drawColor,line width= 0.6pt,line join=round] (209.99,290.75) --
	(209.99,370.88);

\path[draw=drawColor,line width= 0.6pt,line join=round] (254.91,290.75) --
	(254.91,370.88);

\path[draw=drawColor,line width= 0.6pt,line join=round] (299.83,290.75) --
	(299.83,370.88);

\path[draw=drawColor,line width= 0.6pt,line join=round] (344.75,290.75) --
	(344.75,370.88);
\definecolor{fillColor}{RGB}{221,204,119}

\path[fill=fillColor] (171.81,294.39) rectangle (185.29,296.79);

\path[fill=fillColor] (254.91,294.39) rectangle (275.13,294.39);

\path[fill=fillColor] (216.73,294.39) rectangle (230.21,295.71);

\path[fill=fillColor] (306.57,294.39) rectangle (320.05,298.88);

\path[fill=fillColor] ( 81.97,294.39) rectangle ( 95.45,363.52);
\definecolor{fillColor}{RGB}{17,119,51}

\path[fill=fillColor] (158.34,294.39) rectangle (171.81,296.71);

\path[fill=fillColor] (234.70,294.39) rectangle (254.91,294.47);

\path[fill=fillColor] (203.26,294.39) rectangle (216.73,295.09);

\path[fill=fillColor] (293.10,294.39) rectangle (306.57,295.86);

\path[fill=fillColor] ( 68.50,294.39) rectangle ( 81.97,367.24);
\definecolor{fillColor}{RGB}{204,102,119}

\path[fill=fillColor] (144.86,294.39) rectangle (158.34,297.41);

\path[fill=fillColor] (189.78,294.39) rectangle (203.26,300.20);

\path[fill=fillColor] (279.62,294.39) rectangle (293.10,320.63);

\path[fill=fillColor] ( 55.02,294.39) rectangle ( 68.50,336.74);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 48.28,290.75) rectangle (371.71,370.88);
\end{scope}
\begin{scope}
\path[clip] ( 48.28,205.12) rectangle (371.71,285.25);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 48.28,205.12) rectangle (371.71,285.25);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 48.28,218.44) --
	(371.71,218.44);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 48.28,237.79) --
	(371.71,237.79);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 48.28,257.14) --
	(371.71,257.14);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 48.28,276.50) --
	(371.71,276.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 48.28,208.76) --
	(371.71,208.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 48.28,228.11) --
	(371.71,228.11);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 48.28,247.47) --
	(371.71,247.47);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 48.28,266.82) --
	(371.71,266.82);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 75.24,205.12) --
	( 75.24,285.25);

\path[draw=drawColor,line width= 0.6pt,line join=round] (120.16,205.12) --
	(120.16,285.25);

\path[draw=drawColor,line width= 0.6pt,line join=round] (165.08,205.12) --
	(165.08,285.25);

\path[draw=drawColor,line width= 0.6pt,line join=round] (209.99,205.12) --
	(209.99,285.25);

\path[draw=drawColor,line width= 0.6pt,line join=round] (254.91,205.12) --
	(254.91,285.25);

\path[draw=drawColor,line width= 0.6pt,line join=round] (299.83,205.12) --
	(299.83,285.25);

\path[draw=drawColor,line width= 0.6pt,line join=round] (344.75,205.12) --
	(344.75,285.25);
\definecolor{fillColor}{RGB}{221,204,119}

\path[fill=fillColor] (171.81,208.76) rectangle (185.29,221.77);

\path[fill=fillColor] (261.65,208.76) rectangle (275.13,209.92);

\path[fill=fillColor] (216.73,208.76) rectangle (230.21,212.86);

\path[fill=fillColor] (306.57,208.76) rectangle (320.05,216.97);

\path[fill=fillColor] (351.49,208.76) rectangle (364.97,210.62);

\path[fill=fillColor] ( 81.97,208.76) rectangle ( 95.45,246.00);

\path[fill=fillColor] (126.89,208.76) rectangle (140.37,223.70);
\definecolor{fillColor}{RGB}{17,119,51}

\path[fill=fillColor] (158.34,208.76) rectangle (171.81,221.77);

\path[fill=fillColor] (248.18,208.76) rectangle (261.65,209.92);

\path[fill=fillColor] (203.26,208.76) rectangle (216.73,212.24);

\path[fill=fillColor] (293.10,208.76) rectangle (306.57,214.10);

\path[fill=fillColor] (338.02,208.76) rectangle (351.49,209.69);

\path[fill=fillColor] ( 68.50,208.76) rectangle ( 81.97,248.47);

\path[fill=fillColor] (113.42,208.76) rectangle (126.89,225.48);
\definecolor{fillColor}{RGB}{204,102,119}

\path[fill=fillColor] (144.86,208.76) rectangle (158.34,220.53);

\path[fill=fillColor] (234.70,208.76) rectangle (248.18,210.00);

\path[fill=fillColor] (189.78,208.76) rectangle (203.26,215.50);

\path[fill=fillColor] (279.62,208.76) rectangle (293.10,228.97);

\path[fill=fillColor] (324.54,208.76) rectangle (338.02,216.42);

\path[fill=fillColor] ( 55.02,208.76) rectangle ( 68.50,233.92);

\path[fill=fillColor] ( 99.94,208.76) rectangle (113.42,217.12);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 48.28,205.12) rectangle (371.71,285.25);
\end{scope}
\begin{scope}
\path[clip] ( 48.28,119.49) rectangle (371.71,199.62);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 48.28,119.49) rectangle (371.71,199.62);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 48.28,132.81) --
	(371.71,132.81);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 48.28,152.16) --
	(371.71,152.16);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 48.28,171.51) --
	(371.71,171.51);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 48.28,190.87) --
	(371.71,190.87);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 48.28,123.13) --
	(371.71,123.13);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 48.28,142.48) --
	(371.71,142.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 48.28,161.84) --
	(371.71,161.84);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 48.28,181.19) --
	(371.71,181.19);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 75.24,119.49) --
	( 75.24,199.62);

\path[draw=drawColor,line width= 0.6pt,line join=round] (120.16,119.49) --
	(120.16,199.62);

\path[draw=drawColor,line width= 0.6pt,line join=round] (165.08,119.49) --
	(165.08,199.62);

\path[draw=drawColor,line width= 0.6pt,line join=round] (209.99,119.49) --
	(209.99,199.62);

\path[draw=drawColor,line width= 0.6pt,line join=round] (254.91,119.49) --
	(254.91,199.62);

\path[draw=drawColor,line width= 0.6pt,line join=round] (299.83,119.49) --
	(299.83,199.62);

\path[draw=drawColor,line width= 0.6pt,line join=round] (344.75,119.49) --
	(344.75,199.62);
\definecolor{fillColor}{RGB}{221,204,119}

\path[fill=fillColor] (171.81,123.13) rectangle (185.29,132.03);

\path[fill=fillColor] (216.73,123.13) rectangle (230.21,124.99);

\path[fill=fillColor] (306.57,123.13) rectangle (320.05,140.08);

\path[fill=fillColor] (344.75,123.13) rectangle (364.97,123.21);

\path[fill=fillColor] ( 81.97,123.13) rectangle ( 95.45,149.84);

\path[fill=fillColor] (126.89,123.13) rectangle (140.37,146.12);
\definecolor{fillColor}{RGB}{17,119,51}

\path[fill=fillColor] (158.34,123.13) rectangle (171.81,129.86);

\path[fill=fillColor] (203.26,123.13) rectangle (216.73,124.45);

\path[fill=fillColor] (293.10,123.13) rectangle (306.57,137.53);

\path[fill=fillColor] ( 68.50,123.13) rectangle ( 81.97,152.62);

\path[fill=fillColor] (113.42,123.13) rectangle (126.89,148.68);
\definecolor{fillColor}{RGB}{204,102,119}

\path[fill=fillColor] (144.86,123.13) rectangle (158.34,137.61);

\path[fill=fillColor] (189.78,123.13) rectangle (203.26,126.38);

\path[fill=fillColor] (279.62,123.13) rectangle (293.10,146.59);

\path[fill=fillColor] (324.54,123.13) rectangle (344.75,123.28);

\path[fill=fillColor] ( 55.02,123.13) rectangle ( 68.50,142.72);

\path[fill=fillColor] ( 99.94,123.13) rectangle (113.42,139.62);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 48.28,119.49) rectangle (371.71,199.62);
\end{scope}
\begin{scope}
\path[clip] (393.78,547.64) rectangle (717.20,627.77);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (393.78,547.64) rectangle (717.20,627.77);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] (393.78,560.96) --
	(717.20,560.96);

\path[draw=drawColor,line width= 0.3pt,line join=round] (393.78,580.31) --
	(717.20,580.31);

\path[draw=drawColor,line width= 0.3pt,line join=round] (393.78,599.67) --
	(717.20,599.67);

\path[draw=drawColor,line width= 0.3pt,line join=round] (393.78,619.02) --
	(717.20,619.02);

\path[draw=drawColor,line width= 0.6pt,line join=round] (393.78,551.28) --
	(717.20,551.28);

\path[draw=drawColor,line width= 0.6pt,line join=round] (393.78,570.64) --
	(717.20,570.64);

\path[draw=drawColor,line width= 0.6pt,line join=round] (393.78,589.99) --
	(717.20,589.99);

\path[draw=drawColor,line width= 0.6pt,line join=round] (393.78,609.34) --
	(717.20,609.34);

\path[draw=drawColor,line width= 0.6pt,line join=round] (420.73,547.64) --
	(420.73,627.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (465.65,547.64) --
	(465.65,627.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (510.57,547.64) --
	(510.57,627.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (555.49,547.64) --
	(555.49,627.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (600.41,547.64) --
	(600.41,627.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (645.33,547.64) --
	(645.33,627.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (690.25,547.64) --
	(690.25,627.77);
\definecolor{fillColor}{RGB}{221,204,119}

\path[fill=fillColor] (517.31,551.28) rectangle (530.78,579.62);

\path[fill=fillColor] (607.15,551.28) rectangle (620.62,552.60);

\path[fill=fillColor] (562.23,551.28) rectangle (575.70,552.68);

\path[fill=fillColor] (696.99,551.28) rectangle (710.46,558.17);

\path[fill=fillColor] (427.47,551.28) rectangle (440.94,581.71);

\path[fill=fillColor] (472.39,551.28) rectangle (485.86,560.26);
\definecolor{fillColor}{RGB}{17,119,51}

\path[fill=fillColor] (503.83,551.28) rectangle (517.31,579.85);

\path[fill=fillColor] (593.67,551.28) rectangle (607.15,552.60);

\path[fill=fillColor] (548.75,551.28) rectangle (562.23,551.98);

\path[fill=fillColor] (683.51,551.28) rectangle (696.99,556.16);

\path[fill=fillColor] (413.99,551.28) rectangle (427.47,583.33);

\path[fill=fillColor] (458.91,551.28) rectangle (472.39,561.12);
\definecolor{fillColor}{RGB}{204,102,119}

\path[fill=fillColor] (490.36,551.28) rectangle (503.83,578.77);

\path[fill=fillColor] (580.19,551.28) rectangle (593.67,552.37);

\path[fill=fillColor] (535.27,551.28) rectangle (548.75,555.39);

\path[fill=fillColor] (670.03,551.28) rectangle (683.51,565.84);

\path[fill=fillColor] (400.52,551.28) rectangle (413.99,575.67);

\path[fill=fillColor] (445.44,551.28) rectangle (458.91,557.17);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (393.78,547.64) rectangle (717.20,627.77);
\end{scope}
\begin{scope}
\path[clip] (393.78,462.01) rectangle (717.20,542.14);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (393.78,462.01) rectangle (717.20,542.14);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] (393.78,475.33) --
	(717.20,475.33);

\path[draw=drawColor,line width= 0.3pt,line join=round] (393.78,494.68) --
	(717.20,494.68);

\path[draw=drawColor,line width= 0.3pt,line join=round] (393.78,514.04) --
	(717.20,514.04);

\path[draw=drawColor,line width= 0.3pt,line join=round] (393.78,533.39) --
	(717.20,533.39);

\path[draw=drawColor,line width= 0.6pt,line join=round] (393.78,465.65) --
	(717.20,465.65);

\path[draw=drawColor,line width= 0.6pt,line join=round] (393.78,485.01) --
	(717.20,485.01);

\path[draw=drawColor,line width= 0.6pt,line join=round] (393.78,504.36) --
	(717.20,504.36);

\path[draw=drawColor,line width= 0.6pt,line join=round] (393.78,523.71) --
	(717.20,523.71);

\path[draw=drawColor,line width= 0.6pt,line join=round] (420.73,462.01) --
	(420.73,542.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (465.65,462.01) --
	(465.65,542.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (510.57,462.01) --
	(510.57,542.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (555.49,462.01) --
	(555.49,542.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (600.41,462.01) --
	(600.41,542.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (645.33,462.01) --
	(645.33,542.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (690.25,462.01) --
	(690.25,542.14);
\definecolor{fillColor}{RGB}{221,204,119}

\path[fill=fillColor] (517.31,465.65) rectangle (530.78,471.85);

\path[fill=fillColor] (562.23,465.65) rectangle (575.70,471.85);

\path[fill=fillColor] (652.07,465.65) rectangle (665.54,468.67);

\path[fill=fillColor] (696.99,465.65) rectangle (710.46,468.90);

\path[fill=fillColor] (427.47,465.65) rectangle (440.94,510.55);

\path[fill=fillColor] (472.39,465.65) rectangle (485.86,479.59);
\definecolor{fillColor}{RGB}{17,119,51}

\path[fill=fillColor] (503.83,465.65) rectangle (517.31,470.53);

\path[fill=fillColor] (548.75,465.65) rectangle (562.23,469.60);

\path[fill=fillColor] (638.59,465.65) rectangle (652.07,467.67);

\path[fill=fillColor] (683.51,465.65) rectangle (696.99,466.12);

\path[fill=fillColor] (413.99,465.65) rectangle (427.47,515.04);

\path[fill=fillColor] (458.91,465.65) rectangle (472.39,482.37);
\definecolor{fillColor}{RGB}{204,102,119}

\path[fill=fillColor] (490.36,465.65) rectangle (503.83,476.41);

\path[fill=fillColor] (535.27,465.65) rectangle (548.75,479.43);

\path[fill=fillColor] (625.11,465.65) rectangle (638.59,472.23);

\path[fill=fillColor] (670.03,465.65) rectangle (683.51,478.50);

\path[fill=fillColor] (400.52,465.65) rectangle (413.99,494.76);

\path[fill=fillColor] (445.44,465.65) rectangle (458.91,469.99);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (393.78,462.01) rectangle (717.20,542.14);
\end{scope}
\begin{scope}
\path[clip] (393.78,376.38) rectangle (717.20,456.51);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (393.78,376.38) rectangle (717.20,456.51);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] (393.78,389.70) --
	(717.20,389.70);

\path[draw=drawColor,line width= 0.3pt,line join=round] (393.78,409.05) --
	(717.20,409.05);

\path[draw=drawColor,line width= 0.3pt,line join=round] (393.78,428.41) --
	(717.20,428.41);

\path[draw=drawColor,line width= 0.3pt,line join=round] (393.78,447.76) --
	(717.20,447.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (393.78,380.02) --
	(717.20,380.02);

\path[draw=drawColor,line width= 0.6pt,line join=round] (393.78,399.38) --
	(717.20,399.38);

\path[draw=drawColor,line width= 0.6pt,line join=round] (393.78,418.73) --
	(717.20,418.73);

\path[draw=drawColor,line width= 0.6pt,line join=round] (393.78,438.08) --
	(717.20,438.08);

\path[draw=drawColor,line width= 0.6pt,line join=round] (420.73,376.38) --
	(420.73,456.51);

\path[draw=drawColor,line width= 0.6pt,line join=round] (465.65,376.38) --
	(465.65,456.51);

\path[draw=drawColor,line width= 0.6pt,line join=round] (510.57,376.38) --
	(510.57,456.51);

\path[draw=drawColor,line width= 0.6pt,line join=round] (555.49,376.38) --
	(555.49,456.51);

\path[draw=drawColor,line width= 0.6pt,line join=round] (600.41,376.38) --
	(600.41,456.51);

\path[draw=drawColor,line width= 0.6pt,line join=round] (645.33,376.38) --
	(645.33,456.51);

\path[draw=drawColor,line width= 0.6pt,line join=round] (690.25,376.38) --
	(690.25,456.51);
\definecolor{fillColor}{RGB}{221,204,119}

\path[fill=fillColor] (517.31,380.02) rectangle (530.78,386.53);

\path[fill=fillColor] (607.15,380.02) rectangle (620.62,380.41);

\path[fill=fillColor] (562.23,380.02) rectangle (575.70,383.66);

\path[fill=fillColor] (652.07,380.02) rectangle (665.54,386.91);

\path[fill=fillColor] (696.99,380.02) rectangle (710.46,380.87);

\path[fill=fillColor] (427.47,380.02) rectangle (440.94,414.08);

\path[fill=fillColor] (472.39,380.02) rectangle (485.86,405.18);
\definecolor{fillColor}{RGB}{17,119,51}

\path[fill=fillColor] (503.83,380.02) rectangle (517.31,385.67);

\path[fill=fillColor] (593.67,380.02) rectangle (607.15,380.41);

\path[fill=fillColor] (548.75,380.02) rectangle (562.23,382.96);

\path[fill=fillColor] (638.59,380.02) rectangle (652.07,382.42);

\path[fill=fillColor] (683.51,380.02) rectangle (696.99,380.64);

\path[fill=fillColor] (413.99,380.02) rectangle (427.47,416.87);

\path[fill=fillColor] (458.91,380.02) rectangle (472.39,408.59);
\definecolor{fillColor}{RGB}{204,102,119}

\path[fill=fillColor] (490.36,380.02) rectangle (503.83,389.39);

\path[fill=fillColor] (580.19,380.02) rectangle (593.67,380.41);

\path[fill=fillColor] (535.27,380.02) rectangle (548.75,385.98);

\path[fill=fillColor] (625.11,380.02) rectangle (638.59,402.78);

\path[fill=fillColor] (670.03,380.02) rectangle (683.51,381.73);

\path[fill=fillColor] (400.52,380.02) rectangle (413.99,404.10);

\path[fill=fillColor] (445.44,380.02) rectangle (458.91,393.26);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (393.78,376.38) rectangle (717.20,456.51);
\end{scope}
\begin{scope}
\path[clip] (393.78,290.75) rectangle (717.20,370.88);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (393.78,290.75) rectangle (717.20,370.88);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] (393.78,304.07) --
	(717.20,304.07);

\path[draw=drawColor,line width= 0.3pt,line join=round] (393.78,323.42) --
	(717.20,323.42);

\path[draw=drawColor,line width= 0.3pt,line join=round] (393.78,342.78) --
	(717.20,342.78);

\path[draw=drawColor,line width= 0.3pt,line join=round] (393.78,362.13) --
	(717.20,362.13);

\path[draw=drawColor,line width= 0.6pt,line join=round] (393.78,294.39) --
	(717.20,294.39);

\path[draw=drawColor,line width= 0.6pt,line join=round] (393.78,313.74) --
	(717.20,313.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] (393.78,333.10) --
	(717.20,333.10);

\path[draw=drawColor,line width= 0.6pt,line join=round] (393.78,352.45) --
	(717.20,352.45);

\path[draw=drawColor,line width= 0.6pt,line join=round] (420.73,290.75) --
	(420.73,370.88);

\path[draw=drawColor,line width= 0.6pt,line join=round] (465.65,290.75) --
	(465.65,370.88);

\path[draw=drawColor,line width= 0.6pt,line join=round] (510.57,290.75) --
	(510.57,370.88);

\path[draw=drawColor,line width= 0.6pt,line join=round] (555.49,290.75) --
	(555.49,370.88);

\path[draw=drawColor,line width= 0.6pt,line join=round] (600.41,290.75) --
	(600.41,370.88);

\path[draw=drawColor,line width= 0.6pt,line join=round] (645.33,290.75) --
	(645.33,370.88);

\path[draw=drawColor,line width= 0.6pt,line join=round] (690.25,290.75) --
	(690.25,370.88);
\definecolor{fillColor}{RGB}{221,204,119}

\path[fill=fillColor] (517.31,294.39) rectangle (530.78,301.75);

\path[fill=fillColor] (607.15,294.39) rectangle (620.62,298.73);

\path[fill=fillColor] (562.23,294.39) rectangle (575.70,299.97);

\path[fill=fillColor] (652.07,294.39) rectangle (665.54,302.29);

\path[fill=fillColor] (690.25,294.39) rectangle (710.46,294.47);

\path[fill=fillColor] (427.47,294.39) rectangle (440.94,321.49);

\path[fill=fillColor] (472.39,294.39) rectangle (485.86,319.47);
\definecolor{fillColor}{RGB}{17,119,51}

\path[fill=fillColor] (503.83,294.39) rectangle (517.31,301.98);

\path[fill=fillColor] (593.67,294.39) rectangle (607.15,298.26);

\path[fill=fillColor] (548.75,294.39) rectangle (562.23,299.81);

\path[fill=fillColor] (638.59,294.39) rectangle (652.07,297.10);

\path[fill=fillColor] (413.99,294.39) rectangle (427.47,324.04);

\path[fill=fillColor] (458.91,294.39) rectangle (472.39,322.65);
\definecolor{fillColor}{RGB}{204,102,119}

\path[fill=fillColor] (490.36,294.39) rectangle (503.83,300.97);

\path[fill=fillColor] (580.19,294.39) rectangle (593.67,300.20);

\path[fill=fillColor] (535.27,294.39) rectangle (548.75,300.43);

\path[fill=fillColor] (625.11,294.39) rectangle (638.59,319.16);

\path[fill=fillColor] (670.03,294.39) rectangle (690.25,294.86);

\path[fill=fillColor] (400.52,294.39) rectangle (413.99,313.36);

\path[fill=fillColor] (445.44,294.39) rectangle (458.91,309.18);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (393.78,290.75) rectangle (717.20,370.88);
\end{scope}
\begin{scope}
\path[clip] (393.78,205.12) rectangle (717.20,285.25);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (393.78,205.12) rectangle (717.20,285.25);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] (393.78,218.44) --
	(717.20,218.44);

\path[draw=drawColor,line width= 0.3pt,line join=round] (393.78,237.79) --
	(717.20,237.79);

\path[draw=drawColor,line width= 0.3pt,line join=round] (393.78,257.14) --
	(717.20,257.14);

\path[draw=drawColor,line width= 0.3pt,line join=round] (393.78,276.50) --
	(717.20,276.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] (393.78,208.76) --
	(717.20,208.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (393.78,228.11) --
	(717.20,228.11);

\path[draw=drawColor,line width= 0.6pt,line join=round] (393.78,247.47) --
	(717.20,247.47);

\path[draw=drawColor,line width= 0.6pt,line join=round] (393.78,266.82) --
	(717.20,266.82);

\path[draw=drawColor,line width= 0.6pt,line join=round] (420.73,205.12) --
	(420.73,285.25);

\path[draw=drawColor,line width= 0.6pt,line join=round] (465.65,205.12) --
	(465.65,285.25);

\path[draw=drawColor,line width= 0.6pt,line join=round] (510.57,205.12) --
	(510.57,285.25);

\path[draw=drawColor,line width= 0.6pt,line join=round] (555.49,205.12) --
	(555.49,285.25);

\path[draw=drawColor,line width= 0.6pt,line join=round] (600.41,205.12) --
	(600.41,285.25);

\path[draw=drawColor,line width= 0.6pt,line join=round] (645.33,205.12) --
	(645.33,285.25);

\path[draw=drawColor,line width= 0.6pt,line join=round] (690.25,205.12) --
	(690.25,285.25);
\definecolor{fillColor}{RGB}{221,204,119}

\path[fill=fillColor] (517.31,208.76) rectangle (530.78,230.51);

\path[fill=fillColor] (607.15,208.76) rectangle (620.62,212.55);

\path[fill=fillColor] (562.23,208.76) rectangle (575.70,210.46);

\path[fill=fillColor] (652.07,208.76) rectangle (665.54,221.61);

\path[fill=fillColor] (427.47,208.76) rectangle (440.94,236.63);

\path[fill=fillColor] (472.39,208.76) rectangle (485.86,218.28);
\definecolor{fillColor}{RGB}{17,119,51}

\path[fill=fillColor] (503.83,208.76) rectangle (517.31,221.53);

\path[fill=fillColor] (593.67,208.76) rectangle (607.15,211.93);

\path[fill=fillColor] (548.75,208.76) rectangle (562.23,210.31);

\path[fill=fillColor] (638.59,208.76) rectangle (652.07,213.25);

\path[fill=fillColor] (413.99,208.76) rectangle (427.47,245.22);

\path[fill=fillColor] (458.91,208.76) rectangle (472.39,227.65);
\definecolor{fillColor}{RGB}{204,102,119}

\path[fill=fillColor] (490.36,208.76) rectangle (503.83,234.38);

\path[fill=fillColor] (580.19,208.76) rectangle (593.67,212.86);

\path[fill=fillColor] (535.27,208.76) rectangle (548.75,210.46);

\path[fill=fillColor] (625.11,208.76) rectangle (638.59,225.25);

\path[fill=fillColor] (400.52,208.76) rectangle (413.99,232.84);

\path[fill=fillColor] (445.44,208.76) rectangle (458.91,214.10);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (393.78,205.12) rectangle (717.20,285.25);
\end{scope}
\begin{scope}
\path[clip] (377.21,547.64) rectangle (393.78,627.77);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (377.21,547.64) rectangle (393.78,627.77);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (388.52,587.71) {Colombia};
\end{scope}
\begin{scope}
\path[clip] (377.21,462.01) rectangle (393.78,542.14);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (377.21,462.01) rectangle (393.78,542.14);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (388.52,502.08) {Nigeria};
\end{scope}
\begin{scope}
\path[clip] (377.21,376.38) rectangle (393.78,456.51);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (377.21,376.38) rectangle (393.78,456.51);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (388.52,416.45) {Sierra Leone 1};
\end{scope}
\begin{scope}
\path[clip] (377.21,290.75) rectangle (393.78,370.88);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (377.21,290.75) rectangle (393.78,370.88);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (388.52,330.81) {Uganda 2};
\end{scope}
\begin{scope}
\path[clip] (377.21,205.12) rectangle (393.78,285.25);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (377.21,205.12) rectangle (393.78,285.25);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (388.52,245.18) {Russia};
\end{scope}
\begin{scope}
\path[clip] ( 31.71,547.64) rectangle ( 48.28,627.77);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 31.71,547.64) rectangle ( 48.28,627.77);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 43.03,587.71) {Burkina Faso};
\end{scope}
\begin{scope}
\path[clip] ( 31.71,462.01) rectangle ( 48.28,542.14);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 31.71,462.01) rectangle ( 48.28,542.14);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 43.03,502.08) {Nepal};
\end{scope}
\begin{scope}
\path[clip] ( 31.71,376.38) rectangle ( 48.28,456.51);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 31.71,376.38) rectangle ( 48.28,456.51);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 43.03,416.45) {Rwanda};
\end{scope}
\begin{scope}
\path[clip] ( 31.71,290.75) rectangle ( 48.28,370.88);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 31.71,290.75) rectangle ( 48.28,370.88);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 43.03,330.81) {Sierra Leone 2};
\end{scope}
\begin{scope}
\path[clip] ( 31.71,205.12) rectangle ( 48.28,285.25);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 31.71,205.12) rectangle ( 48.28,285.25);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 43.03,245.18) {All LMICs};
\end{scope}
\begin{scope}
\path[clip] ( 31.71,119.49) rectangle ( 48.28,199.62);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 31.71,119.49) rectangle ( 48.28,199.62);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 43.03,159.55) {USA};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 75.24,116.74) --
	( 75.24,119.49);

\path[draw=drawColor,line width= 0.6pt,line join=round] (120.16,116.74) --
	(120.16,119.49);

\path[draw=drawColor,line width= 0.6pt,line join=round] (165.08,116.74) --
	(165.08,119.49);

\path[draw=drawColor,line width= 0.6pt,line join=round] (209.99,116.74) --
	(209.99,119.49);

\path[draw=drawColor,line width= 0.6pt,line join=round] (254.91,116.74) --
	(254.91,119.49);

\path[draw=drawColor,line width= 0.6pt,line join=round] (299.83,116.74) --
	(299.83,119.49);

\path[draw=drawColor,line width= 0.6pt,line join=round] (344.75,116.74) --
	(344.75,119.49);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 78.27,114.54) {Health workers};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (118.43,114.54) {Government or};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (127.94,114.54) {MoH};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (163.35,114.54) {Family or};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (172.86,114.54) {Friends};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (198.77,114.54) {Famous person,};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (208.27,114.54) {religious leader};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (217.78,114.54) {or traditional};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (227.28,114.54) {healers};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (248.44,114.54) {Newspapers,};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (257.95,114.54) {radio or online};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (267.45,114.54) {groups};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (302.86,114.54) {Other};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (343.03,114.54) {Don't know or};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (352.54,114.54) {Refuse};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (420.73,202.37) --
	(420.73,205.12);

\path[draw=drawColor,line width= 0.6pt,line join=round] (465.65,202.37) --
	(465.65,205.12);

\path[draw=drawColor,line width= 0.6pt,line join=round] (510.57,202.37) --
	(510.57,205.12);

\path[draw=drawColor,line width= 0.6pt,line join=round] (555.49,202.37) --
	(555.49,205.12);

\path[draw=drawColor,line width= 0.6pt,line join=round] (600.41,202.37) --
	(600.41,205.12);

\path[draw=drawColor,line width= 0.6pt,line join=round] (645.33,202.37) --
	(645.33,205.12);

\path[draw=drawColor,line width= 0.6pt,line join=round] (690.25,202.37) --
	(690.25,205.12);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (423.76,200.17) {Health workers};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (463.93,200.17) {Government or};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (473.43,200.17) {MoH};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (508.85,200.17) {Family or};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (518.35,200.17) {Friends};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (544.26,200.17) {Famous person,};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (553.77,200.17) {religious leader};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (563.27,200.17) {or traditional};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (572.78,200.17) {healers};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (593.93,200.17) {Newspapers,};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (603.44,200.17) {radio or online};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (612.94,200.17) {groups};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (648.36,200.17) {Other};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (688.53,200.17) {Don't know or};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (698.03,200.17) {Refuse};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,548.25) {0};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,567.61) {25};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,586.96) {50};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,606.31) {75};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96,551.28) --
	( 31.71,551.28);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96,570.64) --
	( 31.71,570.64);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96,589.99) --
	( 31.71,589.99);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96,609.34) --
	( 31.71,609.34);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,462.62) {0};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,481.98) {25};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,501.33) {50};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,520.68) {75};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96,465.65) --
	( 31.71,465.65);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96,485.01) --
	( 31.71,485.01);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96,504.36) --
	( 31.71,504.36);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96,523.71) --
	( 31.71,523.71);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,376.99) {0};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,396.35) {25};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,415.70) {50};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,435.05) {75};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96,380.02) --
	( 31.71,380.02);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96,399.38) --
	( 31.71,399.38);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96,418.73) --
	( 31.71,418.73);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96,438.08) --
	( 31.71,438.08);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,291.36) {0};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,310.71) {25};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,330.07) {50};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,349.42) {75};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96,294.39) --
	( 31.71,294.39);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96,313.74) --
	( 31.71,313.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96,333.10) --
	( 31.71,333.10);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96,352.45) --
	( 31.71,352.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,205.73) {0};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,225.08) {25};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,244.44) {50};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,263.79) {75};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96,208.76) --
	( 31.71,208.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96,228.11) --
	( 31.71,228.11);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96,247.47) --
	( 31.71,247.47);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96,266.82) --
	( 31.71,266.82);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,120.10) {0};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,139.45) {25};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,158.81) {50};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,178.16) {75};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96,123.13) --
	( 31.71,123.13);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96,142.48) --
	( 31.71,142.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96,161.84) --
	( 31.71,161.84);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96,181.19) --
	( 31.71,181.19);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 13.08,373.63) {estimate};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (276.12,  5.50) rectangle (489.36, 30.95);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.10] at (281.62, 14.44) {Answer};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (322.65, 11.00) rectangle (337.10, 25.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{fillColor}{RGB}{204,102,119}

\path[fill=fillColor] (323.36, 11.71) rectangle (336.39, 24.74);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (409.32, 11.00) rectangle (423.78, 25.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{fillColor}{RGB}{17,119,51}

\path[fill=fillColor] (410.03, 11.71) rectangle (423.07, 24.74);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (448.02, 11.00) rectangle (462.48, 25.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{fillColor}{RGB}{221,204,119}

\path[fill=fillColor] (448.73, 11.71) rectangle (461.77, 24.74);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (342.60, 15.20) {No, Don't know};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (429.28, 15.20) {Yes};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (467.98, 15.20) {Any};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.32] at (  5.50,635.84) {Which of the following people would you trust MOST to help you decide whether you would get a COVID-19 vaccine?};
\end{scope}
\end{tikzpicture}

\end{document}