### 2021-03-23 paper_files/figure-latex/genderhist-1.tex

% Created by tikzDevice version 0.12.3.1 on 2021-03-23 16:52:44
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}

\nonstopmode

\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (722.70,650.43);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.00) rectangle (722.70,650.43);
\end{scope}
\begin{scope}
\path[clip] ( 48.28,547.64) rectangle (371.71,627.77);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 48.28,547.64) rectangle (371.71,627.77);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 48.28,561.31) --
	(371.71,561.31);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 48.28,581.37) --
	(371.71,581.37);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 48.28,601.43) --
	(371.71,601.43);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 48.28,621.48) --
	(371.71,621.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 48.28,551.28) --
	(371.71,551.28);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 48.28,571.34) --
	(371.71,571.34);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 48.28,591.40) --
	(371.71,591.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 48.28,611.45) --
	(371.71,611.45);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 75.24,547.64) --
	( 75.24,627.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (120.16,547.64) --
	(120.16,627.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (165.08,547.64) --
	(165.08,627.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (209.99,547.64) --
	(209.99,627.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (254.91,547.64) --
	(254.91,627.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (299.83,547.64) --
	(299.83,627.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (344.75,547.64) --
	(344.75,627.77);
\definecolor{fillColor}{RGB}{221,204,119}

\path[fill=fillColor] (171.81,551.28) rectangle (185.29,566.05);

\path[fill=fillColor] (261.65,551.28) rectangle (275.13,552.81);

\path[fill=fillColor] (216.73,551.28) rectangle (230.21,552.73);

\path[fill=fillColor] (306.57,551.28) rectangle (320.05,560.59);

\path[fill=fillColor] (351.49,551.28) rectangle (364.97,552.65);

\path[fill=fillColor] ( 81.97,551.28) rectangle ( 95.45,592.68);

\path[fill=fillColor] (126.89,551.28) rectangle (140.37,561.63);
\definecolor{fillColor}{RGB}{17,119,51}

\path[fill=fillColor] (158.34,551.28) rectangle (171.81,564.68);

\path[fill=fillColor] (248.18,551.28) rectangle (261.65,552.41);

\path[fill=fillColor] (203.26,551.28) rectangle (216.73,552.65);

\path[fill=fillColor] (293.10,551.28) rectangle (306.57,561.47);

\path[fill=fillColor] (338.02,551.28) rectangle (351.49,552.73);

\path[fill=fillColor] ( 68.50,551.28) rectangle ( 81.97,592.52);

\path[fill=fillColor] (113.42,551.28) rectangle (126.89,562.68);
\definecolor{fillColor}{RGB}{204,102,119}

\path[fill=fillColor] (144.86,551.28) rectangle (158.34,569.17);

\path[fill=fillColor] (234.70,551.28) rectangle (248.18,553.53);

\path[fill=fillColor] (189.78,551.28) rectangle (203.26,552.97);

\path[fill=fillColor] (279.62,551.28) rectangle (293.10,558.58);

\path[fill=fillColor] (324.54,551.28) rectangle (338.02,552.41);

\path[fill=fillColor] ( 55.02,551.28) rectangle ( 68.50,593.08);

\path[fill=fillColor] ( 99.94,551.28) rectangle (113.42,559.31);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 48.28,547.64) rectangle (371.71,627.77);
\end{scope}
\begin{scope}
\path[clip] ( 48.28,462.01) rectangle (371.71,542.14);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 48.28,462.01) rectangle (371.71,542.14);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 48.28,475.68) --
	(371.71,475.68);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 48.28,495.74) --
	(371.71,495.74);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 48.28,515.80) --
	(371.71,515.80);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 48.28,535.85) --
	(371.71,535.85);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 48.28,465.65) --
	(371.71,465.65);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 48.28,485.71) --
	(371.71,485.71);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 48.28,505.77) --
	(371.71,505.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 48.28,525.82) --
	(371.71,525.82);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 75.24,462.01) --
	( 75.24,542.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (120.16,462.01) --
	(120.16,542.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (165.08,462.01) --
	(165.08,542.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (209.99,462.01) --
	(209.99,542.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (254.91,462.01) --
	(254.91,542.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (299.83,462.01) --
	(299.83,542.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (344.75,462.01) --
	(344.75,542.14);
\definecolor{fillColor}{RGB}{221,204,119}

\path[fill=fillColor] (171.81,465.65) rectangle (185.29,494.21);

\path[fill=fillColor] (254.91,465.65) rectangle (275.13,465.97);

\path[fill=fillColor] (216.73,465.65) rectangle (230.21,478.81);

\path[fill=fillColor] (306.57,465.65) rectangle (320.05,466.05);

\path[fill=fillColor] (351.49,465.65) rectangle (364.97,467.58);

\path[fill=fillColor] ( 81.97,465.65) rectangle ( 95.45,500.79);

\path[fill=fillColor] (126.89,465.65) rectangle (140.37,466.30);
\definecolor{fillColor}{RGB}{17,119,51}

\path[fill=fillColor] (158.34,465.65) rectangle (171.81,491.41);

\path[fill=fillColor] (234.70,465.65) rectangle (254.91,466.30);

\path[fill=fillColor] (203.26,465.65) rectangle (216.73,478.97);

\path[fill=fillColor] (293.10,465.65) rectangle (306.57,466.30);

\path[fill=fillColor] (338.02,465.65) rectangle (351.49,466.62);

\path[fill=fillColor] ( 68.50,465.65) rectangle ( 81.97,503.92);

\path[fill=fillColor] (113.42,465.65) rectangle (126.89,466.38);
\definecolor{fillColor}{RGB}{204,102,119}

\path[fill=fillColor] (144.86,465.65) rectangle (158.34,498.15);

\path[fill=fillColor] (189.78,465.65) rectangle (203.26,478.49);

\path[fill=fillColor] (279.62,465.65) rectangle (293.10,465.81);

\path[fill=fillColor] (324.54,465.65) rectangle (338.02,468.86);

\path[fill=fillColor] ( 55.02,465.65) rectangle ( 68.50,496.62);

\path[fill=fillColor] ( 99.94,465.65) rectangle (113.42,466.21);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 48.28,462.01) rectangle (371.71,542.14);
\end{scope}
\begin{scope}
\path[clip] ( 48.28,376.38) rectangle (371.71,456.51);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 48.28,376.38) rectangle (371.71,456.51);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 48.28,390.05) --
	(371.71,390.05);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 48.28,410.11) --
	(371.71,410.11);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 48.28,430.16) --
	(371.71,430.16);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 48.28,450.22) --
	(371.71,450.22);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 48.28,380.02) --
	(371.71,380.02);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 48.28,400.08) --
	(371.71,400.08);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 48.28,420.14) --
	(371.71,420.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 48.28,440.19) --
	(371.71,440.19);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 75.24,376.38) --
	( 75.24,456.51);

\path[draw=drawColor,line width= 0.6pt,line join=round] (120.16,376.38) --
	(120.16,456.51);

\path[draw=drawColor,line width= 0.6pt,line join=round] (165.08,376.38) --
	(165.08,456.51);

\path[draw=drawColor,line width= 0.6pt,line join=round] (209.99,376.38) --
	(209.99,456.51);

\path[draw=drawColor,line width= 0.6pt,line join=round] (254.91,376.38) --
	(254.91,456.51);

\path[draw=drawColor,line width= 0.6pt,line join=round] (299.83,376.38) --
	(299.83,456.51);

\path[draw=drawColor,line width= 0.6pt,line join=round] (344.75,376.38) --
	(344.75,456.51);
\definecolor{fillColor}{RGB}{221,204,119}

\path[fill=fillColor] (171.81,380.02) rectangle (185.29,392.06);

\path[fill=fillColor] (261.65,380.02) rectangle (275.13,380.50);

\path[fill=fillColor] (216.73,380.02) rectangle (230.21,380.90);

\path[fill=fillColor] (306.57,380.02) rectangle (320.05,406.66);

\path[fill=fillColor] (344.75,380.02) rectangle (364.97,380.34);

\path[fill=fillColor] ( 81.97,380.02) rectangle ( 95.45,398.47);

\path[fill=fillColor] (126.89,380.02) rectangle (140.37,401.44);
\definecolor{fillColor}{RGB}{17,119,51}

\path[fill=fillColor] (158.34,380.02) rectangle (171.81,391.41);

\path[fill=fillColor] (248.18,380.02) rectangle (261.65,380.74);

\path[fill=fillColor] (203.26,380.02) rectangle (216.73,380.90);

\path[fill=fillColor] (293.10,380.02) rectangle (306.57,405.86);

\path[fill=fillColor] (324.54,380.02) rectangle (344.75,380.50);

\path[fill=fillColor] ( 68.50,380.02) rectangle ( 81.97,399.36);

\path[fill=fillColor] (113.42,380.02) rectangle (126.89,401.52);
\definecolor{fillColor}{RGB}{204,102,119}

\path[fill=fillColor] (144.86,380.02) rectangle (158.34,393.18);

\path[fill=fillColor] (234.70,380.02) rectangle (248.18,380.18);

\path[fill=fillColor] (189.78,380.02) rectangle (203.26,380.82);

\path[fill=fillColor] (279.62,380.02) rectangle (293.10,408.10);

\path[fill=fillColor] ( 55.02,380.02) rectangle ( 68.50,396.71);

\path[fill=fillColor] ( 99.94,380.02) rectangle (113.42,401.36);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 48.28,376.38) rectangle (371.71,456.51);
\end{scope}
\begin{scope}
\path[clip] ( 48.28,290.75) rectangle (371.71,370.88);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 48.28,290.75) rectangle (371.71,370.88);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 48.28,304.42) --
	(371.71,304.42);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 48.28,324.48) --
	(371.71,324.48);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 48.28,344.53) --
	(371.71,344.53);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 48.28,364.59) --
	(371.71,364.59);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 48.28,294.39) --
	(371.71,294.39);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 48.28,314.45) --
	(371.71,314.45);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 48.28,334.51) --
	(371.71,334.51);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 48.28,354.56) --
	(371.71,354.56);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 75.24,290.75) --
	( 75.24,370.88);

\path[draw=drawColor,line width= 0.6pt,line join=round] (120.16,290.75) --
	(120.16,370.88);

\path[draw=drawColor,line width= 0.6pt,line join=round] (165.08,290.75) --
	(165.08,370.88);

\path[draw=drawColor,line width= 0.6pt,line join=round] (209.99,290.75) --
	(209.99,370.88);

\path[draw=drawColor,line width= 0.6pt,line join=round] (254.91,290.75) --
	(254.91,370.88);

\path[draw=drawColor,line width= 0.6pt,line join=round] (299.83,290.75) --
	(299.83,370.88);

\path[draw=drawColor,line width= 0.6pt,line join=round] (344.75,290.75) --
	(344.75,370.88);
\definecolor{fillColor}{RGB}{221,204,119}

\path[fill=fillColor] (171.81,294.39) rectangle (185.29,296.88);

\path[fill=fillColor] (254.91,294.39) rectangle (275.13,294.39);

\path[fill=fillColor] (216.73,294.39) rectangle (230.21,295.76);

\path[fill=fillColor] (306.57,294.39) rectangle (320.05,299.04);

\path[fill=fillColor] ( 81.97,294.39) rectangle ( 95.45,366.03);
\definecolor{fillColor}{RGB}{17,119,51}

\path[fill=fillColor] (158.34,294.39) rectangle (171.81,296.56);

\path[fill=fillColor] (234.70,294.39) rectangle (254.91,294.47);

\path[fill=fillColor] (203.26,294.39) rectangle (216.73,295.84);

\path[fill=fillColor] (293.10,294.39) rectangle (306.57,299.69);

\path[fill=fillColor] ( 68.50,294.39) rectangle ( 81.97,365.63);
\definecolor{fillColor}{RGB}{204,102,119}

\path[fill=fillColor] (144.86,294.39) rectangle (158.34,297.92);

\path[fill=fillColor] (189.78,294.39) rectangle (203.26,295.59);

\path[fill=fillColor] (279.62,294.39) rectangle (293.10,297.04);

\path[fill=fillColor] ( 55.02,294.39) rectangle ( 68.50,367.24);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 48.28,290.75) rectangle (371.71,370.88);
\end{scope}
\begin{scope}
\path[clip] ( 48.28,205.12) rectangle (371.71,285.25);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 48.28,205.12) rectangle (371.71,285.25);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 48.28,218.79) --
	(371.71,218.79);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 48.28,238.85) --
	(371.71,238.85);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 48.28,258.90) --
	(371.71,258.90);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 48.28,278.96) --
	(371.71,278.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 48.28,208.76) --
	(371.71,208.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 48.28,228.82) --
	(371.71,228.82);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 48.28,248.87) --
	(371.71,248.87);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 48.28,268.93) --
	(371.71,268.93);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 75.24,205.12) --
	( 75.24,285.25);

\path[draw=drawColor,line width= 0.6pt,line join=round] (120.16,205.12) --
	(120.16,285.25);

\path[draw=drawColor,line width= 0.6pt,line join=round] (165.08,205.12) --
	(165.08,285.25);

\path[draw=drawColor,line width= 0.6pt,line join=round] (209.99,205.12) --
	(209.99,285.25);

\path[draw=drawColor,line width= 0.6pt,line join=round] (254.91,205.12) --
	(254.91,285.25);

\path[draw=drawColor,line width= 0.6pt,line join=round] (299.83,205.12) --
	(299.83,285.25);

\path[draw=drawColor,line width= 0.6pt,line join=round] (344.75,205.12) --
	(344.75,285.25);
\definecolor{fillColor}{RGB}{221,204,119}

\path[fill=fillColor] (171.81,208.76) rectangle (185.29,222.24);

\path[fill=fillColor] (261.65,208.76) rectangle (275.13,209.96);

\path[fill=fillColor] (216.73,208.76) rectangle (230.21,213.01);

\path[fill=fillColor] (306.57,208.76) rectangle (320.05,217.26);

\path[fill=fillColor] (351.49,208.76) rectangle (364.97,210.69);

\path[fill=fillColor] ( 81.97,208.76) rectangle ( 95.45,247.27);

\path[fill=fillColor] (126.89,208.76) rectangle (140.37,224.24);
\definecolor{fillColor}{RGB}{17,119,51}

\path[fill=fillColor] (158.34,208.76) rectangle (171.81,221.20);

\path[fill=fillColor] (248.18,208.76) rectangle (261.65,210.12);

\path[fill=fillColor] (203.26,208.76) rectangle (216.73,212.93);

\path[fill=fillColor] (293.10,208.76) rectangle (306.57,217.59);

\path[fill=fillColor] (338.02,208.76) rectangle (351.49,210.77);

\path[fill=fillColor] ( 68.50,208.76) rectangle ( 81.97,247.27);

\path[fill=fillColor] (113.42,208.76) rectangle (126.89,224.97);
\definecolor{fillColor}{RGB}{204,102,119}

\path[fill=fillColor] (144.86,208.76) rectangle (158.34,224.00);

\path[fill=fillColor] (234.70,208.76) rectangle (248.18,209.88);

\path[fill=fillColor] (189.78,208.76) rectangle (203.26,213.17);

\path[fill=fillColor] (279.62,208.76) rectangle (293.10,216.54);

\path[fill=fillColor] (324.54,208.76) rectangle (338.02,210.61);

\path[fill=fillColor] ( 55.02,208.76) rectangle ( 68.50,247.03);

\path[fill=fillColor] ( 99.94,208.76) rectangle (113.42,223.12);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 48.28,205.12) rectangle (371.71,285.25);
\end{scope}
\begin{scope}
\path[clip] ( 48.28,119.49) rectangle (371.71,199.62);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 48.28,119.49) rectangle (371.71,199.62);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 48.28,133.16) --
	(371.71,133.16);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 48.28,153.21) --
	(371.71,153.21);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 48.28,173.27) --
	(371.71,173.27);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 48.28,193.33) --
	(371.71,193.33);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 48.28,123.13) --
	(371.71,123.13);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 48.28,143.19) --
	(371.71,143.19);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 48.28,163.24) --
	(371.71,163.24);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 48.28,183.30) --
	(371.71,183.30);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 75.24,119.49) --
	( 75.24,199.62);

\path[draw=drawColor,line width= 0.6pt,line join=round] (120.16,119.49) --
	(120.16,199.62);

\path[draw=drawColor,line width= 0.6pt,line join=round] (165.08,119.49) --
	(165.08,199.62);

\path[draw=drawColor,line width= 0.6pt,line join=round] (209.99,119.49) --
	(209.99,199.62);

\path[draw=drawColor,line width= 0.6pt,line join=round] (254.91,119.49) --
	(254.91,199.62);

\path[draw=drawColor,line width= 0.6pt,line join=round] (299.83,119.49) --
	(299.83,199.62);

\path[draw=drawColor,line width= 0.6pt,line join=round] (344.75,119.49) --
	(344.75,199.62);
\definecolor{fillColor}{RGB}{221,204,119}

\path[fill=fillColor] (171.81,123.13) rectangle (185.29,132.36);

\path[fill=fillColor] (216.73,123.13) rectangle (230.21,125.06);

\path[fill=fillColor] (306.57,123.13) rectangle (320.05,140.70);

\path[fill=fillColor] (344.75,123.13) rectangle (364.97,123.21);

\path[fill=fillColor] ( 81.97,123.13) rectangle ( 95.45,150.81);

\path[fill=fillColor] (126.89,123.13) rectangle (140.37,146.96);
\definecolor{fillColor}{RGB}{17,119,51}

\path[fill=fillColor] (158.34,123.13) rectangle (171.81,131.71);

\path[fill=fillColor] (203.26,123.13) rectangle (216.73,124.73);

\path[fill=fillColor] (293.10,123.13) rectangle (306.57,142.71);

\path[fill=fillColor] (324.54,123.13) rectangle (344.75,123.21);

\path[fill=fillColor] ( 68.50,123.13) rectangle ( 81.97,149.44);

\path[fill=fillColor] (113.42,123.13) rectangle (126.89,147.20);
\definecolor{fillColor}{RGB}{204,102,119}

\path[fill=fillColor] (144.86,123.13) rectangle (158.34,133.00);

\path[fill=fillColor] (189.78,123.13) rectangle (203.26,125.30);

\path[fill=fillColor] (279.62,123.13) rectangle (293.10,138.61);

\path[fill=fillColor] ( 55.02,123.13) rectangle ( 68.50,152.25);

\path[fill=fillColor] ( 99.94,123.13) rectangle (113.42,146.72);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 48.28,119.49) rectangle (371.71,199.62);
\end{scope}
\begin{scope}
\path[clip] (393.78,547.64) rectangle (717.20,627.77);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (393.78,547.64) rectangle (717.20,627.77);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] (393.78,561.31) --
	(717.20,561.31);

\path[draw=drawColor,line width= 0.3pt,line join=round] (393.78,581.37) --
	(717.20,581.37);

\path[draw=drawColor,line width= 0.3pt,line join=round] (393.78,601.43) --
	(717.20,601.43);

\path[draw=drawColor,line width= 0.3pt,line join=round] (393.78,621.48) --
	(717.20,621.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] (393.78,551.28) --
	(717.20,551.28);

\path[draw=drawColor,line width= 0.6pt,line join=round] (393.78,571.34) --
	(717.20,571.34);

\path[draw=drawColor,line width= 0.6pt,line join=round] (393.78,591.40) --
	(717.20,591.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] (393.78,611.45) --
	(717.20,611.45);

\path[draw=drawColor,line width= 0.6pt,line join=round] (420.73,547.64) --
	(420.73,627.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (465.65,547.64) --
	(465.65,627.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (510.57,547.64) --
	(510.57,627.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (555.49,547.64) --
	(555.49,627.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (600.41,547.64) --
	(600.41,627.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (645.33,547.64) --
	(645.33,627.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (690.25,547.64) --
	(690.25,627.77);
\definecolor{fillColor}{RGB}{221,204,119}

\path[fill=fillColor] (517.31,551.28) rectangle (530.78,580.65);

\path[fill=fillColor] (607.15,551.28) rectangle (620.62,552.65);

\path[fill=fillColor] (562.23,551.28) rectangle (575.70,552.73);

\path[fill=fillColor] (696.99,551.28) rectangle (710.46,558.42);

\path[fill=fillColor] (427.47,551.28) rectangle (440.94,582.81);

\path[fill=fillColor] (472.39,551.28) rectangle (485.86,560.59);
\definecolor{fillColor}{RGB}{17,119,51}

\path[fill=fillColor] (503.83,551.28) rectangle (517.31,580.57);

\path[fill=fillColor] (593.67,551.28) rectangle (607.15,553.37);

\path[fill=fillColor] (548.75,551.28) rectangle (562.23,552.65);

\path[fill=fillColor] (683.51,551.28) rectangle (696.99,559.47);

\path[fill=fillColor] (413.99,551.28) rectangle (427.47,580.09);

\path[fill=fillColor] (458.91,551.28) rectangle (472.39,561.79);
\definecolor{fillColor}{RGB}{204,102,119}

\path[fill=fillColor] (490.36,551.28) rectangle (503.83,580.73);

\path[fill=fillColor] (580.19,551.28) rectangle (593.67,552.17);

\path[fill=fillColor] (535.27,551.28) rectangle (548.75,552.81);

\path[fill=fillColor] (670.03,551.28) rectangle (683.51,557.86);

\path[fill=fillColor] (400.52,551.28) rectangle (413.99,584.50);

\path[fill=fillColor] (445.44,551.28) rectangle (458.91,559.95);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (393.78,547.64) rectangle (717.20,627.77);
\end{scope}
\begin{scope}
\path[clip] (393.78,462.01) rectangle (717.20,542.14);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (393.78,462.01) rectangle (717.20,542.14);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] (393.78,475.68) --
	(717.20,475.68);

\path[draw=drawColor,line width= 0.3pt,line join=round] (393.78,495.74) --
	(717.20,495.74);

\path[draw=drawColor,line width= 0.3pt,line join=round] (393.78,515.80) --
	(717.20,515.80);

\path[draw=drawColor,line width= 0.3pt,line join=round] (393.78,535.85) --
	(717.20,535.85);

\path[draw=drawColor,line width= 0.6pt,line join=round] (393.78,465.65) --
	(717.20,465.65);

\path[draw=drawColor,line width= 0.6pt,line join=round] (393.78,485.71) --
	(717.20,485.71);

\path[draw=drawColor,line width= 0.6pt,line join=round] (393.78,505.77) --
	(717.20,505.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (393.78,525.82) --
	(717.20,525.82);

\path[draw=drawColor,line width= 0.6pt,line join=round] (420.73,462.01) --
	(420.73,542.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (465.65,462.01) --
	(465.65,542.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (510.57,462.01) --
	(510.57,542.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (555.49,462.01) --
	(555.49,542.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (600.41,462.01) --
	(600.41,542.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (645.33,462.01) --
	(645.33,542.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (690.25,462.01) --
	(690.25,542.14);
\definecolor{fillColor}{RGB}{221,204,119}

\path[fill=fillColor] (517.31,465.65) rectangle (530.78,472.07);

\path[fill=fillColor] (562.23,465.65) rectangle (575.70,472.07);

\path[fill=fillColor] (652.07,465.65) rectangle (665.54,468.78);

\path[fill=fillColor] (696.99,465.65) rectangle (710.46,469.02);

\path[fill=fillColor] (427.47,465.65) rectangle (440.94,512.19);

\path[fill=fillColor] (472.39,465.65) rectangle (485.86,480.09);
\definecolor{fillColor}{RGB}{17,119,51}

\path[fill=fillColor] (503.83,465.65) rectangle (517.31,470.63);

\path[fill=fillColor] (548.75,465.65) rectangle (562.23,471.51);

\path[fill=fillColor] (638.59,465.65) rectangle (652.07,469.34);

\path[fill=fillColor] (683.51,465.65) rectangle (696.99,469.18);

\path[fill=fillColor] (413.99,465.65) rectangle (427.47,512.91);

\path[fill=fillColor] (458.91,465.65) rectangle (472.39,480.58);
\definecolor{fillColor}{RGB}{204,102,119}

\path[fill=fillColor] (490.36,465.65) rectangle (503.83,474.40);

\path[fill=fillColor] (535.27,465.65) rectangle (548.75,472.87);

\path[fill=fillColor] (625.11,465.65) rectangle (638.59,467.90);

\path[fill=fillColor] (670.03,465.65) rectangle (683.51,468.70);

\path[fill=fillColor] (400.52,465.65) rectangle (413.99,510.90);

\path[fill=fillColor] (445.44,465.65) rectangle (458.91,479.37);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (393.78,462.01) rectangle (717.20,542.14);
\end{scope}
\begin{scope}
\path[clip] (393.78,376.38) rectangle (717.20,456.51);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (393.78,376.38) rectangle (717.20,456.51);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] (393.78,390.05) --
	(717.20,390.05);

\path[draw=drawColor,line width= 0.3pt,line join=round] (393.78,410.11) --
	(717.20,410.11);

\path[draw=drawColor,line width= 0.3pt,line join=round] (393.78,430.16) --
	(717.20,430.16);

\path[draw=drawColor,line width= 0.3pt,line join=round] (393.78,450.22) --
	(717.20,450.22);

\path[draw=drawColor,line width= 0.6pt,line join=round] (393.78,380.02) --
	(717.20,380.02);

\path[draw=drawColor,line width= 0.6pt,line join=round] (393.78,400.08) --
	(717.20,400.08);

\path[draw=drawColor,line width= 0.6pt,line join=round] (393.78,420.14) --
	(717.20,420.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (393.78,440.19) --
	(717.20,440.19);

\path[draw=drawColor,line width= 0.6pt,line join=round] (420.73,376.38) --
	(420.73,456.51);

\path[draw=drawColor,line width= 0.6pt,line join=round] (465.65,376.38) --
	(465.65,456.51);

\path[draw=drawColor,line width= 0.6pt,line join=round] (510.57,376.38) --
	(510.57,456.51);

\path[draw=drawColor,line width= 0.6pt,line join=round] (555.49,376.38) --
	(555.49,456.51);

\path[draw=drawColor,line width= 0.6pt,line join=round] (600.41,376.38) --
	(600.41,456.51);

\path[draw=drawColor,line width= 0.6pt,line join=round] (645.33,376.38) --
	(645.33,456.51);

\path[draw=drawColor,line width= 0.6pt,line join=round] (690.25,376.38) --
	(690.25,456.51);
\definecolor{fillColor}{RGB}{221,204,119}

\path[fill=fillColor] (517.31,380.02) rectangle (530.78,386.76);

\path[fill=fillColor] (607.15,380.02) rectangle (620.62,380.42);

\path[fill=fillColor] (562.23,380.02) rectangle (575.70,383.79);

\path[fill=fillColor] (652.07,380.02) rectangle (665.54,387.16);

\path[fill=fillColor] (696.99,380.02) rectangle (710.46,380.90);

\path[fill=fillColor] (427.47,380.02) rectangle (440.94,415.32);

\path[fill=fillColor] (472.39,380.02) rectangle (485.86,406.10);
\definecolor{fillColor}{RGB}{17,119,51}

\path[fill=fillColor] (503.83,380.02) rectangle (517.31,385.64);

\path[fill=fillColor] (593.67,380.02) rectangle (607.15,380.26);

\path[fill=fillColor] (548.75,380.02) rectangle (562.23,383.47);

\path[fill=fillColor] (638.59,380.02) rectangle (652.07,387.48);

\path[fill=fillColor] (683.51,380.02) rectangle (696.99,381.07);

\path[fill=fillColor] (413.99,380.02) rectangle (427.47,414.60);

\path[fill=fillColor] (458.91,380.02) rectangle (472.39,407.94);
\definecolor{fillColor}{RGB}{204,102,119}

\path[fill=fillColor] (490.36,380.02) rectangle (503.83,388.77);

\path[fill=fillColor] (580.19,380.02) rectangle (593.67,380.74);

\path[fill=fillColor] (535.27,380.02) rectangle (548.75,384.35);

\path[fill=fillColor] (625.11,380.02) rectangle (638.59,386.52);

\path[fill=fillColor] (670.03,380.02) rectangle (683.51,380.74);

\path[fill=fillColor] (400.52,380.02) rectangle (413.99,416.61);

\path[fill=fillColor] (445.44,380.02) rectangle (458.91,402.81);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (393.78,376.38) rectangle (717.20,456.51);
\end{scope}
\begin{scope}
\path[clip] (393.78,290.75) rectangle (717.20,370.88);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (393.78,290.75) rectangle (717.20,370.88);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] (393.78,304.42) --
	(717.20,304.42);

\path[draw=drawColor,line width= 0.3pt,line join=round] (393.78,324.48) --
	(717.20,324.48);

\path[draw=drawColor,line width= 0.3pt,line join=round] (393.78,344.53) --
	(717.20,344.53);

\path[draw=drawColor,line width= 0.3pt,line join=round] (393.78,364.59) --
	(717.20,364.59);

\path[draw=drawColor,line width= 0.6pt,line join=round] (393.78,294.39) --
	(717.20,294.39);

\path[draw=drawColor,line width= 0.6pt,line join=round] (393.78,314.45) --
	(717.20,314.45);

\path[draw=drawColor,line width= 0.6pt,line join=round] (393.78,334.51) --
	(717.20,334.51);

\path[draw=drawColor,line width= 0.6pt,line join=round] (393.78,354.56) --
	(717.20,354.56);

\path[draw=drawColor,line width= 0.6pt,line join=round] (420.73,290.75) --
	(420.73,370.88);

\path[draw=drawColor,line width= 0.6pt,line join=round] (465.65,290.75) --
	(465.65,370.88);

\path[draw=drawColor,line width= 0.6pt,line join=round] (510.57,290.75) --
	(510.57,370.88);

\path[draw=drawColor,line width= 0.6pt,line join=round] (555.49,290.75) --
	(555.49,370.88);

\path[draw=drawColor,line width= 0.6pt,line join=round] (600.41,290.75) --
	(600.41,370.88);

\path[draw=drawColor,line width= 0.6pt,line join=round] (645.33,290.75) --
	(645.33,370.88);

\path[draw=drawColor,line width= 0.6pt,line join=round] (690.25,290.75) --
	(690.25,370.88);
\definecolor{fillColor}{RGB}{221,204,119}

\path[fill=fillColor] (517.31,294.39) rectangle (530.78,302.01);

\path[fill=fillColor] (607.15,294.39) rectangle (620.62,298.88);

\path[fill=fillColor] (562.23,294.39) rectangle (575.70,300.17);

\path[fill=fillColor] (652.07,294.39) rectangle (665.54,302.57);

\path[fill=fillColor] (696.99,294.39) rectangle (710.46,294.47);

\path[fill=fillColor] (427.47,294.39) rectangle (440.94,322.47);

\path[fill=fillColor] (472.39,294.39) rectangle (485.86,320.39);
\definecolor{fillColor}{RGB}{17,119,51}

\path[fill=fillColor] (503.83,294.39) rectangle (517.31,301.53);

\path[fill=fillColor] (593.67,294.39) rectangle (607.15,299.37);

\path[fill=fillColor] (548.75,294.39) rectangle (562.23,300.09);

\path[fill=fillColor] (638.59,294.39) rectangle (652.07,303.06);

\path[fill=fillColor] (683.51,294.39) rectangle (696.99,294.47);

\path[fill=fillColor] (413.99,294.39) rectangle (427.47,321.75);

\path[fill=fillColor] (458.91,294.39) rectangle (472.39,320.71);
\definecolor{fillColor}{RGB}{204,102,119}

\path[fill=fillColor] (490.36,294.39) rectangle (503.83,302.41);

\path[fill=fillColor] (580.19,294.39) rectangle (593.67,298.32);

\path[fill=fillColor] (535.27,294.39) rectangle (548.75,300.25);

\path[fill=fillColor] (625.11,294.39) rectangle (638.59,302.09);

\path[fill=fillColor] (670.03,294.39) rectangle (683.51,294.47);

\path[fill=fillColor] (400.52,294.39) rectangle (413.99,323.27);

\path[fill=fillColor] (445.44,294.39) rectangle (458.91,320.06);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (393.78,290.75) rectangle (717.20,370.88);
\end{scope}
\begin{scope}
\path[clip] (393.78,205.12) rectangle (717.20,285.25);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (393.78,205.12) rectangle (717.20,285.25);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] (393.78,218.79) --
	(717.20,218.79);

\path[draw=drawColor,line width= 0.3pt,line join=round] (393.78,238.85) --
	(717.20,238.85);

\path[draw=drawColor,line width= 0.3pt,line join=round] (393.78,258.90) --
	(717.20,258.90);

\path[draw=drawColor,line width= 0.3pt,line join=round] (393.78,278.96) --
	(717.20,278.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (393.78,208.76) --
	(717.20,208.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (393.78,228.82) --
	(717.20,228.82);

\path[draw=drawColor,line width= 0.6pt,line join=round] (393.78,248.87) --
	(717.20,248.87);

\path[draw=drawColor,line width= 0.6pt,line join=round] (393.78,268.93) --
	(717.20,268.93);

\path[draw=drawColor,line width= 0.6pt,line join=round] (420.73,205.12) --
	(420.73,285.25);

\path[draw=drawColor,line width= 0.6pt,line join=round] (465.65,205.12) --
	(465.65,285.25);

\path[draw=drawColor,line width= 0.6pt,line join=round] (510.57,205.12) --
	(510.57,285.25);

\path[draw=drawColor,line width= 0.6pt,line join=round] (555.49,205.12) --
	(555.49,285.25);

\path[draw=drawColor,line width= 0.6pt,line join=round] (600.41,205.12) --
	(600.41,285.25);

\path[draw=drawColor,line width= 0.6pt,line join=round] (645.33,205.12) --
	(645.33,285.25);

\path[draw=drawColor,line width= 0.6pt,line join=round] (690.25,205.12) --
	(690.25,285.25);
\definecolor{fillColor}{RGB}{221,204,119}

\path[fill=fillColor] (517.31,208.76) rectangle (530.78,231.30);

\path[fill=fillColor] (607.15,208.76) rectangle (620.62,212.69);

\path[fill=fillColor] (562.23,208.76) rectangle (575.70,210.53);

\path[fill=fillColor] (652.07,208.76) rectangle (665.54,222.08);

\path[fill=fillColor] (427.47,208.76) rectangle (440.94,237.64);

\path[fill=fillColor] (472.39,208.76) rectangle (485.86,218.63);
\definecolor{fillColor}{RGB}{17,119,51}

\path[fill=fillColor] (503.83,208.76) rectangle (517.31,230.66);

\path[fill=fillColor] (593.67,208.76) rectangle (607.15,212.53);

\path[fill=fillColor] (548.75,208.76) rectangle (562.23,210.69);

\path[fill=fillColor] (638.59,208.76) rectangle (652.07,221.04);

\path[fill=fillColor] (413.99,208.76) rectangle (427.47,238.12);

\path[fill=fillColor] (458.91,208.76) rectangle (472.39,219.67);
\definecolor{fillColor}{RGB}{204,102,119}

\path[fill=fillColor] (490.36,208.76) rectangle (503.83,231.87);

\path[fill=fillColor] (580.19,208.76) rectangle (593.67,212.85);

\path[fill=fillColor] (535.27,208.76) rectangle (548.75,210.28);

\path[fill=fillColor] (625.11,208.76) rectangle (638.59,223.12);

\path[fill=fillColor] (400.52,208.76) rectangle (413.99,237.16);

\path[fill=fillColor] (445.44,208.76) rectangle (458.91,217.51);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (393.78,205.12) rectangle (717.20,285.25);
\end{scope}
\begin{scope}
\path[clip] (377.21,547.64) rectangle (393.78,627.77);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (377.21,547.64) rectangle (393.78,627.77);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (388.52,587.71) {Colombia};
\end{scope}
\begin{scope}
\path[clip] (377.21,462.01) rectangle (393.78,542.14);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (377.21,462.01) rectangle (393.78,542.14);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (388.52,502.08) {Nigeria};
\end{scope}
\begin{scope}
\path[clip] (377.21,376.38) rectangle (393.78,456.51);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (377.21,376.38) rectangle (393.78,456.51);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (388.52,416.45) {Sierra Leone 1};
\end{scope}
\begin{scope}
\path[clip] (377.21,290.75) rectangle (393.78,370.88);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (377.21,290.75) rectangle (393.78,370.88);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (388.52,330.81) {Uganda 2};
\end{scope}
\begin{scope}
\path[clip] (377.21,205.12) rectangle (393.78,285.25);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (377.21,205.12) rectangle (393.78,285.25);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (388.52,245.18) {Russia};
\end{scope}
\begin{scope}
\path[clip] ( 31.71,547.64) rectangle ( 48.28,627.77);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 31.71,547.64) rectangle ( 48.28,627.77);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 43.03,587.71) {Burkina Faso};
\end{scope}
\begin{scope}
\path[clip] ( 31.71,462.01) rectangle ( 48.28,542.14);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 31.71,462.01) rectangle ( 48.28,542.14);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 43.03,502.08) {Nepal};
\end{scope}
\begin{scope}
\path[clip] ( 31.71,376.38) rectangle ( 48.28,456.51);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 31.71,376.38) rectangle ( 48.28,456.51);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 43.03,416.45) {Rwanda};
\end{scope}
\begin{scope}
\path[clip] ( 31.71,290.75) rectangle ( 48.28,370.88);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 31.71,290.75) rectangle ( 48.28,370.88);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 43.03,330.81) {Sierra Leone 2};
\end{scope}
\begin{scope}
\path[clip] ( 31.71,205.12) rectangle ( 48.28,285.25);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 31.71,205.12) rectangle ( 48.28,285.25);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 43.03,245.18) {All LMICs};
\end{scope}
\begin{scope}
\path[clip] ( 31.71,119.49) rectangle ( 48.28,199.62);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 31.71,119.49) rectangle ( 48.28,199.62);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 43.03,159.55) {USA};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 75.24,116.74) --
	( 75.24,119.49);

\path[draw=drawColor,line width= 0.6pt,line join=round] (120.16,116.74) --
	(120.16,119.49);

\path[draw=drawColor,line width= 0.6pt,line join=round] (165.08,116.74) --
	(165.08,119.49);

\path[draw=drawColor,line width= 0.6pt,line join=round] (209.99,116.74) --
	(209.99,119.49);

\path[draw=drawColor,line width= 0.6pt,line join=round] (254.91,116.74) --
	(254.91,119.49);

\path[draw=drawColor,line width= 0.6pt,line join=round] (299.83,116.74) --
	(299.83,119.49);

\path[draw=drawColor,line width= 0.6pt,line join=round] (344.75,116.74) --
	(344.75,119.49);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 78.27,114.54) {Health workers};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (118.43,114.54) {Government or};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (127.94,114.54) {MoH};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (163.35,114.54) {Family or};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (172.86,114.54) {Friends};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (198.77,114.54) {Famous person,};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (208.27,114.54) {religious leader};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (217.78,114.54) {or traditional};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (227.28,114.54) {healers};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (248.44,114.54) {Newspapers,};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (257.95,114.54) {radio or online};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (267.45,114.54) {groups};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (302.86,114.54) {Other};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (343.03,114.54) {Don't know or};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (352.54,114.54) {Refuse};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (420.73,202.37) --
	(420.73,205.12);

\path[draw=drawColor,line width= 0.6pt,line join=round] (465.65,202.37) --
	(465.65,205.12);

\path[draw=drawColor,line width= 0.6pt,line join=round] (510.57,202.37) --
	(510.57,205.12);

\path[draw=drawColor,line width= 0.6pt,line join=round] (555.49,202.37) --
	(555.49,205.12);

\path[draw=drawColor,line width= 0.6pt,line join=round] (600.41,202.37) --
	(600.41,205.12);

\path[draw=drawColor,line width= 0.6pt,line join=round] (645.33,202.37) --
	(645.33,205.12);

\path[draw=drawColor,line width= 0.6pt,line join=round] (690.25,202.37) --
	(690.25,205.12);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (423.76,200.17) {Health workers};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (463.93,200.17) {Government or};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (473.43,200.17) {MoH};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (508.85,200.17) {Family or};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (518.35,200.17) {Friends};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (544.26,200.17) {Famous person,};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (553.77,200.17) {religious leader};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (563.27,200.17) {or traditional};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (572.78,200.17) {healers};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (593.93,200.17) {Newspapers,};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (603.44,200.17) {radio or online};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (612.94,200.17) {groups};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (648.36,200.17) {Other};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (688.53,200.17) {Don't know or};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (698.03,200.17) {Refuse};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,548.25) {0};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,568.31) {25};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,588.37) {50};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,608.42) {75};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96,551.28) --
	( 31.71,551.28);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96,571.34) --
	( 31.71,571.34);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96,591.40) --
	( 31.71,591.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96,611.45) --
	( 31.71,611.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,462.62) {0};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,482.68) {25};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,502.74) {50};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,522.79) {75};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96,465.65) --
	( 31.71,465.65);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96,485.71) --
	( 31.71,485.71);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96,505.77) --
	( 31.71,505.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96,525.82) --
	( 31.71,525.82);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,376.99) {0};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,397.05) {25};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,417.11) {50};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,437.16) {75};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96,380.02) --
	( 31.71,380.02);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96,400.08) --
	( 31.71,400.08);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96,420.14) --
	( 31.71,420.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96,440.19) --
	( 31.71,440.19);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,291.36) {0};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,311.42) {25};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,331.47) {50};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,351.53) {75};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96,294.39) --
	( 31.71,294.39);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96,314.45) --
	( 31.71,314.45);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96,334.51) --
	( 31.71,334.51);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96,354.56) --
	( 31.71,354.56);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,205.73) {0};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,225.79) {25};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,245.84) {50};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,265.90) {75};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96,208.76) --
	( 31.71,208.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96,228.82) --
	( 31.71,228.82);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96,248.87) --
	( 31.71,248.87);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96,268.93) --
	( 31.71,268.93);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,120.10) {0};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,140.16) {25};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,160.21) {50};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,180.27) {75};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96,123.13) --
	( 31.71,123.13);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96,143.19) --
	( 31.71,143.19);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96,163.24) --
	( 31.71,163.24);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96,183.30) --
	( 31.71,183.30);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 13.08,373.63) {estimate};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (292.64,  5.50) rectangle (472.84, 30.95);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.10] at (298.14, 14.44) {Answer};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (339.17, 11.00) rectangle (353.62, 25.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{fillColor}{RGB}{204,102,119}

\path[fill=fillColor] (339.88, 11.71) rectangle (352.91, 24.74);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (391.63, 11.00) rectangle (406.08, 25.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{fillColor}{RGB}{17,119,51}

\path[fill=fillColor] (392.34, 11.71) rectangle (405.37, 24.74);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (435.90, 11.00) rectangle (450.35, 25.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{fillColor}{RGB}{221,204,119}

\path[fill=fillColor] (436.61, 11.71) rectangle (449.64, 24.74);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (359.12, 15.20) {Female};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (411.58, 15.20) {Male};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (455.85, 15.20) {All};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.32] at (  5.50,635.84) {If a COVID-19 vaccine becomes available in [country], would you take it?};
\end{scope}
\end{tikzpicture}

\end{document}