### 2021-03-23 paper_files/figure-latex/mainfigure1-1.tex

% Created by tikzDevice version 0.12.3.1 on 2021-03-23 16:52:22
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}

\nonstopmode

\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (650.43,542.02);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.5pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.00) rectangle (650.43,542.02);
\end{scope}
\begin{scope}
\path[clip] (152.00, 76.80) rectangle (271.61,506.36);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (152.00, 76.80) rectangle (271.61,506.36);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] (171.03, 76.80) --
	(171.03,506.36);

\path[draw=drawColor,line width= 0.3pt,line join=round] (198.21, 76.80) --
	(198.21,506.36);

\path[draw=drawColor,line width= 0.3pt,line join=round] (225.40, 76.80) --
	(225.40,506.36);

\path[draw=drawColor,line width= 0.3pt,line join=round] (252.58, 76.80) --
	(252.58,506.36);

\path[draw=drawColor,line width= 0.5pt,line join=round] (152.00, 92.71) --
	(271.61, 92.71);

\path[draw=drawColor,line width= 0.5pt,line join=round] (152.00,119.23) --
	(271.61,119.23);

\path[draw=drawColor,line width= 0.5pt,line join=round] (152.00,145.75) --
	(271.61,145.75);

\path[draw=drawColor,line width= 0.5pt,line join=round] (152.00,172.26) --
	(271.61,172.26);

\path[draw=drawColor,line width= 0.5pt,line join=round] (152.00,198.78) --
	(271.61,198.78);

\path[draw=drawColor,line width= 0.5pt,line join=round] (152.00,225.29) --
	(271.61,225.29);

\path[draw=drawColor,line width= 0.5pt,line join=round] (152.00,251.81) --
	(271.61,251.81);

\path[draw=drawColor,line width= 0.5pt,line join=round] (152.00,278.33) --
	(271.61,278.33);

\path[draw=drawColor,line width= 0.5pt,line join=round] (152.00,304.84) --
	(271.61,304.84);

\path[draw=drawColor,line width= 0.5pt,line join=round] (152.00,331.36) --
	(271.61,331.36);

\path[draw=drawColor,line width= 0.5pt,line join=round] (152.00,357.87) --
	(271.61,357.87);

\path[draw=drawColor,line width= 0.5pt,line join=round] (152.00,384.39) --
	(271.61,384.39);

\path[draw=drawColor,line width= 0.5pt,line join=round] (152.00,410.91) --
	(271.61,410.91);

\path[draw=drawColor,line width= 0.5pt,line join=round] (152.00,437.42) --
	(271.61,437.42);

\path[draw=drawColor,line width= 0.5pt,line join=round] (152.00,463.94) --
	(271.61,463.94);

\path[draw=drawColor,line width= 0.5pt,line join=round] (152.00,490.45) --
	(271.61,490.45);

\path[draw=drawColor,line width= 0.5pt,line join=round] (157.44, 76.80) --
	(157.44,506.36);

\path[draw=drawColor,line width= 0.5pt,line join=round] (184.62, 76.80) --
	(184.62,506.36);

\path[draw=drawColor,line width= 0.5pt,line join=round] (211.80, 76.80) --
	(211.80,506.36);

\path[draw=drawColor,line width= 0.5pt,line join=round] (238.99, 76.80) --
	(238.99,506.36);

\path[draw=drawColor,line width= 0.5pt,line join=round] (266.17, 76.80) --
	(266.17,506.36);
\definecolor{drawColor}{RGB}{204,102,119}

\path[draw=drawColor,line width= 0.6pt,line join=round] (250.51,143.09) --
	(250.51,148.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] (250.51,145.75) --
	(238.88,145.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (238.88,143.09) --
	(238.88,148.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] (233.01,487.80) --
	(233.01,493.10);

\path[draw=drawColor,line width= 0.6pt,line join=round] (233.01,490.45) --
	(226.48,490.45);

\path[draw=drawColor,line width= 0.6pt,line join=round] (226.48,487.80) --
	(226.48,493.10);

\path[draw=drawColor,line width= 0.6pt,line join=round] (241.81,461.29) --
	(241.81,466.59);

\path[draw=drawColor,line width= 0.6pt,line join=round] (241.81,463.94) --
	(235.94,463.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (235.94,461.29) --
	(235.94,466.59);

\path[draw=drawColor,line width= 0.6pt,line join=round] (251.38,434.77) --
	(251.38,440.07);

\path[draw=drawColor,line width= 0.6pt,line join=round] (251.38,437.42) --
	(247.47,437.42);

\path[draw=drawColor,line width= 0.6pt,line join=round] (247.47,434.77) --
	(247.47,440.07);

\path[draw=drawColor,line width= 0.6pt,line join=round] (257.15,408.25) --
	(257.15,413.56);

\path[draw=drawColor,line width= 0.6pt,line join=round] (257.15,410.91) --
	(251.49,410.91);

\path[draw=drawColor,line width= 0.6pt,line join=round] (251.49,408.25) --
	(251.49,413.56);

\path[draw=drawColor,line width= 0.6pt,line join=round] (263.56,381.74) --
	(263.56,387.04);

\path[draw=drawColor,line width= 0.6pt,line join=round] (263.56,384.39) --
	(261.28,384.39);

\path[draw=drawColor,line width= 0.6pt,line join=round] (261.28,381.74) --
	(261.28,387.04);

\path[draw=drawColor,line width= 0.6pt,line join=round] (242.47,355.22) --
	(242.47,360.52);

\path[draw=drawColor,line width= 0.6pt,line join=round] (242.47,357.87) --
	(238.23,357.87);

\path[draw=drawColor,line width= 0.6pt,line join=round] (238.23,355.22) --
	(238.23,360.52);

\path[draw=drawColor,line width= 0.6pt,line join=round] (246.92,328.71) --
	(246.92,334.01);

\path[draw=drawColor,line width= 0.6pt,line join=round] (246.92,331.36) --
	(233.55,331.36);

\path[draw=drawColor,line width= 0.6pt,line join=round] (233.55,328.71) --
	(233.55,334.01);

\path[draw=drawColor,line width= 0.6pt,line join=round] (232.35,302.19) --
	(232.35,307.49);

\path[draw=drawColor,line width= 0.6pt,line join=round] (232.35,304.84) --
	(227.13,304.84);

\path[draw=drawColor,line width= 0.6pt,line join=round] (227.13,302.19) --
	(227.13,307.49);

\path[draw=drawColor,line width= 0.6pt,line join=round] (191.91,116.58) --
	(191.91,121.88);

\path[draw=drawColor,line width= 0.6pt,line join=round] (191.91,119.23) --
	(189.08,119.23);

\path[draw=drawColor,line width= 0.6pt,line join=round] (189.08,116.58) --
	(189.08,121.88);

\path[draw=drawColor,line width= 0.6pt,line join=round] (251.82,275.67) --
	(251.82,280.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (251.82,278.33) --
	(247.58,278.33);

\path[draw=drawColor,line width= 0.6pt,line join=round] (247.58,275.67) --
	(247.58,280.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (244.97,249.16) --
	(244.97,254.46);

\path[draw=drawColor,line width= 0.6pt,line join=round] (244.97,251.81) --
	(239.53,251.81);

\path[draw=drawColor,line width= 0.6pt,line join=round] (239.53,249.16) --
	(239.53,254.46);

\path[draw=drawColor,line width= 0.6pt,line join=round] (254.86,222.64) --
	(254.86,227.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (254.86,225.29) --
	(251.17,225.29);

\path[draw=drawColor,line width= 0.6pt,line join=round] (251.17,222.64) --
	(251.17,227.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (252.25,196.13) --
	(252.25,201.43);

\path[draw=drawColor,line width= 0.6pt,line join=round] (252.25,198.78) --
	(249.21,198.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (249.21,196.13) --
	(249.21,201.43);

\path[draw=drawColor,line width= 0.6pt,line join=round] (243.01,169.61) --
	(243.01,174.91);

\path[draw=drawColor,line width= 0.6pt,line join=round] (243.01,172.26) --
	(238.23,172.26);

\path[draw=drawColor,line width= 0.6pt,line join=round] (238.23,169.61) --
	(238.23,174.91);

\path[draw=drawColor,line width= 0.6pt,line join=round] (230.61, 90.06) --
	(230.61, 95.37);

\path[draw=drawColor,line width= 0.6pt,line join=round] (230.61, 92.71) --
	(224.63, 92.71);

\path[draw=drawColor,line width= 0.6pt,line join=round] (224.63, 90.06) --
	(224.63, 95.37);
\definecolor{fillColor}{RGB}{204,102,119}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (244.75,145.75) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (229.74,490.45) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (238.88,463.94) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (249.43,437.42) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (254.32,410.91) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (262.47,384.39) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (240.29,357.87) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (240.18,331.36) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (229.74,304.84) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (190.49,119.23) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (249.75,278.33) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (242.25,251.81) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (253.01,225.29) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (250.73,198.78) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (240.62,172.26) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (227.68, 92.71) circle (  1.96);
\definecolor{drawColor}{RGB}{169,169,169}

\path[draw=drawColor,line width= 0.6pt,line join=round] (152.00,159.00) -- (271.61,159.00);

\path[draw=drawColor,line width= 0.6pt,line join=round] (152.00,132.49) -- (271.61,132.49);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.5pt,line join=round,line cap=round] (152.00, 76.80) rectangle (271.61,506.36);
\end{scope}
\begin{scope}
\path[clip] (276.61, 76.80) rectangle (396.21,506.36);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (276.61, 76.80) rectangle (396.21,506.36);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] (295.64, 76.80) --
	(295.64,506.36);

\path[draw=drawColor,line width= 0.3pt,line join=round] (322.82, 76.80) --
	(322.82,506.36);

\path[draw=drawColor,line width= 0.3pt,line join=round] (350.00, 76.80) --
	(350.00,506.36);

\path[draw=drawColor,line width= 0.3pt,line join=round] (377.19, 76.80) --
	(377.19,506.36);

\path[draw=drawColor,line width= 0.5pt,line join=round] (276.61, 92.71) --
	(396.21, 92.71);

\path[draw=drawColor,line width= 0.5pt,line join=round] (276.61,119.23) --
	(396.21,119.23);

\path[draw=drawColor,line width= 0.5pt,line join=round] (276.61,145.75) --
	(396.21,145.75);

\path[draw=drawColor,line width= 0.5pt,line join=round] (276.61,172.26) --
	(396.21,172.26);

\path[draw=drawColor,line width= 0.5pt,line join=round] (276.61,198.78) --
	(396.21,198.78);

\path[draw=drawColor,line width= 0.5pt,line join=round] (276.61,225.29) --
	(396.21,225.29);

\path[draw=drawColor,line width= 0.5pt,line join=round] (276.61,251.81) --
	(396.21,251.81);

\path[draw=drawColor,line width= 0.5pt,line join=round] (276.61,278.33) --
	(396.21,278.33);

\path[draw=drawColor,line width= 0.5pt,line join=round] (276.61,304.84) --
	(396.21,304.84);

\path[draw=drawColor,line width= 0.5pt,line join=round] (276.61,331.36) --
	(396.21,331.36);

\path[draw=drawColor,line width= 0.5pt,line join=round] (276.61,357.87) --
	(396.21,357.87);

\path[draw=drawColor,line width= 0.5pt,line join=round] (276.61,384.39) --
	(396.21,384.39);

\path[draw=drawColor,line width= 0.5pt,line join=round] (276.61,410.91) --
	(396.21,410.91);

\path[draw=drawColor,line width= 0.5pt,line join=round] (276.61,437.42) --
	(396.21,437.42);

\path[draw=drawColor,line width= 0.5pt,line join=round] (276.61,463.94) --
	(396.21,463.94);

\path[draw=drawColor,line width= 0.5pt,line join=round] (276.61,490.45) --
	(396.21,490.45);

\path[draw=drawColor,line width= 0.5pt,line join=round] (282.04, 76.80) --
	(282.04,506.36);

\path[draw=drawColor,line width= 0.5pt,line join=round] (309.23, 76.80) --
	(309.23,506.36);

\path[draw=drawColor,line width= 0.5pt,line join=round] (336.41, 76.80) --
	(336.41,506.36);

\path[draw=drawColor,line width= 0.5pt,line join=round] (363.59, 76.80) --
	(363.59,506.36);

\path[draw=drawColor,line width= 0.5pt,line join=round] (390.78, 76.80) --
	(390.78,506.36);
\definecolor{drawColor}{RGB}{153,153,51}

\path[draw=drawColor,line width= 0.6pt,line join=round] (374.36,148.40) --
	(374.36,151.05);

\path[draw=drawColor,line width= 0.6pt,line join=round] (374.36,149.72) --
	(361.75,149.72);

\path[draw=drawColor,line width= 0.6pt,line join=round] (361.75,148.40) --
	(361.75,151.05);
\definecolor{drawColor}{RGB}{68,170,153}

\path[draw=drawColor,line width= 0.6pt,line join=round] (377.73,140.44) --
	(377.73,143.09);

\path[draw=drawColor,line width= 0.6pt,line join=round] (377.73,141.77) --
	(366.20,141.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (366.20,140.44) --
	(366.20,143.09);
\definecolor{drawColor}{RGB}{153,153,51}

\path[draw=drawColor,line width= 0.6pt,line join=round] (355.87,493.10) --
	(355.87,495.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (355.87,494.43) --
	(343.26,494.43);

\path[draw=drawColor,line width= 0.6pt,line join=round] (343.26,493.10) --
	(343.26,495.76);
\definecolor{drawColor}{RGB}{68,170,153}

\path[draw=drawColor,line width= 0.6pt,line join=round] (360.22,485.15) --
	(360.22,487.80);

\path[draw=drawColor,line width= 0.6pt,line join=round] (360.22,486.48) --
	(352.72,486.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] (352.72,485.15) --
	(352.72,487.80);
\definecolor{drawColor}{RGB}{153,153,51}

\path[draw=drawColor,line width= 0.6pt,line join=round] (365.77,466.59) --
	(365.77,469.24);

\path[draw=drawColor,line width= 0.6pt,line join=round] (365.77,467.91) --
	(358.27,467.91);

\path[draw=drawColor,line width= 0.6pt,line join=round] (358.27,466.59) --
	(358.27,469.24);
\definecolor{drawColor}{RGB}{68,170,153}

\path[draw=drawColor,line width= 0.6pt,line join=round] (370.88,458.63) --
	(370.88,461.29);

\path[draw=drawColor,line width= 0.6pt,line join=round] (370.88,459.96) --
	(361.42,459.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (361.42,458.63) --
	(361.42,461.29);
\definecolor{drawColor}{RGB}{153,153,51}

\path[draw=drawColor,line width= 0.6pt,line join=round] (373.49,440.07) --
	(373.49,442.72);

\path[draw=drawColor,line width= 0.6pt,line join=round] (373.49,441.40) --
	(366.42,441.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] (366.42,440.07) --
	(366.42,442.72);
\definecolor{drawColor}{RGB}{68,170,153}

\path[draw=drawColor,line width= 0.6pt,line join=round] (377.19,432.12) --
	(377.19,434.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (377.19,433.44) --
	(373.05,433.44);

\path[draw=drawColor,line width= 0.6pt,line join=round] (373.05,432.12) --
	(373.05,434.77);
\definecolor{drawColor}{RGB}{153,153,51}

\path[draw=drawColor,line width= 0.6pt,line join=round] (379.90,413.56) --
	(379.90,416.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] (379.90,414.88) --
	(371.75,414.88);

\path[draw=drawColor,line width= 0.6pt,line join=round] (371.75,413.56) --
	(371.75,416.21);
\definecolor{drawColor}{RGB}{68,170,153}

\path[draw=drawColor,line width= 0.6pt,line join=round] (384.36,405.60) --
	(384.36,408.25);

\path[draw=drawColor,line width= 0.6pt,line join=round] (384.36,406.93) --
	(378.16,406.93);

\path[draw=drawColor,line width= 0.6pt,line join=round] (378.16,405.60) --
	(378.16,408.25);
\definecolor{drawColor}{RGB}{153,153,51}

\path[draw=drawColor,line width= 0.6pt,line join=round] (388.82,387.04) --
	(388.82,389.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (388.82,388.37) --
	(384.91,388.37);

\path[draw=drawColor,line width= 0.6pt,line join=round] (384.91,387.04) --
	(384.91,389.69);
\definecolor{drawColor}{RGB}{68,170,153}

\path[draw=drawColor,line width= 0.6pt,line join=round] (388.28,379.09) --
	(388.28,381.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] (388.28,380.41) --
	(385.45,380.41);

\path[draw=drawColor,line width= 0.6pt,line join=round] (385.45,379.09) --
	(385.45,381.74);
\definecolor{drawColor}{RGB}{153,153,51}

\path[draw=drawColor,line width= 0.6pt,line join=round] (366.97,360.52) --
	(366.97,363.18);

\path[draw=drawColor,line width= 0.6pt,line join=round] (366.97,361.85) --
	(360.01,361.85);

\path[draw=drawColor,line width= 0.6pt,line join=round] (360.01,360.52) --
	(360.01,363.18);
\definecolor{drawColor}{RGB}{68,170,153}

\path[draw=drawColor,line width= 0.6pt,line join=round] (368.38,352.57) --
	(368.38,355.22);

\path[draw=drawColor,line width= 0.6pt,line join=round] (368.38,353.90) --
	(363.16,353.90);

\path[draw=drawColor,line width= 0.6pt,line join=round] (363.16,352.57) --
	(363.16,355.22);
\definecolor{drawColor}{RGB}{153,153,51}

\path[draw=drawColor,line width= 0.6pt,line join=round] (367.73,334.01) --
	(367.73,336.66);

\path[draw=drawColor,line width= 0.6pt,line join=round] (367.73,335.33) --
	(353.37,335.33);

\path[draw=drawColor,line width= 0.6pt,line join=round] (353.37,334.01) --
	(353.37,336.66);
\definecolor{drawColor}{RGB}{68,170,153}

\path[draw=drawColor,line width= 0.6pt,line join=round] (375.99,326.05) --
	(375.99,328.71);

\path[draw=drawColor,line width= 0.6pt,line join=round] (375.99,327.38) --
	(362.29,327.38);

\path[draw=drawColor,line width= 0.6pt,line join=round] (362.29,326.05) --
	(362.29,328.71);
\definecolor{drawColor}{RGB}{153,153,51}

\path[draw=drawColor,line width= 0.6pt,line join=round] (308.36,121.88) --
	(308.36,124.53);

\path[draw=drawColor,line width= 0.6pt,line join=round] (308.36,123.21) --
	(304.77,123.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] (304.77,121.88) --
	(304.77,124.53);
\definecolor{drawColor}{RGB}{68,170,153}

\path[draw=drawColor,line width= 0.6pt,line join=round] (326.08,113.93) --
	(326.08,116.58);

\path[draw=drawColor,line width= 0.6pt,line join=round] (326.08,115.25) --
	(321.73,115.25);

\path[draw=drawColor,line width= 0.6pt,line join=round] (321.73,113.93) --
	(321.73,116.58);
\definecolor{drawColor}{RGB}{153,153,51}

\path[draw=drawColor,line width= 0.6pt,line join=round] (372.29,280.98) --
	(372.29,283.63);

\path[draw=drawColor,line width= 0.6pt,line join=round] (372.29,282.30) --
	(364.46,282.30);

\path[draw=drawColor,line width= 0.6pt,line join=round] (364.46,280.98) --
	(364.46,283.63);
\definecolor{drawColor}{RGB}{68,170,153}

\path[draw=drawColor,line width= 0.6pt,line join=round] (380.12,273.02) --
	(380.12,275.67);

\path[draw=drawColor,line width= 0.6pt,line join=round] (380.12,274.35) --
	(375.34,274.35);

\path[draw=drawColor,line width= 0.6pt,line join=round] (375.34,273.02) --
	(375.34,275.67);
\definecolor{drawColor}{RGB}{153,153,51}

\path[draw=drawColor,line width= 0.6pt,line join=round] (367.62,254.46) --
	(367.62,257.11);

\path[draw=drawColor,line width= 0.6pt,line join=round] (367.62,255.79) --
	(357.61,255.79);

\path[draw=drawColor,line width= 0.6pt,line join=round] (357.61,254.46) --
	(357.61,257.11);
\definecolor{drawColor}{RGB}{68,170,153}

\path[draw=drawColor,line width= 0.6pt,line join=round] (372.40,246.51) --
	(372.40,249.16);

\path[draw=drawColor,line width= 0.6pt,line join=round] (372.40,247.83) --
	(365.99,247.83);

\path[draw=drawColor,line width= 0.6pt,line join=round] (365.99,246.51) --
	(365.99,249.16);
\definecolor{drawColor}{RGB}{153,153,51}

\path[draw=drawColor,line width= 0.6pt,line join=round] (381.54,227.95) --
	(381.54,230.60);

\path[draw=drawColor,line width= 0.6pt,line join=round] (381.54,229.27) --
	(375.23,229.27);

\path[draw=drawColor,line width= 0.6pt,line join=round] (375.23,227.95) --
	(375.23,230.60);
\definecolor{drawColor}{RGB}{68,170,153}

\path[draw=drawColor,line width= 0.6pt,line join=round] (379.36,219.99) --
	(379.36,222.64);

\path[draw=drawColor,line width= 0.6pt,line join=round] (379.36,221.32) --
	(375.45,221.32);

\path[draw=drawColor,line width= 0.6pt,line join=round] (375.45,219.99) --
	(375.45,222.64);
\definecolor{drawColor}{RGB}{153,153,51}

\path[draw=drawColor,line width= 0.6pt,line join=round] (376.86,196.13) --
	(376.86,201.43);

\path[draw=drawColor,line width= 0.6pt,line join=round] (376.86,198.78) --
	(373.82,198.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (373.82,196.13) --
	(373.82,201.43);

\path[draw=drawColor,line width= 0.6pt,line join=round] (367.18,174.91) --
	(367.18,177.56);

\path[draw=drawColor,line width= 0.6pt,line join=round] (367.18,176.24) --
	(359.79,176.24);

\path[draw=drawColor,line width= 0.6pt,line join=round] (359.79,174.91) --
	(359.79,177.56);
\definecolor{drawColor}{RGB}{68,170,153}

\path[draw=drawColor,line width= 0.6pt,line join=round] (370.01,166.96) --
	(370.01,169.61);

\path[draw=drawColor,line width= 0.6pt,line join=round] (370.01,168.28) --
	(363.81,168.28);

\path[draw=drawColor,line width= 0.6pt,line join=round] (363.81,166.96) --
	(363.81,169.61);
\definecolor{drawColor}{RGB}{153,153,51}

\path[draw=drawColor,line width= 0.6pt,line join=round] (347.39, 95.37) --
	(347.39, 98.02);

\path[draw=drawColor,line width= 0.6pt,line join=round] (347.39, 96.69) --
	(338.69, 96.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (338.69, 95.37) --
	(338.69, 98.02);
\definecolor{drawColor}{RGB}{68,170,153}

\path[draw=drawColor,line width= 0.6pt,line join=round] (365.66, 87.41) --
	(365.66, 90.06);

\path[draw=drawColor,line width= 0.6pt,line join=round] (365.66, 88.74) --
	(357.94, 88.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] (357.94, 87.41) --
	(357.94, 90.06);
\definecolor{drawColor}{RGB}{153,153,51}
\definecolor{fillColor}{RGB}{153,153,51}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (368.05,149.72) circle (  1.96);
\definecolor{drawColor}{RGB}{68,170,153}
\definecolor{fillColor}{RGB}{68,170,153}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (371.97,141.77) circle (  1.96);
\definecolor{drawColor}{RGB}{153,153,51}
\definecolor{fillColor}{RGB}{153,153,51}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (349.57,494.43) circle (  1.96);
\definecolor{drawColor}{RGB}{68,170,153}
\definecolor{fillColor}{RGB}{68,170,153}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (356.42,486.48) circle (  1.96);
\definecolor{drawColor}{RGB}{153,153,51}
\definecolor{fillColor}{RGB}{153,153,51}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (361.96,467.91) circle (  1.96);
\definecolor{drawColor}{RGB}{68,170,153}
\definecolor{fillColor}{RGB}{68,170,153}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (366.10,459.96) circle (  1.96);
\definecolor{drawColor}{RGB}{153,153,51}
\definecolor{fillColor}{RGB}{153,153,51}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (369.90,441.40) circle (  1.96);
\definecolor{drawColor}{RGB}{68,170,153}
\definecolor{fillColor}{RGB}{68,170,153}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (375.12,433.44) circle (  1.96);
\definecolor{drawColor}{RGB}{153,153,51}
\definecolor{fillColor}{RGB}{153,153,51}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (375.77,414.88) circle (  1.96);
\definecolor{drawColor}{RGB}{68,170,153}
\definecolor{fillColor}{RGB}{68,170,153}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (381.32,406.93) circle (  1.96);
\definecolor{drawColor}{RGB}{153,153,51}
\definecolor{fillColor}{RGB}{153,153,51}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (386.86,388.37) circle (  1.96);
\definecolor{drawColor}{RGB}{68,170,153}
\definecolor{fillColor}{RGB}{68,170,153}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (386.86,380.41) circle (  1.96);
\definecolor{drawColor}{RGB}{153,153,51}
\definecolor{fillColor}{RGB}{153,153,51}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (363.49,361.85) circle (  1.96);
\definecolor{drawColor}{RGB}{68,170,153}
\definecolor{fillColor}{RGB}{68,170,153}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (365.77,353.90) circle (  1.96);
\definecolor{drawColor}{RGB}{153,153,51}
\definecolor{fillColor}{RGB}{153,153,51}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (360.55,335.33) circle (  1.96);
\definecolor{drawColor}{RGB}{68,170,153}
\definecolor{fillColor}{RGB}{68,170,153}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (369.14,327.38) circle (  1.96);
\definecolor{drawColor}{RGB}{153,153,51}
\definecolor{fillColor}{RGB}{153,153,51}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (306.62,123.21) circle (  1.96);
\definecolor{drawColor}{RGB}{68,170,153}
\definecolor{fillColor}{RGB}{68,170,153}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (323.91,115.25) circle (  1.96);
\definecolor{drawColor}{RGB}{153,153,51}
\definecolor{fillColor}{RGB}{153,153,51}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (368.38,282.30) circle (  1.96);
\definecolor{drawColor}{RGB}{68,170,153}
\definecolor{fillColor}{RGB}{68,170,153}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (377.73,274.35) circle (  1.96);
\definecolor{drawColor}{RGB}{153,153,51}
\definecolor{fillColor}{RGB}{153,153,51}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (362.62,255.79) circle (  1.96);
\definecolor{drawColor}{RGB}{68,170,153}
\definecolor{fillColor}{RGB}{68,170,153}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (369.14,247.83) circle (  1.96);
\definecolor{drawColor}{RGB}{153,153,51}
\definecolor{fillColor}{RGB}{153,153,51}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (378.38,229.27) circle (  1.96);
\definecolor{drawColor}{RGB}{68,170,153}
\definecolor{fillColor}{RGB}{68,170,153}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (377.40,221.32) circle (  1.96);
\definecolor{drawColor}{RGB}{153,153,51}
\definecolor{fillColor}{RGB}{153,153,51}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (375.34,198.78) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (363.49,176.24) circle (  1.96);
\definecolor{drawColor}{RGB}{68,170,153}
\definecolor{fillColor}{RGB}{68,170,153}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (366.86,168.28) circle (  1.96);
\definecolor{drawColor}{RGB}{153,153,51}
\definecolor{fillColor}{RGB}{153,153,51}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (343.04, 96.69) circle (  1.96);
\definecolor{drawColor}{RGB}{68,170,153}
\definecolor{fillColor}{RGB}{68,170,153}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (361.85, 88.74) circle (  1.96);
\definecolor{drawColor}{RGB}{169,169,169}

\path[draw=drawColor,line width= 0.6pt,line join=round] (276.61,159.00) -- (396.21,159.00);

\path[draw=drawColor,line width= 0.6pt,line join=round] (276.61,132.49) -- (396.21,132.49);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.5pt,line join=round,line cap=round] (276.61, 76.80) rectangle (396.21,506.36);
\end{scope}
\begin{scope}
\path[clip] (401.21, 76.80) rectangle (520.82,506.36);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (401.21, 76.80) rectangle (520.82,506.36);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] (420.24, 76.80) --
	(420.24,506.36);

\path[draw=drawColor,line width= 0.3pt,line join=round] (447.43, 76.80) --
	(447.43,506.36);

\path[draw=drawColor,line width= 0.3pt,line join=round] (474.61, 76.80) --
	(474.61,506.36);

\path[draw=drawColor,line width= 0.3pt,line join=round] (501.79, 76.80) --
	(501.79,506.36);

\path[draw=drawColor,line width= 0.5pt,line join=round] (401.21, 92.71) --
	(520.82, 92.71);

\path[draw=drawColor,line width= 0.5pt,line join=round] (401.21,119.23) --
	(520.82,119.23);

\path[draw=drawColor,line width= 0.5pt,line join=round] (401.21,145.75) --
	(520.82,145.75);

\path[draw=drawColor,line width= 0.5pt,line join=round] (401.21,172.26) --
	(520.82,172.26);

\path[draw=drawColor,line width= 0.5pt,line join=round] (401.21,198.78) --
	(520.82,198.78);

\path[draw=drawColor,line width= 0.5pt,line join=round] (401.21,225.29) --
	(520.82,225.29);

\path[draw=drawColor,line width= 0.5pt,line join=round] (401.21,251.81) --
	(520.82,251.81);

\path[draw=drawColor,line width= 0.5pt,line join=round] (401.21,278.33) --
	(520.82,278.33);

\path[draw=drawColor,line width= 0.5pt,line join=round] (401.21,304.84) --
	(520.82,304.84);

\path[draw=drawColor,line width= 0.5pt,line join=round] (401.21,331.36) --
	(520.82,331.36);

\path[draw=drawColor,line width= 0.5pt,line join=round] (401.21,357.87) --
	(520.82,357.87);

\path[draw=drawColor,line width= 0.5pt,line join=round] (401.21,384.39) --
	(520.82,384.39);

\path[draw=drawColor,line width= 0.5pt,line join=round] (401.21,410.91) --
	(520.82,410.91);

\path[draw=drawColor,line width= 0.5pt,line join=round] (401.21,437.42) --
	(520.82,437.42);

\path[draw=drawColor,line width= 0.5pt,line join=round] (401.21,463.94) --
	(520.82,463.94);

\path[draw=drawColor,line width= 0.5pt,line join=round] (401.21,490.45) --
	(520.82,490.45);

\path[draw=drawColor,line width= 0.5pt,line join=round] (406.65, 76.80) --
	(406.65,506.36);

\path[draw=drawColor,line width= 0.5pt,line join=round] (433.84, 76.80) --
	(433.84,506.36);

\path[draw=drawColor,line width= 0.5pt,line join=round] (461.02, 76.80) --
	(461.02,506.36);

\path[draw=drawColor,line width= 0.5pt,line join=round] (488.20, 76.80) --
	(488.20,506.36);

\path[draw=drawColor,line width= 0.5pt,line join=round] (515.39, 76.80) --
	(515.39,506.36);
\definecolor{drawColor}{RGB}{51,34,136}

\path[draw=drawColor,line width= 0.6pt,line join=round] (497.55,140.44) --
	(497.55,143.09);

\path[draw=drawColor,line width= 0.6pt,line join=round] (497.55,141.77) --
	(484.29,141.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (484.29,140.44) --
	(484.29,143.09);
\definecolor{drawColor}{RGB}{170,68,153}

\path[draw=drawColor,line width= 0.6pt,line join=round] (499.51,148.40) --
	(499.51,151.05);

\path[draw=drawColor,line width= 0.6pt,line join=round] (499.51,149.72) --
	(487.22,149.72);

\path[draw=drawColor,line width= 0.6pt,line join=round] (487.22,148.40) --
	(487.22,151.05);
\definecolor{drawColor}{RGB}{51,34,136}

\path[draw=drawColor,line width= 0.6pt,line join=round] (478.20,485.15) --
	(478.20,487.80);

\path[draw=drawColor,line width= 0.6pt,line join=round] (478.20,486.48) --
	(467.43,486.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] (467.43,485.15) --
	(467.43,487.80);
\definecolor{drawColor}{RGB}{170,68,153}

\path[draw=drawColor,line width= 0.6pt,line join=round] (486.90,493.10) --
	(486.90,495.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (486.90,494.43) --
	(478.85,494.43);

\path[draw=drawColor,line width= 0.6pt,line join=round] (478.85,493.10) --
	(478.85,495.76);
\definecolor{drawColor}{RGB}{51,34,136}

\path[draw=drawColor,line width= 0.6pt,line join=round] (496.36,458.63) --
	(496.36,461.29);

\path[draw=drawColor,line width= 0.6pt,line join=round] (496.36,459.96) --
	(486.68,459.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (486.68,458.63) --
	(486.68,461.29);
\definecolor{drawColor}{RGB}{170,68,153}

\path[draw=drawColor,line width= 0.6pt,line join=round] (490.16,466.59) --
	(490.16,469.24);

\path[draw=drawColor,line width= 0.6pt,line join=round] (490.16,467.91) --
	(482.87,467.91);

\path[draw=drawColor,line width= 0.6pt,line join=round] (482.87,466.59) --
	(482.87,469.24);
\definecolor{drawColor}{RGB}{51,34,136}

\path[draw=drawColor,line width= 0.6pt,line join=round] (510.38,432.12) --
	(510.38,434.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (510.38,433.44) --
	(497.44,433.44);

\path[draw=drawColor,line width= 0.6pt,line join=round] (497.44,432.12) --
	(497.44,434.77);
\definecolor{drawColor}{RGB}{170,68,153}

\path[draw=drawColor,line width= 0.6pt,line join=round] (503.32,440.07) --
	(503.32,442.72);

\path[draw=drawColor,line width= 0.6pt,line join=round] (503.32,441.40) --
	(496.47,441.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] (496.47,440.07) --
	(496.47,442.72);
\definecolor{drawColor}{RGB}{51,34,136}

\path[draw=drawColor,line width= 0.6pt,line join=round] (504.95,405.60) --
	(504.95,408.25);

\path[draw=drawColor,line width= 0.6pt,line join=round] (504.95,406.93) --
	(495.60,406.93);

\path[draw=drawColor,line width= 0.6pt,line join=round] (495.60,405.60) --
	(495.60,408.25);
\definecolor{drawColor}{RGB}{170,68,153}

\path[draw=drawColor,line width= 0.6pt,line join=round] (507.56,413.56) --
	(507.56,416.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] (507.56,414.88) --
	(500.71,414.88);

\path[draw=drawColor,line width= 0.6pt,line join=round] (500.71,413.56) --
	(500.71,416.21);
\definecolor{drawColor}{RGB}{51,34,136}

\path[draw=drawColor,line width= 0.6pt,line join=round] (505.06,326.05) --
	(505.06,328.71);

\path[draw=drawColor,line width= 0.6pt,line join=round] (505.06,327.38) --
	(489.94,327.38);

\path[draw=drawColor,line width= 0.6pt,line join=round] (489.94,326.05) --
	(489.94,328.71);
\definecolor{drawColor}{RGB}{170,68,153}

\path[draw=drawColor,line width= 0.6pt,line join=round] (494.18,334.01) --
	(494.18,336.66);

\path[draw=drawColor,line width= 0.6pt,line join=round] (494.18,335.33) --
	(479.94,335.33);

\path[draw=drawColor,line width= 0.6pt,line join=round] (479.94,334.01) --
	(479.94,336.66);
\definecolor{drawColor}{RGB}{51,34,136}

\path[draw=drawColor,line width= 0.6pt,line join=round] (488.75,299.54) --
	(488.75,302.19);

\path[draw=drawColor,line width= 0.6pt,line join=round] (488.75,300.86) --
	(479.83,300.86);

\path[draw=drawColor,line width= 0.6pt,line join=round] (479.83,299.54) --
	(479.83,302.19);
\definecolor{drawColor}{RGB}{170,68,153}

\path[draw=drawColor,line width= 0.6pt,line join=round] (479.61,307.49) --
	(479.61,310.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (479.61,308.82) --
	(473.20,308.82);

\path[draw=drawColor,line width= 0.6pt,line join=round] (473.20,307.49) --
	(473.20,310.14);
\definecolor{drawColor}{RGB}{51,34,136}

\path[draw=drawColor,line width= 0.6pt,line join=round] (441.99,113.93) --
	(441.99,116.58);

\path[draw=drawColor,line width= 0.6pt,line join=round] (441.99,115.25) --
	(438.84,115.25);

\path[draw=drawColor,line width= 0.6pt,line join=round] (438.84,113.93) --
	(438.84,116.58);
\definecolor{drawColor}{RGB}{170,68,153}

\path[draw=drawColor,line width= 0.6pt,line join=round] (441.45,121.88) --
	(441.45,124.53);

\path[draw=drawColor,line width= 0.6pt,line join=round] (441.45,123.21) --
	(436.34,123.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] (436.34,121.88) --
	(436.34,124.53);
\definecolor{drawColor}{RGB}{51,34,136}

\path[draw=drawColor,line width= 0.6pt,line join=round] (490.59,273.02) --
	(490.59,275.67);

\path[draw=drawColor,line width= 0.6pt,line join=round] (490.59,274.35) --
	(477.87,274.35);

\path[draw=drawColor,line width= 0.6pt,line join=round] (477.87,273.02) --
	(477.87,275.67);
\definecolor{drawColor}{RGB}{170,68,153}

\path[draw=drawColor,line width= 0.6pt,line join=round] (504.19,280.98) --
	(504.19,283.63);

\path[draw=drawColor,line width= 0.6pt,line join=round] (504.19,282.30) --
	(499.95,282.30);

\path[draw=drawColor,line width= 0.6pt,line join=round] (499.95,280.98) --
	(499.95,283.63);
\definecolor{drawColor}{RGB}{51,34,136}

\path[draw=drawColor,line width= 0.6pt,line join=round] (492.23,246.51) --
	(492.23,249.16);

\path[draw=drawColor,line width= 0.6pt,line join=round] (492.23,247.83) --
	(482.87,247.83);

\path[draw=drawColor,line width= 0.6pt,line join=round] (482.87,246.51) --
	(482.87,249.16);
\definecolor{drawColor}{RGB}{170,68,153}

\path[draw=drawColor,line width= 0.6pt,line join=round] (497.23,254.46) --
	(497.23,257.11);

\path[draw=drawColor,line width= 0.6pt,line join=round] (497.23,255.79) --
	(490.38,255.79);

\path[draw=drawColor,line width= 0.6pt,line join=round] (490.38,254.46) --
	(490.38,257.11);
\definecolor{drawColor}{RGB}{51,34,136}

\path[draw=drawColor,line width= 0.6pt,line join=round] (507.23,219.99) --
	(507.23,222.64);

\path[draw=drawColor,line width= 0.6pt,line join=round] (507.23,221.32) --
	(499.08,221.32);

\path[draw=drawColor,line width= 0.6pt,line join=round] (499.08,219.99) --
	(499.08,222.64);
\definecolor{drawColor}{RGB}{170,68,153}

\path[draw=drawColor,line width= 0.6pt,line join=round] (504.08,227.95) --
	(504.08,230.60);

\path[draw=drawColor,line width= 0.6pt,line join=round] (504.08,229.27) --
	(500.16,229.27);

\path[draw=drawColor,line width= 0.6pt,line join=round] (500.16,227.95) --
	(500.16,230.60);
\definecolor{drawColor}{RGB}{51,34,136}

\path[draw=drawColor,line width= 0.6pt,line join=round] (500.05,193.47) --
	(500.05,196.13);

\path[draw=drawColor,line width= 0.6pt,line join=round] (500.05,194.80) --
	(487.55,194.80);

\path[draw=drawColor,line width= 0.6pt,line join=round] (487.55,193.47) --
	(487.55,196.13);
\definecolor{drawColor}{RGB}{170,68,153}

\path[draw=drawColor,line width= 0.6pt,line join=round] (500.71,201.43) --
	(500.71,204.08);

\path[draw=drawColor,line width= 0.6pt,line join=round] (500.71,202.75) --
	(497.12,202.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (497.12,201.43) --
	(497.12,204.08);
\definecolor{drawColor}{RGB}{51,34,136}

\path[draw=drawColor,line width= 0.6pt,line join=round] (485.92,166.96) --
	(485.92,169.61);

\path[draw=drawColor,line width= 0.6pt,line join=round] (485.92,168.28) --
	(476.57,168.28);

\path[draw=drawColor,line width= 0.6pt,line join=round] (476.57,166.96) --
	(476.57,169.61);
\definecolor{drawColor}{RGB}{170,68,153}

\path[draw=drawColor,line width= 0.6pt,line join=round] (496.03,174.91) --
	(496.03,177.56);

\path[draw=drawColor,line width= 0.6pt,line join=round] (496.03,176.24) --
	(490.70,176.24);

\path[draw=drawColor,line width= 0.6pt,line join=round] (490.70,174.91) --
	(490.70,177.56);
\definecolor{drawColor}{RGB}{51,34,136}

\path[draw=drawColor,line width= 0.6pt,line join=round] (488.20, 87.41) --
	(488.20, 90.06);

\path[draw=drawColor,line width= 0.6pt,line join=round] (488.20, 88.74) --
	(482.22, 88.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] (482.22, 87.41) --
	(482.22, 90.06);
\definecolor{drawColor}{RGB}{170,68,153}

\path[draw=drawColor,line width= 0.6pt,line join=round] (468.63, 95.37) --
	(468.63, 98.02);

\path[draw=drawColor,line width= 0.6pt,line join=round] (468.63, 96.69) --
	(456.67, 96.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (456.67, 95.37) --
	(456.67, 98.02);
\definecolor{drawColor}{RGB}{51,34,136}
\definecolor{fillColor}{RGB}{51,34,136}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (490.92,141.77) circle (  1.96);
\definecolor{drawColor}{RGB}{170,68,153}
\definecolor{fillColor}{RGB}{170,68,153}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (493.42,149.72) circle (  1.96);
\definecolor{drawColor}{RGB}{51,34,136}
\definecolor{fillColor}{RGB}{51,34,136}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (472.76,486.48) circle (  1.96);
\definecolor{drawColor}{RGB}{170,68,153}
\definecolor{fillColor}{RGB}{170,68,153}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (482.87,494.43) circle (  1.96);
\definecolor{drawColor}{RGB}{51,34,136}
\definecolor{fillColor}{RGB}{51,34,136}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (491.57,459.96) circle (  1.96);
\definecolor{drawColor}{RGB}{170,68,153}
\definecolor{fillColor}{RGB}{170,68,153}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (486.46,467.91) circle (  1.96);
\definecolor{drawColor}{RGB}{51,34,136}
\definecolor{fillColor}{RGB}{51,34,136}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (503.86,433.44) circle (  1.96);
\definecolor{drawColor}{RGB}{170,68,153}
\definecolor{fillColor}{RGB}{170,68,153}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (499.95,441.40) circle (  1.96);
\definecolor{drawColor}{RGB}{51,34,136}
\definecolor{fillColor}{RGB}{51,34,136}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (500.27,406.93) circle (  1.96);
\definecolor{drawColor}{RGB}{170,68,153}
\definecolor{fillColor}{RGB}{170,68,153}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (504.19,414.88) circle (  1.96);
\definecolor{drawColor}{RGB}{51,34,136}
\definecolor{fillColor}{RGB}{51,34,136}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (497.55,327.38) circle (  1.96);
\definecolor{drawColor}{RGB}{170,68,153}
\definecolor{fillColor}{RGB}{170,68,153}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (487.11,335.33) circle (  1.96);
\definecolor{drawColor}{RGB}{51,34,136}
\definecolor{fillColor}{RGB}{51,34,136}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (484.29,300.86) circle (  1.96);
\definecolor{drawColor}{RGB}{170,68,153}
\definecolor{fillColor}{RGB}{170,68,153}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (476.46,308.82) circle (  1.96);
\definecolor{drawColor}{RGB}{51,34,136}
\definecolor{fillColor}{RGB}{51,34,136}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (440.36,115.25) circle (  1.96);
\definecolor{drawColor}{RGB}{170,68,153}
\definecolor{fillColor}{RGB}{170,68,153}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (438.84,123.21) circle (  1.96);
\definecolor{drawColor}{RGB}{51,34,136}
\definecolor{fillColor}{RGB}{51,34,136}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (484.29,274.35) circle (  1.96);
\definecolor{drawColor}{RGB}{170,68,153}
\definecolor{fillColor}{RGB}{170,68,153}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (502.01,282.30) circle (  1.96);
\definecolor{drawColor}{RGB}{51,34,136}
\definecolor{fillColor}{RGB}{51,34,136}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (487.55,247.83) circle (  1.96);
\definecolor{drawColor}{RGB}{170,68,153}
\definecolor{fillColor}{RGB}{170,68,153}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (493.86,255.79) circle (  1.96);
\definecolor{drawColor}{RGB}{51,34,136}
\definecolor{fillColor}{RGB}{51,34,136}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (503.21,221.32) circle (  1.96);
\definecolor{drawColor}{RGB}{170,68,153}
\definecolor{fillColor}{RGB}{170,68,153}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (502.12,229.27) circle (  1.96);
\definecolor{drawColor}{RGB}{51,34,136}
\definecolor{fillColor}{RGB}{51,34,136}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (493.75,194.80) circle (  1.96);
\definecolor{drawColor}{RGB}{170,68,153}
\definecolor{fillColor}{RGB}{170,68,153}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (498.86,202.75) circle (  1.96);
\definecolor{drawColor}{RGB}{51,34,136}
\definecolor{fillColor}{RGB}{51,34,136}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (481.24,168.28) circle (  1.96);
\definecolor{drawColor}{RGB}{170,68,153}
\definecolor{fillColor}{RGB}{170,68,153}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (493.42,176.24) circle (  1.96);
\definecolor{drawColor}{RGB}{51,34,136}
\definecolor{fillColor}{RGB}{51,34,136}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (485.27, 88.74) circle (  1.96);
\definecolor{drawColor}{RGB}{170,68,153}
\definecolor{fillColor}{RGB}{170,68,153}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (462.65, 96.69) circle (  1.96);
\definecolor{drawColor}{RGB}{169,169,169}

\path[draw=drawColor,line width= 0.6pt,line join=round] (401.21,159.00) -- (520.82,159.00);

\path[draw=drawColor,line width= 0.6pt,line join=round] (401.21,132.49) -- (520.82,132.49);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.5pt,line join=round,line cap=round] (401.21, 76.80) rectangle (520.82,506.36);
\end{scope}
\begin{scope}
\path[clip] (525.82, 76.80) rectangle (645.43,506.36);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (525.82, 76.80) rectangle (645.43,506.36);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] (544.85, 76.80) --
	(544.85,506.36);

\path[draw=drawColor,line width= 0.3pt,line join=round] (572.03, 76.80) --
	(572.03,506.36);

\path[draw=drawColor,line width= 0.3pt,line join=round] (599.22, 76.80) --
	(599.22,506.36);

\path[draw=drawColor,line width= 0.3pt,line join=round] (626.40, 76.80) --
	(626.40,506.36);

\path[draw=drawColor,line width= 0.5pt,line join=round] (525.82, 92.71) --
	(645.43, 92.71);

\path[draw=drawColor,line width= 0.5pt,line join=round] (525.82,119.23) --
	(645.43,119.23);

\path[draw=drawColor,line width= 0.5pt,line join=round] (525.82,145.75) --
	(645.43,145.75);

\path[draw=drawColor,line width= 0.5pt,line join=round] (525.82,172.26) --
	(645.43,172.26);

\path[draw=drawColor,line width= 0.5pt,line join=round] (525.82,198.78) --
	(645.43,198.78);

\path[draw=drawColor,line width= 0.5pt,line join=round] (525.82,225.29) --
	(645.43,225.29);

\path[draw=drawColor,line width= 0.5pt,line join=round] (525.82,251.81) --
	(645.43,251.81);

\path[draw=drawColor,line width= 0.5pt,line join=round] (525.82,278.33) --
	(645.43,278.33);

\path[draw=drawColor,line width= 0.5pt,line join=round] (525.82,304.84) --
	(645.43,304.84);

\path[draw=drawColor,line width= 0.5pt,line join=round] (525.82,331.36) --
	(645.43,331.36);

\path[draw=drawColor,line width= 0.5pt,line join=round] (525.82,357.87) --
	(645.43,357.87);

\path[draw=drawColor,line width= 0.5pt,line join=round] (525.82,384.39) --
	(645.43,384.39);

\path[draw=drawColor,line width= 0.5pt,line join=round] (525.82,410.91) --
	(645.43,410.91);

\path[draw=drawColor,line width= 0.5pt,line join=round] (525.82,437.42) --
	(645.43,437.42);

\path[draw=drawColor,line width= 0.5pt,line join=round] (525.82,463.94) --
	(645.43,463.94);

\path[draw=drawColor,line width= 0.5pt,line join=round] (525.82,490.45) --
	(645.43,490.45);

\path[draw=drawColor,line width= 0.5pt,line join=round] (531.26, 76.80) --
	(531.26,506.36);

\path[draw=drawColor,line width= 0.5pt,line join=round] (558.44, 76.80) --
	(558.44,506.36);

\path[draw=drawColor,line width= 0.5pt,line join=round] (585.63, 76.80) --
	(585.63,506.36);

\path[draw=drawColor,line width= 0.5pt,line join=round] (612.81, 76.80) --
	(612.81,506.36);

\path[draw=drawColor,line width= 0.5pt,line join=round] (639.99, 76.80) --
	(639.99,506.36);
\definecolor{drawColor}{RGB}{17,119,51}

\path[draw=drawColor,line width= 0.6pt,line join=round] (625.31,148.40) --
	(625.31,151.05);

\path[draw=drawColor,line width= 0.6pt,line join=round] (625.31,149.72) --
	(614.01,149.72);

\path[draw=drawColor,line width= 0.6pt,line join=round] (614.01,148.40) --
	(614.01,151.05);
\definecolor{drawColor}{RGB}{221,204,119}

\path[draw=drawColor,line width= 0.6pt,line join=round] (624.77,140.44) --
	(624.77,143.09);

\path[draw=drawColor,line width= 0.6pt,line join=round] (624.77,141.77) --
	(610.20,141.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (610.20,140.44) --
	(610.20,143.09);
\definecolor{drawColor}{RGB}{17,119,51}

\path[draw=drawColor,line width= 0.6pt,line join=round] (612.59,493.10) --
	(612.59,495.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (612.59,494.43) --
	(593.46,494.43);

\path[draw=drawColor,line width= 0.6pt,line join=round] (593.46,493.10) --
	(593.46,495.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (616.07,466.59) --
	(616.07,469.24);

\path[draw=drawColor,line width= 0.6pt,line join=round] (616.07,467.91) --
	(608.35,467.91);

\path[draw=drawColor,line width= 0.6pt,line join=round] (608.35,466.59) --
	(608.35,469.24);
\definecolor{drawColor}{RGB}{221,204,119}

\path[draw=drawColor,line width= 0.6pt,line join=round] (621.07,458.63) --
	(621.07,461.29);

\path[draw=drawColor,line width= 0.6pt,line join=round] (621.07,459.96) --
	(601.94,459.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (601.94,458.63) --
	(601.94,461.29);
\definecolor{drawColor}{RGB}{17,119,51}

\path[draw=drawColor,line width= 0.6pt,line join=round] (625.21,440.07) --
	(625.21,442.72);

\path[draw=drawColor,line width= 0.6pt,line join=round] (625.21,441.40) --
	(621.29,441.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] (621.29,440.07) --
	(621.29,442.72);
\definecolor{drawColor}{RGB}{221,204,119}

\path[draw=drawColor,line width= 0.6pt,line join=round] (628.47,432.12) --
	(628.47,434.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (628.47,433.44) --
	(619.12,433.44);

\path[draw=drawColor,line width= 0.6pt,line join=round] (619.12,432.12) --
	(619.12,434.77);
\definecolor{drawColor}{RGB}{17,119,51}

\path[draw=drawColor,line width= 0.6pt,line join=round] (630.53,413.56) --
	(630.53,416.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] (630.53,414.88) --
	(624.23,414.88);

\path[draw=drawColor,line width= 0.6pt,line join=round] (624.23,413.56) --
	(624.23,416.21);
\definecolor{drawColor}{RGB}{221,204,119}

\path[draw=drawColor,line width= 0.6pt,line join=round] (634.99,405.60) --
	(634.99,408.25);

\path[draw=drawColor,line width= 0.6pt,line join=round] (634.99,406.93) --
	(627.05,406.93);

\path[draw=drawColor,line width= 0.6pt,line join=round] (627.05,405.60) --
	(627.05,408.25);
\definecolor{drawColor}{RGB}{17,119,51}

\path[draw=drawColor,line width= 0.6pt,line join=round] (637.82,387.04) --
	(637.82,389.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (637.82,388.37) --
	(635.21,388.37);

\path[draw=drawColor,line width= 0.6pt,line join=round] (635.21,387.04) --
	(635.21,389.69);
\definecolor{drawColor}{RGB}{221,204,119}

\path[draw=drawColor,line width= 0.6pt,line join=round] (636.95,379.09) --
	(636.95,381.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] (636.95,380.41) --
	(629.55,380.41);

\path[draw=drawColor,line width= 0.6pt,line join=round] (629.55,379.09) --
	(629.55,381.74);
\definecolor{drawColor}{RGB}{17,119,51}

\path[draw=drawColor,line width= 0.6pt,line join=round] (616.40,360.52) --
	(616.40,363.18);

\path[draw=drawColor,line width= 0.6pt,line join=round] (616.40,361.85) --
	(612.05,361.85);

\path[draw=drawColor,line width= 0.6pt,line join=round] (612.05,360.52) --
	(612.05,363.18);
\definecolor{drawColor}{RGB}{221,204,119}

\path[draw=drawColor,line width= 0.6pt,line join=round] (622.16,352.57) --
	(622.16,355.22);

\path[draw=drawColor,line width= 0.6pt,line join=round] (622.16,353.90) --
	(602.81,353.90);

\path[draw=drawColor,line width= 0.6pt,line join=round] (602.81,352.57) --
	(602.81,355.22);
\definecolor{drawColor}{RGB}{17,119,51}

\path[draw=drawColor,line width= 0.6pt,line join=round] (620.64,334.01) --
	(620.64,336.66);

\path[draw=drawColor,line width= 0.6pt,line join=round] (620.64,335.33) --
	(607.26,335.33);

\path[draw=drawColor,line width= 0.6pt,line join=round] (607.26,334.01) --
	(607.26,336.66);
\definecolor{drawColor}{RGB}{221,204,119}

\path[draw=drawColor,line width= 0.6pt,line join=round] (636.41,326.05) --
	(636.41,328.71);

\path[draw=drawColor,line width= 0.6pt,line join=round] (636.41,327.38) --
	(601.83,327.38);

\path[draw=drawColor,line width= 0.6pt,line join=round] (601.83,326.05) --
	(601.83,328.71);
\definecolor{drawColor}{RGB}{17,119,51}

\path[draw=drawColor,line width= 0.6pt,line join=round] (563.55,121.88) --
	(563.55,124.53);

\path[draw=drawColor,line width= 0.6pt,line join=round] (563.55,123.21) --
	(560.73,123.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] (560.73,121.88) --
	(560.73,124.53);
\definecolor{drawColor}{RGB}{221,204,119}

\path[draw=drawColor,line width= 0.6pt,line join=round] (579.10,113.93) --
	(579.10,116.58);

\path[draw=drawColor,line width= 0.6pt,line join=round] (579.10,115.25) --
	(570.29,115.25);

\path[draw=drawColor,line width= 0.6pt,line join=round] (570.29,113.93) --
	(570.29,116.58);
\definecolor{drawColor}{RGB}{17,119,51}

\path[draw=drawColor,line width= 0.6pt,line join=round] (626.08,280.98) --
	(626.08,283.63);

\path[draw=drawColor,line width= 0.6pt,line join=round] (626.08,282.30) --
	(621.83,282.30);

\path[draw=drawColor,line width= 0.6pt,line join=round] (621.83,280.98) --
	(621.83,283.63);
\definecolor{drawColor}{RGB}{221,204,119}

\path[draw=drawColor,line width= 0.6pt,line join=round] (626.18,273.02) --
	(626.18,275.67);

\path[draw=drawColor,line width= 0.6pt,line join=round] (626.18,274.35) --
	(595.63,274.35);

\path[draw=drawColor,line width= 0.6pt,line join=round] (595.63,273.02) --
	(595.63,275.67);
\definecolor{drawColor}{RGB}{17,119,51}

\path[draw=drawColor,line width= 0.6pt,line join=round] (619.12,254.46) --
	(619.12,257.11);

\path[draw=drawColor,line width= 0.6pt,line join=round] (619.12,255.79) --
	(613.57,255.79);

\path[draw=drawColor,line width= 0.6pt,line join=round] (613.57,254.46) --
	(613.57,257.11);
\definecolor{drawColor}{RGB}{221,204,119}

\path[draw=drawColor,line width= 0.6pt,line join=round] (625.42,246.51) --
	(625.42,249.16);

\path[draw=drawColor,line width= 0.6pt,line join=round] (625.42,247.83) --
	(598.02,247.83);

\path[draw=drawColor,line width= 0.6pt,line join=round] (598.02,246.51) --
	(598.02,249.16);
\definecolor{drawColor}{RGB}{17,119,51}

\path[draw=drawColor,line width= 0.6pt,line join=round] (628.36,227.95) --
	(628.36,230.60);

\path[draw=drawColor,line width= 0.6pt,line join=round] (628.36,229.27) --
	(624.23,229.27);

\path[draw=drawColor,line width= 0.6pt,line join=round] (624.23,227.95) --
	(624.23,230.60);
\definecolor{drawColor}{RGB}{221,204,119}

\path[draw=drawColor,line width= 0.6pt,line join=round] (631.95,219.99) --
	(631.95,222.64);

\path[draw=drawColor,line width= 0.6pt,line join=round] (631.95,221.32) --
	(626.29,221.32);

\path[draw=drawColor,line width= 0.6pt,line join=round] (626.29,219.99) --
	(626.29,222.64);
\definecolor{drawColor}{RGB}{17,119,51}

\path[draw=drawColor,line width= 0.6pt,line join=round] (626.08,196.13) --
	(626.08,201.43);

\path[draw=drawColor,line width= 0.6pt,line join=round] (626.08,198.78) --
	(623.03,198.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (623.03,196.13) --
	(623.03,201.43);

\path[draw=drawColor,line width= 0.6pt,line join=round] (617.49,174.91) --
	(617.49,177.56);

\path[draw=drawColor,line width= 0.6pt,line join=round] (617.49,176.24) --
	(612.27,176.24);

\path[draw=drawColor,line width= 0.6pt,line join=round] (612.27,174.91) --
	(612.27,177.56);
\definecolor{drawColor}{RGB}{221,204,119}

\path[draw=drawColor,line width= 0.6pt,line join=round] (618.25,166.96) --
	(618.25,169.61);

\path[draw=drawColor,line width= 0.6pt,line join=round] (618.25,168.28) --
	(604.44,168.28);

\path[draw=drawColor,line width= 0.6pt,line join=round] (604.44,166.96) --
	(604.44,169.61);
\definecolor{drawColor}{RGB}{17,119,51}

\path[draw=drawColor,line width= 0.6pt,line join=round] (602.15, 95.37) --
	(602.15, 98.02);

\path[draw=drawColor,line width= 0.6pt,line join=round] (602.15, 96.69) --
	(594.76, 96.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (594.76, 95.37) --
	(594.76, 98.02);
\definecolor{drawColor}{RGB}{221,204,119}

\path[draw=drawColor,line width= 0.6pt,line join=round] (611.61, 87.41) --
	(611.61, 90.06);

\path[draw=drawColor,line width= 0.6pt,line join=round] (611.61, 88.74) --
	(601.72, 88.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] (601.72, 87.41) --
	(601.72, 90.06);
\definecolor{drawColor}{RGB}{17,119,51}
\definecolor{fillColor}{RGB}{17,119,51}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (619.66,149.72) circle (  1.96);
\definecolor{drawColor}{RGB}{221,204,119}
\definecolor{fillColor}{RGB}{221,204,119}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (617.49,141.77) circle (  1.96);
\definecolor{drawColor}{RGB}{17,119,51}
\definecolor{fillColor}{RGB}{17,119,51}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (603.02,494.43) circle (  1.96);
\definecolor{drawColor}{RGB}{221,204,119}
\definecolor{fillColor}{RGB}{221,204,119}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (597.59,486.48) circle (  1.96);
\definecolor{drawColor}{RGB}{17,119,51}
\definecolor{fillColor}{RGB}{17,119,51}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (612.27,467.91) circle (  1.96);
\definecolor{drawColor}{RGB}{221,204,119}
\definecolor{fillColor}{RGB}{221,204,119}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (611.50,459.96) circle (  1.96);
\definecolor{drawColor}{RGB}{17,119,51}
\definecolor{fillColor}{RGB}{17,119,51}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (623.25,441.40) circle (  1.96);
\definecolor{drawColor}{RGB}{221,204,119}
\definecolor{fillColor}{RGB}{221,204,119}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (623.79,433.44) circle (  1.96);
\definecolor{drawColor}{RGB}{17,119,51}
\definecolor{fillColor}{RGB}{17,119,51}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (627.38,414.88) circle (  1.96);
\definecolor{drawColor}{RGB}{221,204,119}
\definecolor{fillColor}{RGB}{221,204,119}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (630.97,406.93) circle (  1.96);
\definecolor{drawColor}{RGB}{17,119,51}
\definecolor{fillColor}{RGB}{17,119,51}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (636.51,388.37) circle (  1.96);
\definecolor{drawColor}{RGB}{221,204,119}
\definecolor{fillColor}{RGB}{221,204,119}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (633.25,380.41) circle (  1.96);
\definecolor{drawColor}{RGB}{17,119,51}
\definecolor{fillColor}{RGB}{17,119,51}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (614.22,361.85) circle (  1.96);
\definecolor{drawColor}{RGB}{221,204,119}
\definecolor{fillColor}{RGB}{221,204,119}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (612.48,353.90) circle (  1.96);
\definecolor{drawColor}{RGB}{17,119,51}
\definecolor{fillColor}{RGB}{17,119,51}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (613.90,335.33) circle (  1.96);
\definecolor{drawColor}{RGB}{221,204,119}
\definecolor{fillColor}{RGB}{221,204,119}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (619.12,327.38) circle (  1.96);
\definecolor{drawColor}{RGB}{17,119,51}
\definecolor{fillColor}{RGB}{17,119,51}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (562.14,123.21) circle (  1.96);
\definecolor{drawColor}{RGB}{221,204,119}
\definecolor{fillColor}{RGB}{221,204,119}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (574.75,115.25) circle (  1.96);
\definecolor{drawColor}{RGB}{17,119,51}
\definecolor{fillColor}{RGB}{17,119,51}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (623.90,282.30) circle (  1.96);
\definecolor{drawColor}{RGB}{221,204,119}
\definecolor{fillColor}{RGB}{221,204,119}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (610.96,274.35) circle (  1.96);
\definecolor{drawColor}{RGB}{17,119,51}
\definecolor{fillColor}{RGB}{17,119,51}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (616.29,255.79) circle (  1.96);
\definecolor{drawColor}{RGB}{221,204,119}
\definecolor{fillColor}{RGB}{221,204,119}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (611.72,247.83) circle (  1.96);
\definecolor{drawColor}{RGB}{17,119,51}
\definecolor{fillColor}{RGB}{17,119,51}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (626.29,229.27) circle (  1.96);
\definecolor{drawColor}{RGB}{221,204,119}
\definecolor{fillColor}{RGB}{221,204,119}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (629.12,221.32) circle (  1.96);
\definecolor{drawColor}{RGB}{17,119,51}
\definecolor{fillColor}{RGB}{17,119,51}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (624.55,198.78) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (614.88,176.24) circle (  1.96);
\definecolor{drawColor}{RGB}{221,204,119}
\definecolor{fillColor}{RGB}{221,204,119}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (611.40,168.28) circle (  1.96);
\definecolor{drawColor}{RGB}{17,119,51}
\definecolor{fillColor}{RGB}{17,119,51}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (598.46, 96.69) circle (  1.96);
\definecolor{drawColor}{RGB}{221,204,119}
\definecolor{fillColor}{RGB}{221,204,119}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (606.72, 88.74) circle (  1.96);
\definecolor{drawColor}{RGB}{169,169,169}

\path[draw=drawColor,line width= 0.6pt,line join=round] (525.82,159.00) -- (645.43,159.00);

\path[draw=drawColor,line width= 0.6pt,line join=round] (525.82,132.49) -- (645.43,132.49);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.5pt,line join=round,line cap=round] (525.82, 76.80) rectangle (645.43,506.36);
\end{scope}
\begin{scope}
\path[clip] (152.00,506.36) rectangle (271.61,521.43);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.5pt,line join=round,line cap=round,fill=fillColor] (152.00,506.36) rectangle (271.61,521.43);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (211.80,511.14) {All};
\end{scope}
\begin{scope}
\path[clip] (276.61,506.36) rectangle (396.21,521.43);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.5pt,line join=round,line cap=round,fill=fillColor] (276.61,506.36) rectangle (396.21,521.43);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (336.41,511.14) {By gender};
\end{scope}
\begin{scope}
\path[clip] (401.21,506.36) rectangle (520.82,521.43);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.5pt,line join=round,line cap=round,fill=fillColor] (401.21,506.36) rectangle (520.82,521.43);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (461.02,511.14) {By education};
\end{scope}
\begin{scope}
\path[clip] (525.82,506.36) rectangle (645.43,521.43);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.5pt,line join=round,line cap=round,fill=fillColor] (525.82,506.36) rectangle (645.43,521.43);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (585.63,511.14) {By age};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.5pt,line join=round] (157.44, 74.30) --
	(157.44, 76.80);

\path[draw=drawColor,line width= 0.5pt,line join=round] (184.62, 74.30) --
	(184.62, 76.80);

\path[draw=drawColor,line width= 0.5pt,line join=round] (211.80, 74.30) --
	(211.80, 76.80);

\path[draw=drawColor,line width= 0.5pt,line join=round] (238.99, 74.30) --
	(238.99, 76.80);

\path[draw=drawColor,line width= 0.5pt,line join=round] (266.17, 74.30) --
	(266.17, 76.80);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (157.44, 66.79) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (184.62, 66.79) {25};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (211.80, 66.79) {50};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (238.99, 66.79) {75};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (266.17, 66.79) {100};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.5pt,line join=round] (282.04, 74.30) --
	(282.04, 76.80);

\path[draw=drawColor,line width= 0.5pt,line join=round] (309.23, 74.30) --
	(309.23, 76.80);

\path[draw=drawColor,line width= 0.5pt,line join=round] (336.41, 74.30) --
	(336.41, 76.80);

\path[draw=drawColor,line width= 0.5pt,line join=round] (363.59, 74.30) --
	(363.59, 76.80);

\path[draw=drawColor,line width= 0.5pt,line join=round] (390.78, 74.30) --
	(390.78, 76.80);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (282.04, 66.79) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (309.23, 66.79) {25};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (336.41, 66.79) {50};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (363.59, 66.79) {75};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (390.78, 66.79) {100};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.5pt,line join=round] (406.65, 74.30) --
	(406.65, 76.80);

\path[draw=drawColor,line width= 0.5pt,line join=round] (433.84, 74.30) --
	(433.84, 76.80);

\path[draw=drawColor,line width= 0.5pt,line join=round] (461.02, 74.30) --
	(461.02, 76.80);

\path[draw=drawColor,line width= 0.5pt,line join=round] (488.20, 74.30) --
	(488.20, 76.80);

\path[draw=drawColor,line width= 0.5pt,line join=round] (515.39, 74.30) --
	(515.39, 76.80);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (406.65, 66.79) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (433.84, 66.79) {25};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (461.02, 66.79) {50};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (488.20, 66.79) {75};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (515.39, 66.79) {100};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.5pt,line join=round] (531.26, 74.30) --
	(531.26, 76.80);

\path[draw=drawColor,line width= 0.5pt,line join=round] (558.44, 74.30) --
	(558.44, 76.80);

\path[draw=drawColor,line width= 0.5pt,line join=round] (585.63, 74.30) --
	(585.63, 76.80);

\path[draw=drawColor,line width= 0.5pt,line join=round] (612.81, 74.30) --
	(612.81, 76.80);

\path[draw=drawColor,line width= 0.5pt,line join=round] (639.99, 74.30) --
	(639.99, 76.80);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (531.26, 66.79) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (558.44, 66.79) {25};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (585.63, 66.79) {50};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (612.81, 66.79) {75};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (639.99, 66.79) {100};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 16.33, 94.28) {USA};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 16.33, 85.64) {(National, 1959)};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 16.33,120.79) {Russia};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 16.33,112.15) {(Subnational, 61 regions, 22125)};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 16.33,142.99) {All LMICs};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 16.33,173.83) {Uganda 2};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 16.33,165.19) {(Subnational, 1 district, 1366)};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 16.33,200.34) {Uganda 1};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 16.33,191.70) {(Subnational, 13 districts, 3362)};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 16.33,226.86) {Sierra Leone 2};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 16.33,218.22) {(National, 2110)};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 16.33,253.37) {Sierra Leone 1};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 16.33,244.73) {(National, 1070)};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 16.33,279.89) {Rwanda};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 16.33,271.25) {(National, 1355)};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 16.33,306.41) {Pakistan 2};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 16.33,297.77) {(Subnational, 1 province, 1492)};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 16.33,332.92) {Pakistan 1};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 16.33,324.28) {(Subnational, 2 districts, 1633)};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 16.33,359.44) {Nigeria};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 16.33,350.80) {(Subnational, 1 state, 1868)};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 16.33,385.95) {Nepal};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 16.33,377.31) {(Subnational, 2 districts, 1389)};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 16.33,412.47) {Mozambique};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 16.33,403.83) {(Subnational, 2 cities, 862)};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 16.33,438.99) {India};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 16.33,430.35) {(Subnational, Slums in 2 cities, 3348)};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 16.33,465.50) {Colombia};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 16.33,456.86) {(National, 1012)};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 16.33,492.02) {Burkina Faso};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 16.33,483.38) {(National, 977)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.5pt,line join=round] (149.50, 92.71) --
	(152.00, 92.71);

\path[draw=drawColor,line width= 0.5pt,line join=round] (149.50,119.23) --
	(152.00,119.23);

\path[draw=drawColor,line width= 0.5pt,line join=round] (149.50,145.75) --
	(152.00,145.75);

\path[draw=drawColor,line width= 0.5pt,line join=round] (149.50,172.26) --
	(152.00,172.26);

\path[draw=drawColor,line width= 0.5pt,line join=round] (149.50,198.78) --
	(152.00,198.78);

\path[draw=drawColor,line width= 0.5pt,line join=round] (149.50,225.29) --
	(152.00,225.29);

\path[draw=drawColor,line width= 0.5pt,line join=round] (149.50,251.81) --
	(152.00,251.81);

\path[draw=drawColor,line width= 0.5pt,line join=round] (149.50,278.33) --
	(152.00,278.33);

\path[draw=drawColor,line width= 0.5pt,line join=round] (149.50,304.84) --
	(152.00,304.84);

\path[draw=drawColor,line width= 0.5pt,line join=round] (149.50,331.36) --
	(152.00,331.36);

\path[draw=drawColor,line width= 0.5pt,line join=round] (149.50,357.87) --
	(152.00,357.87);

\path[draw=drawColor,line width= 0.5pt,line join=round] (149.50,384.39) --
	(152.00,384.39);

\path[draw=drawColor,line width= 0.5pt,line join=round] (149.50,410.91) --
	(152.00,410.91);

\path[draw=drawColor,line width= 0.5pt,line join=round] (149.50,437.42) --
	(152.00,437.42);

\path[draw=drawColor,line width= 0.5pt,line join=round] (149.50,463.94) --
	(152.00,463.94);

\path[draw=drawColor,line width= 0.5pt,line join=round] (149.50,490.45) --
	(152.00,490.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (398.71, 55.85) {estimate};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (258.71,  5.00) rectangle (538.72, 43.91);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (263.71, 21.01) {Subgroups};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (314.33, 24.45) rectangle (328.79, 38.91);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{drawColor}{RGB}{153,153,51}

\path[draw=drawColor,line width= 0.6pt,line join=round] (315.78, 31.68) -- (327.34, 31.68);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{drawColor}{RGB}{153,153,51}
\definecolor{fillColor}{RGB}{153,153,51}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (321.56, 31.68) circle (  1.96);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (314.33, 10.00) rectangle (328.79, 24.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{drawColor}{RGB}{68,170,153}

\path[draw=drawColor,line width= 0.6pt,line join=round] (315.78, 17.23) -- (327.34, 17.23);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{drawColor}{RGB}{68,170,153}
\definecolor{fillColor}{RGB}{68,170,153}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (321.56, 17.23) circle (  1.96);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (363.34, 24.45) rectangle (377.79, 38.91);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{drawColor}{RGB}{170,68,153}

\path[draw=drawColor,line width= 0.6pt,line join=round] (364.78, 31.68) -- (376.35, 31.68);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{drawColor}{RGB}{170,68,153}
\definecolor{fillColor}{RGB}{170,68,153}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (370.56, 31.68) circle (  1.96);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (363.34, 10.00) rectangle (377.79, 24.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{drawColor}{RGB}{51,34,136}

\path[draw=drawColor,line width= 0.6pt,line join=round] (364.78, 17.23) -- (376.35, 17.23);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{drawColor}{RGB}{51,34,136}
\definecolor{fillColor}{RGB}{51,34,136}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (370.56, 17.23) circle (  1.96);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (462.93, 24.45) rectangle (477.38, 38.91);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{drawColor}{RGB}{17,119,51}

\path[draw=drawColor,line width= 0.6pt,line join=round] (464.37, 31.68) -- (475.94, 31.68);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{drawColor}{RGB}{17,119,51}
\definecolor{fillColor}{RGB}{17,119,51}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (470.16, 31.68) circle (  1.96);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (462.93, 10.00) rectangle (477.38, 24.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{drawColor}{RGB}{221,204,119}

\path[draw=drawColor,line width= 0.6pt,line join=round] (464.37, 17.23) -- (475.94, 17.23);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{drawColor}{RGB}{221,204,119}
\definecolor{fillColor}{RGB}{221,204,119}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (470.16, 17.23) circle (  1.96);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (503.83, 24.45) rectangle (518.28, 38.91);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{drawColor}{RGB}{204,102,119}

\path[draw=drawColor,line width= 0.6pt,line join=round] (505.27, 31.68) -- (516.84, 31.68);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{drawColor}{RGB}{204,102,119}
\definecolor{fillColor}{RGB}{204,102,119}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (511.05, 31.68) circle (  1.96);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.80] at (333.79, 28.93) {Female};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.80] at (333.79, 14.47) {Male};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.80] at (382.79, 28.93) {Up to Secondary};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.80] at (382.79, 14.47) {More than Secondary};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.80] at (482.38, 28.93) {$< 55$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.80] at (482.38, 14.47) {$55 +$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.80] at (523.28, 28.93) {All};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.20] at (  5.00,528.76) {If a COVID-19 vaccine becomes available in [country], would you take it?};
\end{scope}
\end{tikzpicture}

\end{document}