### 2021-05-04 paper_files/figure-latex/fig2paper-1.tex

% Created by tikzDevice version 0.12.3.1 on 2021-05-04 13:12:53
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}

\nonstopmode

\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (722.70,542.02);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.00) rectangle (722.70,542.02);
\end{scope}
\begin{scope}
\path[clip] ( 78.32, 67.14) rectangle (137.25,464.78);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 78.32, 67.14) rectangle (137.25,464.78);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 88.61, 67.14) --
	( 88.61,464.78);

\path[draw=drawColor,line width= 0.3pt,line join=round] (101.74, 67.14) --
	(101.74,464.78);

\path[draw=drawColor,line width= 0.3pt,line join=round] (114.88, 67.14) --
	(114.88,464.78);

\path[draw=drawColor,line width= 0.3pt,line join=round] (128.01, 67.14) --
	(128.01,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 78.32, 83.94) --
	(137.25, 83.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 78.32,111.94) --
	(137.25,111.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 78.32,139.95) --
	(137.25,139.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 78.32,167.95) --
	(137.25,167.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 78.32,195.95) --
	(137.25,195.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 78.32,223.96) --
	(137.25,223.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 78.32,251.96) --
	(137.25,251.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 78.32,279.96) --
	(137.25,279.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 78.32,307.96) --
	(137.25,307.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 78.32,335.97) --
	(137.25,335.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 78.32,363.97) --
	(137.25,363.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 78.32,391.97) --
	(137.25,391.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 78.32,419.98) --
	(137.25,419.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 78.32,447.98) --
	(137.25,447.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 82.05, 67.14) --
	( 82.05,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 95.18, 67.14) --
	( 95.18,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (108.31, 67.14) --
	(108.31,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (121.44, 67.14) --
	(121.44,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (134.58, 67.14) --
	(134.58,464.78);
\definecolor{drawColor}{RGB}{204,102,119}

\path[draw=drawColor,line width= 0.6pt,line join=round] (111.62,137.15) --
	(111.62,142.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (111.62,139.95) --
	( 95.34,139.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 95.34,137.15) --
	( 95.34,142.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (106.37,445.18) --
	(106.37,450.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (106.37,447.98) --
	(100.69,447.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (100.69,445.18) --
	(100.69,450.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (101.80,417.18) --
	(101.80,422.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (101.80,419.98) --
	( 94.86,419.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 94.86,417.18) --
	( 94.86,422.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 91.61,361.17) --
	( 91.61,366.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 91.61,363.97) --
	( 82.20,363.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 82.20,361.17) --
	( 82.20,366.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 95.44,333.17) --
	( 95.44,338.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 95.44,335.97) --
	( 91.24,335.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 91.24,333.17) --
	( 91.24,338.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 98.23,305.16) --
	( 98.23,310.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 98.23,307.96) --
	( 89.98,307.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 89.98,305.16) --
	( 89.98,310.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (102.22,109.14) --
	(102.22,114.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] (102.22,111.94) --
	(100.54,111.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (100.54,109.14) --
	(100.54,114.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] (108.47,277.16) --
	(108.47,282.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (108.47,279.96) --
	( 96.18,279.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 96.18,277.16) --
	( 96.18,282.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (113.51,249.16) --
	(113.51,254.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (113.51,251.96) --
	(106.79,251.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (106.79,249.16) --
	(106.79,254.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (116.56,221.16) --
	(116.56,226.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (116.56,223.96) --
	(108.36,223.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (108.36,221.16) --
	(108.36,226.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (129.11,193.15) --
	(129.11,198.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (129.11,195.95) --
	(124.44,195.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (124.44,193.15) --
	(124.44,198.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (109.62,165.15) --
	(109.62,170.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (109.62,167.95) --
	(104.21,167.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (104.21,165.15) --
	(104.21,170.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (126.17, 81.14) --
	(126.17, 86.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] (126.17, 83.94) --
	(121.23, 83.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (121.23, 81.14) --
	(121.23, 86.74);
\definecolor{fillColor}{RGB}{204,102,119}

\path[fill=fillColor] (103.48,139.95) circle (  3.57);

\path[fill=fillColor] (103.53,447.98) circle (  3.57);

\path[fill=fillColor] ( 98.33,419.98) circle (  3.57);

\path[fill=fillColor] ( 86.93,363.97) circle (  1.43);

\path[fill=fillColor] ( 93.34,335.97) circle (  3.57);

\path[fill=fillColor] ( 94.13,307.96) circle (  3.57);

\path[fill=fillColor] (101.38,111.94) circle (  3.57);

\path[fill=fillColor] (102.32,279.96) circle (  3.57);

\path[fill=fillColor] (110.15,251.96) circle (  3.57);

\path[fill=fillColor] (112.46,223.96) circle (  3.57);

\path[fill=fillColor] (126.75,195.95) circle (  3.57);

\path[fill=fillColor] (106.89,167.95) circle (  3.57);

\path[fill=fillColor] (123.70, 83.94) circle (  3.57);
\definecolor{drawColor}{RGB}{169,169,169}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 78.32,153.95) -- (137.25,153.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 78.32,125.95) -- (137.25,125.95);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 78.32, 67.14) rectangle (137.25,464.78);
\end{scope}
\begin{scope}
\path[clip] (142.75, 67.14) rectangle (201.69,464.78);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (142.75, 67.14) rectangle (201.69,464.78);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] (153.05, 67.14) --
	(153.05,464.78);

\path[draw=drawColor,line width= 0.3pt,line join=round] (166.18, 67.14) --
	(166.18,464.78);

\path[draw=drawColor,line width= 0.3pt,line join=round] (179.32, 67.14) --
	(179.32,464.78);

\path[draw=drawColor,line width= 0.3pt,line join=round] (192.45, 67.14) --
	(192.45,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (142.75, 83.94) --
	(201.69, 83.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (142.75,111.94) --
	(201.69,111.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (142.75,139.95) --
	(201.69,139.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (142.75,167.95) --
	(201.69,167.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (142.75,195.95) --
	(201.69,195.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (142.75,223.96) --
	(201.69,223.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (142.75,251.96) --
	(201.69,251.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (142.75,279.96) --
	(201.69,279.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (142.75,307.96) --
	(201.69,307.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (142.75,335.97) --
	(201.69,335.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (142.75,363.97) --
	(201.69,363.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (142.75,391.97) --
	(201.69,391.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (142.75,419.98) --
	(201.69,419.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (142.75,447.98) --
	(201.69,447.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (146.48, 67.14) --
	(146.48,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (159.62, 67.14) --
	(159.62,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (172.75, 67.14) --
	(172.75,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (185.88, 67.14) --
	(185.88,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (199.01, 67.14) --
	(199.01,464.78);
\definecolor{drawColor}{RGB}{221,204,119}

\path[draw=drawColor,line width= 0.6pt,line join=round] (160.40,137.15) --
	(160.40,142.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (160.40,139.95) --
	(151.05,139.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (151.05,137.15) --
	(151.05,142.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (152.26,445.18) --
	(152.26,450.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (152.26,447.98) --
	(149.11,447.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (149.11,445.18) --
	(149.11,450.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (158.78,417.18) --
	(158.78,422.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (158.78,419.98) --
	(153.16,419.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (153.16,417.18) --
	(153.16,422.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (155.05,361.17) --
	(155.05,366.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (155.05,363.97) --
	(146.27,363.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (146.27,361.17) --
	(146.27,366.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (162.45,333.17) --
	(162.45,338.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (162.45,335.97) --
	(157.94,335.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (157.94,333.17) --
	(157.94,338.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (161.93,305.16) --
	(161.93,310.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (161.93,307.96) --
	(154.00,307.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (154.00,305.16) --
	(154.00,310.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (154.42,109.14) --
	(154.42,114.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] (154.42,111.94) --
	(153.21,111.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (153.21,109.14) --
	(153.21,114.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] (155.57,277.16) --
	(155.57,282.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (155.57,279.96) --
	(147.96,279.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (147.96,277.16) --
	(147.96,282.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (169.70,249.16) --
	(169.70,254.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (169.70,251.96) --
	(163.08,251.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (163.08,249.16) --
	(163.08,254.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (153.94,165.15) --
	(153.94,170.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (153.94,167.95) --
	(150.21,167.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (150.21,165.15) --
	(150.21,170.75);
\definecolor{fillColor}{RGB}{221,204,119}

\path[fill=fillColor] (155.73,139.95) circle (  3.57);

\path[fill=fillColor] (150.69,447.98) circle (  3.57);

\path[fill=fillColor] (155.99,419.98) circle (  3.57);

\path[fill=fillColor] (150.63,363.97) circle (  1.43);

\path[fill=fillColor] (160.19,335.97) circle (  3.57);

\path[fill=fillColor] (157.99,307.96) circle (  3.57);

\path[fill=fillColor] (153.79,111.94) circle (  3.57);

\path[fill=fillColor] (151.79,279.96) circle (  3.57);

\path[fill=fillColor] (166.39,251.96) circle (  3.57);

\path[fill=fillColor] (152.11,167.95) circle (  3.57);
\definecolor{drawColor}{RGB}{169,169,169}

\path[draw=drawColor,line width= 0.6pt,line join=round] (142.75,153.95) -- (201.69,153.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (142.75,125.95) -- (201.69,125.95);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (142.75, 67.14) rectangle (201.69,464.78);
\end{scope}
\begin{scope}
\path[clip] (207.19, 67.14) rectangle (266.13,464.78);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (207.19, 67.14) rectangle (266.13,464.78);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] (217.49, 67.14) --
	(217.49,464.78);

\path[draw=drawColor,line width= 0.3pt,line join=round] (230.62, 67.14) --
	(230.62,464.78);

\path[draw=drawColor,line width= 0.3pt,line join=round] (243.75, 67.14) --
	(243.75,464.78);

\path[draw=drawColor,line width= 0.3pt,line join=round] (256.89, 67.14) --
	(256.89,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (207.19, 83.94) --
	(266.13, 83.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (207.19,111.94) --
	(266.13,111.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (207.19,139.95) --
	(266.13,139.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (207.19,167.95) --
	(266.13,167.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (207.19,195.95) --
	(266.13,195.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (207.19,223.96) --
	(266.13,223.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (207.19,251.96) --
	(266.13,251.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (207.19,279.96) --
	(266.13,279.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (207.19,307.96) --
	(266.13,307.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (207.19,335.97) --
	(266.13,335.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (207.19,363.97) --
	(266.13,363.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (207.19,391.97) --
	(266.13,391.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (207.19,419.98) --
	(266.13,419.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (207.19,447.98) --
	(266.13,447.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (210.92, 67.14) --
	(210.92,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (224.06, 67.14) --
	(224.06,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (237.19, 67.14) --
	(237.19,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (250.32, 67.14) --
	(250.32,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (263.45, 67.14) --
	(263.45,464.78);
\definecolor{drawColor}{RGB}{17,119,51}

\path[draw=drawColor,line width= 0.6pt,line join=round] (220.80,137.15) --
	(220.80,142.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (220.80,139.95) --
	(214.28,139.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (214.28,137.15) --
	(214.28,142.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (216.28,445.18) --
	(216.28,450.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (216.28,447.98) --
	(213.29,447.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (213.29,445.18) --
	(213.29,450.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (217.23,417.18) --
	(217.23,422.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (217.23,419.98) --
	(212.97,419.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (212.97,417.18) --
	(212.97,422.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (214.13,389.17) --
	(214.13,394.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (214.13,391.97) --
	(210.56,391.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (210.56,389.17) --
	(210.56,394.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (228.84,361.17) --
	(228.84,366.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (228.84,363.97) --
	(214.44,363.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (214.44,361.17) --
	(214.44,366.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (221.11,333.17) --
	(221.11,338.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (221.11,335.97) --
	(217.38,335.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (217.38,333.17) --
	(217.38,338.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (230.83,305.16) --
	(230.83,310.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (230.83,307.96) --
	(221.90,307.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (221.90,305.16) --
	(221.90,310.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (214.13,109.14) --
	(214.13,114.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] (214.13,111.94) --
	(213.34,111.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (213.34,109.14) --
	(213.34,114.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] (225.68,277.16) --
	(225.68,282.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (225.68,279.96) --
	(215.81,279.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (215.81,277.16) --
	(215.81,282.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (221.01,249.16) --
	(221.01,254.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (221.01,251.96) --
	(216.23,251.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (216.23,249.16) --
	(216.23,254.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (214.02,193.15) --
	(214.02,198.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (214.02,195.95) --
	(211.82,195.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (211.82,193.15) --
	(211.82,198.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (214.76,165.15) --
	(214.76,170.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (214.76,167.95) --
	(212.34,167.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (212.34,165.15) --
	(212.34,170.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (234.56, 81.14) --
	(234.56, 86.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] (234.56, 83.94) --
	(228.52, 83.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (228.52, 81.14) --
	(228.52, 86.74);
\definecolor{fillColor}{RGB}{17,119,51}

\path[fill=fillColor] (217.54,139.95) circle (  3.57);

\path[fill=fillColor] (214.81,447.98) circle (  3.57);

\path[fill=fillColor] (215.13,419.98) circle (  3.57);

\path[fill=fillColor] (212.34,391.97) circle (  1.43);

\path[fill=fillColor] (221.64,363.97) circle (  3.57);

\path[fill=fillColor] (219.28,335.97) circle (  3.57);

\path[fill=fillColor] (226.37,307.96) circle (  3.57);

\path[fill=fillColor] (213.76,111.94) circle (  3.57);

\path[fill=fillColor] (220.75,279.96) circle (  3.57);

\path[fill=fillColor] (218.59,251.96) circle (  3.57);

\path[fill=fillColor] (212.92,195.95) circle (  3.57);

\path[fill=fillColor] (213.55,167.95) circle (  3.57);

\path[fill=fillColor] (231.57, 83.94) circle (  3.57);
\definecolor{drawColor}{RGB}{169,169,169}

\path[draw=drawColor,line width= 0.6pt,line join=round] (207.19,153.95) -- (266.13,153.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (207.19,125.95) -- (266.13,125.95);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (207.19, 67.14) rectangle (266.13,464.78);
\end{scope}
\begin{scope}
\path[clip] (271.63, 67.14) rectangle (330.57,464.78);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (271.63, 67.14) rectangle (330.57,464.78);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] (281.93, 67.14) --
	(281.93,464.78);

\path[draw=drawColor,line width= 0.3pt,line join=round] (295.06, 67.14) --
	(295.06,464.78);

\path[draw=drawColor,line width= 0.3pt,line join=round] (308.19, 67.14) --
	(308.19,464.78);

\path[draw=drawColor,line width= 0.3pt,line join=round] (321.32, 67.14) --
	(321.32,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (271.63, 83.94) --
	(330.57, 83.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (271.63,111.94) --
	(330.57,111.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (271.63,139.95) --
	(330.57,139.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (271.63,167.95) --
	(330.57,167.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (271.63,195.95) --
	(330.57,195.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (271.63,223.96) --
	(330.57,223.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (271.63,251.96) --
	(330.57,251.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (271.63,279.96) --
	(330.57,279.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (271.63,307.96) --
	(330.57,307.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (271.63,335.97) --
	(330.57,335.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (271.63,363.97) --
	(330.57,363.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (271.63,391.97) --
	(330.57,391.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (271.63,419.98) --
	(330.57,419.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (271.63,447.98) --
	(330.57,447.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (275.36, 67.14) --
	(275.36,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (288.49, 67.14) --
	(288.49,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (301.63, 67.14) --
	(301.63,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (314.76, 67.14) --
	(314.76,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (327.89, 67.14) --
	(327.89,464.78);
\definecolor{drawColor}{RGB}{51,34,136}

\path[draw=drawColor,line width= 0.6pt,line join=round] (288.34,137.15) --
	(288.34,142.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (288.34,139.95) --
	(282.61,139.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (282.61,137.15) --
	(282.61,142.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (287.86,445.18) --
	(287.86,450.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (287.86,447.98) --
	(283.29,447.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (283.29,445.18) --
	(283.29,450.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (282.98,417.18) --
	(282.98,422.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (282.98,419.98) --
	(278.46,419.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (278.46,417.18) --
	(278.46,422.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (296.79,389.17) --
	(296.79,394.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (296.79,391.97) --
	(285.13,391.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (285.13,389.17) --
	(285.13,394.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (289.65,361.17) --
	(289.65,366.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (289.65,363.97) --
	(277.04,363.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (277.04,361.17) --
	(277.04,366.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (281.72,333.17) --
	(281.72,338.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (281.72,335.97) --
	(278.72,335.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (278.72,333.17) --
	(278.72,338.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (293.22,305.16) --
	(293.22,310.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (293.22,307.96) --
	(284.82,307.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (284.82,305.16) --
	(284.82,310.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (291.70,109.14) --
	(291.70,114.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] (291.70,111.94) --
	(290.12,111.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (290.12,109.14) --
	(290.12,114.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] (291.86,277.16) --
	(291.86,282.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (291.86,279.96) --
	(281.45,279.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (281.45,277.16) --
	(281.45,282.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (281.09,249.16) --
	(281.09,254.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (281.09,251.96) --
	(277.57,251.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (277.57,249.16) --
	(277.57,254.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (287.29,221.16) --
	(287.29,226.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (287.29,223.96) --
	(281.61,223.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (281.61,221.16) --
	(281.61,226.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (290.70,193.15) --
	(290.70,198.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (290.70,195.95) --
	(285.45,195.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (285.45,193.15) --
	(285.45,198.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (294.38,165.15) --
	(294.38,170.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (294.38,167.95) --
	(288.97,167.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (288.97,165.15) --
	(288.97,170.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (302.99, 81.14) --
	(302.99, 86.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] (302.99, 83.94) --
	(296.90, 83.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (296.90, 81.14) --
	(296.90, 86.74);
\definecolor{fillColor}{RGB}{51,34,136}

\path[fill=fillColor] (285.45,139.95) circle (  3.57);

\path[fill=fillColor] (285.60,447.98) circle (  3.57);

\path[fill=fillColor] (280.72,419.98) circle (  3.57);

\path[fill=fillColor] (290.96,391.97) circle (  3.57);

\path[fill=fillColor] (283.35,363.97) circle (  3.57);

\path[fill=fillColor] (280.25,335.97) circle (  3.57);

\path[fill=fillColor] (289.02,307.96) circle (  3.57);

\path[fill=fillColor] (290.91,111.94) circle (  3.57);

\path[fill=fillColor] (286.66,279.96) circle (  3.57);

\path[fill=fillColor] (279.30,251.96) circle (  3.57);

\path[fill=fillColor] (284.45,223.96) circle (  3.57);

\path[fill=fillColor] (288.07,195.95) circle (  3.57);

\path[fill=fillColor] (291.65,167.95) circle (  3.57);

\path[fill=fillColor] (299.95, 83.94) circle (  3.57);
\definecolor{drawColor}{RGB}{169,169,169}

\path[draw=drawColor,line width= 0.6pt,line join=round] (271.63,153.95) -- (330.57,153.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (271.63,125.95) -- (330.57,125.95);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (271.63, 67.14) rectangle (330.57,464.78);
\end{scope}
\begin{scope}
\path[clip] (336.07, 67.14) rectangle (395.01,464.78);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (336.07, 67.14) rectangle (395.01,464.78);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] (346.37, 67.14) --
	(346.37,464.78);

\path[draw=drawColor,line width= 0.3pt,line join=round] (359.50, 67.14) --
	(359.50,464.78);

\path[draw=drawColor,line width= 0.3pt,line join=round] (372.63, 67.14) --
	(372.63,464.78);

\path[draw=drawColor,line width= 0.3pt,line join=round] (385.76, 67.14) --
	(385.76,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (336.07, 83.94) --
	(395.01, 83.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (336.07,111.94) --
	(395.01,111.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (336.07,139.95) --
	(395.01,139.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (336.07,167.95) --
	(395.01,167.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (336.07,195.95) --
	(395.01,195.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (336.07,223.96) --
	(395.01,223.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (336.07,251.96) --
	(395.01,251.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (336.07,279.96) --
	(395.01,279.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (336.07,307.96) --
	(395.01,307.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (336.07,335.97) --
	(395.01,335.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (336.07,363.97) --
	(395.01,363.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (336.07,391.97) --
	(395.01,391.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (336.07,419.98) --
	(395.01,419.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (336.07,447.98) --
	(395.01,447.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (339.80, 67.14) --
	(339.80,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (352.93, 67.14) --
	(352.93,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (366.06, 67.14) --
	(366.06,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (379.20, 67.14) --
	(379.20,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (392.33, 67.14) --
	(392.33,464.78);
\definecolor{drawColor}{RGB}{170,68,153}

\path[draw=drawColor,line width= 0.6pt,line join=round] (347.63,137.15) --
	(347.63,142.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (347.63,139.95) --
	(341.06,139.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (341.06,137.15) --
	(341.06,142.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (348.83,445.18) --
	(348.83,450.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (348.83,447.98) --
	(344.95,447.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (344.95,445.18) --
	(344.95,450.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (342.06,417.18) --
	(342.06,422.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (342.06,419.98) --
	(339.96,419.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (339.96,417.18) --
	(339.96,422.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (354.14,361.17) --
	(354.14,366.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (354.14,363.97) --
	(341.90,363.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (341.90,361.17) --
	(341.90,366.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (356.24,305.16) --
	(356.24,310.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (356.24,307.96) --
	(346.52,307.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (346.52,305.16) --
	(346.52,310.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (343.63,109.14) --
	(343.63,114.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] (343.63,111.94) --
	(342.74,111.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (342.74,109.14) --
	(342.74,114.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] (345.79,277.16) --
	(345.79,282.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (345.79,279.96) --
	(339.85,279.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (339.85,277.16) --
	(339.85,282.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (343.37,249.16) --
	(343.37,254.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (343.37,251.96) --
	(340.64,251.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (340.64,249.16) --
	(340.64,254.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (341.48,193.15) --
	(341.48,198.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (341.48,195.95) --
	(339.90,195.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (339.90,193.15) --
	(339.90,198.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (343.06,165.15) --
	(343.06,170.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (343.06,167.95) --
	(340.80,167.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (340.80,165.15) --
	(340.80,170.75);
\definecolor{fillColor}{RGB}{170,68,153}

\path[fill=fillColor] (344.37,139.95) circle (  3.57);

\path[fill=fillColor] (346.89,447.98) circle (  3.57);

\path[fill=fillColor] (341.01,419.98) circle (  1.43);

\path[fill=fillColor] (348.05,363.97) circle (  3.57);

\path[fill=fillColor] (351.41,307.96) circle (  3.57);

\path[fill=fillColor] (343.16,111.94) circle (  3.57);

\path[fill=fillColor] (342.85,279.96) circle (  1.43);

\path[fill=fillColor] (342.01,251.96) circle (  3.57);

\path[fill=fillColor] (340.69,195.95) circle (  1.43);

\path[fill=fillColor] (341.95,167.95) circle (  3.57);
\definecolor{drawColor}{RGB}{169,169,169}

\path[draw=drawColor,line width= 0.6pt,line join=round] (336.07,153.95) -- (395.01,153.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (336.07,125.95) -- (395.01,125.95);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (336.07, 67.14) rectangle (395.01,464.78);
\end{scope}
\begin{scope}
\path[clip] (400.51, 67.14) rectangle (459.45,464.78);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (400.51, 67.14) rectangle (459.45,464.78);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] (410.80, 67.14) --
	(410.80,464.78);

\path[draw=drawColor,line width= 0.3pt,line join=round] (423.94, 67.14) --
	(423.94,464.78);

\path[draw=drawColor,line width= 0.3pt,line join=round] (437.07, 67.14) --
	(437.07,464.78);

\path[draw=drawColor,line width= 0.3pt,line join=round] (450.20, 67.14) --
	(450.20,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (400.51, 83.94) --
	(459.45, 83.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (400.51,111.94) --
	(459.45,111.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (400.51,139.95) --
	(459.45,139.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (400.51,167.95) --
	(459.45,167.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (400.51,195.95) --
	(459.45,195.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (400.51,223.96) --
	(459.45,223.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (400.51,251.96) --
	(459.45,251.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (400.51,279.96) --
	(459.45,279.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (400.51,307.96) --
	(459.45,307.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (400.51,335.97) --
	(459.45,335.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (400.51,363.97) --
	(459.45,363.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (400.51,391.97) --
	(459.45,391.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (400.51,419.98) --
	(459.45,419.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (400.51,447.98) --
	(459.45,447.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (404.24, 67.14) --
	(404.24,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (417.37, 67.14) --
	(417.37,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (430.50, 67.14) --
	(430.50,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (443.64, 67.14) --
	(443.64,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (456.77, 67.14) --
	(456.77,464.78);
\definecolor{drawColor}{RGB}{68,170,153}

\path[draw=drawColor,line width= 0.6pt,line join=round] (407.81,137.15) --
	(407.81,142.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (407.81,139.95) --
	(405.13,139.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (405.13,137.15) --
	(405.13,142.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (407.18,445.18) --
	(407.18,450.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (407.18,447.98) --
	(405.03,447.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (405.03,445.18) --
	(405.03,450.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (405.18,417.18) --
	(405.18,422.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (405.18,419.98) --
	(403.92,419.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (403.92,417.18) --
	(403.92,422.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (409.81,361.17) --
	(409.81,366.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (409.81,363.97) --
	(403.24,363.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (403.24,361.17) --
	(403.24,366.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (413.38,305.16) --
	(413.38,310.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (413.38,307.96) --
	(407.13,307.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (407.13,305.16) --
	(407.13,310.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (406.50,109.14) --
	(406.50,114.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] (406.50,111.94) --
	(405.87,111.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (405.87,109.14) --
	(405.87,114.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] (411.17,277.16) --
	(411.17,282.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (411.17,279.96) --
	(404.71,279.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (404.71,277.16) --
	(404.71,282.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (406.97,249.16) --
	(406.97,254.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (406.97,251.96) --
	(404.66,251.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (404.66,249.16) --
	(404.66,254.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (408.34,221.16) --
	(408.34,226.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (408.34,223.96) --
	(405.55,223.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (405.55,221.16) --
	(405.55,226.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (405.92,193.15) --
	(405.92,198.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (405.92,195.95) --
	(404.34,195.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (404.34,193.15) --
	(404.34,198.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (405.76,165.15) --
	(405.76,170.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (405.76,167.95) --
	(404.34,167.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (404.34,165.15) --
	(404.34,170.75);
\definecolor{fillColor}{RGB}{68,170,153}

\path[fill=fillColor] (406.50,139.95) circle (  3.57);

\path[fill=fillColor] (406.08,447.98) circle (  3.57);

\path[fill=fillColor] (404.55,419.98) circle (  1.43);

\path[fill=fillColor] (406.55,363.97) circle (  1.43);

\path[fill=fillColor] (410.28,307.96) circle (  3.57);

\path[fill=fillColor] (406.18,111.94) circle (  3.57);

\path[fill=fillColor] (407.92,279.96) circle (  1.43);

\path[fill=fillColor] (405.81,251.96) circle (  3.57);

\path[fill=fillColor] (406.92,223.96) circle (  3.57);

\path[fill=fillColor] (405.13,195.95) circle (  1.43);

\path[fill=fillColor] (405.08,167.95) circle (  3.57);
\definecolor{drawColor}{RGB}{169,169,169}

\path[draw=drawColor,line width= 0.6pt,line join=round] (400.51,153.95) -- (459.45,153.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (400.51,125.95) -- (459.45,125.95);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (400.51, 67.14) rectangle (459.45,464.78);
\end{scope}
\begin{scope}
\path[clip] (464.95, 67.14) rectangle (523.88,464.78);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (464.95, 67.14) rectangle (523.88,464.78);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] (475.24, 67.14) --
	(475.24,464.78);

\path[draw=drawColor,line width= 0.3pt,line join=round] (488.37, 67.14) --
	(488.37,464.78);

\path[draw=drawColor,line width= 0.3pt,line join=round] (501.51, 67.14) --
	(501.51,464.78);

\path[draw=drawColor,line width= 0.3pt,line join=round] (514.64, 67.14) --
	(514.64,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (464.95, 83.94) --
	(523.88, 83.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (464.95,111.94) --
	(523.88,111.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (464.95,139.95) --
	(523.88,139.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (464.95,167.95) --
	(523.88,167.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (464.95,195.95) --
	(523.88,195.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (464.95,223.96) --
	(523.88,223.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (464.95,251.96) --
	(523.88,251.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (464.95,279.96) --
	(523.88,279.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (464.95,307.96) --
	(523.88,307.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (464.95,335.97) --
	(523.88,335.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (464.95,363.97) --
	(523.88,363.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (464.95,391.97) --
	(523.88,391.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (464.95,419.98) --
	(523.88,419.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (464.95,447.98) --
	(523.88,447.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (468.68, 67.14) --
	(468.68,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (481.81, 67.14) --
	(481.81,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (494.94, 67.14) --
	(494.94,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (508.07, 67.14) --
	(508.07,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (521.21, 67.14) --
	(521.21,464.78);
\definecolor{drawColor}{RGB}{153,153,51}

\path[draw=drawColor,line width= 0.6pt,line join=round] (470.41,137.15) --
	(470.41,142.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (470.41,139.95) --
	(468.57,139.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (468.57,137.15) --
	(468.57,142.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (470.15,445.18) --
	(470.15,450.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (470.15,447.98) --
	(468.78,447.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (468.78,445.18) --
	(468.78,450.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (469.36,417.18) --
	(469.36,422.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (469.36,419.98) --
	(468.47,419.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (468.47,417.18) --
	(468.47,422.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (471.57,361.17) --
	(471.57,366.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (471.57,363.97) --
	(467.68,363.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (467.68,361.17) --
	(467.68,366.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (469.04,333.17) --
	(469.04,338.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (469.04,335.97) --
	(468.57,335.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (468.57,333.17) --
	(468.57,338.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (474.56,109.14) --
	(474.56,114.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] (474.56,111.94) --
	(473.51,111.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (473.51,109.14) --
	(473.51,114.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] (474.51,277.16) --
	(474.51,282.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (474.51,279.96) --
	(468.73,279.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (468.73,277.16) --
	(468.73,282.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (469.83,249.16) --
	(469.83,254.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (469.83,251.96) --
	(468.47,251.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (468.47,249.16) --
	(468.47,254.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (469.15,165.15) --
	(469.15,170.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (469.15,167.95) --
	(468.52,167.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (468.52,165.15) --
	(468.52,170.75);
\definecolor{fillColor}{RGB}{153,153,51}

\path[fill=fillColor] (469.46,139.95) circle (  3.57);

\path[fill=fillColor] (469.46,447.98) circle (  1.43);

\path[fill=fillColor] (468.89,419.98) circle (  1.43);

\path[fill=fillColor] (469.62,363.97) circle (  1.43);

\path[fill=fillColor] (468.78,335.97) circle (  1.43);

\path[fill=fillColor] (474.03,111.94) circle (  3.57);

\path[fill=fillColor] (471.62,279.96) circle (  1.43);

\path[fill=fillColor] (469.15,251.96) circle (  1.43);

\path[fill=fillColor] (468.83,167.95) circle (  1.43);
\definecolor{drawColor}{RGB}{169,169,169}

\path[draw=drawColor,line width= 0.6pt,line join=round] (464.95,153.95) -- (523.88,153.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (464.95,125.95) -- (523.88,125.95);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (464.95, 67.14) rectangle (523.88,464.78);
\end{scope}
\begin{scope}
\path[clip] (529.38, 67.14) rectangle (588.32,464.78);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (529.38, 67.14) rectangle (588.32,464.78);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] (539.68, 67.14) --
	(539.68,464.78);

\path[draw=drawColor,line width= 0.3pt,line join=round] (552.81, 67.14) --
	(552.81,464.78);

\path[draw=drawColor,line width= 0.3pt,line join=round] (565.95, 67.14) --
	(565.95,464.78);

\path[draw=drawColor,line width= 0.3pt,line join=round] (579.08, 67.14) --
	(579.08,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (529.38, 83.94) --
	(588.32, 83.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (529.38,111.94) --
	(588.32,111.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (529.38,139.95) --
	(588.32,139.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (529.38,167.95) --
	(588.32,167.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (529.38,195.95) --
	(588.32,195.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (529.38,223.96) --
	(588.32,223.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (529.38,251.96) --
	(588.32,251.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (529.38,279.96) --
	(588.32,279.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (529.38,307.96) --
	(588.32,307.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (529.38,335.97) --
	(588.32,335.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (529.38,363.97) --
	(588.32,363.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (529.38,391.97) --
	(588.32,391.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (529.38,419.98) --
	(588.32,419.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (529.38,447.98) --
	(588.32,447.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (533.11, 67.14) --
	(533.11,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (546.25, 67.14) --
	(546.25,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (559.38, 67.14) --
	(559.38,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (572.51, 67.14) --
	(572.51,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (585.64, 67.14) --
	(585.64,464.78);
\definecolor{drawColor}{RGB}{136,34,85}

\path[draw=drawColor,line width= 0.6pt,line join=round] (534.80,137.15) --
	(534.80,142.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (534.80,139.95) --
	(532.80,139.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (532.80,137.15) --
	(532.80,142.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (534.11,445.18) --
	(534.11,450.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (534.11,447.98) --
	(533.06,447.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (533.06,445.18) --
	(533.06,450.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (533.90,417.18) --
	(533.90,422.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (533.90,419.98) --
	(532.85,419.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (532.85,417.18) --
	(532.85,422.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (533.85,109.14) --
	(533.85,114.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] (533.85,111.94) --
	(533.48,111.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (533.48,109.14) --
	(533.48,114.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] (536.53,249.16) --
	(536.53,254.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (536.53,251.96) --
	(533.85,251.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (533.85,249.16) --
	(533.85,254.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (534.27,193.15) --
	(534.27,198.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (534.27,195.95) --
	(533.06,195.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (533.06,193.15) --
	(533.06,198.75);
\definecolor{fillColor}{RGB}{136,34,85}

\path[fill=fillColor] (533.80,139.95) circle (  3.57);

\path[fill=fillColor] (533.59,447.98) circle (  1.43);

\path[fill=fillColor] (533.38,419.98) circle (  1.43);

\path[fill=fillColor] (533.64,111.94) circle (  3.57);

\path[fill=fillColor] (535.22,251.96) circle (  3.57);

\path[fill=fillColor] (533.64,195.95) circle (  1.43);
\definecolor{drawColor}{RGB}{169,169,169}

\path[draw=drawColor,line width= 0.6pt,line join=round] (529.38,153.95) -- (588.32,153.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (529.38,125.95) -- (588.32,125.95);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (529.38, 67.14) rectangle (588.32,464.78);
\end{scope}
\begin{scope}
\path[clip] (593.82, 67.14) rectangle (652.76,464.78);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (593.82, 67.14) rectangle (652.76,464.78);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] (604.12, 67.14) --
	(604.12,464.78);

\path[draw=drawColor,line width= 0.3pt,line join=round] (617.25, 67.14) --
	(617.25,464.78);

\path[draw=drawColor,line width= 0.3pt,line join=round] (630.38, 67.14) --
	(630.38,464.78);

\path[draw=drawColor,line width= 0.3pt,line join=round] (643.52, 67.14) --
	(643.52,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (593.82, 83.94) --
	(652.76, 83.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (593.82,111.94) --
	(652.76,111.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (593.82,139.95) --
	(652.76,139.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (593.82,167.95) --
	(652.76,167.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (593.82,195.95) --
	(652.76,195.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (593.82,223.96) --
	(652.76,223.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (593.82,251.96) --
	(652.76,251.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (593.82,279.96) --
	(652.76,279.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (593.82,307.96) --
	(652.76,307.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (593.82,335.97) --
	(652.76,335.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (593.82,363.97) --
	(652.76,363.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (593.82,391.97) --
	(652.76,391.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (593.82,419.98) --
	(652.76,419.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (593.82,447.98) --
	(652.76,447.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (597.55, 67.14) --
	(597.55,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (610.69, 67.14) --
	(610.69,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (623.82, 67.14) --
	(623.82,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (636.95, 67.14) --
	(636.95,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (650.08, 67.14) --
	(650.08,464.78);
\definecolor{drawColor}{RGB}{102,17,0}

\path[draw=drawColor,line width= 0.6pt,line join=round] (606.48,137.15) --
	(606.48,142.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (606.48,139.95) --
	(600.76,139.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (600.76,137.15) --
	(600.76,142.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (609.16,445.18) --
	(609.16,450.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (609.16,447.98) --
	(604.75,447.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (604.75,445.18) --
	(604.75,450.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (605.01,417.18) --
	(605.01,422.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (605.01,419.98) --
	(600.60,419.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (600.60,417.18) --
	(600.60,422.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (601.34,361.17) --
	(601.34,366.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (601.34,363.97) --
	(596.76,363.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (596.76,361.17) --
	(596.76,366.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (601.23,333.17) --
	(601.23,338.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (601.23,335.97) --
	(599.02,335.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (599.02,333.17) --
	(599.02,338.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (607.74,305.16) --
	(607.74,310.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (607.74,307.96) --
	(601.28,307.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (601.28,305.16) --
	(601.28,310.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (609.53,109.14) --
	(609.53,114.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] (609.53,111.94) --
	(608.11,111.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (608.11,109.14) --
	(608.11,114.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] (613.89,277.16) --
	(613.89,282.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (613.89,279.96) --
	(603.59,279.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (603.59,277.16) --
	(603.59,282.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (610.95,249.16) --
	(610.95,254.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (610.95,251.96) --
	(605.48,251.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (605.48,249.16) --
	(605.48,254.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (600.55,221.16) --
	(600.55,226.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (600.55,223.96) --
	(598.24,223.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (598.24,221.16) --
	(598.24,226.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (604.75,165.15) --
	(604.75,170.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (604.75,167.95) --
	(601.23,167.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (601.23,165.15) --
	(601.23,170.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (602.12, 81.14) --
	(602.12, 86.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] (602.12, 83.94) --
	(599.34, 83.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (599.34, 81.14) --
	(599.34, 86.74);
\definecolor{fillColor}{RGB}{102,17,0}

\path[fill=fillColor] (603.65,139.95) circle (  3.57);

\path[fill=fillColor] (606.96,447.98) circle (  3.57);

\path[fill=fillColor] (602.81,419.98) circle (  3.57);

\path[fill=fillColor] (599.02,363.97) circle (  1.43);

\path[fill=fillColor] (600.13,335.97) circle (  3.57);

\path[fill=fillColor] (604.49,307.96) circle (  3.57);

\path[fill=fillColor] (608.79,111.94) circle (  3.57);

\path[fill=fillColor] (608.74,279.96) circle (  3.57);

\path[fill=fillColor] (608.22,251.96) circle (  3.57);

\path[fill=fillColor] (599.39,223.96) circle (  3.57);

\path[fill=fillColor] (602.96,167.95) circle (  3.57);

\path[fill=fillColor] (600.70, 83.94) circle (  3.57);
\definecolor{drawColor}{RGB}{169,169,169}

\path[draw=drawColor,line width= 0.6pt,line join=round] (593.82,153.95) -- (652.76,153.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (593.82,125.95) -- (652.76,125.95);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (593.82, 67.14) rectangle (652.76,464.78);
\end{scope}
\begin{scope}
\path[clip] (658.26, 67.14) rectangle (717.20,464.78);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (658.26, 67.14) rectangle (717.20,464.78);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] (668.56, 67.14) --
	(668.56,464.78);

\path[draw=drawColor,line width= 0.3pt,line join=round] (681.69, 67.14) --
	(681.69,464.78);

\path[draw=drawColor,line width= 0.3pt,line join=round] (694.82, 67.14) --
	(694.82,464.78);

\path[draw=drawColor,line width= 0.3pt,line join=round] (707.95, 67.14) --
	(707.95,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (658.26, 83.94) --
	(717.20, 83.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (658.26,111.94) --
	(717.20,111.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (658.26,139.95) --
	(717.20,139.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (658.26,167.95) --
	(717.20,167.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (658.26,195.95) --
	(717.20,195.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (658.26,223.96) --
	(717.20,223.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (658.26,251.96) --
	(717.20,251.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (658.26,279.96) --
	(717.20,279.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (658.26,307.96) --
	(717.20,307.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (658.26,335.97) --
	(717.20,335.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (658.26,363.97) --
	(717.20,363.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (658.26,391.97) --
	(717.20,391.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (658.26,419.98) --
	(717.20,419.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (658.26,447.98) --
	(717.20,447.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (661.99, 67.14) --
	(661.99,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (675.12, 67.14) --
	(675.12,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (688.26, 67.14) --
	(688.26,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (701.39, 67.14) --
	(701.39,464.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (714.52, 67.14) --
	(714.52,464.78);
\definecolor{drawColor}{RGB}{102,153,204}

\path[draw=drawColor,line width= 0.6pt,line join=round] (674.44,137.15) --
	(674.44,142.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (674.44,139.95) --
	(667.77,139.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (667.77,137.15) --
	(667.77,142.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (668.19,445.18) --
	(668.19,450.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (668.19,447.98) --
	(664.93,447.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (664.93,445.18) --
	(664.93,450.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (682.06,417.18) --
	(682.06,422.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (682.06,419.98) --
	(675.18,419.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (675.18,417.18) --
	(675.18,422.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (678.28,389.17) --
	(678.28,394.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (678.28,391.97) --
	(668.40,391.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (668.40,389.17) --
	(668.40,394.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (674.34,361.17) --
	(674.34,366.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (674.34,363.97) --
	(662.41,363.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (662.41,361.17) --
	(662.41,366.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (678.33,333.17) --
	(678.33,338.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (678.33,335.97) --
	(673.81,335.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (673.81,333.17) --
	(673.81,338.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (677.12,305.16) --
	(677.12,310.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (677.12,307.96) --
	(667.45,307.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (667.45,305.16) --
	(667.45,310.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (665.04,109.14) --
	(665.04,114.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] (665.04,111.94) --
	(664.30,111.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (664.30,109.14) --
	(664.30,114.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] (681.06,277.16) --
	(681.06,282.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (681.06,279.96) --
	(670.03,279.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (670.03,277.16) --
	(670.03,282.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (666.56,249.16) --
	(666.56,254.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (666.56,251.96) --
	(663.46,251.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (663.46,249.16) --
	(663.46,254.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (677.91,221.16) --
	(677.91,226.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (677.91,223.96) --
	(672.13,223.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (672.13,221.16) --
	(672.13,226.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (667.77,193.15) --
	(667.77,198.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (667.77,195.95) --
	(664.57,195.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (664.57,193.15) --
	(664.57,198.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (666.46,165.15) --
	(666.46,170.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (666.46,167.95) --
	(663.78,167.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (663.78,165.15) --
	(663.78,170.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (690.83, 81.14) --
	(690.83, 86.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] (690.83, 83.94) --
	(684.74, 83.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (684.74, 81.14) --
	(684.74, 86.74);
\definecolor{fillColor}{RGB}{102,153,204}

\path[fill=fillColor] (671.08,139.95) circle (  3.57);

\path[fill=fillColor] (666.56,447.98) circle (  3.57);

\path[fill=fillColor] (678.59,419.98) circle (  3.57);

\path[fill=fillColor] (673.34,391.97) circle (  3.57);

\path[fill=fillColor] (668.35,363.97) circle (  3.57);

\path[fill=fillColor] (676.07,335.97) circle (  3.57);

\path[fill=fillColor] (672.29,307.96) circle (  3.57);

\path[fill=fillColor] (664.67,111.94) circle (  3.57);

\path[fill=fillColor] (675.54,279.96) circle (  3.57);

\path[fill=fillColor] (664.99,251.96) circle (  3.57);

\path[fill=fillColor] (675.02,223.96) circle (  3.57);

\path[fill=fillColor] (666.19,195.95) circle (  3.57);

\path[fill=fillColor] (665.14,167.95) circle (  3.57);

\path[fill=fillColor] (687.78, 83.94) circle (  3.57);
\definecolor{drawColor}{RGB}{169,169,169}

\path[draw=drawColor,line width= 0.6pt,line join=round] (658.26,153.95) -- (717.20,153.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (658.26,125.95) -- (717.20,125.95);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (658.26, 67.14) rectangle (717.20,464.78);
\end{scope}
\begin{scope}
\path[clip] ( 78.32,464.78) rectangle (137.25,519.37);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 78.32,464.78) rectangle (137.25,519.37);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (107.79,498.55) {Concerned};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (107.79,489.04) {about side};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (107.79,479.54) {effects};
\end{scope}
\begin{scope}
\path[clip] (142.75,464.78) rectangle (201.69,519.37);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (142.75,464.78) rectangle (201.69,519.37);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (172.22,508.05) {Concerned};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (172.22,498.55) {about getting};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (172.22,489.04) {coronavirus};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (172.22,479.54) {from the};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (172.22,470.04) {vaccine};
\end{scope}
\begin{scope}
\path[clip] (207.19,464.78) rectangle (266.13,519.37);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (207.19,464.78) rectangle (266.13,519.37);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (236.66,498.55) {Not concerned};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (236.66,489.04) {about getting};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (236.66,479.54) {seriously ill};
\end{scope}
\begin{scope}
\path[clip] (271.63,464.78) rectangle (330.57,519.37);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (271.63,464.78) rectangle (330.57,519.37);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (301.10,498.55) {Doesn't think};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (301.10,489.04) {vaccines are};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (301.10,479.54) {effective};
\end{scope}
\begin{scope}
\path[clip] (336.07,464.78) rectangle (395.01,519.37);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (336.07,464.78) rectangle (395.01,519.37);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (365.54,508.05) {Doesn't think};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (365.54,498.55) {Coronavirus};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (365.54,489.04) {outbreak is as};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (365.54,479.54) {serious as};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (365.54,470.04) {people say};
\end{scope}
\begin{scope}
\path[clip] (400.51,464.78) rectangle (459.45,519.37);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (400.51,464.78) rectangle (459.45,519.37);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (429.98,493.80) {Doesn't like};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (429.98,484.29) {needles};
\end{scope}
\begin{scope}
\path[clip] (464.95,464.78) rectangle (523.88,519.37);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (464.95,464.78) rectangle (523.88,519.37);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (494.42,493.80) {Allergic to};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (494.42,484.29) {vaccines};
\end{scope}
\begin{scope}
\path[clip] (529.38,464.78) rectangle (588.32,519.37);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (529.38,464.78) rectangle (588.32,519.37);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (558.85,498.55) {Won't have};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (558.85,489.04) {time to get};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (558.85,479.54) {vaccinated};
\end{scope}
\begin{scope}
\path[clip] (593.82,464.78) rectangle (652.76,519.37);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (593.82,464.78) rectangle (652.76,519.37);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (623.29,498.55) {Mentions a};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (623.29,489.04) {conspiracy};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (623.29,479.54) {theory};
\end{scope}
\begin{scope}
\path[clip] (658.26,464.78) rectangle (717.20,519.37);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (658.26,464.78) rectangle (717.20,519.37);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (687.73,489.04) {Other reasons};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 82.05, 64.39) --
	( 82.05, 67.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 95.18, 64.39) --
	( 95.18, 67.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (108.31, 64.39) --
	(108.31, 67.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (121.44, 64.39) --
	(121.44, 67.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (134.58, 64.39) --
	(134.58, 67.14);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 82.05, 56.13) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 95.18, 56.13) {25};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (108.31, 56.13) {50};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (121.44, 56.13) {75};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (134.58, 56.13) {100};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (146.48, 64.39) --
	(146.48, 67.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (159.62, 64.39) --
	(159.62, 67.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (172.75, 64.39) --
	(172.75, 67.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (185.88, 64.39) --
	(185.88, 67.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (199.01, 64.39) --
	(199.01, 67.14);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (146.48, 56.13) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (159.62, 56.13) {25};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (172.75, 56.13) {50};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (185.88, 56.13) {75};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (199.01, 56.13) {100};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (210.92, 64.39) --
	(210.92, 67.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (224.06, 64.39) --
	(224.06, 67.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (237.19, 64.39) --
	(237.19, 67.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (250.32, 64.39) --
	(250.32, 67.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (263.45, 64.39) --
	(263.45, 67.14);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (210.92, 56.13) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (224.06, 56.13) {25};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (237.19, 56.13) {50};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (250.32, 56.13) {75};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (263.45, 56.13) {100};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (275.36, 64.39) --
	(275.36, 67.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (288.49, 64.39) --
	(288.49, 67.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (301.63, 64.39) --
	(301.63, 67.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (314.76, 64.39) --
	(314.76, 67.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (327.89, 64.39) --
	(327.89, 67.14);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (275.36, 56.13) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (288.49, 56.13) {25};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (301.63, 56.13) {50};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (314.76, 56.13) {75};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (327.89, 56.13) {100};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (339.80, 64.39) --
	(339.80, 67.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (352.93, 64.39) --
	(352.93, 67.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (366.06, 64.39) --
	(366.06, 67.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (379.20, 64.39) --
	(379.20, 67.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (392.33, 64.39) --
	(392.33, 67.14);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (339.80, 56.13) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (352.93, 56.13) {25};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (366.06, 56.13) {50};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (379.20, 56.13) {75};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (392.33, 56.13) {100};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (404.24, 64.39) --
	(404.24, 67.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (417.37, 64.39) --
	(417.37, 67.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (430.50, 64.39) --
	(430.50, 67.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (443.64, 64.39) --
	(443.64, 67.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (456.77, 64.39) --
	(456.77, 67.14);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (404.24, 56.13) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (417.37, 56.13) {25};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (430.50, 56.13) {50};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (443.64, 56.13) {75};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (456.77, 56.13) {100};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (468.68, 64.39) --
	(468.68, 67.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (481.81, 64.39) --
	(481.81, 67.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (494.94, 64.39) --
	(494.94, 67.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (508.07, 64.39) --
	(508.07, 67.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (521.21, 64.39) --
	(521.21, 67.14);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (468.68, 56.13) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (481.81, 56.13) {25};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (494.94, 56.13) {50};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (508.07, 56.13) {75};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (521.21, 56.13) {100};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (533.11, 64.39) --
	(533.11, 67.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (546.25, 64.39) --
	(546.25, 67.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (559.38, 64.39) --
	(559.38, 67.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (572.51, 64.39) --
	(572.51, 67.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (585.64, 64.39) --
	(585.64, 67.14);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (533.11, 56.13) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (546.25, 56.13) {25};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (559.38, 56.13) {50};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (572.51, 56.13) {75};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (585.64, 56.13) {100};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (597.55, 64.39) --
	(597.55, 67.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (610.69, 64.39) --
	(610.69, 67.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (623.82, 64.39) --
	(623.82, 67.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (636.95, 64.39) --
	(636.95, 67.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (650.08, 64.39) --
	(650.08, 67.14);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (597.55, 56.13) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (610.69, 56.13) {25};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (623.82, 56.13) {50};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (636.95, 56.13) {75};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (650.08, 56.13) {100};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (661.99, 64.39) --
	(661.99, 67.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (675.12, 64.39) --
	(675.12, 67.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (688.26, 64.39) --
	(688.26, 67.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (701.39, 64.39) --
	(701.39, 67.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (714.52, 64.39) --
	(714.52, 67.14);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (661.99, 56.13) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (675.12, 56.13) {25};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (688.26, 56.13) {50};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (701.39, 56.13) {75};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (714.52, 56.13) {100};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96, 80.91) {USA};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,108.91) {Russia};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,136.92) {All LMICs};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,164.92) {Uganda 2};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,192.92) {Uganda 1};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,220.93) {Sierra Leone 2};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,248.93) {Sierra Leone 1};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,276.93) {Rwanda};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,304.93) {Pakistan 1};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,332.94) {Nigeria};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,360.94) {Nepal};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,388.94) {Mozambique};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,416.95) {Colombia};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,444.95) {Burkina Faso};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 75.57, 83.94) --
	( 78.32, 83.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 75.57,111.94) --
	( 78.32,111.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 75.57,139.95) --
	( 78.32,139.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 75.57,167.95) --
	( 78.32,167.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 75.57,195.95) --
	( 78.32,195.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 75.57,223.96) --
	( 78.32,223.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 75.57,251.96) --
	( 78.32,251.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 75.57,279.96) --
	( 78.32,279.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 75.57,307.96) --
	( 78.32,307.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 75.57,335.97) --
	( 78.32,335.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 75.57,363.97) --
	( 78.32,363.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 75.57,391.97) --
	( 78.32,391.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 75.57,419.98) --
	( 78.32,419.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 75.57,447.98) --
	( 78.32,447.98);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (397.76, 44.09) {estimate};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (284.57,  5.50) rectangle (510.95, 30.95);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.10] at (290.07, 14.44) {Number of observations};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (409.70, 11.00) rectangle (424.15, 25.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] (416.92, 18.23) circle (  1.43);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (465.45, 11.00) rectangle (479.91, 25.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{fillColor}{RGB}{0,0,0}

\path[fill=fillColor] (472.68, 18.23) circle (  3.57);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (429.65, 15.20) {[50,500)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (485.41, 15.20) {500+};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.32] at (  5.50,527.43) {Why would you not take the COVID-19 vaccine?};
\end{scope}
\end{tikzpicture}

\end{document}