### 2021-05-04 paper_files/figure-latex/fig3paper-1.tex

% Created by tikzDevice version 0.12.3.1 on 2021-05-04 13:13:00
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}

\nonstopmode

\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (722.70,650.43);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.00) rectangle (722.70,650.43);
\end{scope}
\begin{scope}
\path[clip] ( 48.28,541.57) rectangle (371.71,627.77);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 48.28,541.57) rectangle (371.71,627.77);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 48.28,556.45) --
	(371.71,556.45);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 48.28,578.39) --
	(371.71,578.39);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 48.28,600.33) --
	(371.71,600.33);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 48.28,622.27) --
	(371.71,622.27);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 48.28,545.48) --
	(371.71,545.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 48.28,567.42) --
	(371.71,567.42);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 48.28,589.36) --
	(371.71,589.36);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 48.28,611.30) --
	(371.71,611.30);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 75.24,541.57) --
	( 75.24,627.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (120.16,541.57) --
	(120.16,627.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (165.08,541.57) --
	(165.08,627.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (209.99,541.57) --
	(209.99,627.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (254.91,541.57) --
	(254.91,627.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (299.83,541.57) --
	(299.83,627.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (344.75,541.57) --
	(344.75,627.77);
\definecolor{fillColor}{RGB}{221,204,119}

\path[fill=fillColor] (144.86,545.48) rectangle (185.29,561.63);

\path[fill=fillColor] (234.70,545.48) rectangle (275.13,547.15);

\path[fill=fillColor] (189.78,545.48) rectangle (230.21,547.06);

\path[fill=fillColor] (279.62,545.48) rectangle (320.05,555.66);

\path[fill=fillColor] (324.54,545.48) rectangle (364.97,546.98);

\path[fill=fillColor] ( 55.02,545.48) rectangle ( 95.45,590.77);

\path[fill=fillColor] ( 99.94,545.48) rectangle (140.37,556.81);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 48.28,541.57) rectangle (371.71,627.77);
\end{scope}
\begin{scope}
\path[clip] ( 48.28,449.86) rectangle (371.71,536.07);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 48.28,449.86) rectangle (371.71,536.07);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 48.28,464.75) --
	(371.71,464.75);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 48.28,486.69) --
	(371.71,486.69);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 48.28,508.63) --
	(371.71,508.63);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 48.28,530.57) --
	(371.71,530.57);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 48.28,453.78) --
	(371.71,453.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 48.28,475.72) --
	(371.71,475.72);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 48.28,497.66) --
	(371.71,497.66);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 48.28,519.60) --
	(371.71,519.60);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 75.24,449.86) --
	( 75.24,536.07);

\path[draw=drawColor,line width= 0.6pt,line join=round] (120.16,449.86) --
	(120.16,536.07);

\path[draw=drawColor,line width= 0.6pt,line join=round] (165.08,449.86) --
	(165.08,536.07);

\path[draw=drawColor,line width= 0.6pt,line join=round] (209.99,449.86) --
	(209.99,536.07);

\path[draw=drawColor,line width= 0.6pt,line join=round] (254.91,449.86) --
	(254.91,536.07);

\path[draw=drawColor,line width= 0.6pt,line join=round] (299.83,449.86) --
	(299.83,536.07);

\path[draw=drawColor,line width= 0.6pt,line join=round] (344.75,449.86) --
	(344.75,536.07);
\definecolor{fillColor}{RGB}{221,204,119}

\path[fill=fillColor] (144.86,453.78) rectangle (185.29,485.02);

\path[fill=fillColor] (234.70,453.78) rectangle (275.13,454.13);

\path[fill=fillColor] (189.78,453.78) rectangle (230.21,467.91);

\path[fill=fillColor] (279.62,453.78) rectangle (320.05,454.22);

\path[fill=fillColor] (324.54,453.78) rectangle (364.97,455.88);

\path[fill=fillColor] ( 55.02,453.78) rectangle ( 95.45,492.57);

\path[fill=fillColor] ( 99.94,453.78) rectangle (140.37,454.48);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 48.28,449.86) rectangle (371.71,536.07);
\end{scope}
\begin{scope}
\path[clip] ( 48.28,358.15) rectangle (371.71,444.36);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 48.28,358.15) rectangle (371.71,444.36);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 48.28,373.04) --
	(371.71,373.04);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 48.28,394.98) --
	(371.71,394.98);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 48.28,416.92) --
	(371.71,416.92);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 48.28,438.86) --
	(371.71,438.86);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 48.28,362.07) --
	(371.71,362.07);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 48.28,384.01) --
	(371.71,384.01);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 48.28,405.95) --
	(371.71,405.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 48.28,427.89) --
	(371.71,427.89);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 75.24,358.15) --
	( 75.24,444.36);

\path[draw=drawColor,line width= 0.6pt,line join=round] (120.16,358.15) --
	(120.16,444.36);

\path[draw=drawColor,line width= 0.6pt,line join=round] (165.08,358.15) --
	(165.08,444.36);

\path[draw=drawColor,line width= 0.6pt,line join=round] (209.99,358.15) --
	(209.99,444.36);

\path[draw=drawColor,line width= 0.6pt,line join=round] (254.91,358.15) --
	(254.91,444.36);

\path[draw=drawColor,line width= 0.6pt,line join=round] (299.83,358.15) --
	(299.83,444.36);

\path[draw=drawColor,line width= 0.6pt,line join=round] (344.75,358.15) --
	(344.75,444.36);
\definecolor{fillColor}{RGB}{221,204,119}

\path[fill=fillColor] (144.86,362.07) rectangle (185.29,375.24);

\path[fill=fillColor] (234.70,362.07) rectangle (275.13,362.60);

\path[fill=fillColor] (189.78,362.07) rectangle (230.21,363.04);

\path[fill=fillColor] (279.62,362.07) rectangle (320.05,391.21);

\path[fill=fillColor] (324.54,362.07) rectangle (364.97,362.42);

\path[fill=fillColor] ( 55.02,362.07) rectangle ( 95.45,382.26);

\path[fill=fillColor] ( 99.94,362.07) rectangle (140.37,385.50);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 48.28,358.15) rectangle (371.71,444.36);
\end{scope}
\begin{scope}
\path[clip] ( 48.28,266.45) rectangle (371.71,352.65);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 48.28,266.45) rectangle (371.71,352.65);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 48.28,281.33) --
	(371.71,281.33);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 48.28,303.27) --
	(371.71,303.27);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 48.28,325.21) --
	(371.71,325.21);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 48.28,347.15) --
	(371.71,347.15);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 48.28,270.36) --
	(371.71,270.36);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 48.28,292.30) --
	(371.71,292.30);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 48.28,314.24) --
	(371.71,314.24);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 48.28,336.18) --
	(371.71,336.18);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 75.24,266.45) --
	( 75.24,352.65);

\path[draw=drawColor,line width= 0.6pt,line join=round] (120.16,266.45) --
	(120.16,352.65);

\path[draw=drawColor,line width= 0.6pt,line join=round] (165.08,266.45) --
	(165.08,352.65);

\path[draw=drawColor,line width= 0.6pt,line join=round] (209.99,266.45) --
	(209.99,352.65);

\path[draw=drawColor,line width= 0.6pt,line join=round] (254.91,266.45) --
	(254.91,352.65);

\path[draw=drawColor,line width= 0.6pt,line join=round] (299.83,266.45) --
	(299.83,352.65);

\path[draw=drawColor,line width= 0.6pt,line join=round] (344.75,266.45) --
	(344.75,352.65);
\definecolor{fillColor}{RGB}{221,204,119}

\path[fill=fillColor] (144.86,270.36) rectangle (185.29,273.09);

\path[fill=fillColor] (234.70,270.36) rectangle (275.13,270.36);

\path[fill=fillColor] (189.78,270.36) rectangle (230.21,271.86);

\path[fill=fillColor] (279.62,270.36) rectangle (320.05,275.46);

\path[fill=fillColor] ( 55.02,270.36) rectangle ( 95.45,348.73);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 48.28,266.45) rectangle (371.71,352.65);
\end{scope}
\begin{scope}
\path[clip] ( 48.28,174.74) rectangle (371.71,260.95);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 48.28,174.74) rectangle (371.71,260.95);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 48.28,189.63) --
	(371.71,189.63);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 48.28,211.57) --
	(371.71,211.57);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 48.28,233.51) --
	(371.71,233.51);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 48.28,255.45) --
	(371.71,255.45);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 48.28,178.66) --
	(371.71,178.66);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 48.28,200.60) --
	(371.71,200.60);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 48.28,222.54) --
	(371.71,222.54);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 48.28,244.48) --
	(371.71,244.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 75.24,174.74) --
	( 75.24,260.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (120.16,174.74) --
	(120.16,260.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (165.08,174.74) --
	(165.08,260.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (209.99,174.74) --
	(209.99,260.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (254.91,174.74) --
	(254.91,260.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (299.83,174.74) --
	(299.83,260.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (344.75,174.74) --
	(344.75,260.95);
\definecolor{fillColor}{RGB}{221,204,119}

\path[fill=fillColor] (144.86,178.66) rectangle (185.29,193.40);

\path[fill=fillColor] (234.70,178.66) rectangle (275.13,179.97);

\path[fill=fillColor] (189.78,178.66) rectangle (230.21,183.31);

\path[fill=fillColor] (279.62,178.66) rectangle (320.05,187.96);

\path[fill=fillColor] (324.54,178.66) rectangle (364.97,180.76);

\path[fill=fillColor] ( 55.02,178.66) rectangle ( 95.45,220.87);

\path[fill=fillColor] ( 99.94,178.66) rectangle (140.37,195.60);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 48.28,174.74) rectangle (371.71,260.95);
\end{scope}
\begin{scope}
\path[clip] ( 48.28, 83.03) rectangle (371.71,169.24);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 48.28, 83.03) rectangle (371.71,169.24);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 48.28, 97.92) --
	(371.71, 97.92);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 48.28,119.86) --
	(371.71,119.86);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 48.28,141.80) --
	(371.71,141.80);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 48.28,163.74) --
	(371.71,163.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 48.28, 86.95) --
	(371.71, 86.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 48.28,108.89) --
	(371.71,108.89);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 48.28,130.83) --
	(371.71,130.83);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 48.28,152.77) --
	(371.71,152.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 75.24, 83.03) --
	( 75.24,169.24);

\path[draw=drawColor,line width= 0.6pt,line join=round] (120.16, 83.03) --
	(120.16,169.24);

\path[draw=drawColor,line width= 0.6pt,line join=round] (165.08, 83.03) --
	(165.08,169.24);

\path[draw=drawColor,line width= 0.6pt,line join=round] (209.99, 83.03) --
	(209.99,169.24);

\path[draw=drawColor,line width= 0.6pt,line join=round] (254.91, 83.03) --
	(254.91,169.24);

\path[draw=drawColor,line width= 0.6pt,line join=round] (299.83, 83.03) --
	(299.83,169.24);

\path[draw=drawColor,line width= 0.6pt,line join=round] (344.75, 83.03) --
	(344.75,169.24);
\definecolor{fillColor}{RGB}{221,204,119}

\path[fill=fillColor] (144.86, 86.95) rectangle (185.29, 97.04);

\path[fill=fillColor] (189.78, 86.95) rectangle (230.21, 89.06);

\path[fill=fillColor] (279.62, 86.95) rectangle (320.05,106.17);

\path[fill=fillColor] (324.54, 86.95) rectangle (364.97, 87.04);

\path[fill=fillColor] ( 55.02, 86.95) rectangle ( 95.45,117.23);

\path[fill=fillColor] ( 99.94, 86.95) rectangle (140.37,113.02);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 48.28, 83.03) rectangle (371.71,169.24);
\end{scope}
\begin{scope}
\path[clip] (393.78,541.57) rectangle (717.20,627.77);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (393.78,541.57) rectangle (717.20,627.77);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] (393.78,556.45) --
	(717.20,556.45);

\path[draw=drawColor,line width= 0.3pt,line join=round] (393.78,578.39) --
	(717.20,578.39);

\path[draw=drawColor,line width= 0.3pt,line join=round] (393.78,600.33) --
	(717.20,600.33);

\path[draw=drawColor,line width= 0.3pt,line join=round] (393.78,622.27) --
	(717.20,622.27);

\path[draw=drawColor,line width= 0.6pt,line join=round] (393.78,545.48) --
	(717.20,545.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] (393.78,567.42) --
	(717.20,567.42);

\path[draw=drawColor,line width= 0.6pt,line join=round] (393.78,589.36) --
	(717.20,589.36);

\path[draw=drawColor,line width= 0.6pt,line join=round] (393.78,611.30) --
	(717.20,611.30);

\path[draw=drawColor,line width= 0.6pt,line join=round] (420.73,541.57) --
	(420.73,627.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (465.65,541.57) --
	(465.65,627.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (510.57,541.57) --
	(510.57,627.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (555.49,541.57) --
	(555.49,627.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (600.41,541.57) --
	(600.41,627.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (645.33,541.57) --
	(645.33,627.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (690.25,541.57) --
	(690.25,627.77);
\definecolor{fillColor}{RGB}{221,204,119}

\path[fill=fillColor] (490.36,545.48) rectangle (530.78,577.60);

\path[fill=fillColor] (580.19,545.48) rectangle (620.62,546.98);

\path[fill=fillColor] (535.27,545.48) rectangle (575.70,547.06);

\path[fill=fillColor] (670.03,545.48) rectangle (710.46,553.30);

\path[fill=fillColor] (400.52,545.48) rectangle (440.94,579.97);

\path[fill=fillColor] (445.44,545.48) rectangle (485.86,555.66);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (393.78,541.57) rectangle (717.20,627.77);
\end{scope}
\begin{scope}
\path[clip] (393.78,449.86) rectangle (717.20,536.07);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (393.78,449.86) rectangle (717.20,536.07);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] (393.78,464.75) --
	(717.20,464.75);

\path[draw=drawColor,line width= 0.3pt,line join=round] (393.78,486.69) --
	(717.20,486.69);

\path[draw=drawColor,line width= 0.3pt,line join=round] (393.78,508.63) --
	(717.20,508.63);

\path[draw=drawColor,line width= 0.3pt,line join=round] (393.78,530.57) --
	(717.20,530.57);

\path[draw=drawColor,line width= 0.6pt,line join=round] (393.78,453.78) --
	(717.20,453.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (393.78,475.72) --
	(717.20,475.72);

\path[draw=drawColor,line width= 0.6pt,line join=round] (393.78,497.66) --
	(717.20,497.66);

\path[draw=drawColor,line width= 0.6pt,line join=round] (393.78,519.60) --
	(717.20,519.60);

\path[draw=drawColor,line width= 0.6pt,line join=round] (420.73,449.86) --
	(420.73,536.07);

\path[draw=drawColor,line width= 0.6pt,line join=round] (465.65,449.86) --
	(465.65,536.07);

\path[draw=drawColor,line width= 0.6pt,line join=round] (510.57,449.86) --
	(510.57,536.07);

\path[draw=drawColor,line width= 0.6pt,line join=round] (555.49,449.86) --
	(555.49,536.07);

\path[draw=drawColor,line width= 0.6pt,line join=round] (600.41,449.86) --
	(600.41,536.07);

\path[draw=drawColor,line width= 0.6pt,line join=round] (645.33,449.86) --
	(645.33,536.07);

\path[draw=drawColor,line width= 0.6pt,line join=round] (690.25,449.86) --
	(690.25,536.07);
\definecolor{fillColor}{RGB}{221,204,119}

\path[fill=fillColor] (490.36,453.78) rectangle (530.78,460.80);

\path[fill=fillColor] (535.27,453.78) rectangle (575.70,460.80);

\path[fill=fillColor] (625.11,453.78) rectangle (665.54,457.20);

\path[fill=fillColor] (670.03,453.78) rectangle (710.46,457.46);

\path[fill=fillColor] (400.52,453.78) rectangle (440.94,504.68);

\path[fill=fillColor] (445.44,453.78) rectangle (485.86,469.57);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (393.78,449.86) rectangle (717.20,536.07);
\end{scope}
\begin{scope}
\path[clip] (393.78,358.15) rectangle (717.20,444.36);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (393.78,358.15) rectangle (717.20,444.36);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] (393.78,373.04) --
	(717.20,373.04);

\path[draw=drawColor,line width= 0.3pt,line join=round] (393.78,394.98) --
	(717.20,394.98);

\path[draw=drawColor,line width= 0.3pt,line join=round] (393.78,416.92) --
	(717.20,416.92);

\path[draw=drawColor,line width= 0.3pt,line join=round] (393.78,438.86) --
	(717.20,438.86);

\path[draw=drawColor,line width= 0.6pt,line join=round] (393.78,362.07) --
	(717.20,362.07);

\path[draw=drawColor,line width= 0.6pt,line join=round] (393.78,384.01) --
	(717.20,384.01);

\path[draw=drawColor,line width= 0.6pt,line join=round] (393.78,405.95) --
	(717.20,405.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (393.78,427.89) --
	(717.20,427.89);

\path[draw=drawColor,line width= 0.6pt,line join=round] (420.73,358.15) --
	(420.73,444.36);

\path[draw=drawColor,line width= 0.6pt,line join=round] (465.65,358.15) --
	(465.65,444.36);

\path[draw=drawColor,line width= 0.6pt,line join=round] (510.57,358.15) --
	(510.57,444.36);

\path[draw=drawColor,line width= 0.6pt,line join=round] (555.49,358.15) --
	(555.49,444.36);

\path[draw=drawColor,line width= 0.6pt,line join=round] (600.41,358.15) --
	(600.41,444.36);

\path[draw=drawColor,line width= 0.6pt,line join=round] (645.33,358.15) --
	(645.33,444.36);

\path[draw=drawColor,line width= 0.6pt,line join=round] (690.25,358.15) --
	(690.25,444.36);
\definecolor{fillColor}{RGB}{221,204,119}

\path[fill=fillColor] (490.36,362.07) rectangle (530.78,369.44);

\path[fill=fillColor] (580.19,362.07) rectangle (620.62,362.51);

\path[fill=fillColor] (535.27,362.07) rectangle (575.70,366.20);

\path[fill=fillColor] (625.11,362.07) rectangle (665.54,369.88);

\path[fill=fillColor] (670.03,362.07) rectangle (710.46,363.04);

\path[fill=fillColor] (400.52,362.07) rectangle (440.94,400.69);

\path[fill=fillColor] (445.44,362.07) rectangle (485.86,390.59);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (393.78,358.15) rectangle (717.20,444.36);
\end{scope}
\begin{scope}
\path[clip] (393.78,266.45) rectangle (717.20,352.65);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (393.78,266.45) rectangle (717.20,352.65);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] (393.78,281.33) --
	(717.20,281.33);

\path[draw=drawColor,line width= 0.3pt,line join=round] (393.78,303.27) --
	(717.20,303.27);

\path[draw=drawColor,line width= 0.3pt,line join=round] (393.78,325.21) --
	(717.20,325.21);

\path[draw=drawColor,line width= 0.3pt,line join=round] (393.78,347.15) --
	(717.20,347.15);

\path[draw=drawColor,line width= 0.6pt,line join=round] (393.78,270.36) --
	(717.20,270.36);

\path[draw=drawColor,line width= 0.6pt,line join=round] (393.78,292.30) --
	(717.20,292.30);

\path[draw=drawColor,line width= 0.6pt,line join=round] (393.78,314.24) --
	(717.20,314.24);

\path[draw=drawColor,line width= 0.6pt,line join=round] (393.78,336.18) --
	(717.20,336.18);

\path[draw=drawColor,line width= 0.6pt,line join=round] (420.73,266.45) --
	(420.73,352.65);

\path[draw=drawColor,line width= 0.6pt,line join=round] (465.65,266.45) --
	(465.65,352.65);

\path[draw=drawColor,line width= 0.6pt,line join=round] (510.57,266.45) --
	(510.57,352.65);

\path[draw=drawColor,line width= 0.6pt,line join=round] (555.49,266.45) --
	(555.49,352.65);

\path[draw=drawColor,line width= 0.6pt,line join=round] (600.41,266.45) --
	(600.41,352.65);

\path[draw=drawColor,line width= 0.6pt,line join=round] (645.33,266.45) --
	(645.33,352.65);

\path[draw=drawColor,line width= 0.6pt,line join=round] (690.25,266.45) --
	(690.25,352.65);
\definecolor{fillColor}{RGB}{221,204,119}

\path[fill=fillColor] (490.36,270.36) rectangle (530.78,278.70);

\path[fill=fillColor] (580.19,270.36) rectangle (620.62,275.28);

\path[fill=fillColor] (535.27,270.36) rectangle (575.70,276.68);

\path[fill=fillColor] (625.11,270.36) rectangle (665.54,279.32);

\path[fill=fillColor] (670.03,270.36) rectangle (710.46,270.45);

\path[fill=fillColor] (400.52,270.36) rectangle (440.94,301.08);

\path[fill=fillColor] (445.44,270.36) rectangle (485.86,298.80);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (393.78,266.45) rectangle (717.20,352.65);
\end{scope}
\begin{scope}
\path[clip] (393.78,174.74) rectangle (717.20,260.95);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (393.78,174.74) rectangle (717.20,260.95);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] (393.78,189.63) --
	(717.20,189.63);

\path[draw=drawColor,line width= 0.3pt,line join=round] (393.78,211.57) --
	(717.20,211.57);

\path[draw=drawColor,line width= 0.3pt,line join=round] (393.78,233.51) --
	(717.20,233.51);

\path[draw=drawColor,line width= 0.3pt,line join=round] (393.78,255.45) --
	(717.20,255.45);

\path[draw=drawColor,line width= 0.6pt,line join=round] (393.78,178.66) --
	(717.20,178.66);

\path[draw=drawColor,line width= 0.6pt,line join=round] (393.78,200.60) --
	(717.20,200.60);

\path[draw=drawColor,line width= 0.6pt,line join=round] (393.78,222.54) --
	(717.20,222.54);

\path[draw=drawColor,line width= 0.6pt,line join=round] (393.78,244.48) --
	(717.20,244.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] (420.73,174.74) --
	(420.73,260.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (465.65,174.74) --
	(465.65,260.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (510.57,174.74) --
	(510.57,260.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (555.49,174.74) --
	(555.49,260.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (600.41,174.74) --
	(600.41,260.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (645.33,174.74) --
	(645.33,260.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (690.25,174.74) --
	(690.25,260.95);
\definecolor{fillColor}{RGB}{221,204,119}

\path[fill=fillColor] (490.36,178.66) rectangle (530.78,203.32);

\path[fill=fillColor] (580.19,178.66) rectangle (620.62,182.96);

\path[fill=fillColor] (535.27,178.66) rectangle (575.70,180.59);

\path[fill=fillColor] (625.11,178.66) rectangle (665.54,193.23);

\path[fill=fillColor] (400.52,178.66) rectangle (440.94,210.25);

\path[fill=fillColor] (445.44,178.66) rectangle (485.86,189.45);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (393.78,174.74) rectangle (717.20,260.95);
\end{scope}
\begin{scope}
\path[clip] (377.21,541.57) rectangle (393.78,627.77);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (377.21,541.57) rectangle (393.78,627.77);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (388.52,584.67) {Colombia};
\end{scope}
\begin{scope}
\path[clip] (377.21,449.86) rectangle (393.78,536.07);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (377.21,449.86) rectangle (393.78,536.07);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (388.52,492.96) {Nigeria};
\end{scope}
\begin{scope}
\path[clip] (377.21,358.15) rectangle (393.78,444.36);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (377.21,358.15) rectangle (393.78,444.36);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (388.52,401.26) {Sierra Leone 1};
\end{scope}
\begin{scope}
\path[clip] (377.21,266.45) rectangle (393.78,352.65);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (377.21,266.45) rectangle (393.78,352.65);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (388.52,309.55) {Uganda 2};
\end{scope}
\begin{scope}
\path[clip] (377.21,174.74) rectangle (393.78,260.95);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (377.21,174.74) rectangle (393.78,260.95);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (388.52,217.84) {Russia};
\end{scope}
\begin{scope}
\path[clip] ( 31.71,541.57) rectangle ( 48.28,627.77);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 31.71,541.57) rectangle ( 48.28,627.77);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 43.03,584.67) {Burkina Faso};
\end{scope}
\begin{scope}
\path[clip] ( 31.71,449.86) rectangle ( 48.28,536.07);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 31.71,449.86) rectangle ( 48.28,536.07);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 43.03,492.96) {Nepal};
\end{scope}
\begin{scope}
\path[clip] ( 31.71,358.15) rectangle ( 48.28,444.36);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 31.71,358.15) rectangle ( 48.28,444.36);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 43.03,401.26) {Rwanda};
\end{scope}
\begin{scope}
\path[clip] ( 31.71,266.45) rectangle ( 48.28,352.65);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 31.71,266.45) rectangle ( 48.28,352.65);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 43.03,309.55) {Sierra Leone 2};
\end{scope}
\begin{scope}
\path[clip] ( 31.71,174.74) rectangle ( 48.28,260.95);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 31.71,174.74) rectangle ( 48.28,260.95);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 43.03,217.84) {All LMICs};
\end{scope}
\begin{scope}
\path[clip] ( 31.71, 83.03) rectangle ( 48.28,169.24);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 31.71, 83.03) rectangle ( 48.28,169.24);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 43.03,126.14) {USA};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 75.24, 80.28) --
	( 75.24, 83.03);

\path[draw=drawColor,line width= 0.6pt,line join=round] (120.16, 80.28) --
	(120.16, 83.03);

\path[draw=drawColor,line width= 0.6pt,line join=round] (165.08, 80.28) --
	(165.08, 83.03);

\path[draw=drawColor,line width= 0.6pt,line join=round] (209.99, 80.28) --
	(209.99, 83.03);

\path[draw=drawColor,line width= 0.6pt,line join=round] (254.91, 80.28) --
	(254.91, 83.03);

\path[draw=drawColor,line width= 0.6pt,line join=round] (299.83, 80.28) --
	(299.83, 83.03);

\path[draw=drawColor,line width= 0.6pt,line join=round] (344.75, 80.28) --
	(344.75, 83.03);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 78.27, 78.08) {Health workers};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (118.43, 78.08) {Government or};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (127.94, 78.08) {MoH};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (163.35, 78.08) {Family or};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (172.86, 78.08) {Friends};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (198.77, 78.08) {Famous person,};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (208.27, 78.08) {religious leader};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (217.78, 78.08) {or traditional};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (227.28, 78.08) {healers};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (248.44, 78.08) {Newspapers,};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (257.95, 78.08) {radio or online};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (267.45, 78.08) {groups};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (302.86, 78.08) {Other};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (343.03, 78.08) {Don't know or};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (352.54, 78.08) {Refuse};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (420.73,171.99) --
	(420.73,174.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] (465.65,171.99) --
	(465.65,174.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] (510.57,171.99) --
	(510.57,174.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] (555.49,171.99) --
	(555.49,174.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] (600.41,171.99) --
	(600.41,174.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] (645.33,171.99) --
	(645.33,174.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] (690.25,171.99) --
	(690.25,174.74);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (423.76,169.79) {Health workers};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (463.93,169.79) {Government or};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (473.43,169.79) {MoH};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (508.85,169.79) {Family or};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (518.35,169.79) {Friends};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (544.26,169.79) {Famous person,};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (553.77,169.79) {religious leader};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (563.27,169.79) {or traditional};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (572.78,169.79) {healers};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (593.93,169.79) {Newspapers,};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (603.44,169.79) {radio or online};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (612.94,169.79) {groups};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (648.36,169.79) {Other};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (688.53,169.79) {Don't know or};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (698.03,169.79) {Refuse};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,542.45) {0};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,564.39) {25};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,586.33) {50};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,608.27) {75};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96,545.48) --
	( 31.71,545.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96,567.42) --
	( 31.71,567.42);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96,589.36) --
	( 31.71,589.36);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96,611.30) --
	( 31.71,611.30);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,450.75) {0};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,472.69) {25};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,494.63) {50};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,516.57) {75};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96,453.78) --
	( 31.71,453.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96,475.72) --
	( 31.71,475.72);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96,497.66) --
	( 31.71,497.66);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96,519.60) --
	( 31.71,519.60);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,359.04) {0};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,380.98) {25};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,402.92) {50};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,424.86) {75};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96,362.07) --
	( 31.71,362.07);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96,384.01) --
	( 31.71,384.01);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96,405.95) --
	( 31.71,405.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96,427.89) --
	( 31.71,427.89);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,267.33) {0};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,289.27) {25};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,311.21) {50};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,333.15) {75};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96,270.36) --
	( 31.71,270.36);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96,292.30) --
	( 31.71,292.30);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96,314.24) --
	( 31.71,314.24);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96,336.18) --
	( 31.71,336.18);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,175.63) {0};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,197.57) {25};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,219.51) {50};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,241.45) {75};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96,178.66) --
	( 31.71,178.66);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96,200.60) --
	( 31.71,200.60);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96,222.54) --
	( 31.71,222.54);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96,244.48) --
	( 31.71,244.48);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96, 83.92) {0};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,105.86) {25};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,127.80) {50};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 17.96,149.74) {75};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96, 86.95) --
	( 31.71, 86.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96,108.89) --
	( 31.71,108.89);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96,130.83) --
	( 31.71,130.83);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.96,152.77) --
	( 31.71,152.77);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 13.08,355.40) {estimate};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.32] at (  5.50,635.84) {Which of the following people would you trust MOST to help you decide whether you would get a COVID-19 vaccine?};
\end{scope}
\end{tikzpicture}

\end{document}