### 2021-06-22 paper_files/figure-latex/fig2paper-1.tex

% Created by tikzDevice version 0.12.3.1 on 2021-06-22 11:48:16
% !TEX encoding = UTF-8 Unicode
% !TEX program = xelatex

\documentclass[fontsize=7pt,DIV=calc]{scrartcl}

\nonstopmode

\usepackage{tikz}

\usepackage[active,tightpage,xetex]{preview}

\usepackage{fontspec,xunicode}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}

\setmainfont{Arial}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (722.70,542.02);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.00) rectangle (722.70,542.03);
\end{scope}
\begin{scope}
\path[clip] ( 64.26, 25.44) rectangle (124.60,470.61);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 64.26, 25.44) rectangle (124.60,470.61);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 74.80, 25.44) --
	( 74.80,470.61);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 88.25, 25.44) --
	( 88.25,470.61);

\path[draw=drawColor,line width= 0.3pt,line join=round] (101.69, 25.44) --
	(101.69,470.61);

\path[draw=drawColor,line width= 0.3pt,line join=round] (115.14, 25.44) --
	(115.14,470.61);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 64.26, 44.25) --
	(124.60, 44.25);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 64.26, 75.60) --
	(124.60, 75.60);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 64.26,106.95) --
	(124.60,106.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 64.26,138.30) --
	(124.60,138.30);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 64.26,169.65) --
	(124.60,169.65);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 64.26,201.00) --
	(124.60,201.00);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 64.26,232.35) --
	(124.60,232.35);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 64.26,263.70) --
	(124.60,263.70);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 64.26,295.05) --
	(124.60,295.05);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 64.26,326.40) --
	(124.60,326.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 64.26,357.75) --
	(124.60,357.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 64.26,389.10) --
	(124.60,389.10);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 64.26,420.45) --
	(124.60,420.45);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 64.26,451.80) --
	(124.60,451.80);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 68.08, 25.44) --
	( 68.08,470.61);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 81.52, 25.44) --
	( 81.52,470.61);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 94.97, 25.44) --
	( 94.97,470.61);

\path[draw=drawColor,line width= 0.6pt,line join=round] (108.41, 25.44) --
	(108.41,470.61);

\path[draw=drawColor,line width= 0.6pt,line join=round] (121.86, 25.44) --
	(121.86,470.61);
\definecolor{drawColor}{RGB}{204,102,119}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 98.36,103.81) --
	( 98.36,110.08);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 98.36,106.95) --
	( 81.68,106.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 81.68,103.81) --
	( 81.68,110.08);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 92.98,448.66) --
	( 92.98,454.93);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 92.98,451.80) --
	( 87.17,451.80);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 87.17,448.66) --
	( 87.17,454.93);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 88.30,417.31) --
	( 88.30,423.58);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 88.30,420.45) --
	( 81.20,420.45);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 81.20,417.31) --
	( 81.20,423.58);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 77.87,354.61) --
	( 77.87,360.88);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 77.87,357.75) --
	( 68.24,357.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 68.24,354.61) --
	( 68.24,360.88);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 81.79,323.26) --
	( 81.79,329.53);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 81.79,326.40) --
	( 77.49,326.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 77.49,323.26) --
	( 77.49,329.53);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 84.64,291.91) --
	( 84.64,298.18);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 84.64,295.05) --
	( 76.20,295.05);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 76.20,291.91) --
	( 76.20,298.18);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 88.73, 72.46) --
	( 88.73, 78.73);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 88.73, 75.60) --
	( 87.01, 75.60);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 87.01, 72.46) --
	( 87.01, 78.73);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 95.13,260.56) --
	( 95.13,266.83);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 95.13,263.70) --
	( 82.54,263.70);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 82.54,260.56) --
	( 82.54,266.83);

\path[draw=drawColor,line width= 0.6pt,line join=round] (100.29,229.21) --
	(100.29,235.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] (100.29,232.35) --
	( 93.41,232.35);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 93.41,229.21) --
	( 93.41,235.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] (103.41,197.86) --
	(103.41,204.13);

\path[draw=drawColor,line width= 0.6pt,line join=round] (103.41,201.00) --
	( 95.02,201.00);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 95.02,197.86) --
	( 95.02,204.13);

\path[draw=drawColor,line width= 0.6pt,line join=round] (116.27,166.51) --
	(116.27,172.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (116.27,169.65) --
	(111.48,169.65);

\path[draw=drawColor,line width= 0.6pt,line join=round] (111.48,166.51) --
	(111.48,172.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 96.31,135.16) --
	( 96.31,141.43);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 96.31,138.30) --
	( 90.77,138.30);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 90.77,135.16) --
	( 90.77,141.43);

\path[draw=drawColor,line width= 0.6pt,line join=round] (113.25, 41.12) --
	(113.25, 47.39);

\path[draw=drawColor,line width= 0.6pt,line join=round] (113.25, 44.25) --
	(108.20, 44.25);

\path[draw=drawColor,line width= 0.6pt,line join=round] (108.20, 41.12) --
	(108.20, 47.39);
\definecolor{fillColor}{RGB}{204,102,119}

\path[fill=fillColor] ( 90.02,106.95) circle (  1.96);

\path[fill=fillColor] ( 90.07,451.80) circle (  1.96);

\path[fill=fillColor] ( 84.75,420.45) circle (  1.96);

\path[fill=fillColor] ( 73.08,357.75) circle (  1.96);

\path[fill=fillColor] ( 79.64,326.40) circle (  1.96);

\path[fill=fillColor] ( 80.45,295.05) circle (  1.96);

\path[fill=fillColor] ( 87.87, 75.60) circle (  1.96);

\path[fill=fillColor] ( 88.84,263.70) circle (  1.96);

\path[fill=fillColor] ( 96.85,232.35) circle (  1.96);

\path[fill=fillColor] ( 99.22,201.00) circle (  1.96);

\path[fill=fillColor] (113.85,169.65) circle (  1.96);

\path[fill=fillColor] ( 93.52,138.30) circle (  1.96);

\path[fill=fillColor] (110.73, 44.25) circle (  1.96);
\definecolor{drawColor}{RGB}{169,169,169}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 64.26,122.62) -- (124.60,122.62);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 64.26, 91.27) -- (124.60, 91.27);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 64.26, 25.44) rectangle (124.60,470.61);
\end{scope}
\begin{scope}
\path[clip] (130.10, 25.44) rectangle (190.45,470.61);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (130.10, 25.44) rectangle (190.45,470.61);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] (140.64, 25.44) --
	(140.64,470.61);

\path[draw=drawColor,line width= 0.3pt,line join=round] (154.09, 25.44) --
	(154.09,470.61);

\path[draw=drawColor,line width= 0.3pt,line join=round] (167.54, 25.44) --
	(167.54,470.61);

\path[draw=drawColor,line width= 0.3pt,line join=round] (180.98, 25.44) --
	(180.98,470.61);

\path[draw=drawColor,line width= 0.6pt,line join=round] (130.10, 44.25) --
	(190.45, 44.25);

\path[draw=drawColor,line width= 0.6pt,line join=round] (130.10, 75.60) --
	(190.45, 75.60);

\path[draw=drawColor,line width= 0.6pt,line join=round] (130.10,106.95) --
	(190.45,106.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (130.10,138.30) --
	(190.45,138.30);

\path[draw=drawColor,line width= 0.6pt,line join=round] (130.10,169.65) --
	(190.45,169.65);

\path[draw=drawColor,line width= 0.6pt,line join=round] (130.10,201.00) --
	(190.45,201.00);

\path[draw=drawColor,line width= 0.6pt,line join=round] (130.10,232.35) --
	(190.45,232.35);

\path[draw=drawColor,line width= 0.6pt,line join=round] (130.10,263.70) --
	(190.45,263.70);

\path[draw=drawColor,line width= 0.6pt,line join=round] (130.10,295.05) --
	(190.45,295.05);

\path[draw=drawColor,line width= 0.6pt,line join=round] (130.10,326.40) --
	(190.45,326.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] (130.10,357.75) --
	(190.45,357.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (130.10,389.10) --
	(190.45,389.10);

\path[draw=drawColor,line width= 0.6pt,line join=round] (130.10,420.45) --
	(190.45,420.45);

\path[draw=drawColor,line width= 0.6pt,line join=round] (130.10,451.80) --
	(190.45,451.80);

\path[draw=drawColor,line width= 0.6pt,line join=round] (133.92, 25.44) --
	(133.92,470.61);

\path[draw=drawColor,line width= 0.6pt,line join=round] (147.37, 25.44) --
	(147.37,470.61);

\path[draw=drawColor,line width= 0.6pt,line join=round] (160.81, 25.44) --
	(160.81,470.61);

\path[draw=drawColor,line width= 0.6pt,line join=round] (174.26, 25.44) --
	(174.26,470.61);

\path[draw=drawColor,line width= 0.6pt,line join=round] (187.70, 25.44) --
	(187.70,470.61);
\definecolor{drawColor}{RGB}{221,204,119}

\path[draw=drawColor,line width= 0.6pt,line join=round] (148.17,103.81) --
	(148.17,110.08);

\path[draw=drawColor,line width= 0.6pt,line join=round] (148.17,106.95) --
	(138.60,106.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (138.60,103.81) --
	(138.60,110.08);

\path[draw=drawColor,line width= 0.6pt,line join=round] (139.84,448.66) --
	(139.84,454.93);

\path[draw=drawColor,line width= 0.6pt,line join=round] (139.84,451.80) --
	(136.61,451.80);

\path[draw=drawColor,line width= 0.6pt,line join=round] (136.61,448.66) --
	(136.61,454.93);

\path[draw=drawColor,line width= 0.6pt,line join=round] (146.51,417.31) --
	(146.51,423.58);

\path[draw=drawColor,line width= 0.6pt,line join=round] (146.51,420.45) --
	(140.75,420.45);

\path[draw=drawColor,line width= 0.6pt,line join=round] (140.75,417.31) --
	(140.75,423.58);

\path[draw=drawColor,line width= 0.6pt,line join=round] (142.69,354.61) --
	(142.69,360.88);

\path[draw=drawColor,line width= 0.6pt,line join=round] (142.69,357.75) --
	(133.71,357.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (133.71,354.61) --
	(133.71,360.88);

\path[draw=drawColor,line width= 0.6pt,line join=round] (150.27,323.26) --
	(150.27,329.53);

\path[draw=drawColor,line width= 0.6pt,line join=round] (150.27,326.40) --
	(145.65,326.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] (145.65,323.26) --
	(145.65,329.53);

\path[draw=drawColor,line width= 0.6pt,line join=round] (149.73,291.91) --
	(149.73,298.18);

\path[draw=drawColor,line width= 0.6pt,line join=round] (149.73,295.05) --
	(141.61,295.05);

\path[draw=drawColor,line width= 0.6pt,line join=round] (141.61,291.91) --
	(141.61,298.18);

\path[draw=drawColor,line width= 0.6pt,line join=round] (142.04, 72.46) --
	(142.04, 78.73);

\path[draw=drawColor,line width= 0.6pt,line join=round] (142.04, 75.60) --
	(140.81, 75.60);

\path[draw=drawColor,line width= 0.6pt,line join=round] (140.81, 72.46) --
	(140.81, 78.73);

\path[draw=drawColor,line width= 0.6pt,line join=round] (143.23,260.56) --
	(143.23,266.83);

\path[draw=drawColor,line width= 0.6pt,line join=round] (143.23,263.70) --
	(135.43,263.70);

\path[draw=drawColor,line width= 0.6pt,line join=round] (135.43,260.56) --
	(135.43,266.83);

\path[draw=drawColor,line width= 0.6pt,line join=round] (157.69,229.21) --
	(157.69,235.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] (157.69,232.35) --
	(150.92,232.35);

\path[draw=drawColor,line width= 0.6pt,line join=round] (150.92,229.21) --
	(150.92,235.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] (141.56,135.16) --
	(141.56,141.43);

\path[draw=drawColor,line width= 0.6pt,line join=round] (141.56,138.30) --
	(137.74,138.30);

\path[draw=drawColor,line width= 0.6pt,line join=round] (137.74,135.16) --
	(137.74,141.43);
\definecolor{fillColor}{RGB}{221,204,119}

\path[fill=fillColor] (143.39,106.95) circle (  1.96);

\path[fill=fillColor] (138.22,451.80) circle (  1.96);

\path[fill=fillColor] (143.66,420.45) circle (  1.96);

\path[fill=fillColor] (138.17,357.75) circle (  1.96);

\path[fill=fillColor] (147.96,326.40) circle (  1.96);

\path[fill=fillColor] (145.70,295.05) circle (  1.96);

\path[fill=fillColor] (141.40, 75.60) circle (  1.96);

\path[fill=fillColor] (139.35,263.70) circle (  1.96);

\path[fill=fillColor] (154.30,232.35) circle (  1.96);

\path[fill=fillColor] (139.68,138.30) circle (  1.96);
\definecolor{drawColor}{RGB}{169,169,169}

\path[draw=drawColor,line width= 0.6pt,line join=round] (130.10,122.62) -- (190.45,122.62);

\path[draw=drawColor,line width= 0.6pt,line join=round] (130.10, 91.27) -- (190.45, 91.27);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (130.10, 25.44) rectangle (190.45,470.61);
\end{scope}
\begin{scope}
\path[clip] (195.95, 25.44) rectangle (256.29,470.61);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (195.95, 25.44) rectangle (256.29,470.61);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] (206.49, 25.44) --
	(206.49,470.61);

\path[draw=drawColor,line width= 0.3pt,line join=round] (219.93, 25.44) --
	(219.93,470.61);

\path[draw=drawColor,line width= 0.3pt,line join=round] (233.38, 25.44) --
	(233.38,470.61);

\path[draw=drawColor,line width= 0.3pt,line join=round] (246.83, 25.44) --
	(246.83,470.61);

\path[draw=drawColor,line width= 0.6pt,line join=round] (195.95, 44.25) --
	(256.29, 44.25);

\path[draw=drawColor,line width= 0.6pt,line join=round] (195.95, 75.60) --
	(256.29, 75.60);

\path[draw=drawColor,line width= 0.6pt,line join=round] (195.95,106.95) --
	(256.29,106.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (195.95,138.30) --
	(256.29,138.30);

\path[draw=drawColor,line width= 0.6pt,line join=round] (195.95,169.65) --
	(256.29,169.65);

\path[draw=drawColor,line width= 0.6pt,line join=round] (195.95,201.00) --
	(256.29,201.00);

\path[draw=drawColor,line width= 0.6pt,line join=round] (195.95,232.35) --
	(256.29,232.35);

\path[draw=drawColor,line width= 0.6pt,line join=round] (195.95,263.70) --
	(256.29,263.70);

\path[draw=drawColor,line width= 0.6pt,line join=round] (195.95,295.05) --
	(256.29,295.05);

\path[draw=drawColor,line width= 0.6pt,line join=round] (195.95,326.40) --
	(256.29,326.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] (195.95,357.75) --
	(256.29,357.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (195.95,389.10) --
	(256.29,389.10);

\path[draw=drawColor,line width= 0.6pt,line join=round] (195.95,420.45) --
	(256.29,420.45);

\path[draw=drawColor,line width= 0.6pt,line join=round] (195.95,451.80) --
	(256.29,451.80);

\path[draw=drawColor,line width= 0.6pt,line join=round] (199.77, 25.44) --
	(199.77,470.61);

\path[draw=drawColor,line width= 0.6pt,line join=round] (213.21, 25.44) --
	(213.21,470.61);

\path[draw=drawColor,line width= 0.6pt,line join=round] (226.66, 25.44) --
	(226.66,470.61);

\path[draw=drawColor,line width= 0.6pt,line join=round] (240.10, 25.44) --
	(240.10,470.61);

\path[draw=drawColor,line width= 0.6pt,line join=round] (253.55, 25.44) --
	(253.55,470.61);
\definecolor{drawColor}{RGB}{17,119,51}

\path[draw=drawColor,line width= 0.6pt,line join=round] (209.88,103.81) --
	(209.88,110.08);

\path[draw=drawColor,line width= 0.6pt,line join=round] (209.88,106.95) --
	(203.21,106.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (203.21,103.81) --
	(203.21,110.08);

\path[draw=drawColor,line width= 0.6pt,line join=round] (205.25,448.66) --
	(205.25,454.93);

\path[draw=drawColor,line width= 0.6pt,line join=round] (205.25,451.80) --
	(202.19,451.80);

\path[draw=drawColor,line width= 0.6pt,line join=round] (202.19,448.66) --
	(202.19,454.93);

\path[draw=drawColor,line width= 0.6pt,line join=round] (206.22,417.31) --
	(206.22,423.58);

\path[draw=drawColor,line width= 0.6pt,line join=round] (206.22,420.45) --
	(201.86,420.45);

\path[draw=drawColor,line width= 0.6pt,line join=round] (201.86,417.31) --
	(201.86,423.58);

\path[draw=drawColor,line width= 0.6pt,line join=round] (203.05,385.96) --
	(203.05,392.23);

\path[draw=drawColor,line width= 0.6pt,line join=round] (203.05,389.10) --
	(199.39,389.10);

\path[draw=drawColor,line width= 0.6pt,line join=round] (199.39,385.96) --
	(199.39,392.23);

\path[draw=drawColor,line width= 0.6pt,line join=round] (218.11,354.61) --
	(218.11,360.88);

\path[draw=drawColor,line width= 0.6pt,line join=round] (218.11,357.75) --
	(203.37,357.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (203.37,354.61) --
	(203.37,360.88);

\path[draw=drawColor,line width= 0.6pt,line join=round] (210.20,323.26) --
	(210.20,329.53);

\path[draw=drawColor,line width= 0.6pt,line join=round] (210.20,326.40) --
	(206.38,326.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] (206.38,323.26) --
	(206.38,329.53);

\path[draw=drawColor,line width= 0.6pt,line join=round] (220.15,291.91) --
	(220.15,298.18);

\path[draw=drawColor,line width= 0.6pt,line join=round] (220.15,295.05) --
	(211.01,295.05);

\path[draw=drawColor,line width= 0.6pt,line join=round] (211.01,291.91) --
	(211.01,298.18);

\path[draw=drawColor,line width= 0.6pt,line join=round] (203.05, 72.46) --
	(203.05, 78.73);

\path[draw=drawColor,line width= 0.6pt,line join=round] (203.05, 75.60) --
	(202.24, 75.60);

\path[draw=drawColor,line width= 0.6pt,line join=round] (202.24, 72.46) --
	(202.24, 78.73);

\path[draw=drawColor,line width= 0.6pt,line join=round] (214.88,260.56) --
	(214.88,266.83);

\path[draw=drawColor,line width= 0.6pt,line join=round] (214.88,263.70) --
	(204.77,263.70);

\path[draw=drawColor,line width= 0.6pt,line join=round] (204.77,260.56) --
	(204.77,266.83);

\path[draw=drawColor,line width= 0.6pt,line join=round] (210.09,229.21) --
	(210.09,235.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] (210.09,232.35) --
	(205.20,232.35);

\path[draw=drawColor,line width= 0.6pt,line join=round] (205.20,229.21) --
	(205.20,235.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] (202.94,166.51) --
	(202.94,172.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (202.94,169.65) --
	(200.68,169.65);

\path[draw=drawColor,line width= 0.6pt,line join=round] (200.68,166.51) --
	(200.68,172.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (203.69,135.16) --
	(203.69,141.43);

\path[draw=drawColor,line width= 0.6pt,line join=round] (203.69,138.30) --
	(201.22,138.30);

\path[draw=drawColor,line width= 0.6pt,line join=round] (201.22,135.16) --
	(201.22,141.43);

\path[draw=drawColor,line width= 0.6pt,line join=round] (223.97, 41.12) --
	(223.97, 47.39);

\path[draw=drawColor,line width= 0.6pt,line join=round] (223.97, 44.25) --
	(217.78, 44.25);

\path[draw=drawColor,line width= 0.6pt,line join=round] (217.78, 41.12) --
	(217.78, 47.39);
\definecolor{fillColor}{RGB}{17,119,51}

\path[fill=fillColor] (206.54,106.95) circle (  1.96);

\path[fill=fillColor] (203.75,451.80) circle (  1.96);

\path[fill=fillColor] (204.07,420.45) circle (  1.96);

\path[fill=fillColor] (201.22,389.10) circle (  1.96);

\path[fill=fillColor] (210.74,357.75) circle (  1.96);

\path[fill=fillColor] (208.32,326.40) circle (  1.96);

\path[fill=fillColor] (215.58,295.05) circle (  1.96);

\path[fill=fillColor] (202.67, 75.60) circle (  1.96);

\path[fill=fillColor] (209.82,263.70) circle (  1.96);

\path[fill=fillColor] (207.62,232.35) circle (  1.96);

\path[fill=fillColor] (201.81,169.65) circle (  1.96);

\path[fill=fillColor] (202.45,138.30) circle (  1.96);

\path[fill=fillColor] (220.90, 44.25) circle (  1.96);
\definecolor{drawColor}{RGB}{169,169,169}

\path[draw=drawColor,line width= 0.6pt,line join=round] (195.95,122.62) -- (256.29,122.62);

\path[draw=drawColor,line width= 0.6pt,line join=round] (195.95, 91.27) -- (256.29, 91.27);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (195.95, 25.44) rectangle (256.29,470.61);
\end{scope}
\begin{scope}
\path[clip] (261.79, 25.44) rectangle (322.14,470.61);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (261.79, 25.44) rectangle (322.14,470.61);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] (272.33, 25.44) --
	(272.33,470.61);

\path[draw=drawColor,line width= 0.3pt,line join=round] (285.78, 25.44) --
	(285.78,470.61);

\path[draw=drawColor,line width= 0.3pt,line join=round] (299.22, 25.44) --
	(299.22,470.61);

\path[draw=drawColor,line width= 0.3pt,line join=round] (312.67, 25.44) --
	(312.67,470.61);

\path[draw=drawColor,line width= 0.6pt,line join=round] (261.79, 44.25) --
	(322.14, 44.25);

\path[draw=drawColor,line width= 0.6pt,line join=round] (261.79, 75.60) --
	(322.14, 75.60);

\path[draw=drawColor,line width= 0.6pt,line join=round] (261.79,106.95) --
	(322.14,106.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (261.79,138.30) --
	(322.14,138.30);

\path[draw=drawColor,line width= 0.6pt,line join=round] (261.79,169.65) --
	(322.14,169.65);

\path[draw=drawColor,line width= 0.6pt,line join=round] (261.79,201.00) --
	(322.14,201.00);

\path[draw=drawColor,line width= 0.6pt,line join=round] (261.79,232.35) --
	(322.14,232.35);

\path[draw=drawColor,line width= 0.6pt,line join=round] (261.79,263.70) --
	(322.14,263.70);

\path[draw=drawColor,line width= 0.6pt,line join=round] (261.79,295.05) --
	(322.14,295.05);

\path[draw=drawColor,line width= 0.6pt,line join=round] (261.79,326.40) --
	(322.14,326.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] (261.79,357.75) --
	(322.14,357.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (261.79,389.10) --
	(322.14,389.10);

\path[draw=drawColor,line width= 0.6pt,line join=round] (261.79,420.45) --
	(322.14,420.45);

\path[draw=drawColor,line width= 0.6pt,line join=round] (261.79,451.80) --
	(322.14,451.80);

\path[draw=drawColor,line width= 0.6pt,line join=round] (265.61, 25.44) --
	(265.61,470.61);

\path[draw=drawColor,line width= 0.6pt,line join=round] (279.06, 25.44) --
	(279.06,470.61);

\path[draw=drawColor,line width= 0.6pt,line join=round] (292.50, 25.44) --
	(292.50,470.61);

\path[draw=drawColor,line width= 0.6pt,line join=round] (305.95, 25.44) --
	(305.95,470.61);

\path[draw=drawColor,line width= 0.6pt,line join=round] (319.39, 25.44) --
	(319.39,470.61);
\definecolor{drawColor}{RGB}{51,34,136}

\path[draw=drawColor,line width= 0.6pt,line join=round] (278.89,103.81) --
	(278.89,110.08);

\path[draw=drawColor,line width= 0.6pt,line join=round] (278.89,106.95) --
	(273.03,106.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (273.03,103.81) --
	(273.03,110.08);

\path[draw=drawColor,line width= 0.6pt,line join=round] (278.41,448.66) --
	(278.41,454.93);

\path[draw=drawColor,line width= 0.6pt,line join=round] (278.41,451.80) --
	(273.73,451.80);

\path[draw=drawColor,line width= 0.6pt,line join=round] (273.73,448.66) --
	(273.73,454.93);

\path[draw=drawColor,line width= 0.6pt,line join=round] (273.41,417.31) --
	(273.41,423.58);

\path[draw=drawColor,line width= 0.6pt,line join=round] (273.41,420.45) --
	(268.78,420.45);

\path[draw=drawColor,line width= 0.6pt,line join=round] (268.78,417.31) --
	(268.78,423.58);

\path[draw=drawColor,line width= 0.6pt,line join=round] (287.55,385.96) --
	(287.55,392.23);

\path[draw=drawColor,line width= 0.6pt,line join=round] (287.55,389.10) --
	(275.61,389.10);

\path[draw=drawColor,line width= 0.6pt,line join=round] (275.61,385.96) --
	(275.61,392.23);

\path[draw=drawColor,line width= 0.6pt,line join=round] (280.24,354.61) --
	(280.24,360.88);

\path[draw=drawColor,line width= 0.6pt,line join=round] (280.24,357.75) --
	(267.33,357.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (267.33,354.61) --
	(267.33,360.88);

\path[draw=drawColor,line width= 0.6pt,line join=round] (272.12,323.26) --
	(272.12,329.53);

\path[draw=drawColor,line width= 0.6pt,line join=round] (272.12,326.40) --
	(269.05,326.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] (269.05,323.26) --
	(269.05,329.53);

\path[draw=drawColor,line width= 0.6pt,line join=round] (283.90,291.91) --
	(283.90,298.18);

\path[draw=drawColor,line width= 0.6pt,line join=round] (283.90,295.05) --
	(275.29,295.05);

\path[draw=drawColor,line width= 0.6pt,line join=round] (275.29,291.91) --
	(275.29,298.18);

\path[draw=drawColor,line width= 0.6pt,line join=round] (282.34, 72.46) --
	(282.34, 78.73);

\path[draw=drawColor,line width= 0.6pt,line join=round] (282.34, 75.60) --
	(280.72, 75.60);

\path[draw=drawColor,line width= 0.6pt,line join=round] (280.72, 72.46) --
	(280.72, 78.73);

\path[draw=drawColor,line width= 0.6pt,line join=round] (282.50,260.56) --
	(282.50,266.83);

\path[draw=drawColor,line width= 0.6pt,line join=round] (282.50,263.70) --
	(271.85,263.70);

\path[draw=drawColor,line width= 0.6pt,line join=round] (271.85,260.56) --
	(271.85,266.83);

\path[draw=drawColor,line width= 0.6pt,line join=round] (271.47,229.21) --
	(271.47,235.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] (271.47,232.35) --
	(267.87,232.35);

\path[draw=drawColor,line width= 0.6pt,line join=round] (267.87,229.21) --
	(267.87,235.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] (277.82,197.86) --
	(277.82,204.13);

\path[draw=drawColor,line width= 0.6pt,line join=round] (277.82,201.00) --
	(272.01,201.00);

\path[draw=drawColor,line width= 0.6pt,line join=round] (272.01,197.86) --
	(272.01,204.13);

\path[draw=drawColor,line width= 0.6pt,line join=round] (281.31,166.51) --
	(281.31,172.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (281.31,169.65) --
	(275.94,169.65);

\path[draw=drawColor,line width= 0.6pt,line join=round] (275.94,166.51) --
	(275.94,172.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (285.08,135.16) --
	(285.08,141.43);

\path[draw=drawColor,line width= 0.6pt,line join=round] (285.08,138.30) --
	(279.54,138.30);

\path[draw=drawColor,line width= 0.6pt,line join=round] (279.54,135.16) --
	(279.54,141.43);

\path[draw=drawColor,line width= 0.6pt,line join=round] (293.90, 41.12) --
	(293.90, 47.39);

\path[draw=drawColor,line width= 0.6pt,line join=round] (293.90, 44.25) --
	(287.66, 44.25);

\path[draw=drawColor,line width= 0.6pt,line join=round] (287.66, 41.12) --
	(287.66, 47.39);
\definecolor{fillColor}{RGB}{51,34,136}

\path[fill=fillColor] (275.94,106.95) circle (  1.96);

\path[fill=fillColor] (276.10,451.80) circle (  1.96);

\path[fill=fillColor] (271.10,420.45) circle (  1.96);

\path[fill=fillColor] (281.58,389.10) circle (  1.96);

\path[fill=fillColor] (273.78,357.75) circle (  1.96);

\path[fill=fillColor] (270.61,326.40) circle (  1.96);

\path[fill=fillColor] (279.59,295.05) circle (  1.96);

\path[fill=fillColor] (281.53, 75.60) circle (  1.96);

\path[fill=fillColor] (277.17,263.70) circle (  1.96);

\path[fill=fillColor] (269.64,232.35) circle (  1.96);

\path[fill=fillColor] (274.91,201.00) circle (  1.96);

\path[fill=fillColor] (278.62,169.65) circle (  1.96);

\path[fill=fillColor] (282.28,138.30) circle (  1.96);

\path[fill=fillColor] (290.78, 44.25) circle (  1.96);
\definecolor{drawColor}{RGB}{169,169,169}

\path[draw=drawColor,line width= 0.6pt,line join=round] (261.79,122.62) -- (322.14,122.62);

\path[draw=drawColor,line width= 0.6pt,line join=round] (261.79, 91.27) -- (322.14, 91.27);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (261.79, 25.44) rectangle (322.14,470.61);
\end{scope}
\begin{scope}
\path[clip] (327.64, 25.44) rectangle (387.98,470.61);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (327.64, 25.44) rectangle (387.98,470.61);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] (338.18, 25.44) --
	(338.18,470.61);

\path[draw=drawColor,line width= 0.3pt,line join=round] (351.62, 25.44) --
	(351.62,470.61);

\path[draw=drawColor,line width= 0.3pt,line join=round] (365.07, 25.44) --
	(365.07,470.61);

\path[draw=drawColor,line width= 0.3pt,line join=round] (378.51, 25.44) --
	(378.51,470.61);

\path[draw=drawColor,line width= 0.6pt,line join=round] (327.64, 44.25) --
	(387.98, 44.25);

\path[draw=drawColor,line width= 0.6pt,line join=round] (327.64, 75.60) --
	(387.98, 75.60);

\path[draw=drawColor,line width= 0.6pt,line join=round] (327.64,106.95) --
	(387.98,106.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (327.64,138.30) --
	(387.98,138.30);

\path[draw=drawColor,line width= 0.6pt,line join=round] (327.64,169.65) --
	(387.98,169.65);

\path[draw=drawColor,line width= 0.6pt,line join=round] (327.64,201.00) --
	(387.98,201.00);

\path[draw=drawColor,line width= 0.6pt,line join=round] (327.64,232.35) --
	(387.98,232.35);

\path[draw=drawColor,line width= 0.6pt,line join=round] (327.64,263.70) --
	(387.98,263.70);

\path[draw=drawColor,line width= 0.6pt,line join=round] (327.64,295.05) --
	(387.98,295.05);

\path[draw=drawColor,line width= 0.6pt,line join=round] (327.64,326.40) --
	(387.98,326.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] (327.64,357.75) --
	(387.98,357.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (327.64,389.10) --
	(387.98,389.10);

\path[draw=drawColor,line width= 0.6pt,line join=round] (327.64,420.45) --
	(387.98,420.45);

\path[draw=drawColor,line width= 0.6pt,line join=round] (327.64,451.80) --
	(387.98,451.80);

\path[draw=drawColor,line width= 0.6pt,line join=round] (331.45, 25.44) --
	(331.45,470.61);

\path[draw=drawColor,line width= 0.6pt,line join=round] (344.90, 25.44) --
	(344.90,470.61);

\path[draw=drawColor,line width= 0.6pt,line join=round] (358.34, 25.44) --
	(358.34,470.61);

\path[draw=drawColor,line width= 0.6pt,line join=round] (371.79, 25.44) --
	(371.79,470.61);

\path[draw=drawColor,line width= 0.6pt,line join=round] (385.24, 25.44) --
	(385.24,470.61);
\definecolor{drawColor}{RGB}{170,68,153}

\path[draw=drawColor,line width= 0.6pt,line join=round] (339.47,103.81) --
	(339.47,110.08);

\path[draw=drawColor,line width= 0.6pt,line join=round] (339.47,106.95) --
	(332.74,106.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (332.74,103.81) --
	(332.74,110.08);

\path[draw=drawColor,line width= 0.6pt,line join=round] (340.70,448.66) --
	(340.70,454.93);

\path[draw=drawColor,line width= 0.6pt,line join=round] (340.70,451.80) --
	(336.72,451.80);

\path[draw=drawColor,line width= 0.6pt,line join=round] (336.72,448.66) --
	(336.72,454.93);

\path[draw=drawColor,line width= 0.6pt,line join=round] (333.77,417.31) --
	(333.77,423.58);

\path[draw=drawColor,line width= 0.6pt,line join=round] (333.77,420.45) --
	(331.61,420.45);

\path[draw=drawColor,line width= 0.6pt,line join=round] (331.61,417.31) --
	(331.61,423.58);

\path[draw=drawColor,line width= 0.6pt,line join=round] (346.14,354.61) --
	(346.14,360.88);

\path[draw=drawColor,line width= 0.6pt,line join=round] (346.14,357.75) --
	(333.60,357.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (333.60,354.61) --
	(333.60,360.88);

\path[draw=drawColor,line width= 0.6pt,line join=round] (348.29,291.91) --
	(348.29,298.18);

\path[draw=drawColor,line width= 0.6pt,line join=round] (348.29,295.05) --
	(338.34,295.05);

\path[draw=drawColor,line width= 0.6pt,line join=round] (338.34,291.91) --
	(338.34,298.18);

\path[draw=drawColor,line width= 0.6pt,line join=round] (335.38, 72.46) --
	(335.38, 78.73);

\path[draw=drawColor,line width= 0.6pt,line join=round] (335.38, 75.60) --
	(334.47, 75.60);

\path[draw=drawColor,line width= 0.6pt,line join=round] (334.47, 72.46) --
	(334.47, 78.73);

\path[draw=drawColor,line width= 0.6pt,line join=round] (337.58,260.56) --
	(337.58,266.83);

\path[draw=drawColor,line width= 0.6pt,line join=round] (337.58,263.70) --
	(331.51,263.70);

\path[draw=drawColor,line width= 0.6pt,line join=round] (331.51,260.56) --
	(331.51,266.83);

\path[draw=drawColor,line width= 0.6pt,line join=round] (335.11,229.21) --
	(335.11,235.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] (335.11,232.35) --
	(332.31,232.35);

\path[draw=drawColor,line width= 0.6pt,line join=round] (332.31,229.21) --
	(332.31,235.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] (333.17,166.51) --
	(333.17,172.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (333.17,169.65) --
	(331.56,169.65);

\path[draw=drawColor,line width= 0.6pt,line join=round] (331.56,166.51) --
	(331.56,172.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (334.79,135.16) --
	(334.79,141.43);

\path[draw=drawColor,line width= 0.6pt,line join=round] (334.79,138.30) --
	(332.48,138.30);

\path[draw=drawColor,line width= 0.6pt,line join=round] (332.48,135.16) --
	(332.48,141.43);
\definecolor{fillColor}{RGB}{170,68,153}

\path[fill=fillColor] (336.13,106.95) circle (  1.96);

\path[fill=fillColor] (338.71,451.80) circle (  1.96);

\path[fill=fillColor] (332.69,420.45) circle (  1.96);

\path[fill=fillColor] (339.90,357.75) circle (  1.96);

\path[fill=fillColor] (343.34,295.05) circle (  1.96);

\path[fill=fillColor] (334.90, 75.60) circle (  1.96);

\path[fill=fillColor] (334.57,263.70) circle (  1.96);

\path[fill=fillColor] (333.71,232.35) circle (  1.96);

\path[fill=fillColor] (332.37,169.65) circle (  1.96);

\path[fill=fillColor] (333.66,138.30) circle (  1.96);
\definecolor{drawColor}{RGB}{169,169,169}

\path[draw=drawColor,line width= 0.6pt,line join=round] (327.64,122.62) -- (387.98,122.62);

\path[draw=drawColor,line width= 0.6pt,line join=round] (327.64, 91.27) -- (387.98, 91.27);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (327.64, 25.44) rectangle (387.98,470.61);
\end{scope}
\begin{scope}
\path[clip] (393.48, 25.44) rectangle (453.82,470.61);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (393.48, 25.44) rectangle (453.82,470.61);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] (404.02, 25.44) --
	(404.02,470.61);

\path[draw=drawColor,line width= 0.3pt,line join=round] (417.47, 25.44) --
	(417.47,470.61);

\path[draw=drawColor,line width= 0.3pt,line join=round] (430.91, 25.44) --
	(430.91,470.61);

\path[draw=drawColor,line width= 0.3pt,line join=round] (444.36, 25.44) --
	(444.36,470.61);

\path[draw=drawColor,line width= 0.6pt,line join=round] (393.48, 44.25) --
	(453.82, 44.25);

\path[draw=drawColor,line width= 0.6pt,line join=round] (393.48, 75.60) --
	(453.82, 75.60);

\path[draw=drawColor,line width= 0.6pt,line join=round] (393.48,106.95) --
	(453.82,106.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (393.48,138.30) --
	(453.82,138.30);

\path[draw=drawColor,line width= 0.6pt,line join=round] (393.48,169.65) --
	(453.82,169.65);

\path[draw=drawColor,line width= 0.6pt,line join=round] (393.48,201.00) --
	(453.82,201.00);

\path[draw=drawColor,line width= 0.6pt,line join=round] (393.48,232.35) --
	(453.82,232.35);

\path[draw=drawColor,line width= 0.6pt,line join=round] (393.48,263.70) --
	(453.82,263.70);

\path[draw=drawColor,line width= 0.6pt,line join=round] (393.48,295.05) --
	(453.82,295.05);

\path[draw=drawColor,line width= 0.6pt,line join=round] (393.48,326.40) --
	(453.82,326.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] (393.48,357.75) --
	(453.82,357.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (393.48,389.10) --
	(453.82,389.10);

\path[draw=drawColor,line width= 0.6pt,line join=round] (393.48,420.45) --
	(453.82,420.45);

\path[draw=drawColor,line width= 0.6pt,line join=round] (393.48,451.80) --
	(453.82,451.80);

\path[draw=drawColor,line width= 0.6pt,line join=round] (397.30, 25.44) --
	(397.30,470.61);

\path[draw=drawColor,line width= 0.6pt,line join=round] (410.74, 25.44) --
	(410.74,470.61);

\path[draw=drawColor,line width= 0.6pt,line join=round] (424.19, 25.44) --
	(424.19,470.61);

\path[draw=drawColor,line width= 0.6pt,line join=round] (437.63, 25.44) --
	(437.63,470.61);

\path[draw=drawColor,line width= 0.6pt,line join=round] (451.08, 25.44) --
	(451.08,470.61);
\definecolor{drawColor}{RGB}{68,170,153}

\path[draw=drawColor,line width= 0.6pt,line join=round] (400.95,103.81) --
	(400.95,110.08);

\path[draw=drawColor,line width= 0.6pt,line join=round] (400.95,106.95) --
	(398.21,106.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (398.21,103.81) --
	(398.21,110.08);

\path[draw=drawColor,line width= 0.6pt,line join=round] (400.31,448.66) --
	(400.31,454.93);

\path[draw=drawColor,line width= 0.6pt,line join=round] (400.31,451.80) --
	(398.10,451.80);

\path[draw=drawColor,line width= 0.6pt,line join=round] (398.10,448.66) --
	(398.10,454.93);

\path[draw=drawColor,line width= 0.6pt,line join=round] (398.27,417.31) --
	(398.27,423.58);

\path[draw=drawColor,line width= 0.6pt,line join=round] (398.27,420.45) --
	(396.98,420.45);

\path[draw=drawColor,line width= 0.6pt,line join=round] (396.98,417.31) --
	(396.98,423.58);

\path[draw=drawColor,line width= 0.6pt,line join=round] (403.00,354.61) --
	(403.00,360.88);

\path[draw=drawColor,line width= 0.6pt,line join=round] (403.00,357.75) --
	(396.28,357.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (396.28,354.61) --
	(396.28,360.88);

\path[draw=drawColor,line width= 0.6pt,line join=round] (406.66,291.91) --
	(406.66,298.18);

\path[draw=drawColor,line width= 0.6pt,line join=round] (406.66,295.05) --
	(400.26,295.05);

\path[draw=drawColor,line width= 0.6pt,line join=round] (400.26,291.91) --
	(400.26,298.18);

\path[draw=drawColor,line width= 0.6pt,line join=round] (399.61, 72.46) --
	(399.61, 78.73);

\path[draw=drawColor,line width= 0.6pt,line join=round] (399.61, 75.60) --
	(398.97, 75.60);

\path[draw=drawColor,line width= 0.6pt,line join=round] (398.97, 72.46) --
	(398.97, 78.73);

\path[draw=drawColor,line width= 0.6pt,line join=round] (404.40,260.56) --
	(404.40,266.83);

\path[draw=drawColor,line width= 0.6pt,line join=round] (404.40,263.70) --
	(397.78,263.70);

\path[draw=drawColor,line width= 0.6pt,line join=round] (397.78,260.56) --
	(397.78,266.83);

\path[draw=drawColor,line width= 0.6pt,line join=round] (400.09,229.21) --
	(400.09,235.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] (400.09,232.35) --
	(397.73,232.35);

\path[draw=drawColor,line width= 0.6pt,line join=round] (397.73,229.21) --
	(397.73,235.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] (401.49,197.86) --
	(401.49,204.13);

\path[draw=drawColor,line width= 0.6pt,line join=round] (401.49,201.00) --
	(398.64,201.00);

\path[draw=drawColor,line width= 0.6pt,line join=round] (398.64,197.86) --
	(398.64,204.13);

\path[draw=drawColor,line width= 0.6pt,line join=round] (399.02,166.51) --
	(399.02,172.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (399.02,169.65) --
	(397.41,169.65);

\path[draw=drawColor,line width= 0.6pt,line join=round] (397.41,166.51) --
	(397.41,172.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (398.86,135.16) --
	(398.86,141.43);

\path[draw=drawColor,line width= 0.6pt,line join=round] (398.86,138.30) --
	(397.41,138.30);

\path[draw=drawColor,line width= 0.6pt,line join=round] (397.41,135.16) --
	(397.41,141.43);
\definecolor{fillColor}{RGB}{68,170,153}

\path[fill=fillColor] (399.61,106.95) circle (  1.96);

\path[fill=fillColor] (399.18,451.80) circle (  1.96);

\path[fill=fillColor] (397.62,420.45) circle (  1.96);

\path[fill=fillColor] (399.66,357.75) circle (  1.96);

\path[fill=fillColor] (403.48,295.05) circle (  1.96);

\path[fill=fillColor] (399.29, 75.60) circle (  1.96);

\path[fill=fillColor] (401.06,263.70) circle (  1.96);

\path[fill=fillColor] (398.91,232.35) circle (  1.96);

\path[fill=fillColor] (400.04,201.00) circle (  1.96);

\path[fill=fillColor] (398.21,169.65) circle (  1.96);

\path[fill=fillColor] (398.16,138.30) circle (  1.96);
\definecolor{drawColor}{RGB}{169,169,169}

\path[draw=drawColor,line width= 0.6pt,line join=round] (393.48,122.62) -- (453.82,122.62);

\path[draw=drawColor,line width= 0.6pt,line join=round] (393.48, 91.27) -- (453.82, 91.27);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (393.48, 25.44) rectangle (453.82,470.61);
\end{scope}
\begin{scope}
\path[clip] (459.32, 25.44) rectangle (519.67,470.61);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (459.32, 25.44) rectangle (519.67,470.61);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] (469.86, 25.44) --
	(469.86,470.61);

\path[draw=drawColor,line width= 0.3pt,line join=round] (483.31, 25.44) --
	(483.31,470.61);

\path[draw=drawColor,line width= 0.3pt,line join=round] (496.76, 25.44) --
	(496.76,470.61);

\path[draw=drawColor,line width= 0.3pt,line join=round] (510.20, 25.44) --
	(510.20,470.61);

\path[draw=drawColor,line width= 0.6pt,line join=round] (459.32, 44.25) --
	(519.67, 44.25);

\path[draw=drawColor,line width= 0.6pt,line join=round] (459.32, 75.60) --
	(519.67, 75.60);

\path[draw=drawColor,line width= 0.6pt,line join=round] (459.32,106.95) --
	(519.67,106.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (459.32,138.30) --
	(519.67,138.30);

\path[draw=drawColor,line width= 0.6pt,line join=round] (459.32,169.65) --
	(519.67,169.65);

\path[draw=drawColor,line width= 0.6pt,line join=round] (459.32,201.00) --
	(519.67,201.00);

\path[draw=drawColor,line width= 0.6pt,line join=round] (459.32,232.35) --
	(519.67,232.35);

\path[draw=drawColor,line width= 0.6pt,line join=round] (459.32,263.70) --
	(519.67,263.70);

\path[draw=drawColor,line width= 0.6pt,line join=round] (459.32,295.05) --
	(519.67,295.05);

\path[draw=drawColor,line width= 0.6pt,line join=round] (459.32,326.40) --
	(519.67,326.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] (459.32,357.75) --
	(519.67,357.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (459.32,389.10) --
	(519.67,389.10);

\path[draw=drawColor,line width= 0.6pt,line join=round] (459.32,420.45) --
	(519.67,420.45);

\path[draw=drawColor,line width= 0.6pt,line join=round] (459.32,451.80) --
	(519.67,451.80);

\path[draw=drawColor,line width= 0.6pt,line join=round] (463.14, 25.44) --
	(463.14,470.61);

\path[draw=drawColor,line width= 0.6pt,line join=round] (476.59, 25.44) --
	(476.59,470.61);

\path[draw=drawColor,line width= 0.6pt,line join=round] (490.03, 25.44) --
	(490.03,470.61);

\path[draw=drawColor,line width= 0.6pt,line join=round] (503.48, 25.44) --
	(503.48,470.61);

\path[draw=drawColor,line width= 0.6pt,line join=round] (516.92, 25.44) --
	(516.92,470.61);
\definecolor{drawColor}{RGB}{153,153,51}

\path[draw=drawColor,line width= 0.6pt,line join=round] (464.92,103.81) --
	(464.92,110.08);

\path[draw=drawColor,line width= 0.6pt,line join=round] (464.92,106.95) --
	(463.03,106.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (463.03,103.81) --
	(463.03,110.08);

\path[draw=drawColor,line width= 0.6pt,line join=round] (464.65,448.66) --
	(464.65,454.93);

\path[draw=drawColor,line width= 0.6pt,line join=round] (464.65,451.80) --
	(463.25,451.80);

\path[draw=drawColor,line width= 0.6pt,line join=round] (463.25,448.66) --
	(463.25,454.93);

\path[draw=drawColor,line width= 0.6pt,line join=round] (463.84,417.31) --
	(463.84,423.58);

\path[draw=drawColor,line width= 0.6pt,line join=round] (463.84,420.45) --
	(462.93,420.45);

\path[draw=drawColor,line width= 0.6pt,line join=round] (462.93,417.31) --
	(462.93,423.58);

\path[draw=drawColor,line width= 0.6pt,line join=round] (466.10,354.61) --
	(466.10,360.88);

\path[draw=drawColor,line width= 0.6pt,line join=round] (466.10,357.75) --
	(462.12,357.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (462.12,354.61) --
	(462.12,360.88);

\path[draw=drawColor,line width= 0.6pt,line join=round] (463.52,323.26) --
	(463.52,329.53);

\path[draw=drawColor,line width= 0.6pt,line join=round] (463.52,326.40) --
	(463.03,326.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] (463.03,323.26) --
	(463.03,329.53);

\path[draw=drawColor,line width= 0.6pt,line join=round] (469.17, 72.46) --
	(469.17, 78.73);

\path[draw=drawColor,line width= 0.6pt,line join=round] (469.17, 75.60) --
	(468.09, 75.60);

\path[draw=drawColor,line width= 0.6pt,line join=round] (468.09, 72.46) --
	(468.09, 78.73);

\path[draw=drawColor,line width= 0.6pt,line join=round] (469.11,260.56) --
	(469.11,266.83);

\path[draw=drawColor,line width= 0.6pt,line join=round] (469.11,263.70) --
	(463.20,263.70);

\path[draw=drawColor,line width= 0.6pt,line join=round] (463.20,260.56) --
	(463.20,266.83);

\path[draw=drawColor,line width= 0.6pt,line join=round] (464.33,229.21) --
	(464.33,235.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] (464.33,232.35) --
	(462.93,232.35);

\path[draw=drawColor,line width= 0.6pt,line join=round] (462.93,229.21) --
	(462.93,235.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] (463.63,135.16) --
	(463.63,141.43);

\path[draw=drawColor,line width= 0.6pt,line join=round] (463.63,138.30) --
	(462.98,138.30);

\path[draw=drawColor,line width= 0.6pt,line join=round] (462.98,135.16) --
	(462.98,141.43);
\definecolor{fillColor}{RGB}{153,153,51}

\path[fill=fillColor] (463.95,106.95) circle (  1.96);

\path[fill=fillColor] (463.95,451.80) circle (  1.96);

\path[fill=fillColor] (463.36,420.45) circle (  1.96);

\path[fill=fillColor] (464.11,357.75) circle (  1.96);

\path[fill=fillColor] (463.25,326.40) circle (  1.96);

\path[fill=fillColor] (468.63, 75.60) circle (  1.96);

\path[fill=fillColor] (466.15,263.70) circle (  1.96);

\path[fill=fillColor] (463.63,232.35) circle (  1.96);

\path[fill=fillColor] (463.30,138.30) circle (  1.96);
\definecolor{drawColor}{RGB}{169,169,169}

\path[draw=drawColor,line width= 0.6pt,line join=round] (459.32,122.62) -- (519.67,122.62);

\path[draw=drawColor,line width= 0.6pt,line join=round] (459.32, 91.27) -- (519.67, 91.27);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (459.32, 25.44) rectangle (519.67,470.61);
\end{scope}
\begin{scope}
\path[clip] (525.17, 25.44) rectangle (585.51,470.61);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (525.17, 25.44) rectangle (585.51,470.61);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] (535.71, 25.44) --
	(535.71,470.61);

\path[draw=drawColor,line width= 0.3pt,line join=round] (549.15, 25.44) --
	(549.15,470.61);

\path[draw=drawColor,line width= 0.3pt,line join=round] (562.60, 25.44) --
	(562.60,470.61);

\path[draw=drawColor,line width= 0.3pt,line join=round] (576.05, 25.44) --
	(576.05,470.61);

\path[draw=drawColor,line width= 0.6pt,line join=round] (525.17, 44.25) --
	(585.51, 44.25);

\path[draw=drawColor,line width= 0.6pt,line join=round] (525.17, 75.60) --
	(585.51, 75.60);

\path[draw=drawColor,line width= 0.6pt,line join=round] (525.17,106.95) --
	(585.51,106.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (525.17,138.30) --
	(585.51,138.30);

\path[draw=drawColor,line width= 0.6pt,line join=round] (525.17,169.65) --
	(585.51,169.65);

\path[draw=drawColor,line width= 0.6pt,line join=round] (525.17,201.00) --
	(585.51,201.00);

\path[draw=drawColor,line width= 0.6pt,line join=round] (525.17,232.35) --
	(585.51,232.35);

\path[draw=drawColor,line width= 0.6pt,line join=round] (525.17,263.70) --
	(585.51,263.70);

\path[draw=drawColor,line width= 0.6pt,line join=round] (525.17,295.05) --
	(585.51,295.05);

\path[draw=drawColor,line width= 0.6pt,line join=round] (525.17,326.40) --
	(585.51,326.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] (525.17,357.75) --
	(585.51,357.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (525.17,389.10) --
	(585.51,389.10);

\path[draw=drawColor,line width= 0.6pt,line join=round] (525.17,420.45) --
	(585.51,420.45);

\path[draw=drawColor,line width= 0.6pt,line join=round] (525.17,451.80) --
	(585.51,451.80);

\path[draw=drawColor,line width= 0.6pt,line join=round] (528.99, 25.44) --
	(528.99,470.61);

\path[draw=drawColor,line width= 0.6pt,line join=round] (542.43, 25.44) --
	(542.43,470.61);

\path[draw=drawColor,line width= 0.6pt,line join=round] (555.88, 25.44) --
	(555.88,470.61);

\path[draw=drawColor,line width= 0.6pt,line join=round] (569.32, 25.44) --
	(569.32,470.61);

\path[draw=drawColor,line width= 0.6pt,line join=round] (582.77, 25.44) --
	(582.77,470.61);
\definecolor{drawColor}{RGB}{136,34,85}

\path[draw=drawColor,line width= 0.6pt,line join=round] (530.71,103.81) --
	(530.71,110.08);

\path[draw=drawColor,line width= 0.6pt,line join=round] (530.71,106.95) --
	(528.66,106.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (528.66,103.81) --
	(528.66,110.08);

\path[draw=drawColor,line width= 0.6pt,line join=round] (530.01,448.66) --
	(530.01,454.93);

\path[draw=drawColor,line width= 0.6pt,line join=round] (530.01,451.80) --
	(528.93,451.80);

\path[draw=drawColor,line width= 0.6pt,line join=round] (528.93,448.66) --
	(528.93,454.93);

\path[draw=drawColor,line width= 0.6pt,line join=round] (529.79,417.31) --
	(529.79,423.58);

\path[draw=drawColor,line width= 0.6pt,line join=round] (529.79,420.45) --
	(528.72,420.45);

\path[draw=drawColor,line width= 0.6pt,line join=round] (528.72,417.31) --
	(528.72,423.58);

\path[draw=drawColor,line width= 0.6pt,line join=round] (529.74, 72.46) --
	(529.74, 78.73);

\path[draw=drawColor,line width= 0.6pt,line join=round] (529.74, 75.60) --
	(529.36, 75.60);

\path[draw=drawColor,line width= 0.6pt,line join=round] (529.36, 72.46) --
	(529.36, 78.73);

\path[draw=drawColor,line width= 0.6pt,line join=round] (532.48,229.21) --
	(532.48,235.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] (532.48,232.35) --
	(529.74,232.35);

\path[draw=drawColor,line width= 0.6pt,line join=round] (529.74,229.21) --
	(529.74,235.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] (530.17,166.51) --
	(530.17,172.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (530.17,169.65) --
	(528.93,169.65);

\path[draw=drawColor,line width= 0.6pt,line join=round] (528.93,166.51) --
	(528.93,172.78);
\definecolor{fillColor}{RGB}{136,34,85}

\path[fill=fillColor] (529.69,106.95) circle (  1.96);

\path[fill=fillColor] (529.47,451.80) circle (  1.96);

\path[fill=fillColor] (529.25,420.45) circle (  1.96);

\path[fill=fillColor] (529.52, 75.60) circle (  1.96);

\path[fill=fillColor] (531.14,232.35) circle (  1.96);

\path[fill=fillColor] (529.52,169.65) circle (  1.96);
\definecolor{drawColor}{RGB}{169,169,169}

\path[draw=drawColor,line width= 0.6pt,line join=round] (525.17,122.62) -- (585.51,122.62);

\path[draw=drawColor,line width= 0.6pt,line join=round] (525.17, 91.27) -- (585.51, 91.27);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (525.17, 25.44) rectangle (585.51,470.61);
\end{scope}
\begin{scope}
\path[clip] (591.01, 25.44) rectangle (651.36,470.61);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (591.01, 25.44) rectangle (651.36,470.61);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] (601.55, 25.44) --
	(601.55,470.61);

\path[draw=drawColor,line width= 0.3pt,line join=round] (615.00, 25.44) --
	(615.00,470.61);

\path[draw=drawColor,line width= 0.3pt,line join=round] (628.44, 25.44) --
	(628.44,470.61);

\path[draw=drawColor,line width= 0.3pt,line join=round] (641.89, 25.44) --
	(641.89,470.61);

\path[draw=drawColor,line width= 0.6pt,line join=round] (591.01, 44.25) --
	(651.36, 44.25);

\path[draw=drawColor,line width= 0.6pt,line join=round] (591.01, 75.60) --
	(651.36, 75.60);

\path[draw=drawColor,line width= 0.6pt,line join=round] (591.01,106.95) --
	(651.36,106.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (591.01,138.30) --
	(651.36,138.30);

\path[draw=drawColor,line width= 0.6pt,line join=round] (591.01,169.65) --
	(651.36,169.65);

\path[draw=drawColor,line width= 0.6pt,line join=round] (591.01,201.00) --
	(651.36,201.00);

\path[draw=drawColor,line width= 0.6pt,line join=round] (591.01,232.35) --
	(651.36,232.35);

\path[draw=drawColor,line width= 0.6pt,line join=round] (591.01,263.70) --
	(651.36,263.70);

\path[draw=drawColor,line width= 0.6pt,line join=round] (591.01,295.05) --
	(651.36,295.05);

\path[draw=drawColor,line width= 0.6pt,line join=round] (591.01,326.40) --
	(651.36,326.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] (591.01,357.75) --
	(651.36,357.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (591.01,389.10) --
	(651.36,389.10);

\path[draw=drawColor,line width= 0.6pt,line join=round] (591.01,420.45) --
	(651.36,420.45);

\path[draw=drawColor,line width= 0.6pt,line join=round] (591.01,451.80) --
	(651.36,451.80);

\path[draw=drawColor,line width= 0.6pt,line join=round] (594.83, 25.44) --
	(594.83,470.61);

\path[draw=drawColor,line width= 0.6pt,line join=round] (608.28, 25.44) --
	(608.28,470.61);

\path[draw=drawColor,line width= 0.6pt,line join=round] (621.72, 25.44) --
	(621.72,470.61);

\path[draw=drawColor,line width= 0.6pt,line join=round] (635.17, 25.44) --
	(635.17,470.61);

\path[draw=drawColor,line width= 0.6pt,line join=round] (648.61, 25.44) --
	(648.61,470.61);
\definecolor{drawColor}{RGB}{102,17,0}

\path[draw=drawColor,line width= 0.6pt,line join=round] (603.97,103.81) --
	(603.97,110.08);

\path[draw=drawColor,line width= 0.6pt,line join=round] (603.97,106.95) --
	(598.11,106.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (598.11,103.81) --
	(598.11,110.08);

\path[draw=drawColor,line width= 0.6pt,line join=round] (606.72,448.66) --
	(606.72,454.93);

\path[draw=drawColor,line width= 0.6pt,line join=round] (606.72,451.80) --
	(602.20,451.80);

\path[draw=drawColor,line width= 0.6pt,line join=round] (602.20,448.66) --
	(602.20,454.93);

\path[draw=drawColor,line width= 0.6pt,line join=round] (602.47,417.31) --
	(602.47,423.58);

\path[draw=drawColor,line width= 0.6pt,line join=round] (602.47,420.45) --
	(597.95,420.45);

\path[draw=drawColor,line width= 0.6pt,line join=round] (597.95,417.31) --
	(597.95,423.58);

\path[draw=drawColor,line width= 0.6pt,line join=round] (598.70,354.61) --
	(598.70,360.88);

\path[draw=drawColor,line width= 0.6pt,line join=round] (598.70,357.75) --
	(594.02,357.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (594.02,354.61) --
	(594.02,360.88);

\path[draw=drawColor,line width= 0.6pt,line join=round] (598.60,323.26) --
	(598.60,329.53);

\path[draw=drawColor,line width= 0.6pt,line join=round] (598.60,326.40) --
	(596.34,326.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] (596.34,323.26) --
	(596.34,329.53);

\path[draw=drawColor,line width= 0.6pt,line join=round] (605.26,291.91) --
	(605.26,298.18);

\path[draw=drawColor,line width= 0.6pt,line join=round] (605.26,295.05) --
	(598.65,295.05);

\path[draw=drawColor,line width= 0.6pt,line join=round] (598.65,291.91) --
	(598.65,298.18);

\path[draw=drawColor,line width= 0.6pt,line join=round] (607.09, 72.46) --
	(607.09, 78.73);

\path[draw=drawColor,line width= 0.6pt,line join=round] (607.09, 75.60) --
	(605.64, 75.60);

\path[draw=drawColor,line width= 0.6pt,line join=round] (605.64, 72.46) --
	(605.64, 78.73);

\path[draw=drawColor,line width= 0.6pt,line join=round] (611.56,260.56) --
	(611.56,266.83);

\path[draw=drawColor,line width= 0.6pt,line join=round] (611.56,263.70) --
	(601.02,263.70);

\path[draw=drawColor,line width= 0.6pt,line join=round] (601.02,260.56) --
	(601.02,266.83);

\path[draw=drawColor,line width= 0.6pt,line join=round] (608.54,229.21) --
	(608.54,235.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] (608.54,232.35) --
	(602.95,232.35);

\path[draw=drawColor,line width= 0.6pt,line join=round] (602.95,229.21) --
	(602.95,235.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] (597.90,197.86) --
	(597.90,204.13);

\path[draw=drawColor,line width= 0.6pt,line join=round] (597.90,201.00) --
	(595.53,201.00);

\path[draw=drawColor,line width= 0.6pt,line join=round] (595.53,197.86) --
	(595.53,204.13);

\path[draw=drawColor,line width= 0.6pt,line join=round] (602.20,135.16) --
	(602.20,141.43);

\path[draw=drawColor,line width= 0.6pt,line join=round] (602.20,138.30) --
	(598.60,138.30);

\path[draw=drawColor,line width= 0.6pt,line join=round] (598.60,135.16) --
	(598.60,141.43);

\path[draw=drawColor,line width= 0.6pt,line join=round] (599.51, 41.12) --
	(599.51, 47.39);

\path[draw=drawColor,line width= 0.6pt,line join=round] (599.51, 44.25) --
	(596.66, 44.25);

\path[draw=drawColor,line width= 0.6pt,line join=round] (596.66, 41.12) --
	(596.66, 47.39);
\definecolor{fillColor}{RGB}{102,17,0}

\path[fill=fillColor] (601.07,106.95) circle (  1.96);

\path[fill=fillColor] (604.46,451.80) circle (  1.96);

\path[fill=fillColor] (600.21,420.45) circle (  1.96);

\path[fill=fillColor] (596.34,357.75) circle (  1.96);

\path[fill=fillColor] (597.47,326.40) circle (  1.96);

\path[fill=fillColor] (601.93,295.05) circle (  1.96);

\path[fill=fillColor] (606.34, 75.60) circle (  1.96);

\path[fill=fillColor] (606.29,263.70) circle (  1.96);

\path[fill=fillColor] (605.75,232.35) circle (  1.96);

\path[fill=fillColor] (596.71,201.00) circle (  1.96);

\path[fill=fillColor] (600.37,138.30) circle (  1.96);

\path[fill=fillColor] (598.06, 44.25) circle (  1.96);
\definecolor{drawColor}{RGB}{169,169,169}

\path[draw=drawColor,line width= 0.6pt,line join=round] (591.01,122.62) -- (651.36,122.62);

\path[draw=drawColor,line width= 0.6pt,line join=round] (591.01, 91.27) -- (651.36, 91.27);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (591.01, 25.44) rectangle (651.36,470.61);
\end{scope}
\begin{scope}
\path[clip] (656.86, 25.44) rectangle (717.20,470.61);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (656.86, 25.44) rectangle (717.20,470.61);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] (667.40, 25.44) --
	(667.40,470.61);

\path[draw=drawColor,line width= 0.3pt,line join=round] (680.84, 25.44) --
	(680.84,470.61);

\path[draw=drawColor,line width= 0.3pt,line join=round] (694.29, 25.44) --
	(694.29,470.61);

\path[draw=drawColor,line width= 0.3pt,line join=round] (707.73, 25.44) --
	(707.73,470.61);

\path[draw=drawColor,line width= 0.6pt,line join=round] (656.86, 44.25) --
	(717.20, 44.25);

\path[draw=drawColor,line width= 0.6pt,line join=round] (656.86, 75.60) --
	(717.20, 75.60);

\path[draw=drawColor,line width= 0.6pt,line join=round] (656.86,106.95) --
	(717.20,106.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (656.86,138.30) --
	(717.20,138.30);

\path[draw=drawColor,line width= 0.6pt,line join=round] (656.86,169.65) --
	(717.20,169.65);

\path[draw=drawColor,line width= 0.6pt,line join=round] (656.86,201.00) --
	(717.20,201.00);

\path[draw=drawColor,line width= 0.6pt,line join=round] (656.86,232.35) --
	(717.20,232.35);

\path[draw=drawColor,line width= 0.6pt,line join=round] (656.86,263.70) --
	(717.20,263.70);

\path[draw=drawColor,line width= 0.6pt,line join=round] (656.86,295.05) --
	(717.20,295.05);

\path[draw=drawColor,line width= 0.6pt,line join=round] (656.86,326.40) --
	(717.20,326.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] (656.86,357.75) --
	(717.20,357.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (656.86,389.10) --
	(717.20,389.10);

\path[draw=drawColor,line width= 0.6pt,line join=round] (656.86,420.45) --
	(717.20,420.45);

\path[draw=drawColor,line width= 0.6pt,line join=round] (656.86,451.80) --
	(717.20,451.80);

\path[draw=drawColor,line width= 0.6pt,line join=round] (660.67, 25.44) --
	(660.67,470.61);

\path[draw=drawColor,line width= 0.6pt,line join=round] (674.12, 25.44) --
	(674.12,470.61);

\path[draw=drawColor,line width= 0.6pt,line join=round] (687.57, 25.44) --
	(687.57,470.61);

\path[draw=drawColor,line width= 0.6pt,line join=round] (701.01, 25.44) --
	(701.01,470.61);

\path[draw=drawColor,line width= 0.6pt,line join=round] (714.46, 25.44) --
	(714.46,470.61);
\definecolor{drawColor}{RGB}{102,153,204}

\path[draw=drawColor,line width= 0.6pt,line join=round] (673.42,103.81) --
	(673.42,110.08);

\path[draw=drawColor,line width= 0.6pt,line join=round] (673.42,106.95) --
	(666.59,106.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (666.59,103.81) --
	(666.59,110.08);

\path[draw=drawColor,line width= 0.6pt,line join=round] (667.02,448.66) --
	(667.02,454.93);

\path[draw=drawColor,line width= 0.6pt,line join=round] (667.02,451.80) --
	(663.69,451.80);

\path[draw=drawColor,line width= 0.6pt,line join=round] (663.69,448.66) --
	(663.69,454.93);

\path[draw=drawColor,line width= 0.6pt,line join=round] (681.22,417.31) --
	(681.22,423.58);

\path[draw=drawColor,line width= 0.6pt,line join=round] (681.22,420.45) --
	(674.17,420.45);

\path[draw=drawColor,line width= 0.6pt,line join=round] (674.17,417.31) --
	(674.17,423.58);

\path[draw=drawColor,line width= 0.6pt,line join=round] (677.35,385.96) --
	(677.35,392.23);

\path[draw=drawColor,line width= 0.6pt,line join=round] (677.35,389.10) --
	(667.24,389.10);

\path[draw=drawColor,line width= 0.6pt,line join=round] (667.24,385.96) --
	(667.24,392.23);

\path[draw=drawColor,line width= 0.6pt,line join=round] (673.31,354.61) --
	(673.31,360.88);

\path[draw=drawColor,line width= 0.6pt,line join=round] (673.31,357.75) --
	(661.10,357.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (661.10,354.61) --
	(661.10,360.88);

\path[draw=drawColor,line width= 0.6pt,line join=round] (677.40,323.26) --
	(677.40,329.53);

\path[draw=drawColor,line width= 0.6pt,line join=round] (677.40,326.40) --
	(672.78,326.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] (672.78,323.26) --
	(672.78,329.53);

\path[draw=drawColor,line width= 0.6pt,line join=round] (676.16,291.91) --
	(676.16,298.18);

\path[draw=drawColor,line width= 0.6pt,line join=round] (676.16,295.05) --
	(666.27,295.05);

\path[draw=drawColor,line width= 0.6pt,line join=round] (666.27,291.91) --
	(666.27,298.18);

\path[draw=drawColor,line width= 0.6pt,line join=round] (663.79, 72.46) --
	(663.79, 78.73);

\path[draw=drawColor,line width= 0.6pt,line join=round] (663.79, 75.60) --
	(663.04, 75.60);

\path[draw=drawColor,line width= 0.6pt,line join=round] (663.04, 72.46) --
	(663.04, 78.73);

\path[draw=drawColor,line width= 0.6pt,line join=round] (680.20,260.56) --
	(680.20,266.83);

\path[draw=drawColor,line width= 0.6pt,line join=round] (680.20,263.70) --
	(668.90,263.70);

\path[draw=drawColor,line width= 0.6pt,line join=round] (668.90,260.56) --
	(668.90,266.83);

\path[draw=drawColor,line width= 0.6pt,line join=round] (665.35,229.21) --
	(665.35,235.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] (665.35,232.35) --
	(662.18,232.35);

\path[draw=drawColor,line width= 0.6pt,line join=round] (662.18,229.21) --
	(662.18,235.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] (676.97,197.86) --
	(676.97,204.13);

\path[draw=drawColor,line width= 0.6pt,line join=round] (676.97,201.00) --
	(671.05,201.00);

\path[draw=drawColor,line width= 0.6pt,line join=round] (671.05,197.86) --
	(671.05,204.13);

\path[draw=drawColor,line width= 0.6pt,line join=round] (666.59,166.51) --
	(666.59,172.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (666.59,169.65) --
	(663.31,169.65);

\path[draw=drawColor,line width= 0.6pt,line join=round] (663.31,166.51) --
	(663.31,172.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (665.25,135.16) --
	(665.25,141.43);

\path[draw=drawColor,line width= 0.6pt,line join=round] (665.25,138.30) --
	(662.50,138.30);

\path[draw=drawColor,line width= 0.6pt,line join=round] (662.50,135.16) --
	(662.50,141.43);

\path[draw=drawColor,line width= 0.6pt,line join=round] (690.20, 41.12) --
	(690.20, 47.39);

\path[draw=drawColor,line width= 0.6pt,line join=round] (690.20, 44.25) --
	(683.96, 44.25);

\path[draw=drawColor,line width= 0.6pt,line join=round] (683.96, 41.12) --
	(683.96, 47.39);
\definecolor{fillColor}{RGB}{102,153,204}

\path[fill=fillColor] (669.98,106.95) circle (  1.96);

\path[fill=fillColor] (665.35,451.80) circle (  1.96);

\path[fill=fillColor] (677.67,420.45) circle (  1.96);

\path[fill=fillColor] (672.29,389.10) circle (  1.96);

\path[fill=fillColor] (667.18,357.75) circle (  1.96);

\path[fill=fillColor] (675.09,326.40) circle (  1.96);

\path[fill=fillColor] (671.22,295.05) circle (  1.96);

\path[fill=fillColor] (663.42, 75.60) circle (  1.96);

\path[fill=fillColor] (674.55,263.70) circle (  1.96);

\path[fill=fillColor] (663.74,232.35) circle (  1.96);

\path[fill=fillColor] (674.01,201.00) circle (  1.96);

\path[fill=fillColor] (664.98,169.65) circle (  1.96);

\path[fill=fillColor] (663.90,138.30) circle (  1.96);

\path[fill=fillColor] (687.08, 44.25) circle (  1.96);
\definecolor{drawColor}{RGB}{169,169,169}

\path[draw=drawColor,line width= 0.6pt,line join=round] (656.86,122.62) -- (717.20,122.62);

\path[draw=drawColor,line width= 0.6pt,line join=round] (656.86, 91.27) -- (717.20, 91.27);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (656.86, 25.44) rectangle (717.20,470.61);
\end{scope}
\begin{scope}
\path[clip] ( 64.26,470.61) rectangle (124.60,522.86);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 64.26,470.61) rectangle (124.60,522.86);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 94.43,504.12) {Concerned};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 94.43,494.62) {about side};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 94.43,485.11) {effects};
\end{scope}
\begin{scope}
\path[clip] (130.10,470.61) rectangle (190.45,522.86);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (130.10,470.61) rectangle (190.45,522.86);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (160.27,513.62) {Concerned};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (160.27,504.12) {about getting};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (160.27,494.62) {coronavirus};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (160.27,485.11) {from the};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (160.27,475.61) {vaccine};
\end{scope}
\begin{scope}
\path[clip] (195.95,470.61) rectangle (256.29,522.86);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (195.95,470.61) rectangle (256.29,522.86);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (226.12,504.12) {Not concerned};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (226.12,494.62) {about getting};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (226.12,485.11) {seriously ill};
\end{scope}
\begin{scope}
\path[clip] (261.79,470.61) rectangle (322.14,522.86);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (261.79,470.61) rectangle (322.14,522.86);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (291.96,504.12) {Doesn't think};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (291.96,494.62) {vaccines are};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (291.96,485.11) {effective};
\end{scope}
\begin{scope}
\path[clip] (327.64,470.61) rectangle (387.98,522.86);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (327.64,470.61) rectangle (387.98,522.86);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (357.81,513.62) {Doesn't think};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (357.81,504.12) {Coronavirus};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (357.81,494.62) {outbreak is as};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (357.81,485.11) {serious as};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (357.81,475.61) {people say};
\end{scope}
\begin{scope}
\path[clip] (393.48,470.61) rectangle (453.82,522.86);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (393.48,470.61) rectangle (453.82,522.86);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (423.65,499.37) {Doesn't like};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (423.65,489.86) {needles};
\end{scope}
\begin{scope}
\path[clip] (459.32,470.61) rectangle (519.67,522.86);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (459.32,470.61) rectangle (519.67,522.86);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (489.50,499.37) {Allergic to};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (489.50,489.86) {vaccines};
\end{scope}
\begin{scope}
\path[clip] (525.17,470.61) rectangle (585.51,522.86);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (525.17,470.61) rectangle (585.51,522.86);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (555.34,504.12) {Won't have};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (555.34,494.62) {time to get};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (555.34,485.11) {vaccinated};
\end{scope}
\begin{scope}
\path[clip] (591.01,470.61) rectangle (651.36,522.86);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (591.01,470.61) rectangle (651.36,522.86);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (621.18,504.12) {Mentions a};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (621.18,494.62) {conspiracy};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (621.18,485.11) {theory};
\end{scope}
\begin{scope}
\path[clip] (656.86,470.61) rectangle (717.20,522.86);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (656.86,470.61) rectangle (717.20,522.86);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (687.03,494.62) {Other reasons};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 68.08, 22.69) --
	( 68.08, 25.44);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 81.52, 22.69) --
	( 81.52, 25.44);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 94.97, 22.69) --
	( 94.97, 25.44);

\path[draw=drawColor,line width= 0.6pt,line join=round] (108.41, 22.69) --
	(108.41, 25.44);

\path[draw=drawColor,line width= 0.6pt,line join=round] (121.86, 22.69) --
	(121.86, 25.44);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 68.08, 16.25) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 81.52, 16.25) {25};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 94.97, 16.25) {50};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (108.41, 16.25) {75};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (121.86, 16.25) {100};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (133.92, 22.69) --
	(133.92, 25.44);

\path[draw=drawColor,line width= 0.6pt,line join=round] (147.37, 22.69) --
	(147.37, 25.44);

\path[draw=drawColor,line width= 0.6pt,line join=round] (160.81, 22.69) --
	(160.81, 25.44);

\path[draw=drawColor,line width= 0.6pt,line join=round] (174.26, 22.69) --
	(174.26, 25.44);

\path[draw=drawColor,line width= 0.6pt,line join=round] (187.70, 22.69) --
	(187.70, 25.44);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (133.92, 16.25) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (147.37, 16.25) {25};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (160.81, 16.25) {50};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (174.26, 16.25) {75};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (187.70, 16.25) {100};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (199.77, 22.69) --
	(199.77, 25.44);

\path[draw=drawColor,line width= 0.6pt,line join=round] (213.21, 22.69) --
	(213.21, 25.44);

\path[draw=drawColor,line width= 0.6pt,line join=round] (226.66, 22.69) --
	(226.66, 25.44);

\path[draw=drawColor,line width= 0.6pt,line join=round] (240.10, 22.69) --
	(240.10, 25.44);

\path[draw=drawColor,line width= 0.6pt,line join=round] (253.55, 22.69) --
	(253.55, 25.44);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (199.77, 16.25) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (213.21, 16.25) {25};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (226.66, 16.25) {50};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (240.10, 16.25) {75};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (253.55, 16.25) {100};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (265.61, 22.69) --
	(265.61, 25.44);

\path[draw=drawColor,line width= 0.6pt,line join=round] (279.06, 22.69) --
	(279.06, 25.44);

\path[draw=drawColor,line width= 0.6pt,line join=round] (292.50, 22.69) --
	(292.50, 25.44);

\path[draw=drawColor,line width= 0.6pt,line join=round] (305.95, 22.69) --
	(305.95, 25.44);

\path[draw=drawColor,line width= 0.6pt,line join=round] (319.39, 22.69) --
	(319.39, 25.44);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (265.61, 16.25) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (279.06, 16.25) {25};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (292.50, 16.25) {50};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (305.95, 16.25) {75};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (319.39, 16.25) {100};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (331.45, 22.69) --
	(331.45, 25.44);

\path[draw=drawColor,line width= 0.6pt,line join=round] (344.90, 22.69) --
	(344.90, 25.44);

\path[draw=drawColor,line width= 0.6pt,line join=round] (358.34, 22.69) --
	(358.34, 25.44);

\path[draw=drawColor,line width= 0.6pt,line join=round] (371.79, 22.69) --
	(371.79, 25.44);

\path[draw=drawColor,line width= 0.6pt,line join=round] (385.24, 22.69) --
	(385.24, 25.44);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (331.45, 16.25) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (344.90, 16.25) {25};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (358.34, 16.25) {50};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (371.79, 16.25) {75};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (385.24, 16.25) {100};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (397.30, 22.69) --
	(397.30, 25.44);

\path[draw=drawColor,line width= 0.6pt,line join=round] (410.74, 22.69) --
	(410.74, 25.44);

\path[draw=drawColor,line width= 0.6pt,line join=round] (424.19, 22.69) --
	(424.19, 25.44);

\path[draw=drawColor,line width= 0.6pt,line join=round] (437.63, 22.69) --
	(437.63, 25.44);

\path[draw=drawColor,line width= 0.6pt,line join=round] (451.08, 22.69) --
	(451.08, 25.44);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (397.30, 16.25) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (410.74, 16.25) {25};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (424.19, 16.25) {50};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (437.63, 16.25) {75};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (451.08, 16.25) {100};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (463.14, 22.69) --
	(463.14, 25.44);

\path[draw=drawColor,line width= 0.6pt,line join=round] (476.59, 22.69) --
	(476.59, 25.44);

\path[draw=drawColor,line width= 0.6pt,line join=round] (490.03, 22.69) --
	(490.03, 25.44);

\path[draw=drawColor,line width= 0.6pt,line join=round] (503.48, 22.69) --
	(503.48, 25.44);

\path[draw=drawColor,line width= 0.6pt,line join=round] (516.92, 22.69) --
	(516.92, 25.44);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (463.14, 16.25) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (476.59, 16.25) {25};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (490.03, 16.25) {50};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (503.48, 16.25) {75};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (516.92, 16.25) {100};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (528.99, 22.69) --
	(528.99, 25.44);

\path[draw=drawColor,line width= 0.6pt,line join=round] (542.43, 22.69) --
	(542.43, 25.44);

\path[draw=drawColor,line width= 0.6pt,line join=round] (555.88, 22.69) --
	(555.88, 25.44);

\path[draw=drawColor,line width= 0.6pt,line join=round] (569.32, 22.69) --
	(569.32, 25.44);

\path[draw=drawColor,line width= 0.6pt,line join=round] (582.77, 22.69) --
	(582.77, 25.44);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (528.99, 16.25) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (542.43, 16.25) {25};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (555.88, 16.25) {50};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (569.32, 16.25) {75};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (582.77, 16.25) {100};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (594.83, 22.69) --
	(594.83, 25.44);

\path[draw=drawColor,line width= 0.6pt,line join=round] (608.28, 22.69) --
	(608.28, 25.44);

\path[draw=drawColor,line width= 0.6pt,line join=round] (621.72, 22.69) --
	(621.72, 25.44);

\path[draw=drawColor,line width= 0.6pt,line join=round] (635.17, 22.69) --
	(635.17, 25.44);

\path[draw=drawColor,line width= 0.6pt,line join=round] (648.61, 22.69) --
	(648.61, 25.44);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (594.83, 16.25) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (608.28, 16.25) {25};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (621.72, 16.25) {50};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (635.17, 16.25) {75};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (648.61, 16.25) {100};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (660.67, 22.69) --
	(660.67, 25.44);

\path[draw=drawColor,line width= 0.6pt,line join=round] (674.12, 22.69) --
	(674.12, 25.44);

\path[draw=drawColor,line width= 0.6pt,line join=round] (687.57, 22.69) --
	(687.57, 25.44);

\path[draw=drawColor,line width= 0.6pt,line join=round] (701.01, 22.69) --
	(701.01, 25.44);

\path[draw=drawColor,line width= 0.6pt,line join=round] (714.46, 22.69) --
	(714.46, 25.44);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (660.67, 16.25) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (674.12, 16.25) {25};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (687.57, 16.25) {50};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (701.01, 16.25) {75};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (714.46, 16.25) {100};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05, 46.88) {USA};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05, 37.38) {(n=462)};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05, 78.23) {Russia};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05, 68.73) {(n=16238)};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,104.83) {All LMICs};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,140.93) {Uganda 2};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,131.43) {(n=319)};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,172.28) {Uganda 1};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,162.78) {(n=289)};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,203.63) {Sierra Leone 2};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,194.13) {(n=254)};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,234.98) {Sierra Leone 1};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,225.48) {(n=234)};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,266.33) {Rwanda};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,256.83) {(n=70)};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,297.68) {Pakistan 1};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,288.18) {(n=441)};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,329.03) {Nigeria};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,319.53) {(n=410)};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,360.38) {Nepal};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,350.88) {(n=48)};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,391.73) {Mozambique};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,382.23) {(n=74)};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,423.08) {Colombia};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,413.58) {(n=202)};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,454.43) {Burkina Faso};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,444.93) {(n=325)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 61.51, 44.25) --
	( 64.26, 44.25);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 61.51, 75.60) --
	( 64.26, 75.60);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 61.51,106.95) --
	( 64.26,106.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 61.51,138.30) --
	( 64.26,138.30);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 61.51,169.65) --
	( 64.26,169.65);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 61.51,201.00) --
	( 64.26,201.00);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 61.51,232.35) --
	( 64.26,232.35);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 61.51,263.70) --
	( 64.26,263.70);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 61.51,295.05) --
	( 64.26,295.05);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 61.51,326.40) --
	( 64.26,326.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 61.51,357.75) --
	( 64.26,357.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 61.51,389.10) --
	( 64.26,389.10);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 61.51,420.45) --
	( 64.26,420.45);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 61.51,451.80) --
	( 64.26,451.80);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (390.73,  7.00) {estimate};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,542.02);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.32] at (  5.50,530.16) {Why would you not take the COVID-19 vaccine?};
\end{scope}
\end{tikzpicture}

\end{document}