### 2021-06-22 paper_files/figure-latex/genderhist-1.tex

% Created by tikzDevice version 0.12.3.1 on 2021-06-22 11:48:37
% !TEX encoding = UTF-8 Unicode
% !TEX program = xelatex

\documentclass[fontsize=7pt,DIV=calc]{scrartcl}

\nonstopmode

\usepackage{tikz}

\usepackage[active,tightpage,xetex]{preview}

\usepackage{fontspec,xunicode}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}

\setmainfont{Arial}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (722.70,650.43);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.00) rectangle (722.70,650.43);
\end{scope}
\begin{scope}
\path[clip] ( 41.25,548.04) rectangle (369.36,631.27);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 41.25,548.04) rectangle (369.36,631.27);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 41.25,562.24) --
	(369.36,562.24);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 41.25,583.07) --
	(369.36,583.07);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 41.25,603.90) --
	(369.36,603.90);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 41.25,624.74) --
	(369.36,624.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 41.25,551.82) --
	(369.36,551.82);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 41.25,572.66) --
	(369.36,572.66);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 41.25,593.49) --
	(369.36,593.49);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 41.25,614.32) --
	(369.36,614.32);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 68.60,548.04) --
	( 68.60,631.27);

\path[draw=drawColor,line width= 0.6pt,line join=round] (114.17,548.04) --
	(114.17,631.27);

\path[draw=drawColor,line width= 0.6pt,line join=round] (159.74,548.04) --
	(159.74,631.27);

\path[draw=drawColor,line width= 0.6pt,line join=round] (205.31,548.04) --
	(205.31,631.27);

\path[draw=drawColor,line width= 0.6pt,line join=round] (250.88,548.04) --
	(250.88,631.27);

\path[draw=drawColor,line width= 0.6pt,line join=round] (296.45,548.04) --
	(296.45,631.27);

\path[draw=drawColor,line width= 0.6pt,line join=round] (342.02,548.04) --
	(342.02,631.27);
\definecolor{fillColor}{RGB}{221,204,119}

\path[fill=fillColor] (166.57,551.82) rectangle (180.24,567.16);

\path[fill=fillColor] (257.71,551.82) rectangle (271.38,553.41);

\path[fill=fillColor] (212.14,551.82) rectangle (225.81,553.32);

\path[fill=fillColor] (303.28,551.82) rectangle (316.95,561.49);

\path[fill=fillColor] (348.85,551.82) rectangle (362.52,553.24);

\path[fill=fillColor] ( 75.43,551.82) rectangle ( 89.10,594.82);

\path[fill=fillColor] (121.00,551.82) rectangle (134.67,562.57);
\definecolor{fillColor}{RGB}{17,119,51}

\path[fill=fillColor] (152.90,551.82) rectangle (166.57,565.74);

\path[fill=fillColor] (244.04,551.82) rectangle (257.71,552.99);

\path[fill=fillColor] (198.47,551.82) rectangle (212.14,553.24);

\path[fill=fillColor] (289.61,551.82) rectangle (303.28,562.41);

\path[fill=fillColor] (335.18,551.82) rectangle (348.85,553.32);

\path[fill=fillColor] ( 61.76,551.82) rectangle ( 75.43,594.65);

\path[fill=fillColor] (107.33,551.82) rectangle (121.00,563.66);
\definecolor{fillColor}{RGB}{204,102,119}

\path[fill=fillColor] (139.23,551.82) rectangle (152.90,570.41);

\path[fill=fillColor] (230.37,551.82) rectangle (244.04,554.16);

\path[fill=fillColor] (184.80,551.82) rectangle (198.47,553.57);

\path[fill=fillColor] (275.94,551.82) rectangle (289.61,559.41);

\path[fill=fillColor] (321.51,551.82) rectangle (335.18,552.99);

\path[fill=fillColor] ( 48.09,551.82) rectangle ( 61.76,595.24);

\path[fill=fillColor] ( 93.66,551.82) rectangle (107.33,560.16);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 41.25,548.04) rectangle (369.36,631.27);
\end{scope}
\begin{scope}
\path[clip] ( 41.25,459.31) rectangle (369.36,542.54);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 41.25,459.31) rectangle (369.36,542.54);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 41.25,473.51) --
	(369.36,473.51);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 41.25,494.34) --
	(369.36,494.34);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 41.25,515.17) --
	(369.36,515.17);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 41.25,536.01) --
	(369.36,536.01);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 41.25,463.09) --
	(369.36,463.09);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 41.25,483.92) --
	(369.36,483.92);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 41.25,504.76) --
	(369.36,504.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 41.25,525.59) --
	(369.36,525.59);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 68.60,459.31) --
	( 68.60,542.54);

\path[draw=drawColor,line width= 0.6pt,line join=round] (114.17,459.31) --
	(114.17,542.54);

\path[draw=drawColor,line width= 0.6pt,line join=round] (159.74,459.31) --
	(159.74,542.54);

\path[draw=drawColor,line width= 0.6pt,line join=round] (205.31,459.31) --
	(205.31,542.54);

\path[draw=drawColor,line width= 0.6pt,line join=round] (250.88,459.31) --
	(250.88,542.54);

\path[draw=drawColor,line width= 0.6pt,line join=round] (296.45,459.31) --
	(296.45,542.54);

\path[draw=drawColor,line width= 0.6pt,line join=round] (342.02,459.31) --
	(342.02,542.54);
\definecolor{fillColor}{RGB}{221,204,119}

\path[fill=fillColor] (166.57,463.09) rectangle (180.24,492.76);

\path[fill=fillColor] (250.88,463.09) rectangle (271.38,463.43);

\path[fill=fillColor] (212.14,463.09) rectangle (225.81,476.76);

\path[fill=fillColor] (303.28,463.09) rectangle (316.95,463.51);

\path[fill=fillColor] (348.85,463.09) rectangle (362.52,465.09);

\path[fill=fillColor] ( 75.43,463.09) rectangle ( 89.10,499.59);

\path[fill=fillColor] (121.00,463.09) rectangle (134.67,463.76);
\definecolor{fillColor}{RGB}{17,119,51}

\path[fill=fillColor] (152.90,463.09) rectangle (166.57,489.84);

\path[fill=fillColor] (230.37,463.09) rectangle (250.88,463.76);

\path[fill=fillColor] (198.47,463.09) rectangle (212.14,476.92);

\path[fill=fillColor] (289.61,463.09) rectangle (303.28,463.76);

\path[fill=fillColor] (335.18,463.09) rectangle (348.85,464.09);

\path[fill=fillColor] ( 61.76,463.09) rectangle ( 75.43,502.84);

\path[fill=fillColor] (107.33,463.09) rectangle (121.00,463.84);
\definecolor{fillColor}{RGB}{204,102,119}

\path[fill=fillColor] (139.23,463.09) rectangle (152.90,496.84);

\path[fill=fillColor] (184.80,463.09) rectangle (198.47,476.42);

\path[fill=fillColor] (275.94,463.09) rectangle (289.61,463.26);

\path[fill=fillColor] (321.51,463.09) rectangle (335.18,466.43);

\path[fill=fillColor] ( 48.09,463.09) rectangle ( 61.76,495.26);

\path[fill=fillColor] ( 93.66,463.09) rectangle (107.33,463.68);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 41.25,459.31) rectangle (369.36,542.54);
\end{scope}
\begin{scope}
\path[clip] ( 41.25,370.58) rectangle (369.36,453.81);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 41.25,370.58) rectangle (369.36,453.81);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 41.25,384.78) --
	(369.36,384.78);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 41.25,405.61) --
	(369.36,405.61);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 41.25,426.44) --
	(369.36,426.44);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 41.25,447.28) --
	(369.36,447.28);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 41.25,374.36) --
	(369.36,374.36);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 41.25,395.19) --
	(369.36,395.19);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 41.25,416.03) --
	(369.36,416.03);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 41.25,436.86) --
	(369.36,436.86);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 68.60,370.58) --
	( 68.60,453.81);

\path[draw=drawColor,line width= 0.6pt,line join=round] (114.17,370.58) --
	(114.17,453.81);

\path[draw=drawColor,line width= 0.6pt,line join=round] (159.74,370.58) --
	(159.74,453.81);

\path[draw=drawColor,line width= 0.6pt,line join=round] (205.31,370.58) --
	(205.31,453.81);

\path[draw=drawColor,line width= 0.6pt,line join=round] (250.88,370.58) --
	(250.88,453.81);

\path[draw=drawColor,line width= 0.6pt,line join=round] (296.45,370.58) --
	(296.45,453.81);

\path[draw=drawColor,line width= 0.6pt,line join=round] (342.02,370.58) --
	(342.02,453.81);
\definecolor{fillColor}{RGB}{221,204,119}

\path[fill=fillColor] (166.57,374.36) rectangle (180.24,386.86);

\path[fill=fillColor] (257.71,374.36) rectangle (271.38,374.86);

\path[fill=fillColor] (212.14,374.36) rectangle (225.81,375.28);

\path[fill=fillColor] (303.28,374.36) rectangle (316.95,402.03);

\path[fill=fillColor] (342.02,374.36) rectangle (362.52,374.69);

\path[fill=fillColor] ( 75.43,374.36) rectangle ( 89.10,393.53);

\path[fill=fillColor] (121.00,374.36) rectangle (134.67,396.61);
\definecolor{fillColor}{RGB}{17,119,51}

\path[fill=fillColor] (152.90,374.36) rectangle (166.57,386.19);

\path[fill=fillColor] (244.04,374.36) rectangle (257.71,375.11);

\path[fill=fillColor] (198.47,374.36) rectangle (212.14,375.28);

\path[fill=fillColor] (289.61,374.36) rectangle (303.28,401.19);

\path[fill=fillColor] (321.51,374.36) rectangle (342.02,374.86);

\path[fill=fillColor] ( 61.76,374.36) rectangle ( 75.43,394.44);

\path[fill=fillColor] (107.33,374.36) rectangle (121.00,396.69);
\definecolor{fillColor}{RGB}{204,102,119}

\path[fill=fillColor] (139.23,374.36) rectangle (152.90,388.03);

\path[fill=fillColor] (230.37,374.36) rectangle (244.04,374.53);

\path[fill=fillColor] (184.80,374.36) rectangle (198.47,375.19);

\path[fill=fillColor] (275.94,374.36) rectangle (289.61,403.53);

\path[fill=fillColor] ( 48.09,374.36) rectangle ( 61.76,391.69);

\path[fill=fillColor] ( 93.66,374.36) rectangle (107.33,396.53);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 41.25,370.58) rectangle (369.36,453.81);
\end{scope}
\begin{scope}
\path[clip] ( 41.25,281.85) rectangle (369.36,365.08);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 41.25,281.85) rectangle (369.36,365.08);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 41.25,296.05) --
	(369.36,296.05);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 41.25,316.88) --
	(369.36,316.88);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 41.25,337.71) --
	(369.36,337.71);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 41.25,358.55) --
	(369.36,358.55);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 41.25,285.63) --
	(369.36,285.63);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 41.25,306.46) --
	(369.36,306.46);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 41.25,327.30) --
	(369.36,327.30);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 41.25,348.13) --
	(369.36,348.13);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 68.60,281.85) --
	( 68.60,365.08);

\path[draw=drawColor,line width= 0.6pt,line join=round] (114.17,281.85) --
	(114.17,365.08);

\path[draw=drawColor,line width= 0.6pt,line join=round] (159.74,281.85) --
	(159.74,365.08);

\path[draw=drawColor,line width= 0.6pt,line join=round] (205.31,281.85) --
	(205.31,365.08);

\path[draw=drawColor,line width= 0.6pt,line join=round] (250.88,281.85) --
	(250.88,365.08);

\path[draw=drawColor,line width= 0.6pt,line join=round] (296.45,281.85) --
	(296.45,365.08);

\path[draw=drawColor,line width= 0.6pt,line join=round] (342.02,281.85) --
	(342.02,365.08);
\definecolor{fillColor}{RGB}{221,204,119}

\path[fill=fillColor] (166.57,285.63) rectangle (180.24,288.21);

\path[fill=fillColor] (250.88,285.63) rectangle (271.38,285.63);

\path[fill=fillColor] (212.14,285.63) rectangle (225.81,287.05);

\path[fill=fillColor] (303.28,285.63) rectangle (316.95,290.46);

\path[fill=fillColor] ( 75.43,285.63) rectangle ( 89.10,360.05);
\definecolor{fillColor}{RGB}{17,119,51}

\path[fill=fillColor] (152.90,285.63) rectangle (166.57,287.88);

\path[fill=fillColor] (230.37,285.63) rectangle (250.88,285.71);

\path[fill=fillColor] (198.47,285.63) rectangle (212.14,287.13);

\path[fill=fillColor] (289.61,285.63) rectangle (303.28,291.13);

\path[fill=fillColor] ( 61.76,285.63) rectangle ( 75.43,359.63);
\definecolor{fillColor}{RGB}{204,102,119}

\path[fill=fillColor] (139.23,285.63) rectangle (152.90,289.30);

\path[fill=fillColor] (184.80,285.63) rectangle (198.47,286.88);

\path[fill=fillColor] (275.94,285.63) rectangle (289.61,288.38);

\path[fill=fillColor] ( 48.09,285.63) rectangle ( 61.76,361.30);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 41.25,281.85) rectangle (369.36,365.08);
\end{scope}
\begin{scope}
\path[clip] ( 41.25,193.12) rectangle (369.36,276.35);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 41.25,193.12) rectangle (369.36,276.35);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 41.25,207.32) --
	(369.36,207.32);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 41.25,228.15) --
	(369.36,228.15);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 41.25,248.98) --
	(369.36,248.98);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 41.25,269.81) --
	(369.36,269.81);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 41.25,196.90) --
	(369.36,196.90);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 41.25,217.73) --
	(369.36,217.73);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 41.25,238.57) --
	(369.36,238.57);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 41.25,259.40) --
	(369.36,259.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 68.60,193.12) --
	( 68.60,276.35);

\path[draw=drawColor,line width= 0.6pt,line join=round] (114.17,193.12) --
	(114.17,276.35);

\path[draw=drawColor,line width= 0.6pt,line join=round] (159.74,193.12) --
	(159.74,276.35);

\path[draw=drawColor,line width= 0.6pt,line join=round] (205.31,193.12) --
	(205.31,276.35);

\path[draw=drawColor,line width= 0.6pt,line join=round] (250.88,193.12) --
	(250.88,276.35);

\path[draw=drawColor,line width= 0.6pt,line join=round] (296.45,193.12) --
	(296.45,276.35);

\path[draw=drawColor,line width= 0.6pt,line join=round] (342.02,193.12) --
	(342.02,276.35);
\definecolor{fillColor}{RGB}{221,204,119}

\path[fill=fillColor] (166.57,196.90) rectangle (180.24,210.90);

\path[fill=fillColor] (257.71,196.90) rectangle (271.38,198.15);

\path[fill=fillColor] (212.14,196.90) rectangle (225.81,201.32);

\path[fill=fillColor] (303.28,196.90) rectangle (316.95,205.73);

\path[fill=fillColor] (348.85,196.90) rectangle (362.52,198.90);

\path[fill=fillColor] ( 75.43,196.90) rectangle ( 89.10,236.90);

\path[fill=fillColor] (121.00,196.90) rectangle (134.67,212.98);
\definecolor{fillColor}{RGB}{17,119,51}

\path[fill=fillColor] (152.90,196.90) rectangle (166.57,209.82);

\path[fill=fillColor] (244.04,196.90) rectangle (257.71,198.32);

\path[fill=fillColor] (198.47,196.90) rectangle (212.14,201.23);

\path[fill=fillColor] (289.61,196.90) rectangle (303.28,206.07);

\path[fill=fillColor] (335.18,196.90) rectangle (348.85,198.98);

\path[fill=fillColor] ( 61.76,196.90) rectangle ( 75.43,236.90);

\path[fill=fillColor] (107.33,196.90) rectangle (121.00,213.73);
\definecolor{fillColor}{RGB}{204,102,119}

\path[fill=fillColor] (139.23,196.90) rectangle (152.90,212.73);

\path[fill=fillColor] (230.37,196.90) rectangle (244.04,198.07);

\path[fill=fillColor] (184.80,196.90) rectangle (198.47,201.48);

\path[fill=fillColor] (275.94,196.90) rectangle (289.61,204.98);

\path[fill=fillColor] (321.51,196.90) rectangle (335.18,198.82);

\path[fill=fillColor] ( 48.09,196.90) rectangle ( 61.76,236.65);

\path[fill=fillColor] ( 93.66,196.90) rectangle (107.33,211.82);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 41.25,193.12) rectangle (369.36,276.35);
\end{scope}
\begin{scope}
\path[clip] ( 41.25,104.39) rectangle (369.36,187.62);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 41.25,104.39) rectangle (369.36,187.62);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 41.25,118.59) --
	(369.36,118.59);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 41.25,139.42) --
	(369.36,139.42);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 41.25,160.25) --
	(369.36,160.25);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 41.25,181.08) --
	(369.36,181.08);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 41.25,108.17) --
	(369.36,108.17);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 41.25,129.00) --
	(369.36,129.00);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 41.25,149.84) --
	(369.36,149.84);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 41.25,170.67) --
	(369.36,170.67);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 68.60,104.39) --
	( 68.60,187.62);

\path[draw=drawColor,line width= 0.6pt,line join=round] (114.17,104.39) --
	(114.17,187.62);

\path[draw=drawColor,line width= 0.6pt,line join=round] (159.74,104.39) --
	(159.74,187.62);

\path[draw=drawColor,line width= 0.6pt,line join=round] (205.31,104.39) --
	(205.31,187.62);

\path[draw=drawColor,line width= 0.6pt,line join=round] (250.88,104.39) --
	(250.88,187.62);

\path[draw=drawColor,line width= 0.6pt,line join=round] (296.45,104.39) --
	(296.45,187.62);

\path[draw=drawColor,line width= 0.6pt,line join=round] (342.02,104.39) --
	(342.02,187.62);
\definecolor{fillColor}{RGB}{221,204,119}

\path[fill=fillColor] (166.57,108.17) rectangle (180.24,117.75);

\path[fill=fillColor] (212.14,108.17) rectangle (225.81,110.17);

\path[fill=fillColor] (303.28,108.17) rectangle (316.95,126.42);

\path[fill=fillColor] (342.02,108.17) rectangle (362.52,108.25);

\path[fill=fillColor] ( 75.43,108.17) rectangle ( 89.10,136.92);

\path[fill=fillColor] (121.00,108.17) rectangle (134.67,132.92);
\definecolor{fillColor}{RGB}{17,119,51}

\path[fill=fillColor] (152.90,108.17) rectangle (166.57,117.09);

\path[fill=fillColor] (198.47,108.17) rectangle (212.14,109.84);

\path[fill=fillColor] (289.61,108.17) rectangle (303.28,128.50);

\path[fill=fillColor] (321.51,108.17) rectangle (342.02,108.25);

\path[fill=fillColor] ( 61.76,108.17) rectangle ( 75.43,135.50);

\path[fill=fillColor] (107.33,108.17) rectangle (121.00,133.17);
\definecolor{fillColor}{RGB}{204,102,119}

\path[fill=fillColor] (139.23,108.17) rectangle (152.90,118.42);

\path[fill=fillColor] (184.80,108.17) rectangle (198.47,110.42);

\path[fill=fillColor] (275.94,108.17) rectangle (289.61,124.25);

\path[fill=fillColor] ( 48.09,108.17) rectangle ( 61.76,138.42);

\path[fill=fillColor] ( 93.66,108.17) rectangle (107.33,132.67);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 41.25,104.39) rectangle (369.36,187.62);
\end{scope}
\begin{scope}
\path[clip] (389.10,548.04) rectangle (717.20,631.27);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (389.10,548.04) rectangle (717.20,631.27);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] (389.10,562.24) --
	(717.20,562.24);

\path[draw=drawColor,line width= 0.3pt,line join=round] (389.10,583.07) --
	(717.20,583.07);

\path[draw=drawColor,line width= 0.3pt,line join=round] (389.10,603.90) --
	(717.20,603.90);

\path[draw=drawColor,line width= 0.3pt,line join=round] (389.10,624.74) --
	(717.20,624.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] (389.10,551.82) --
	(717.20,551.82);

\path[draw=drawColor,line width= 0.6pt,line join=round] (389.10,572.66) --
	(717.20,572.66);

\path[draw=drawColor,line width= 0.6pt,line join=round] (389.10,593.49) --
	(717.20,593.49);

\path[draw=drawColor,line width= 0.6pt,line join=round] (389.10,614.32) --
	(717.20,614.32);

\path[draw=drawColor,line width= 0.6pt,line join=round] (416.44,548.04) --
	(416.44,631.27);

\path[draw=drawColor,line width= 0.6pt,line join=round] (462.01,548.04) --
	(462.01,631.27);

\path[draw=drawColor,line width= 0.6pt,line join=round] (507.58,548.04) --
	(507.58,631.27);

\path[draw=drawColor,line width= 0.6pt,line join=round] (553.15,548.04) --
	(553.15,631.27);

\path[draw=drawColor,line width= 0.6pt,line join=round] (598.72,548.04) --
	(598.72,631.27);

\path[draw=drawColor,line width= 0.6pt,line join=round] (644.29,548.04) --
	(644.29,631.27);

\path[draw=drawColor,line width= 0.6pt,line join=round] (689.86,548.04) --
	(689.86,631.27);
\definecolor{fillColor}{RGB}{221,204,119}

\path[fill=fillColor] (514.41,551.82) rectangle (528.09,582.32);

\path[fill=fillColor] (605.55,551.82) rectangle (619.22,553.24);

\path[fill=fillColor] (559.98,551.82) rectangle (573.65,553.32);

\path[fill=fillColor] (696.69,551.82) rectangle (710.36,559.24);

\path[fill=fillColor] (423.27,551.82) rectangle (436.95,584.57);

\path[fill=fillColor] (468.84,551.82) rectangle (482.52,561.49);
\definecolor{fillColor}{RGB}{17,119,51}

\path[fill=fillColor] (500.74,551.82) rectangle (514.41,582.24);

\path[fill=fillColor] (591.88,551.82) rectangle (605.55,553.99);

\path[fill=fillColor] (546.31,551.82) rectangle (559.98,553.24);

\path[fill=fillColor] (683.02,551.82) rectangle (696.69,560.32);

\path[fill=fillColor] (409.60,551.82) rectangle (423.27,581.74);

\path[fill=fillColor] (455.17,551.82) rectangle (468.84,562.74);
\definecolor{fillColor}{RGB}{204,102,119}

\path[fill=fillColor] (487.07,551.82) rectangle (500.74,582.40);

\path[fill=fillColor] (578.21,551.82) rectangle (591.88,552.74);

\path[fill=fillColor] (532.64,551.82) rectangle (546.31,553.41);

\path[fill=fillColor] (669.35,551.82) rectangle (683.02,558.66);

\path[fill=fillColor] (395.93,551.82) rectangle (409.60,586.32);

\path[fill=fillColor] (441.50,551.82) rectangle (455.17,560.82);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (389.10,548.04) rectangle (717.20,631.27);
\end{scope}
\begin{scope}
\path[clip] (389.10,459.31) rectangle (717.20,542.54);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (389.10,459.31) rectangle (717.20,542.54);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] (389.10,473.51) --
	(717.20,473.51);

\path[draw=drawColor,line width= 0.3pt,line join=round] (389.10,494.34) --
	(717.20,494.34);

\path[draw=drawColor,line width= 0.3pt,line join=round] (389.10,515.17) --
	(717.20,515.17);

\path[draw=drawColor,line width= 0.3pt,line join=round] (389.10,536.01) --
	(717.20,536.01);

\path[draw=drawColor,line width= 0.6pt,line join=round] (389.10,463.09) --
	(717.20,463.09);

\path[draw=drawColor,line width= 0.6pt,line join=round] (389.10,483.92) --
	(717.20,483.92);

\path[draw=drawColor,line width= 0.6pt,line join=round] (389.10,504.76) --
	(717.20,504.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (389.10,525.59) --
	(717.20,525.59);

\path[draw=drawColor,line width= 0.6pt,line join=round] (416.44,459.31) --
	(416.44,542.54);

\path[draw=drawColor,line width= 0.6pt,line join=round] (462.01,459.31) --
	(462.01,542.54);

\path[draw=drawColor,line width= 0.6pt,line join=round] (507.58,459.31) --
	(507.58,542.54);

\path[draw=drawColor,line width= 0.6pt,line join=round] (553.15,459.31) --
	(553.15,542.54);

\path[draw=drawColor,line width= 0.6pt,line join=round] (598.72,459.31) --
	(598.72,542.54);

\path[draw=drawColor,line width= 0.6pt,line join=round] (644.29,459.31) --
	(644.29,542.54);

\path[draw=drawColor,line width= 0.6pt,line join=round] (689.86,459.31) --
	(689.86,542.54);
\definecolor{fillColor}{RGB}{221,204,119}

\path[fill=fillColor] (514.41,463.09) rectangle (528.09,469.76);

\path[fill=fillColor] (559.98,463.09) rectangle (573.65,469.76);

\path[fill=fillColor] (651.12,463.09) rectangle (664.79,466.34);

\path[fill=fillColor] (696.69,463.09) rectangle (710.36,466.59);

\path[fill=fillColor] (423.27,463.09) rectangle (436.95,511.42);

\path[fill=fillColor] (468.84,463.09) rectangle (482.52,478.09);
\definecolor{fillColor}{RGB}{17,119,51}

\path[fill=fillColor] (500.74,463.09) rectangle (514.41,468.26);

\path[fill=fillColor] (546.31,463.09) rectangle (559.98,469.18);

\path[fill=fillColor] (637.45,463.09) rectangle (651.12,466.93);

\path[fill=fillColor] (683.02,463.09) rectangle (696.69,466.76);

\path[fill=fillColor] (409.60,463.09) rectangle (423.27,512.17);

\path[fill=fillColor] (455.17,463.09) rectangle (468.84,478.59);
\definecolor{fillColor}{RGB}{204,102,119}

\path[fill=fillColor] (487.07,463.09) rectangle (500.74,472.18);

\path[fill=fillColor] (532.64,463.09) rectangle (546.31,470.59);

\path[fill=fillColor] (623.78,463.09) rectangle (637.45,465.43);

\path[fill=fillColor] (669.35,463.09) rectangle (683.02,466.26);

\path[fill=fillColor] (395.93,463.09) rectangle (409.60,510.09);

\path[fill=fillColor] (441.50,463.09) rectangle (455.17,477.34);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (389.10,459.31) rectangle (717.20,542.54);
\end{scope}
\begin{scope}
\path[clip] (389.10,370.58) rectangle (717.20,453.81);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (389.10,370.58) rectangle (717.20,453.81);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] (389.10,384.78) --
	(717.20,384.78);

\path[draw=drawColor,line width= 0.3pt,line join=round] (389.10,405.61) --
	(717.20,405.61);

\path[draw=drawColor,line width= 0.3pt,line join=round] (389.10,426.44) --
	(717.20,426.44);

\path[draw=drawColor,line width= 0.3pt,line join=round] (389.10,447.28) --
	(717.20,447.28);

\path[draw=drawColor,line width= 0.6pt,line join=round] (389.10,374.36) --
	(717.20,374.36);

\path[draw=drawColor,line width= 0.6pt,line join=round] (389.10,395.19) --
	(717.20,395.19);

\path[draw=drawColor,line width= 0.6pt,line join=round] (389.10,416.03) --
	(717.20,416.03);

\path[draw=drawColor,line width= 0.6pt,line join=round] (389.10,436.86) --
	(717.20,436.86);

\path[draw=drawColor,line width= 0.6pt,line join=round] (416.44,370.58) --
	(416.44,453.81);

\path[draw=drawColor,line width= 0.6pt,line join=round] (462.01,370.58) --
	(462.01,453.81);

\path[draw=drawColor,line width= 0.6pt,line join=round] (507.58,370.58) --
	(507.58,453.81);

\path[draw=drawColor,line width= 0.6pt,line join=round] (553.15,370.58) --
	(553.15,453.81);

\path[draw=drawColor,line width= 0.6pt,line join=round] (598.72,370.58) --
	(598.72,453.81);

\path[draw=drawColor,line width= 0.6pt,line join=round] (644.29,370.58) --
	(644.29,453.81);

\path[draw=drawColor,line width= 0.6pt,line join=round] (689.86,370.58) --
	(689.86,453.81);
\definecolor{fillColor}{RGB}{221,204,119}

\path[fill=fillColor] (514.41,374.36) rectangle (528.09,381.36);

\path[fill=fillColor] (605.55,374.36) rectangle (619.22,374.78);

\path[fill=fillColor] (559.98,374.36) rectangle (573.65,378.28);

\path[fill=fillColor] (651.12,374.36) rectangle (664.79,381.78);

\path[fill=fillColor] (696.69,374.36) rectangle (710.36,375.28);

\path[fill=fillColor] (423.27,374.36) rectangle (436.95,411.03);

\path[fill=fillColor] (468.84,374.36) rectangle (482.52,401.44);
\definecolor{fillColor}{RGB}{17,119,51}

\path[fill=fillColor] (500.74,374.36) rectangle (514.41,380.19);

\path[fill=fillColor] (591.88,374.36) rectangle (605.55,374.61);

\path[fill=fillColor] (546.31,374.36) rectangle (559.98,377.94);

\path[fill=fillColor] (637.45,374.36) rectangle (651.12,382.11);

\path[fill=fillColor] (683.02,374.36) rectangle (696.69,375.44);

\path[fill=fillColor] (409.60,374.36) rectangle (423.27,410.28);

\path[fill=fillColor] (455.17,374.36) rectangle (468.84,403.36);
\definecolor{fillColor}{RGB}{204,102,119}

\path[fill=fillColor] (487.07,374.36) rectangle (500.74,383.44);

\path[fill=fillColor] (578.21,374.36) rectangle (591.88,375.11);

\path[fill=fillColor] (532.64,374.36) rectangle (546.31,378.86);

\path[fill=fillColor] (623.78,374.36) rectangle (637.45,381.11);

\path[fill=fillColor] (669.35,374.36) rectangle (683.02,375.11);

\path[fill=fillColor] (395.93,374.36) rectangle (409.60,412.36);

\path[fill=fillColor] (441.50,374.36) rectangle (455.17,398.03);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (389.10,370.58) rectangle (717.20,453.81);
\end{scope}
\begin{scope}
\path[clip] (389.10,281.85) rectangle (717.20,365.08);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (389.10,281.85) rectangle (717.20,365.08);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] (389.10,296.05) --
	(717.20,296.05);

\path[draw=drawColor,line width= 0.3pt,line join=round] (389.10,316.88) --
	(717.20,316.88);

\path[draw=drawColor,line width= 0.3pt,line join=round] (389.10,337.71) --
	(717.20,337.71);

\path[draw=drawColor,line width= 0.3pt,line join=round] (389.10,358.55) --
	(717.20,358.55);

\path[draw=drawColor,line width= 0.6pt,line join=round] (389.10,285.63) --
	(717.20,285.63);

\path[draw=drawColor,line width= 0.6pt,line join=round] (389.10,306.46) --
	(717.20,306.46);

\path[draw=drawColor,line width= 0.6pt,line join=round] (389.10,327.30) --
	(717.20,327.30);

\path[draw=drawColor,line width= 0.6pt,line join=round] (389.10,348.13) --
	(717.20,348.13);

\path[draw=drawColor,line width= 0.6pt,line join=round] (416.44,281.85) --
	(416.44,365.08);

\path[draw=drawColor,line width= 0.6pt,line join=round] (462.01,281.85) --
	(462.01,365.08);

\path[draw=drawColor,line width= 0.6pt,line join=round] (507.58,281.85) --
	(507.58,365.08);

\path[draw=drawColor,line width= 0.6pt,line join=round] (553.15,281.85) --
	(553.15,365.08);

\path[draw=drawColor,line width= 0.6pt,line join=round] (598.72,281.85) --
	(598.72,365.08);

\path[draw=drawColor,line width= 0.6pt,line join=round] (644.29,281.85) --
	(644.29,365.08);

\path[draw=drawColor,line width= 0.6pt,line join=round] (689.86,281.85) --
	(689.86,365.08);
\definecolor{fillColor}{RGB}{221,204,119}

\path[fill=fillColor] (514.41,285.63) rectangle (528.09,293.55);

\path[fill=fillColor] (605.55,285.63) rectangle (619.22,290.30);

\path[fill=fillColor] (559.98,285.63) rectangle (573.65,291.63);

\path[fill=fillColor] (651.12,285.63) rectangle (664.79,294.13);

\path[fill=fillColor] (696.69,285.63) rectangle (710.36,285.71);

\path[fill=fillColor] (423.27,285.63) rectangle (436.95,314.80);

\path[fill=fillColor] (468.84,285.63) rectangle (482.52,312.63);
\definecolor{fillColor}{RGB}{17,119,51}

\path[fill=fillColor] (500.74,285.63) rectangle (514.41,293.05);

\path[fill=fillColor] (591.88,285.63) rectangle (605.55,290.80);

\path[fill=fillColor] (546.31,285.63) rectangle (559.98,291.55);

\path[fill=fillColor] (637.45,285.63) rectangle (651.12,294.63);

\path[fill=fillColor] (683.02,285.63) rectangle (696.69,285.71);

\path[fill=fillColor] (409.60,285.63) rectangle (423.27,314.05);

\path[fill=fillColor] (455.17,285.63) rectangle (468.84,312.96);
\definecolor{fillColor}{RGB}{204,102,119}

\path[fill=fillColor] (487.07,285.63) rectangle (500.74,293.96);

\path[fill=fillColor] (578.21,285.63) rectangle (591.88,289.71);

\path[fill=fillColor] (532.64,285.63) rectangle (546.31,291.71);

\path[fill=fillColor] (623.78,285.63) rectangle (637.45,293.63);

\path[fill=fillColor] (669.35,285.63) rectangle (683.02,285.71);

\path[fill=fillColor] (395.93,285.63) rectangle (409.60,315.63);

\path[fill=fillColor] (441.50,285.63) rectangle (455.17,312.30);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (389.10,281.85) rectangle (717.20,365.08);
\end{scope}
\begin{scope}
\path[clip] (389.10,193.12) rectangle (717.20,276.35);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (389.10,193.12) rectangle (717.20,276.35);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] (389.10,207.32) --
	(717.20,207.32);

\path[draw=drawColor,line width= 0.3pt,line join=round] (389.10,228.15) --
	(717.20,228.15);

\path[draw=drawColor,line width= 0.3pt,line join=round] (389.10,248.98) --
	(717.20,248.98);

\path[draw=drawColor,line width= 0.3pt,line join=round] (389.10,269.81) --
	(717.20,269.81);

\path[draw=drawColor,line width= 0.6pt,line join=round] (389.10,196.90) --
	(717.20,196.90);

\path[draw=drawColor,line width= 0.6pt,line join=round] (389.10,217.73) --
	(717.20,217.73);

\path[draw=drawColor,line width= 0.6pt,line join=round] (389.10,238.57) --
	(717.20,238.57);

\path[draw=drawColor,line width= 0.6pt,line join=round] (389.10,259.40) --
	(717.20,259.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] (416.44,193.12) --
	(416.44,276.35);

\path[draw=drawColor,line width= 0.6pt,line join=round] (462.01,193.12) --
	(462.01,276.35);

\path[draw=drawColor,line width= 0.6pt,line join=round] (507.58,193.12) --
	(507.58,276.35);

\path[draw=drawColor,line width= 0.6pt,line join=round] (553.15,193.12) --
	(553.15,276.35);

\path[draw=drawColor,line width= 0.6pt,line join=round] (598.72,193.12) --
	(598.72,276.35);

\path[draw=drawColor,line width= 0.6pt,line join=round] (644.29,193.12) --
	(644.29,276.35);

\path[draw=drawColor,line width= 0.6pt,line join=round] (689.86,193.12) --
	(689.86,276.35);
\definecolor{fillColor}{RGB}{221,204,119}

\path[fill=fillColor] (514.41,196.90) rectangle (528.09,220.32);

\path[fill=fillColor] (605.55,196.90) rectangle (619.22,200.98);

\path[fill=fillColor] (559.98,196.90) rectangle (573.65,198.73);

\path[fill=fillColor] (651.12,196.90) rectangle (664.79,210.73);

\path[fill=fillColor] (423.27,196.90) rectangle (436.95,226.90);

\path[fill=fillColor] (468.84,196.90) rectangle (482.52,207.15);
\definecolor{fillColor}{RGB}{17,119,51}

\path[fill=fillColor] (500.74,196.90) rectangle (514.41,219.65);

\path[fill=fillColor] (591.88,196.90) rectangle (605.55,200.82);

\path[fill=fillColor] (546.31,196.90) rectangle (559.98,198.90);

\path[fill=fillColor] (637.45,196.90) rectangle (651.12,209.65);

\path[fill=fillColor] (409.60,196.90) rectangle (423.27,227.40);

\path[fill=fillColor] (455.17,196.90) rectangle (468.84,208.23);
\definecolor{fillColor}{RGB}{204,102,119}

\path[fill=fillColor] (487.07,196.90) rectangle (500.74,220.90);

\path[fill=fillColor] (578.21,196.90) rectangle (591.88,201.15);

\path[fill=fillColor] (532.64,196.90) rectangle (546.31,198.48);

\path[fill=fillColor] (623.78,196.90) rectangle (637.45,211.82);

\path[fill=fillColor] (395.93,196.90) rectangle (409.60,226.40);

\path[fill=fillColor] (441.50,196.90) rectangle (455.17,205.98);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (389.10,193.12) rectangle (717.20,276.35);
\end{scope}
\begin{scope}
\path[clip] (374.86,548.04) rectangle (389.10,631.27);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (374.86,548.04) rectangle (389.10,631.27);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (384.10,589.65) {Colombia};
\end{scope}
\begin{scope}
\path[clip] (374.86,459.31) rectangle (389.10,542.54);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (374.86,459.31) rectangle (389.10,542.54);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (384.10,500.92) {Nigeria};
\end{scope}
\begin{scope}
\path[clip] (374.86,370.58) rectangle (389.10,453.81);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (374.86,370.58) rectangle (389.10,453.81);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (384.10,412.19) {Sierra Leone 1};
\end{scope}
\begin{scope}
\path[clip] (374.86,281.85) rectangle (389.10,365.08);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (374.86,281.85) rectangle (389.10,365.08);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (384.10,323.46) {Uganda 2};
\end{scope}
\begin{scope}
\path[clip] (374.86,193.12) rectangle (389.10,276.35);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (374.86,193.12) rectangle (389.10,276.35);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (384.10,234.73) {Russia};
\end{scope}
\begin{scope}
\path[clip] ( 27.01,548.04) rectangle ( 41.25,631.27);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 27.01,548.04) rectangle ( 41.25,631.27);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 36.26,589.65) {Burkina Faso};
\end{scope}
\begin{scope}
\path[clip] ( 27.01,459.31) rectangle ( 41.25,542.54);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 27.01,459.31) rectangle ( 41.25,542.54);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 36.26,500.92) {Nepal};
\end{scope}
\begin{scope}
\path[clip] ( 27.01,370.58) rectangle ( 41.25,453.81);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 27.01,370.58) rectangle ( 41.25,453.81);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 36.26,412.19) {Rwanda};
\end{scope}
\begin{scope}
\path[clip] ( 27.01,281.85) rectangle ( 41.25,365.08);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 27.01,281.85) rectangle ( 41.25,365.08);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 36.26,323.46) {Sierra Leone 2};
\end{scope}
\begin{scope}
\path[clip] ( 27.01,193.12) rectangle ( 41.25,276.35);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 27.01,193.12) rectangle ( 41.25,276.35);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 36.26,234.73) {All LMICs};
\end{scope}
\begin{scope}
\path[clip] ( 27.01,104.39) rectangle ( 41.25,187.62);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 27.01,104.39) rectangle ( 41.25,187.62);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 36.26,146.00) {USA};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 68.60,101.64) --
	( 68.60,104.39);

\path[draw=drawColor,line width= 0.6pt,line join=round] (114.17,101.64) --
	(114.17,104.39);

\path[draw=drawColor,line width= 0.6pt,line join=round] (159.74,101.64) --
	(159.74,104.39);

\path[draw=drawColor,line width= 0.6pt,line join=round] (205.31,101.64) --
	(205.31,104.39);

\path[draw=drawColor,line width= 0.6pt,line join=round] (250.88,101.64) --
	(250.88,104.39);

\path[draw=drawColor,line width= 0.6pt,line join=round] (296.45,101.64) --
	(296.45,104.39);

\path[draw=drawColor,line width= 0.6pt,line join=round] (342.02,101.64) --
	(342.02,104.39);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 70.72, 99.44) {Health workers};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (111.53, 99.44) {Government or};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (121.04, 99.44) {MoH};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (157.10, 99.44) {Family or};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (166.61, 99.44) {Friends};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (193.17, 99.44) {Famous person,};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (202.67, 99.44) {religious leader};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (212.18, 99.44) {or traditional};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (221.68, 99.44) {healers};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (243.49, 99.44) {Newspapers,};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (253.00, 99.44) {radio or online};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (262.50, 99.44) {groups};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (298.57, 99.44) {Other};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (339.38, 99.44) {Don't know or};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (348.89, 99.44) {Refuse};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (416.44,190.37) --
	(416.44,193.12);

\path[draw=drawColor,line width= 0.6pt,line join=round] (462.01,190.37) --
	(462.01,193.12);

\path[draw=drawColor,line width= 0.6pt,line join=round] (507.58,190.37) --
	(507.58,193.12);

\path[draw=drawColor,line width= 0.6pt,line join=round] (553.15,190.37) --
	(553.15,193.12);

\path[draw=drawColor,line width= 0.6pt,line join=round] (598.72,190.37) --
	(598.72,193.12);

\path[draw=drawColor,line width= 0.6pt,line join=round] (644.29,190.37) --
	(644.29,193.12);

\path[draw=drawColor,line width= 0.6pt,line join=round] (689.86,190.37) --
	(689.86,193.12);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (418.56,188.17) {Health workers};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (459.38,188.17) {Government or};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (468.88,188.17) {MoH};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (504.95,188.17) {Family or};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (514.45,188.17) {Friends};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (541.01,188.17) {Famous person,};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (550.52,188.17) {religious leader};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (560.02,188.17) {or traditional};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (569.53,188.17) {healers};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (591.34,188.17) {Newspapers,};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (600.84,188.17) {radio or online};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (610.34,188.17) {groups};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (646.41,188.17) {Other};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (687.23,188.17) {Don't know or};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (696.73,188.17) {Refuse};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,549.70) {0};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,570.53) {25};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,591.37) {50};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,612.20) {75};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.26,551.82) --
	( 27.01,551.82);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.26,572.66) --
	( 27.01,572.66);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.26,593.49) --
	( 27.01,593.49);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.26,614.32) --
	( 27.01,614.32);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,460.97) {0};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,481.80) {25};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,502.64) {50};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,523.47) {75};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.26,463.09) --
	( 27.01,463.09);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.26,483.92) --
	( 27.01,483.92);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.26,504.76) --
	( 27.01,504.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.26,525.59) --
	( 27.01,525.59);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,372.24) {0};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,393.07) {25};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,413.91) {50};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,434.74) {75};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.26,374.36) --
	( 27.01,374.36);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.26,395.19) --
	( 27.01,395.19);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.26,416.03) --
	( 27.01,416.03);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.26,436.86) --
	( 27.01,436.86);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,283.51) {0};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,304.34) {25};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,325.17) {50};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,346.01) {75};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.26,285.63) --
	( 27.01,285.63);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.26,306.46) --
	( 27.01,306.46);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.26,327.30) --
	( 27.01,327.30);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.26,348.13) --
	( 27.01,348.13);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,194.78) {0};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,215.61) {25};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,236.44) {50};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,257.28) {75};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.26,196.90) --
	( 27.01,196.90);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.26,217.73) --
	( 27.01,217.73);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.26,238.57) --
	( 27.01,238.57);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.26,259.40) --
	( 27.01,259.40);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,106.05) {0};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,126.88) {25};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,147.71) {50};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,168.55) {75};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.26,108.17) --
	( 27.01,108.17);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.26,129.00) --
	( 27.01,129.00);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.26,149.84) --
	( 27.01,149.84);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.26,170.67) --
	( 27.01,170.67);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 10.80,367.83) {estimate};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (298.62,  5.50) rectangle (459.84, 30.95);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.10] at (304.12, 15.58) {Answer};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (337.84, 11.00) rectangle (352.29, 25.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{fillColor}{RGB}{204,102,119}

\path[fill=fillColor] (338.55, 11.71) rectangle (351.58, 24.74);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (384.80, 11.00) rectangle (399.25, 25.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{fillColor}{RGB}{17,119,51}

\path[fill=fillColor] (385.51, 11.71) rectangle (398.54, 24.74);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (425.21, 11.00) rectangle (439.66, 25.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{fillColor}{RGB}{221,204,119}

\path[fill=fillColor] (425.92, 11.71) rectangle (438.95, 24.74);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (357.79, 16.11) {Female};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (404.75, 16.11) {Male};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (445.16, 16.11) {All};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.32] at (  5.50,638.57) {Which of the following people would you trust MOST to help you decide whether you would get a COVID-19 vaccine?};
\end{scope}
\end{tikzpicture}

\end{document}