### 2021-06-22 paper_files/figure-latex/mainfigure1-1.tex

% Created by tikzDevice version 0.12.3.1 on 2021-06-22 11:48:04
% !TEX encoding = UTF-8 Unicode
% !TEX program = xelatex

\documentclass[fontsize=7pt,DIV=calc]{scrartcl}

\nonstopmode

\usepackage{tikz}

\usepackage[active,tightpage,xetex]{preview}

\usepackage{fontspec,xunicode}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}

\setmainfont{Arial}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (650.43,542.02);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.5pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.00) rectangle (650.43,542.02);
\end{scope}
\begin{scope}
\path[clip] (123.21, 72.04) rectangle (250.02,511.66);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (123.21, 72.04) rectangle (250.02,511.66);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] (143.39, 72.04) --
	(143.39,511.66);

\path[draw=drawColor,line width= 0.3pt,line join=round] (172.21, 72.04) --
	(172.21,511.66);

\path[draw=drawColor,line width= 0.3pt,line join=round] (201.03, 72.04) --
	(201.03,511.66);

\path[draw=drawColor,line width= 0.3pt,line join=round] (229.84, 72.04) --
	(229.84,511.66);

\path[draw=drawColor,line width= 0.5pt,line join=round] (123.21, 87.37) --
	(250.02, 87.37);

\path[draw=drawColor,line width= 0.5pt,line join=round] (123.21,112.93) --
	(250.02,112.93);

\path[draw=drawColor,line width= 0.5pt,line join=round] (123.21,138.49) --
	(250.02,138.49);

\path[draw=drawColor,line width= 0.5pt,line join=round] (123.21,164.05) --
	(250.02,164.05);

\path[draw=drawColor,line width= 0.5pt,line join=round] (123.21,189.61) --
	(250.02,189.61);

\path[draw=drawColor,line width= 0.5pt,line join=round] (123.21,215.17) --
	(250.02,215.17);

\path[draw=drawColor,line width= 0.5pt,line join=round] (123.21,240.73) --
	(250.02,240.73);

\path[draw=drawColor,line width= 0.5pt,line join=round] (123.21,266.29) --
	(250.02,266.29);

\path[draw=drawColor,line width= 0.5pt,line join=round] (123.21,291.85) --
	(250.02,291.85);

\path[draw=drawColor,line width= 0.5pt,line join=round] (123.21,317.41) --
	(250.02,317.41);

\path[draw=drawColor,line width= 0.5pt,line join=round] (123.21,342.97) --
	(250.02,342.97);

\path[draw=drawColor,line width= 0.5pt,line join=round] (123.21,368.53) --
	(250.02,368.53);

\path[draw=drawColor,line width= 0.5pt,line join=round] (123.21,394.09) --
	(250.02,394.09);

\path[draw=drawColor,line width= 0.5pt,line join=round] (123.21,419.65) --
	(250.02,419.65);

\path[draw=drawColor,line width= 0.5pt,line join=round] (123.21,445.21) --
	(250.02,445.21);

\path[draw=drawColor,line width= 0.5pt,line join=round] (123.21,470.77) --
	(250.02,470.77);

\path[draw=drawColor,line width= 0.5pt,line join=round] (123.21,496.33) --
	(250.02,496.33);

\path[draw=drawColor,line width= 0.5pt,line join=round] (128.98, 72.04) --
	(128.98,511.66);

\path[draw=drawColor,line width= 0.5pt,line join=round] (157.80, 72.04) --
	(157.80,511.66);

\path[draw=drawColor,line width= 0.5pt,line join=round] (186.62, 72.04) --
	(186.62,511.66);

\path[draw=drawColor,line width= 0.5pt,line join=round] (215.44, 72.04) --
	(215.44,511.66);

\path[draw=drawColor,line width= 0.5pt,line join=round] (244.25, 72.04) --
	(244.25,511.66);
\definecolor{drawColor}{RGB}{204,102,119}

\path[draw=drawColor,line width= 0.6pt,line join=round] (231.57,135.93) --
	(231.57,141.05);

\path[draw=drawColor,line width= 0.6pt,line join=round] (231.57,138.49) --
	(207.25,138.49);

\path[draw=drawColor,line width= 0.6pt,line join=round] (207.25,135.93) --
	(207.25,141.05);

\path[draw=drawColor,line width= 0.6pt,line join=round] (227.65,161.49) --
	(227.65,166.61);

\path[draw=drawColor,line width= 0.6pt,line join=round] (227.65,164.05) --
	(215.32,164.05);

\path[draw=drawColor,line width= 0.6pt,line join=round] (215.32,161.49) --
	(215.32,166.61);

\path[draw=drawColor,line width= 0.6pt,line join=round] (209.10,493.77) --
	(209.10,498.88);

\path[draw=drawColor,line width= 0.6pt,line join=round] (209.10,496.33) --
	(202.18,496.33);

\path[draw=drawColor,line width= 0.6pt,line join=round] (202.18,493.77) --
	(202.18,498.88);

\path[draw=drawColor,line width= 0.6pt,line join=round] (218.43,468.21) --
	(218.43,473.32);

\path[draw=drawColor,line width= 0.6pt,line join=round] (218.43,470.77) --
	(212.21,470.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (212.21,468.21) --
	(212.21,473.32);

\path[draw=drawColor,line width= 0.6pt,line join=round] (228.46,442.65) --
	(228.46,447.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (228.46,445.21) --
	(223.85,445.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] (223.85,442.65) --
	(223.85,447.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (234.69,417.09) --
	(234.69,422.20);

\path[draw=drawColor,line width= 0.6pt,line join=round] (234.69,419.65) --
	(228.69,419.65);

\path[draw=drawColor,line width= 0.6pt,line join=round] (228.69,417.09) --
	(228.69,422.20);

\path[draw=drawColor,line width= 0.6pt,line join=round] (241.49,391.53) --
	(241.49,396.64);

\path[draw=drawColor,line width= 0.6pt,line join=round] (241.49,394.09) --
	(239.07,394.09);

\path[draw=drawColor,line width= 0.6pt,line join=round] (239.07,391.53) --
	(239.07,396.64);

\path[draw=drawColor,line width= 0.6pt,line join=round] (219.12,365.97) --
	(219.12,371.08);

\path[draw=drawColor,line width= 0.6pt,line join=round] (219.12,368.53) --
	(214.63,368.53);

\path[draw=drawColor,line width= 0.6pt,line join=round] (214.63,365.97) --
	(214.63,371.08);

\path[draw=drawColor,line width= 0.6pt,line join=round] (223.85,340.41) --
	(223.85,345.52);

\path[draw=drawColor,line width= 0.6pt,line join=round] (223.85,342.97) --
	(209.67,342.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (209.67,340.41) --
	(209.67,345.52);

\path[draw=drawColor,line width= 0.6pt,line join=round] (208.40,314.85) --
	(208.40,319.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (208.40,317.41) --
	(202.87,317.41);

\path[draw=drawColor,line width= 0.6pt,line join=round] (202.87,314.85) --
	(202.87,319.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (165.52,110.37) --
	(165.52,115.49);

\path[draw=drawColor,line width= 0.6pt,line join=round] (165.52,112.93) --
	(162.52,112.93);

\path[draw=drawColor,line width= 0.6pt,line join=round] (162.52,110.37) --
	(162.52,115.49);

\path[draw=drawColor,line width= 0.6pt,line join=round] (229.04,289.29) --
	(229.04,294.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] (229.04,291.85) --
	(224.54,291.85);

\path[draw=drawColor,line width= 0.6pt,line join=round] (224.54,289.29) --
	(224.54,294.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] (221.78,263.73) --
	(221.78,268.84);

\path[draw=drawColor,line width= 0.6pt,line join=round] (221.78,266.29) --
	(216.01,266.29);

\path[draw=drawColor,line width= 0.6pt,line join=round] (216.01,263.73) --
	(216.01,268.84);

\path[draw=drawColor,line width= 0.6pt,line join=round] (232.27,238.17) --
	(232.27,243.28);

\path[draw=drawColor,line width= 0.6pt,line join=round] (232.27,240.73) --
	(228.35,240.73);

\path[draw=drawColor,line width= 0.6pt,line join=round] (228.35,238.17) --
	(228.35,243.28);

\path[draw=drawColor,line width= 0.6pt,line join=round] (229.50,212.61) --
	(229.50,217.73);

\path[draw=drawColor,line width= 0.6pt,line join=round] (229.50,215.17) --
	(226.27,215.17);

\path[draw=drawColor,line width= 0.6pt,line join=round] (226.27,212.61) --
	(226.27,217.73);

\path[draw=drawColor,line width= 0.6pt,line join=round] (219.70,187.05) --
	(219.70,192.17);

\path[draw=drawColor,line width= 0.6pt,line join=round] (219.70,189.61) --
	(214.63,189.61);

\path[draw=drawColor,line width= 0.6pt,line join=round] (214.63,187.05) --
	(214.63,192.17);

\path[draw=drawColor,line width= 0.6pt,line join=round] (206.56, 84.82) --
	(206.56, 89.93);

\path[draw=drawColor,line width= 0.6pt,line join=round] (206.56, 87.37) --
	(200.22, 87.37);

\path[draw=drawColor,line width= 0.6pt,line join=round] (200.22, 84.82) --
	(200.22, 89.93);
\definecolor{fillColor}{RGB}{204,102,119}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (219.35,138.49) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (221.55,164.05) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (205.64,496.33) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (215.32,470.77) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (226.16,445.21) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (231.69,419.65) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (240.34,394.09) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (216.82,368.53) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (216.70,342.97) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (205.64,317.41) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (164.02,112.93) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (226.85,291.85) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (218.89,266.29) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (230.31,240.73) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (227.89,215.17) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (217.16,189.61) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (203.45, 87.37) circle (  1.96);
\definecolor{drawColor}{RGB}{169,169,169}

\path[draw=drawColor,line width= 0.6pt,line join=round] (123.21,176.83) -- (250.02,176.83);

\path[draw=drawColor,line width= 0.6pt,line join=round] (123.21,151.27) -- (250.02,151.27);

\path[draw=drawColor,line width= 0.6pt,line join=round] (123.21,125.71) -- (250.02,125.71);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.5pt,line join=round,line cap=round] (123.21, 72.04) rectangle (250.02,511.66);
\end{scope}
\begin{scope}
\path[clip] (255.02, 72.04) rectangle (381.82,511.66);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (255.02, 72.04) rectangle (381.82,511.66);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] (275.19, 72.04) --
	(275.19,511.66);

\path[draw=drawColor,line width= 0.3pt,line join=round] (304.01, 72.04) --
	(304.01,511.66);

\path[draw=drawColor,line width= 0.3pt,line join=round] (332.83, 72.04) --
	(332.83,511.66);

\path[draw=drawColor,line width= 0.3pt,line join=round] (361.65, 72.04) --
	(361.65,511.66);

\path[draw=drawColor,line width= 0.5pt,line join=round] (255.02, 87.37) --
	(381.82, 87.37);

\path[draw=drawColor,line width= 0.5pt,line join=round] (255.02,112.93) --
	(381.82,112.93);

\path[draw=drawColor,line width= 0.5pt,line join=round] (255.02,138.49) --
	(381.82,138.49);

\path[draw=drawColor,line width= 0.5pt,line join=round] (255.02,164.05) --
	(381.82,164.05);

\path[draw=drawColor,line width= 0.5pt,line join=round] (255.02,189.61) --
	(381.82,189.61);

\path[draw=drawColor,line width= 0.5pt,line join=round] (255.02,215.17) --
	(381.82,215.17);

\path[draw=drawColor,line width= 0.5pt,line join=round] (255.02,240.73) --
	(381.82,240.73);

\path[draw=drawColor,line width= 0.5pt,line join=round] (255.02,266.29) --
	(381.82,266.29);

\path[draw=drawColor,line width= 0.5pt,line join=round] (255.02,291.85) --
	(381.82,291.85);

\path[draw=drawColor,line width= 0.5pt,line join=round] (255.02,317.41) --
	(381.82,317.41);

\path[draw=drawColor,line width= 0.5pt,line join=round] (255.02,342.97) --
	(381.82,342.97);

\path[draw=drawColor,line width= 0.5pt,line join=round] (255.02,368.53) --
	(381.82,368.53);

\path[draw=drawColor,line width= 0.5pt,line join=round] (255.02,394.09) --
	(381.82,394.09);

\path[draw=drawColor,line width= 0.5pt,line join=round] (255.02,419.65) --
	(381.82,419.65);

\path[draw=drawColor,line width= 0.5pt,line join=round] (255.02,445.21) --
	(381.82,445.21);

\path[draw=drawColor,line width= 0.5pt,line join=round] (255.02,470.77) --
	(381.82,470.77);

\path[draw=drawColor,line width= 0.5pt,line join=round] (255.02,496.33) --
	(381.82,496.33);

\path[draw=drawColor,line width= 0.5pt,line join=round] (260.78, 72.04) --
	(260.78,511.66);

\path[draw=drawColor,line width= 0.5pt,line join=round] (289.60, 72.04) --
	(289.60,511.66);

\path[draw=drawColor,line width= 0.5pt,line join=round] (318.42, 72.04) --
	(318.42,511.66);

\path[draw=drawColor,line width= 0.5pt,line join=round] (347.24, 72.04) --
	(347.24,511.66);

\path[draw=drawColor,line width= 0.5pt,line join=round] (376.06, 72.04) --
	(376.06,511.66);
\definecolor{drawColor}{RGB}{136,34,85}

\path[draw=drawColor,line width= 0.6pt,line join=round] (361.65,141.05) --
	(361.65,143.60);

\path[draw=drawColor,line width= 0.6pt,line join=round] (361.65,142.32) --
	(334.10,142.32);

\path[draw=drawColor,line width= 0.6pt,line join=round] (334.10,141.05) --
	(334.10,143.60);
\definecolor{drawColor}{RGB}{153,153,51}

\path[draw=drawColor,line width= 0.6pt,line join=round] (364.99,133.38) --
	(364.99,135.93);

\path[draw=drawColor,line width= 0.6pt,line join=round] (364.99,134.66) --
	(341.71,134.66);

\path[draw=drawColor,line width= 0.6pt,line join=round] (341.71,133.38) --
	(341.71,135.93);
\definecolor{drawColor}{RGB}{136,34,85}

\path[draw=drawColor,line width= 0.6pt,line join=round] (358.77,166.61) --
	(358.77,169.16);

\path[draw=drawColor,line width= 0.6pt,line join=round] (358.77,167.88) --
	(345.39,167.88);

\path[draw=drawColor,line width= 0.6pt,line join=round] (345.39,166.61) --
	(345.39,169.16);
\definecolor{drawColor}{RGB}{153,153,51}

\path[draw=drawColor,line width= 0.6pt,line join=round] (362.11,158.94) --
	(362.11,161.49);

\path[draw=drawColor,line width= 0.6pt,line join=round] (362.11,160.22) --
	(350.01,160.22);

\path[draw=drawColor,line width= 0.6pt,line join=round] (350.01,158.94) --
	(350.01,161.49);
\definecolor{drawColor}{RGB}{136,34,85}

\path[draw=drawColor,line width= 0.6pt,line join=round] (339.05,498.88) --
	(339.05,501.44);

\path[draw=drawColor,line width= 0.6pt,line join=round] (339.05,500.16) --
	(325.68,500.16);

\path[draw=drawColor,line width= 0.6pt,line join=round] (325.68,498.88) --
	(325.68,501.44);
\definecolor{drawColor}{RGB}{153,153,51}

\path[draw=drawColor,line width= 0.6pt,line join=round] (343.67,491.21) --
	(343.67,493.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (343.67,492.49) --
	(335.71,492.49);

\path[draw=drawColor,line width= 0.6pt,line join=round] (335.71,491.21) --
	(335.71,493.77);
\definecolor{drawColor}{RGB}{136,34,85}

\path[draw=drawColor,line width= 0.6pt,line join=round] (349.54,473.32) --
	(349.54,475.88);

\path[draw=drawColor,line width= 0.6pt,line join=round] (349.54,474.60) --
	(341.59,474.60);

\path[draw=drawColor,line width= 0.6pt,line join=round] (341.59,473.32) --
	(341.59,475.88);
\definecolor{drawColor}{RGB}{153,153,51}

\path[draw=drawColor,line width= 0.6pt,line join=round] (354.96,465.65) --
	(354.96,468.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] (354.96,466.93) --
	(344.93,466.93);

\path[draw=drawColor,line width= 0.6pt,line join=round] (344.93,465.65) --
	(344.93,468.21);
\definecolor{drawColor}{RGB}{136,34,85}

\path[draw=drawColor,line width= 0.6pt,line join=round] (359.69,447.76) --
	(359.69,450.32);

\path[draw=drawColor,line width= 0.6pt,line join=round] (359.69,449.04) --
	(351.85,449.04);

\path[draw=drawColor,line width= 0.6pt,line join=round] (351.85,447.76) --
	(351.85,450.32);
\definecolor{drawColor}{RGB}{153,153,51}

\path[draw=drawColor,line width= 0.6pt,line join=round] (361.07,440.09) --
	(361.07,442.65);

\path[draw=drawColor,line width= 0.6pt,line join=round] (361.07,441.37) --
	(355.89,441.37);

\path[draw=drawColor,line width= 0.6pt,line join=round] (355.89,440.09) --
	(355.89,442.65);
\definecolor{drawColor}{RGB}{136,34,85}

\path[draw=drawColor,line width= 0.6pt,line join=round] (364.53,422.20) --
	(364.53,424.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (364.53,423.48) --
	(355.89,423.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] (355.89,422.20) --
	(355.89,424.76);
\definecolor{drawColor}{RGB}{153,153,51}

\path[draw=drawColor,line width= 0.6pt,line join=round] (369.26,414.53) --
	(369.26,417.09);

\path[draw=drawColor,line width= 0.6pt,line join=round] (369.26,415.81) --
	(362.69,415.81);

\path[draw=drawColor,line width= 0.6pt,line join=round] (362.69,414.53) --
	(362.69,417.09);
\definecolor{drawColor}{RGB}{136,34,85}

\path[draw=drawColor,line width= 0.6pt,line join=round] (373.98,396.64) --
	(373.98,399.20);

\path[draw=drawColor,line width= 0.6pt,line join=round] (373.98,397.92) --
	(369.83,397.92);

\path[draw=drawColor,line width= 0.6pt,line join=round] (369.83,396.64) --
	(369.83,399.20);
\definecolor{drawColor}{RGB}{153,153,51}

\path[draw=drawColor,line width= 0.6pt,line join=round] (373.41,388.98) --
	(373.41,391.53);

\path[draw=drawColor,line width= 0.6pt,line join=round] (373.41,390.25) --
	(370.41,390.25);

\path[draw=drawColor,line width= 0.6pt,line join=round] (370.41,388.98) --
	(370.41,391.53);
\definecolor{drawColor}{RGB}{136,34,85}

\path[draw=drawColor,line width= 0.6pt,line join=round] (350.81,371.08) --
	(350.81,373.64);

\path[draw=drawColor,line width= 0.6pt,line join=round] (350.81,372.36) --
	(343.44,372.36);

\path[draw=drawColor,line width= 0.6pt,line join=round] (343.44,371.08) --
	(343.44,373.64);
\definecolor{drawColor}{RGB}{153,153,51}

\path[draw=drawColor,line width= 0.6pt,line join=round] (352.31,363.42) --
	(352.31,365.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (352.31,364.69) --
	(346.78,364.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (346.78,363.42) --
	(346.78,365.97);
\definecolor{drawColor}{RGB}{136,34,85}

\path[draw=drawColor,line width= 0.6pt,line join=round] (351.62,345.52) --
	(351.62,348.08);

\path[draw=drawColor,line width= 0.6pt,line join=round] (351.62,346.80) --
	(336.40,346.80);

\path[draw=drawColor,line width= 0.6pt,line join=round] (336.40,345.52) --
	(336.40,348.08);
\definecolor{drawColor}{RGB}{153,153,51}

\path[draw=drawColor,line width= 0.6pt,line join=round] (360.38,337.86) --
	(360.38,340.41);

\path[draw=drawColor,line width= 0.6pt,line join=round] (360.38,339.13) --
	(345.86,339.13);

\path[draw=drawColor,line width= 0.6pt,line join=round] (345.86,337.86) --
	(345.86,340.41);
\definecolor{drawColor}{RGB}{136,34,85}

\path[draw=drawColor,line width= 0.6pt,line join=round] (288.68,115.49) --
	(288.68,118.04);

\path[draw=drawColor,line width= 0.6pt,line join=round] (288.68,116.76) --
	(284.87,116.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (284.87,115.49) --
	(284.87,118.04);
\definecolor{drawColor}{RGB}{153,153,51}

\path[draw=drawColor,line width= 0.6pt,line join=round] (307.47,107.82) --
	(307.47,110.37);

\path[draw=drawColor,line width= 0.6pt,line join=round] (307.47,109.10) --
	(302.86,109.10);

\path[draw=drawColor,line width= 0.6pt,line join=round] (302.86,107.82) --
	(302.86,110.37);
\definecolor{drawColor}{RGB}{136,34,85}

\path[draw=drawColor,line width= 0.6pt,line join=round] (356.46,294.40) --
	(356.46,296.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (356.46,295.68) --
	(348.16,295.68);

\path[draw=drawColor,line width= 0.6pt,line join=round] (348.16,294.40) --
	(348.16,296.96);
\definecolor{drawColor}{RGB}{153,153,51}

\path[draw=drawColor,line width= 0.6pt,line join=round] (364.76,286.74) --
	(364.76,289.29);

\path[draw=drawColor,line width= 0.6pt,line join=round] (364.76,288.01) --
	(359.69,288.01);

\path[draw=drawColor,line width= 0.6pt,line join=round] (359.69,286.74) --
	(359.69,289.29);
\definecolor{drawColor}{RGB}{136,34,85}

\path[draw=drawColor,line width= 0.6pt,line join=round] (351.50,268.84) --
	(351.50,271.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] (351.50,270.12) --
	(340.90,270.12);

\path[draw=drawColor,line width= 0.6pt,line join=round] (340.90,268.84) --
	(340.90,271.40);
\definecolor{drawColor}{RGB}{153,153,51}

\path[draw=drawColor,line width= 0.6pt,line join=round] (356.58,261.18) --
	(356.58,263.73);

\path[draw=drawColor,line width= 0.6pt,line join=round] (356.58,262.45) --
	(349.78,262.45);

\path[draw=drawColor,line width= 0.6pt,line join=round] (349.78,261.18) --
	(349.78,263.73);
\definecolor{drawColor}{RGB}{136,34,85}

\path[draw=drawColor,line width= 0.6pt,line join=round] (366.26,243.28) --
	(366.26,245.84);

\path[draw=drawColor,line width= 0.6pt,line join=round] (366.26,244.56) --
	(359.57,244.56);

\path[draw=drawColor,line width= 0.6pt,line join=round] (359.57,243.28) --
	(359.57,245.84);
\definecolor{drawColor}{RGB}{153,153,51}

\path[draw=drawColor,line width= 0.6pt,line join=round] (363.95,235.62) --
	(363.95,238.17);

\path[draw=drawColor,line width= 0.6pt,line join=round] (363.95,236.90) --
	(359.80,236.90);

\path[draw=drawColor,line width= 0.6pt,line join=round] (359.80,235.62) --
	(359.80,238.17);
\definecolor{drawColor}{RGB}{136,34,85}

\path[draw=drawColor,line width= 0.6pt,line join=round] (361.30,212.61) --
	(361.30,217.73);

\path[draw=drawColor,line width= 0.6pt,line join=round] (361.30,215.17) --
	(358.08,215.17);

\path[draw=drawColor,line width= 0.6pt,line join=round] (358.08,212.61) --
	(358.08,217.73);

\path[draw=drawColor,line width= 0.6pt,line join=round] (351.04,192.17) --
	(351.04,194.72);

\path[draw=drawColor,line width= 0.6pt,line join=round] (351.04,193.44) --
	(343.20,193.44);

\path[draw=drawColor,line width= 0.6pt,line join=round] (343.20,192.17) --
	(343.20,194.72);
\definecolor{drawColor}{RGB}{153,153,51}

\path[draw=drawColor,line width= 0.6pt,line join=round] (354.04,184.50) --
	(354.04,187.05);

\path[draw=drawColor,line width= 0.6pt,line join=round] (354.04,185.78) --
	(347.47,185.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (347.47,184.50) --
	(347.47,187.05);
\definecolor{drawColor}{RGB}{136,34,85}

\path[draw=drawColor,line width= 0.6pt,line join=round] (330.06, 89.93) --
	(330.06, 92.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] (330.06, 91.21) --
	(320.84, 91.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] (320.84, 89.93) --
	(320.84, 92.48);
\definecolor{drawColor}{RGB}{153,153,51}

\path[draw=drawColor,line width= 0.6pt,line join=round] (349.43, 82.26) --
	(349.43, 84.82);

\path[draw=drawColor,line width= 0.6pt,line join=round] (349.43, 83.54) --
	(341.24, 83.54);

\path[draw=drawColor,line width= 0.6pt,line join=round] (341.24, 82.26) --
	(341.24, 84.82);
\definecolor{drawColor}{RGB}{136,34,85}
\definecolor{fillColor}{RGB}{136,34,85}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (347.82,142.32) circle (  1.96);
\definecolor{drawColor}{RGB}{153,153,51}
\definecolor{fillColor}{RGB}{153,153,51}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (353.35,134.66) circle (  1.96);
\definecolor{drawColor}{RGB}{136,34,85}
\definecolor{fillColor}{RGB}{136,34,85}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (352.08,167.88) circle (  1.96);
\definecolor{drawColor}{RGB}{153,153,51}
\definecolor{fillColor}{RGB}{153,153,51}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (356.00,160.22) circle (  1.96);
\definecolor{drawColor}{RGB}{136,34,85}
\definecolor{fillColor}{RGB}{136,34,85}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (332.37,500.16) circle (  1.96);
\definecolor{drawColor}{RGB}{153,153,51}
\definecolor{fillColor}{RGB}{153,153,51}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (339.63,492.49) circle (  1.96);
\definecolor{drawColor}{RGB}{136,34,85}
\definecolor{fillColor}{RGB}{136,34,85}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (345.51,474.60) circle (  1.96);
\definecolor{drawColor}{RGB}{153,153,51}
\definecolor{fillColor}{RGB}{153,153,51}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (349.89,466.93) circle (  1.96);
\definecolor{drawColor}{RGB}{136,34,85}
\definecolor{fillColor}{RGB}{136,34,85}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (355.77,449.04) circle (  1.96);
\definecolor{drawColor}{RGB}{153,153,51}
\definecolor{fillColor}{RGB}{153,153,51}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (358.42,441.37) circle (  1.96);
\definecolor{drawColor}{RGB}{136,34,85}
\definecolor{fillColor}{RGB}{136,34,85}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (360.15,423.48) circle (  1.96);
\definecolor{drawColor}{RGB}{153,153,51}
\definecolor{fillColor}{RGB}{153,153,51}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (366.03,415.81) circle (  1.96);
\definecolor{drawColor}{RGB}{136,34,85}
\definecolor{fillColor}{RGB}{136,34,85}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (371.91,397.92) circle (  1.96);
\definecolor{drawColor}{RGB}{153,153,51}
\definecolor{fillColor}{RGB}{153,153,51}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (371.91,390.25) circle (  1.96);
\definecolor{drawColor}{RGB}{136,34,85}
\definecolor{fillColor}{RGB}{136,34,85}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (347.12,372.36) circle (  1.96);
\definecolor{drawColor}{RGB}{153,153,51}
\definecolor{fillColor}{RGB}{153,153,51}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (349.54,364.69) circle (  1.96);
\definecolor{drawColor}{RGB}{136,34,85}
\definecolor{fillColor}{RGB}{136,34,85}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (344.01,346.80) circle (  1.96);
\definecolor{drawColor}{RGB}{153,153,51}
\definecolor{fillColor}{RGB}{153,153,51}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (353.12,339.13) circle (  1.96);
\definecolor{drawColor}{RGB}{136,34,85}
\definecolor{fillColor}{RGB}{136,34,85}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (286.83,116.76) circle (  1.96);
\definecolor{drawColor}{RGB}{153,153,51}
\definecolor{fillColor}{RGB}{153,153,51}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (305.16,109.10) circle (  1.96);
\definecolor{drawColor}{RGB}{136,34,85}
\definecolor{fillColor}{RGB}{136,34,85}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (352.31,295.68) circle (  1.96);
\definecolor{drawColor}{RGB}{153,153,51}
\definecolor{fillColor}{RGB}{153,153,51}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (362.23,288.01) circle (  1.96);
\definecolor{drawColor}{RGB}{136,34,85}
\definecolor{fillColor}{RGB}{136,34,85}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (346.20,270.12) circle (  1.96);
\definecolor{drawColor}{RGB}{153,153,51}
\definecolor{fillColor}{RGB}{153,153,51}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (353.12,262.45) circle (  1.96);
\definecolor{drawColor}{RGB}{136,34,85}
\definecolor{fillColor}{RGB}{136,34,85}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (362.92,244.56) circle (  1.96);
\definecolor{drawColor}{RGB}{153,153,51}
\definecolor{fillColor}{RGB}{153,153,51}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (361.88,236.90) circle (  1.96);
\definecolor{drawColor}{RGB}{136,34,85}
\definecolor{fillColor}{RGB}{136,34,85}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (359.69,215.17) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (347.12,193.44) circle (  1.96);
\definecolor{drawColor}{RGB}{153,153,51}
\definecolor{fillColor}{RGB}{153,153,51}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (350.70,185.78) circle (  1.96);
\definecolor{drawColor}{RGB}{136,34,85}
\definecolor{fillColor}{RGB}{136,34,85}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (325.45, 91.21) circle (  1.96);
\definecolor{drawColor}{RGB}{153,153,51}
\definecolor{fillColor}{RGB}{153,153,51}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (345.39, 83.54) circle (  1.96);
\definecolor{drawColor}{RGB}{169,169,169}

\path[draw=drawColor,line width= 0.6pt,line join=round] (255.02,176.83) -- (381.82,176.83);

\path[draw=drawColor,line width= 0.6pt,line join=round] (255.02,151.27) -- (381.82,151.27);

\path[draw=drawColor,line width= 0.6pt,line join=round] (255.02,125.71) -- (381.82,125.71);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.5pt,line join=round,line cap=round] (255.02, 72.04) rectangle (381.82,511.66);
\end{scope}
\begin{scope}
\path[clip] (386.82, 72.04) rectangle (513.63,511.66);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (386.82, 72.04) rectangle (513.63,511.66);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] (407.00, 72.04) --
	(407.00,511.66);

\path[draw=drawColor,line width= 0.3pt,line join=round] (435.81, 72.04) --
	(435.81,511.66);

\path[draw=drawColor,line width= 0.3pt,line join=round] (464.63, 72.04) --
	(464.63,511.66);

\path[draw=drawColor,line width= 0.3pt,line join=round] (493.45, 72.04) --
	(493.45,511.66);

\path[draw=drawColor,line width= 0.5pt,line join=round] (386.82, 87.37) --
	(513.63, 87.37);

\path[draw=drawColor,line width= 0.5pt,line join=round] (386.82,112.93) --
	(513.63,112.93);

\path[draw=drawColor,line width= 0.5pt,line join=round] (386.82,138.49) --
	(513.63,138.49);

\path[draw=drawColor,line width= 0.5pt,line join=round] (386.82,164.05) --
	(513.63,164.05);

\path[draw=drawColor,line width= 0.5pt,line join=round] (386.82,189.61) --
	(513.63,189.61);

\path[draw=drawColor,line width= 0.5pt,line join=round] (386.82,215.17) --
	(513.63,215.17);

\path[draw=drawColor,line width= 0.5pt,line join=round] (386.82,240.73) --
	(513.63,240.73);

\path[draw=drawColor,line width= 0.5pt,line join=round] (386.82,266.29) --
	(513.63,266.29);

\path[draw=drawColor,line width= 0.5pt,line join=round] (386.82,291.85) --
	(513.63,291.85);

\path[draw=drawColor,line width= 0.5pt,line join=round] (386.82,317.41) --
	(513.63,317.41);

\path[draw=drawColor,line width= 0.5pt,line join=round] (386.82,342.97) --
	(513.63,342.97);

\path[draw=drawColor,line width= 0.5pt,line join=round] (386.82,368.53) --
	(513.63,368.53);

\path[draw=drawColor,line width= 0.5pt,line join=round] (386.82,394.09) --
	(513.63,394.09);

\path[draw=drawColor,line width= 0.5pt,line join=round] (386.82,419.65) --
	(513.63,419.65);

\path[draw=drawColor,line width= 0.5pt,line join=round] (386.82,445.21) --
	(513.63,445.21);

\path[draw=drawColor,line width= 0.5pt,line join=round] (386.82,470.77) --
	(513.63,470.77);

\path[draw=drawColor,line width= 0.5pt,line join=round] (386.82,496.33) --
	(513.63,496.33);

\path[draw=drawColor,line width= 0.5pt,line join=round] (392.59, 72.04) --
	(392.59,511.66);

\path[draw=drawColor,line width= 0.5pt,line join=round] (421.41, 72.04) --
	(421.41,511.66);

\path[draw=drawColor,line width= 0.5pt,line join=round] (450.22, 72.04) --
	(450.22,511.66);

\path[draw=drawColor,line width= 0.5pt,line join=round] (479.04, 72.04) --
	(479.04,511.66);

\path[draw=drawColor,line width= 0.5pt,line join=round] (507.86, 72.04) --
	(507.86,511.66);
\definecolor{drawColor}{RGB}{170,68,153}

\path[draw=drawColor,line width= 0.6pt,line join=round] (493.22,133.38) --
	(493.22,135.93);

\path[draw=drawColor,line width= 0.6pt,line join=round] (493.22,134.66) --
	(464.17,134.66);

\path[draw=drawColor,line width= 0.6pt,line join=round] (464.17,133.38) --
	(464.17,135.93);
\definecolor{drawColor}{RGB}{68,170,153}

\path[draw=drawColor,line width= 0.6pt,line join=round] (496.22,141.05) --
	(496.22,143.60);

\path[draw=drawColor,line width= 0.6pt,line join=round] (496.22,142.32) --
	(473.05,142.32);

\path[draw=drawColor,line width= 0.6pt,line join=round] (473.05,141.05) --
	(473.05,143.60);
\definecolor{drawColor}{RGB}{170,68,153}

\path[draw=drawColor,line width= 0.6pt,line join=round] (488.73,158.94) --
	(488.73,161.49);

\path[draw=drawColor,line width= 0.6pt,line join=round] (488.73,160.22) --
	(474.89,160.22);

\path[draw=drawColor,line width= 0.6pt,line join=round] (474.89,158.94) --
	(474.89,161.49);
\definecolor{drawColor}{RGB}{68,170,153}

\path[draw=drawColor,line width= 0.6pt,line join=round] (491.03,166.61) --
	(491.03,169.16);

\path[draw=drawColor,line width= 0.6pt,line join=round] (491.03,167.88) --
	(478.01,167.88);

\path[draw=drawColor,line width= 0.6pt,line join=round] (478.01,166.61) --
	(478.01,169.16);
\definecolor{drawColor}{RGB}{170,68,153}

\path[draw=drawColor,line width= 0.6pt,line join=round] (468.44,491.21) --
	(468.44,493.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (468.44,492.49) --
	(457.03,492.49);

\path[draw=drawColor,line width= 0.6pt,line join=round] (457.03,491.21) --
	(457.03,493.77);
\definecolor{drawColor}{RGB}{68,170,153}

\path[draw=drawColor,line width= 0.6pt,line join=round] (477.66,498.88) --
	(477.66,501.44);

\path[draw=drawColor,line width= 0.6pt,line join=round] (477.66,500.16) --
	(469.13,500.16);

\path[draw=drawColor,line width= 0.6pt,line join=round] (469.13,498.88) --
	(469.13,501.44);
\definecolor{drawColor}{RGB}{170,68,153}

\path[draw=drawColor,line width= 0.6pt,line join=round] (487.69,465.65) --
	(487.69,468.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] (487.69,466.93) --
	(477.43,466.93);

\path[draw=drawColor,line width= 0.6pt,line join=round] (477.43,465.65) --
	(477.43,468.21);
\definecolor{drawColor}{RGB}{68,170,153}

\path[draw=drawColor,line width= 0.6pt,line join=round] (481.12,473.32) --
	(481.12,475.88);

\path[draw=drawColor,line width= 0.6pt,line join=round] (481.12,474.60) --
	(473.39,474.60);

\path[draw=drawColor,line width= 0.6pt,line join=round] (473.39,473.32) --
	(473.39,475.88);
\definecolor{drawColor}{RGB}{170,68,153}

\path[draw=drawColor,line width= 0.6pt,line join=round] (501.64,440.09) --
	(501.64,442.65);

\path[draw=drawColor,line width= 0.6pt,line join=round] (501.64,441.37) --
	(486.08,441.37);

\path[draw=drawColor,line width= 0.6pt,line join=round] (486.08,440.09) --
	(486.08,442.65);
\definecolor{drawColor}{RGB}{68,170,153}

\path[draw=drawColor,line width= 0.6pt,line join=round] (495.41,447.76) --
	(495.41,450.32);

\path[draw=drawColor,line width= 0.6pt,line join=round] (495.41,449.04) --
	(487.69,449.04);

\path[draw=drawColor,line width= 0.6pt,line join=round] (487.69,447.76) --
	(487.69,450.32);
\definecolor{drawColor}{RGB}{170,68,153}

\path[draw=drawColor,line width= 0.6pt,line join=round] (496.80,414.53) --
	(496.80,417.09);

\path[draw=drawColor,line width= 0.6pt,line join=round] (496.80,415.81) --
	(486.88,415.81);

\path[draw=drawColor,line width= 0.6pt,line join=round] (486.88,414.53) --
	(486.88,417.09);
\definecolor{drawColor}{RGB}{68,170,153}

\path[draw=drawColor,line width= 0.6pt,line join=round] (499.56,422.20) --
	(499.56,424.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (499.56,423.48) --
	(492.30,423.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] (492.30,422.20) --
	(492.30,424.76);
\definecolor{drawColor}{RGB}{170,68,153}

\path[draw=drawColor,line width= 0.6pt,line join=round] (496.91,337.86) --
	(496.91,340.41);

\path[draw=drawColor,line width= 0.6pt,line join=round] (496.91,339.13) --
	(480.89,339.13);

\path[draw=drawColor,line width= 0.6pt,line join=round] (480.89,337.86) --
	(480.89,340.41);
\definecolor{drawColor}{RGB}{68,170,153}

\path[draw=drawColor,line width= 0.6pt,line join=round] (485.38,345.52) --
	(485.38,348.08);

\path[draw=drawColor,line width= 0.6pt,line join=round] (485.38,346.80) --
	(470.28,346.80);

\path[draw=drawColor,line width= 0.6pt,line join=round] (470.28,345.52) --
	(470.28,348.08);
\definecolor{drawColor}{RGB}{170,68,153}

\path[draw=drawColor,line width= 0.6pt,line join=round] (479.62,312.30) --
	(479.62,314.85);

\path[draw=drawColor,line width= 0.6pt,line join=round] (479.62,313.57) --
	(470.17,313.57);

\path[draw=drawColor,line width= 0.6pt,line join=round] (470.17,312.30) --
	(470.17,314.85);
\definecolor{drawColor}{RGB}{68,170,153}

\path[draw=drawColor,line width= 0.6pt,line join=round] (469.94,319.96) --
	(469.94,322.52);

\path[draw=drawColor,line width= 0.6pt,line join=round] (469.94,321.24) --
	(463.14,321.24);

\path[draw=drawColor,line width= 0.6pt,line join=round] (463.14,319.96) --
	(463.14,322.52);
\definecolor{drawColor}{RGB}{170,68,153}

\path[draw=drawColor,line width= 0.6pt,line join=round] (430.05,107.82) --
	(430.05,110.37);

\path[draw=drawColor,line width= 0.6pt,line join=round] (430.05,109.10) --
	(426.71,109.10);

\path[draw=drawColor,line width= 0.6pt,line join=round] (426.71,107.82) --
	(426.71,110.37);
\definecolor{drawColor}{RGB}{68,170,153}

\path[draw=drawColor,line width= 0.6pt,line join=round] (429.47,115.49) --
	(429.47,118.04);

\path[draw=drawColor,line width= 0.6pt,line join=round] (429.47,116.76) --
	(424.06,116.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (424.06,115.49) --
	(424.06,118.04);
\definecolor{drawColor}{RGB}{170,68,153}

\path[draw=drawColor,line width= 0.6pt,line join=round] (481.58,286.74) --
	(481.58,289.29);

\path[draw=drawColor,line width= 0.6pt,line join=round] (481.58,288.01) --
	(468.09,288.01);

\path[draw=drawColor,line width= 0.6pt,line join=round] (468.09,286.74) --
	(468.09,289.29);
\definecolor{drawColor}{RGB}{68,170,153}

\path[draw=drawColor,line width= 0.6pt,line join=round] (495.99,294.40) --
	(495.99,296.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (495.99,295.68) --
	(491.49,295.68);

\path[draw=drawColor,line width= 0.6pt,line join=round] (491.49,294.40) --
	(491.49,296.96);
\definecolor{drawColor}{RGB}{170,68,153}

\path[draw=drawColor,line width= 0.6pt,line join=round] (483.31,261.18) --
	(483.31,263.73);

\path[draw=drawColor,line width= 0.6pt,line join=round] (483.31,262.45) --
	(473.39,262.45);

\path[draw=drawColor,line width= 0.6pt,line join=round] (473.39,261.18) --
	(473.39,263.73);
\definecolor{drawColor}{RGB}{68,170,153}

\path[draw=drawColor,line width= 0.6pt,line join=round] (488.61,268.84) --
	(488.61,271.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] (488.61,270.12) --
	(481.35,270.12);

\path[draw=drawColor,line width= 0.6pt,line join=round] (481.35,268.84) --
	(481.35,271.40);
\definecolor{drawColor}{RGB}{170,68,153}

\path[draw=drawColor,line width= 0.6pt,line join=round] (499.22,235.62) --
	(499.22,238.17);

\path[draw=drawColor,line width= 0.6pt,line join=round] (499.22,236.90) --
	(490.57,236.90);

\path[draw=drawColor,line width= 0.6pt,line join=round] (490.57,235.62) --
	(490.57,238.17);
\definecolor{drawColor}{RGB}{68,170,153}

\path[draw=drawColor,line width= 0.6pt,line join=round] (495.87,243.28) --
	(495.87,245.84);

\path[draw=drawColor,line width= 0.6pt,line join=round] (495.87,244.56) --
	(491.72,244.56);

\path[draw=drawColor,line width= 0.6pt,line join=round] (491.72,243.28) --
	(491.72,245.84);
\definecolor{drawColor}{RGB}{170,68,153}

\path[draw=drawColor,line width= 0.6pt,line join=round] (491.61,210.06) --
	(491.61,212.61);

\path[draw=drawColor,line width= 0.6pt,line join=round] (491.61,211.34) --
	(478.35,211.34);

\path[draw=drawColor,line width= 0.6pt,line join=round] (478.35,210.06) --
	(478.35,212.61);
\definecolor{drawColor}{RGB}{68,170,153}

\path[draw=drawColor,line width= 0.6pt,line join=round] (492.30,217.73) --
	(492.30,220.28);

\path[draw=drawColor,line width= 0.6pt,line join=round] (492.30,219.00) --
	(488.50,219.00);

\path[draw=drawColor,line width= 0.6pt,line join=round] (488.50,217.73) --
	(488.50,220.28);
\definecolor{drawColor}{RGB}{170,68,153}

\path[draw=drawColor,line width= 0.6pt,line join=round] (476.62,184.50) --
	(476.62,187.05);

\path[draw=drawColor,line width= 0.6pt,line join=round] (476.62,185.78) --
	(466.71,185.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (466.71,184.50) --
	(466.71,187.05);
\definecolor{drawColor}{RGB}{68,170,153}

\path[draw=drawColor,line width= 0.6pt,line join=round] (487.34,192.17) --
	(487.34,194.72);

\path[draw=drawColor,line width= 0.6pt,line join=round] (487.34,193.44) --
	(481.69,193.44);

\path[draw=drawColor,line width= 0.6pt,line join=round] (481.69,192.17) --
	(481.69,194.72);
\definecolor{drawColor}{RGB}{170,68,153}

\path[draw=drawColor,line width= 0.6pt,line join=round] (479.04, 82.26) --
	(479.04, 84.82);

\path[draw=drawColor,line width= 0.6pt,line join=round] (479.04, 83.54) --
	(472.70, 83.54);

\path[draw=drawColor,line width= 0.6pt,line join=round] (472.70, 82.26) --
	(472.70, 84.82);
\definecolor{drawColor}{RGB}{68,170,153}

\path[draw=drawColor,line width= 0.6pt,line join=round] (458.29, 89.93) --
	(458.29, 92.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] (458.29, 91.21) --
	(445.61, 91.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] (445.61, 89.93) --
	(445.61, 92.48);
\definecolor{drawColor}{RGB}{170,68,153}
\definecolor{fillColor}{RGB}{170,68,153}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (478.70,134.66) circle (  1.96);
\definecolor{drawColor}{RGB}{68,170,153}
\definecolor{fillColor}{RGB}{68,170,153}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (484.58,142.32) circle (  1.96);
\definecolor{drawColor}{RGB}{170,68,153}
\definecolor{fillColor}{RGB}{170,68,153}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (481.81,160.22) circle (  1.96);
\definecolor{drawColor}{RGB}{68,170,153}
\definecolor{fillColor}{RGB}{68,170,153}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (484.58,167.88) circle (  1.96);
\definecolor{drawColor}{RGB}{170,68,153}
\definecolor{fillColor}{RGB}{170,68,153}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (462.67,492.49) circle (  1.96);
\definecolor{drawColor}{RGB}{68,170,153}
\definecolor{fillColor}{RGB}{68,170,153}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (473.39,500.16) circle (  1.96);
\definecolor{drawColor}{RGB}{170,68,153}
\definecolor{fillColor}{RGB}{170,68,153}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (482.62,466.93) circle (  1.96);
\definecolor{drawColor}{RGB}{68,170,153}
\definecolor{fillColor}{RGB}{68,170,153}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (477.20,474.60) circle (  1.96);
\definecolor{drawColor}{RGB}{170,68,153}
\definecolor{fillColor}{RGB}{170,68,153}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (493.80,441.37) circle (  1.96);
\definecolor{drawColor}{RGB}{68,170,153}
\definecolor{fillColor}{RGB}{68,170,153}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (491.61,449.04) circle (  1.96);
\definecolor{drawColor}{RGB}{170,68,153}
\definecolor{fillColor}{RGB}{170,68,153}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (491.84,415.81) circle (  1.96);
\definecolor{drawColor}{RGB}{68,170,153}
\definecolor{fillColor}{RGB}{68,170,153}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (495.99,423.48) circle (  1.96);
\definecolor{drawColor}{RGB}{170,68,153}
\definecolor{fillColor}{RGB}{170,68,153}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (488.96,339.13) circle (  1.96);
\definecolor{drawColor}{RGB}{68,170,153}
\definecolor{fillColor}{RGB}{68,170,153}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (477.89,346.80) circle (  1.96);
\definecolor{drawColor}{RGB}{170,68,153}
\definecolor{fillColor}{RGB}{170,68,153}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (474.89,313.57) circle (  1.96);
\definecolor{drawColor}{RGB}{68,170,153}
\definecolor{fillColor}{RGB}{68,170,153}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (466.59,321.24) circle (  1.96);
\definecolor{drawColor}{RGB}{170,68,153}
\definecolor{fillColor}{RGB}{170,68,153}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (428.32,109.10) circle (  1.96);
\definecolor{drawColor}{RGB}{68,170,153}
\definecolor{fillColor}{RGB}{68,170,153}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (426.71,116.76) circle (  1.96);
\definecolor{drawColor}{RGB}{170,68,153}
\definecolor{fillColor}{RGB}{170,68,153}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (474.89,288.01) circle (  1.96);
\definecolor{drawColor}{RGB}{68,170,153}
\definecolor{fillColor}{RGB}{68,170,153}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (493.68,295.68) circle (  1.96);
\definecolor{drawColor}{RGB}{170,68,153}
\definecolor{fillColor}{RGB}{170,68,153}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (478.35,262.45) circle (  1.96);
\definecolor{drawColor}{RGB}{68,170,153}
\definecolor{fillColor}{RGB}{68,170,153}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (485.04,270.12) circle (  1.96);
\definecolor{drawColor}{RGB}{170,68,153}
\definecolor{fillColor}{RGB}{170,68,153}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (494.95,236.90) circle (  1.96);
\definecolor{drawColor}{RGB}{68,170,153}
\definecolor{fillColor}{RGB}{68,170,153}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (493.80,244.56) circle (  1.96);
\definecolor{drawColor}{RGB}{170,68,153}
\definecolor{fillColor}{RGB}{170,68,153}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (484.92,211.34) circle (  1.96);
\definecolor{drawColor}{RGB}{68,170,153}
\definecolor{fillColor}{RGB}{68,170,153}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (490.34,219.00) circle (  1.96);
\definecolor{drawColor}{RGB}{170,68,153}
\definecolor{fillColor}{RGB}{170,68,153}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (471.67,185.78) circle (  1.96);
\definecolor{drawColor}{RGB}{68,170,153}
\definecolor{fillColor}{RGB}{68,170,153}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (484.58,193.44) circle (  1.96);
\definecolor{drawColor}{RGB}{170,68,153}
\definecolor{fillColor}{RGB}{170,68,153}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (475.93, 83.54) circle (  1.96);
\definecolor{drawColor}{RGB}{68,170,153}
\definecolor{fillColor}{RGB}{68,170,153}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (451.95, 91.21) circle (  1.96);
\definecolor{drawColor}{RGB}{169,169,169}

\path[draw=drawColor,line width= 0.6pt,line join=round] (386.82,176.83) -- (513.63,176.83);

\path[draw=drawColor,line width= 0.6pt,line join=round] (386.82,151.27) -- (513.63,151.27);

\path[draw=drawColor,line width= 0.6pt,line join=round] (386.82,125.71) -- (513.63,125.71);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.5pt,line join=round,line cap=round] (386.82, 72.04) rectangle (513.63,511.66);
\end{scope}
\begin{scope}
\path[clip] (518.63, 72.04) rectangle (645.43,511.66);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (518.63, 72.04) rectangle (645.43,511.66);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] (538.80, 72.04) --
	(538.80,511.66);

\path[draw=drawColor,line width= 0.3pt,line join=round] (567.62, 72.04) --
	(567.62,511.66);

\path[draw=drawColor,line width= 0.3pt,line join=round] (596.44, 72.04) --
	(596.44,511.66);

\path[draw=drawColor,line width= 0.3pt,line join=round] (625.26, 72.04) --
	(625.26,511.66);

\path[draw=drawColor,line width= 0.5pt,line join=round] (518.63, 87.37) --
	(645.43, 87.37);

\path[draw=drawColor,line width= 0.5pt,line join=round] (518.63,112.93) --
	(645.43,112.93);

\path[draw=drawColor,line width= 0.5pt,line join=round] (518.63,138.49) --
	(645.43,138.49);

\path[draw=drawColor,line width= 0.5pt,line join=round] (518.63,164.05) --
	(645.43,164.05);

\path[draw=drawColor,line width= 0.5pt,line join=round] (518.63,189.61) --
	(645.43,189.61);

\path[draw=drawColor,line width= 0.5pt,line join=round] (518.63,215.17) --
	(645.43,215.17);

\path[draw=drawColor,line width= 0.5pt,line join=round] (518.63,240.73) --
	(645.43,240.73);

\path[draw=drawColor,line width= 0.5pt,line join=round] (518.63,266.29) --
	(645.43,266.29);

\path[draw=drawColor,line width= 0.5pt,line join=round] (518.63,291.85) --
	(645.43,291.85);

\path[draw=drawColor,line width= 0.5pt,line join=round] (518.63,317.41) --
	(645.43,317.41);

\path[draw=drawColor,line width= 0.5pt,line join=round] (518.63,342.97) --
	(645.43,342.97);

\path[draw=drawColor,line width= 0.5pt,line join=round] (518.63,368.53) --
	(645.43,368.53);

\path[draw=drawColor,line width= 0.5pt,line join=round] (518.63,394.09) --
	(645.43,394.09);

\path[draw=drawColor,line width= 0.5pt,line join=round] (518.63,419.65) --
	(645.43,419.65);

\path[draw=drawColor,line width= 0.5pt,line join=round] (518.63,445.21) --
	(645.43,445.21);

\path[draw=drawColor,line width= 0.5pt,line join=round] (518.63,470.77) --
	(645.43,470.77);

\path[draw=drawColor,line width= 0.5pt,line join=round] (518.63,496.33) --
	(645.43,496.33);

\path[draw=drawColor,line width= 0.5pt,line join=round] (524.39, 72.04) --
	(524.39,511.66);

\path[draw=drawColor,line width= 0.5pt,line join=round] (553.21, 72.04) --
	(553.21,511.66);

\path[draw=drawColor,line width= 0.5pt,line join=round] (582.03, 72.04) --
	(582.03,511.66);

\path[draw=drawColor,line width= 0.5pt,line join=round] (610.85, 72.04) --
	(610.85,511.66);

\path[draw=drawColor,line width= 0.5pt,line join=round] (639.67, 72.04) --
	(639.67,511.66);
\definecolor{drawColor}{RGB}{51,34,136}

\path[draw=drawColor,line width= 0.6pt,line join=round] (624.33,142.75) --
	(624.33,144.45);

\path[draw=drawColor,line width= 0.6pt,line join=round] (624.33,143.60) --
	(609.00,143.60);

\path[draw=drawColor,line width= 0.6pt,line join=round] (609.00,142.75) --
	(609.00,144.45);
\definecolor{drawColor}{RGB}{17,119,51}

\path[draw=drawColor,line width= 0.6pt,line join=round] (627.22,137.64) --
	(627.22,139.34);

\path[draw=drawColor,line width= 0.6pt,line join=round] (627.22,138.49) --
	(600.13,138.49);

\path[draw=drawColor,line width= 0.6pt,line join=round] (600.13,137.64) --
	(600.13,139.34);
\definecolor{drawColor}{RGB}{221,204,119}

\path[draw=drawColor,line width= 0.6pt,line join=round] (624.91,132.53) --
	(624.91,134.23);

\path[draw=drawColor,line width= 0.6pt,line join=round] (624.91,133.38) --
	(595.52,133.38);

\path[draw=drawColor,line width= 0.6pt,line join=round] (595.52,132.53) --
	(595.52,134.23);
\definecolor{drawColor}{RGB}{51,34,136}

\path[draw=drawColor,line width= 0.6pt,line join=round] (626.64,168.31) --
	(626.64,170.01);

\path[draw=drawColor,line width= 0.6pt,line join=round] (626.64,169.16) --
	(613.04,169.16);

\path[draw=drawColor,line width= 0.6pt,line join=round] (613.04,168.31) --
	(613.04,170.01);
\definecolor{drawColor}{RGB}{17,119,51}

\path[draw=drawColor,line width= 0.6pt,line join=round] (624.22,163.20) --
	(624.22,164.90);

\path[draw=drawColor,line width= 0.6pt,line join=round] (624.22,164.05) --
	(611.54,164.05);

\path[draw=drawColor,line width= 0.6pt,line join=round] (611.54,163.20) --
	(611.54,164.90);
\definecolor{drawColor}{RGB}{221,204,119}

\path[draw=drawColor,line width= 0.6pt,line join=round] (623.18,158.09) --
	(623.18,159.79);

\path[draw=drawColor,line width= 0.6pt,line join=round] (623.18,158.94) --
	(607.97,158.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (607.97,158.09) --
	(607.97,159.79);
\definecolor{drawColor}{RGB}{51,34,136}

\path[draw=drawColor,line width= 0.6pt,line join=round] (632.75,498.88) --
	(632.75,501.44);

\path[draw=drawColor,line width= 0.6pt,line join=round] (632.75,500.16) --
	(591.25,500.16);

\path[draw=drawColor,line width= 0.6pt,line join=round] (591.25,498.88) --
	(591.25,501.44);
\definecolor{drawColor}{RGB}{17,119,51}

\path[draw=drawColor,line width= 0.6pt,line join=round] (609.00,491.21) --
	(609.00,493.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (609.00,492.49) --
	(585.49,492.49);

\path[draw=drawColor,line width= 0.6pt,line join=round] (585.49,491.21) --
	(585.49,493.77);
\definecolor{drawColor}{RGB}{51,34,136}

\path[draw=drawColor,line width= 0.6pt,line join=round] (620.53,475.03) --
	(620.53,476.73);

\path[draw=drawColor,line width= 0.6pt,line join=round] (620.53,475.88) --
	(602.20,475.88);

\path[draw=drawColor,line width= 0.6pt,line join=round] (602.20,475.03) --
	(602.20,476.73);
\definecolor{drawColor}{RGB}{17,119,51}

\path[draw=drawColor,line width= 0.6pt,line join=round] (614.65,469.91) --
	(614.65,471.62);

\path[draw=drawColor,line width= 0.6pt,line join=round] (614.65,470.77) --
	(605.31,470.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (605.31,469.91) --
	(605.31,471.62);
\definecolor{drawColor}{RGB}{221,204,119}

\path[draw=drawColor,line width= 0.6pt,line join=round] (619.61,464.80) --
	(619.61,466.51);

\path[draw=drawColor,line width= 0.6pt,line join=round] (619.61,465.65) --
	(599.32,465.65);

\path[draw=drawColor,line width= 0.6pt,line join=round] (599.32,464.80) --
	(599.32,466.51);
\definecolor{drawColor}{RGB}{51,34,136}

\path[draw=drawColor,line width= 0.6pt,line join=round] (620.76,449.47) --
	(620.76,451.17);

\path[draw=drawColor,line width= 0.6pt,line join=round] (620.76,450.32) --
	(606.93,450.32);

\path[draw=drawColor,line width= 0.6pt,line join=round] (606.93,449.47) --
	(606.93,451.17);
\definecolor{drawColor}{RGB}{17,119,51}

\path[draw=drawColor,line width= 0.6pt,line join=round] (625.03,444.35) --
	(625.03,446.06);

\path[draw=drawColor,line width= 0.6pt,line join=round] (625.03,445.21) --
	(620.53,445.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] (620.53,444.35) --
	(620.53,446.06);
\definecolor{drawColor}{RGB}{221,204,119}

\path[draw=drawColor,line width= 0.6pt,line join=round] (626.41,439.24) --
	(626.41,440.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (626.41,440.09) --
	(613.84,440.09);

\path[draw=drawColor,line width= 0.6pt,line join=round] (613.84,439.24) --
	(613.84,440.95);
\definecolor{drawColor}{RGB}{17,119,51}

\path[draw=drawColor,line width= 0.6pt,line join=round] (629.52,422.20) --
	(629.52,424.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] (629.52,423.48) --
	(622.72,423.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] (622.72,422.20) --
	(622.72,424.76);
\definecolor{drawColor}{RGB}{221,204,119}

\path[draw=drawColor,line width= 0.6pt,line join=round] (634.36,414.53) --
	(634.36,417.09);

\path[draw=drawColor,line width= 0.6pt,line join=round] (634.36,415.81) --
	(625.95,415.81);

\path[draw=drawColor,line width= 0.6pt,line join=round] (625.95,414.53) --
	(625.95,417.09);
\definecolor{drawColor}{RGB}{51,34,136}

\path[draw=drawColor,line width= 0.6pt,line join=round] (639.44,398.35) --
	(639.44,400.05);

\path[draw=drawColor,line width= 0.6pt,line join=round] (639.44,399.20) --
	(634.71,399.20);

\path[draw=drawColor,line width= 0.6pt,line join=round] (634.71,398.35) --
	(634.71,400.05);
\definecolor{drawColor}{RGB}{17,119,51}

\path[draw=drawColor,line width= 0.6pt,line join=round] (637.25,393.23) --
	(637.25,394.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (637.25,394.09) --
	(634.13,394.09);

\path[draw=drawColor,line width= 0.6pt,line join=round] (634.13,393.23) --
	(634.13,394.94);
\definecolor{drawColor}{RGB}{221,204,119}

\path[draw=drawColor,line width= 0.6pt,line join=round] (636.44,388.12) --
	(636.44,389.83);

\path[draw=drawColor,line width= 0.6pt,line join=round] (636.44,388.98) --
	(628.60,388.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (628.60,388.12) --
	(628.60,389.83);
\definecolor{drawColor}{RGB}{51,34,136}

\path[draw=drawColor,line width= 0.6pt,line join=round] (610.50,372.79) --
	(610.50,374.49);

\path[draw=drawColor,line width= 0.6pt,line join=round] (610.50,373.64) --
	(597.36,373.64);

\path[draw=drawColor,line width= 0.6pt,line join=round] (597.36,372.79) --
	(597.36,374.49);
\definecolor{drawColor}{RGB}{17,119,51}

\path[draw=drawColor,line width= 0.6pt,line join=round] (616.27,367.68) --
	(616.27,369.38);

\path[draw=drawColor,line width= 0.6pt,line join=round] (616.27,368.53) --
	(611.42,368.53);

\path[draw=drawColor,line width= 0.6pt,line join=round] (611.42,367.68) --
	(611.42,369.38);
\definecolor{drawColor}{RGB}{221,204,119}

\path[draw=drawColor,line width= 0.6pt,line join=round] (620.76,362.56) --
	(620.76,364.27);

\path[draw=drawColor,line width= 0.6pt,line join=round] (620.76,363.42) --
	(600.24,363.42);

\path[draw=drawColor,line width= 0.6pt,line join=round] (600.24,362.56) --
	(600.24,364.27);
\definecolor{drawColor}{RGB}{51,34,136}

\path[draw=drawColor,line width= 0.6pt,line join=round] (632.86,347.23) --
	(632.86,348.93);

\path[draw=drawColor,line width= 0.6pt,line join=round] (632.86,348.08) --
	(614.77,348.08);

\path[draw=drawColor,line width= 0.6pt,line join=round] (614.77,347.23) --
	(614.77,348.93);
\definecolor{drawColor}{RGB}{17,119,51}

\path[draw=drawColor,line width= 0.6pt,line join=round] (618.69,342.12) --
	(618.69,343.82);

\path[draw=drawColor,line width= 0.6pt,line join=round] (618.69,342.97) --
	(604.28,342.97);

\path[draw=drawColor,line width= 0.6pt,line join=round] (604.28,342.12) --
	(604.28,343.82);
\definecolor{drawColor}{RGB}{221,204,119}

\path[draw=drawColor,line width= 0.6pt,line join=round] (635.86,337.00) --
	(635.86,338.71);

\path[draw=drawColor,line width= 0.6pt,line join=round] (635.86,337.86) --
	(599.20,337.86);

\path[draw=drawColor,line width= 0.6pt,line join=round] (599.20,337.00) --
	(599.20,338.71);
\definecolor{drawColor}{RGB}{51,34,136}

\path[draw=drawColor,line width= 0.6pt,line join=round] (567.85,117.19) --
	(567.85,118.89);

\path[draw=drawColor,line width= 0.6pt,line join=round] (567.85,118.04) --
	(558.05,118.04);

\path[draw=drawColor,line width= 0.6pt,line join=round] (558.05,117.19) --
	(558.05,118.89);
\definecolor{drawColor}{RGB}{17,119,51}

\path[draw=drawColor,line width= 0.6pt,line join=round] (557.70,112.08) --
	(557.70,113.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] (557.70,112.93) --
	(554.59,112.93);

\path[draw=drawColor,line width= 0.6pt,line join=round] (554.59,112.08) --
	(554.59,113.78);
\definecolor{drawColor}{RGB}{221,204,119}

\path[draw=drawColor,line width= 0.6pt,line join=round] (575.11,106.97) --
	(575.11,108.67);

\path[draw=drawColor,line width= 0.6pt,line join=round] (575.11,107.82) --
	(565.77,107.82);

\path[draw=drawColor,line width= 0.6pt,line join=round] (565.77,106.97) --
	(565.77,108.67);
\definecolor{drawColor}{RGB}{51,34,136}

\path[draw=drawColor,line width= 0.6pt,line join=round] (629.41,296.11) --
	(629.41,297.81);

\path[draw=drawColor,line width= 0.6pt,line join=round] (629.41,296.96) --
	(622.37,296.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (622.37,296.11) --
	(622.37,297.81);
\definecolor{drawColor}{RGB}{17,119,51}

\path[draw=drawColor,line width= 0.6pt,line join=round] (623.87,291.00) --
	(623.87,292.70);

\path[draw=drawColor,line width= 0.6pt,line join=round] (623.87,291.85) --
	(618.11,291.85);

\path[draw=drawColor,line width= 0.6pt,line join=round] (618.11,291.00) --
	(618.11,292.70);
\definecolor{drawColor}{RGB}{221,204,119}

\path[draw=drawColor,line width= 0.6pt,line join=round] (625.03,285.88) --
	(625.03,287.59);

\path[draw=drawColor,line width= 0.6pt,line join=round] (625.03,286.74) --
	(592.63,286.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] (592.63,285.88) --
	(592.63,287.59);
\definecolor{drawColor}{RGB}{51,34,136}

\path[draw=drawColor,line width= 0.6pt,line join=round] (620.76,270.55) --
	(620.76,272.25);

\path[draw=drawColor,line width= 0.6pt,line join=round] (620.76,271.40) --
	(607.85,271.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] (607.85,270.55) --
	(607.85,272.25);
\definecolor{drawColor}{RGB}{17,119,51}

\path[draw=drawColor,line width= 0.6pt,line join=round] (617.88,265.44) --
	(617.88,267.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (617.88,266.29) --
	(611.31,266.29);

\path[draw=drawColor,line width= 0.6pt,line join=round] (611.31,265.44) --
	(611.31,267.14);
\definecolor{drawColor}{RGB}{221,204,119}

\path[draw=drawColor,line width= 0.6pt,line join=round] (624.22,260.32) --
	(624.22,262.03);

\path[draw=drawColor,line width= 0.6pt,line join=round] (624.22,261.18) --
	(595.17,261.18);

\path[draw=drawColor,line width= 0.6pt,line join=round] (595.17,260.32) --
	(595.17,262.03);
\definecolor{drawColor}{RGB}{51,34,136}

\path[draw=drawColor,line width= 0.6pt,line join=round] (630.33,244.99) --
	(630.33,246.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (630.33,245.84) --
	(609.58,245.84);

\path[draw=drawColor,line width= 0.6pt,line join=round] (609.58,244.99) --
	(609.58,246.69);
\definecolor{drawColor}{RGB}{17,119,51}

\path[draw=drawColor,line width= 0.6pt,line join=round] (627.56,239.88) --
	(627.56,241.58);

\path[draw=drawColor,line width= 0.6pt,line join=round] (627.56,240.73) --
	(623.07,240.73);

\path[draw=drawColor,line width= 0.6pt,line join=round] (623.07,239.88) --
	(623.07,241.58);
\definecolor{drawColor}{RGB}{221,204,119}

\path[draw=drawColor,line width= 0.6pt,line join=round] (631.14,234.77) --
	(631.14,236.47);

\path[draw=drawColor,line width= 0.6pt,line join=round] (631.14,235.62) --
	(625.14,235.62);

\path[draw=drawColor,line width= 0.6pt,line join=round] (625.14,234.77) --
	(625.14,236.47);
\definecolor{drawColor}{RGB}{51,34,136}

\path[draw=drawColor,line width= 0.6pt,line join=round] (626.29,217.73) --
	(626.29,220.28);

\path[draw=drawColor,line width= 0.6pt,line join=round] (626.29,219.00) --
	(619.72,219.00);

\path[draw=drawColor,line width= 0.6pt,line join=round] (619.72,217.73) --
	(619.72,220.28);
\definecolor{drawColor}{RGB}{17,119,51}

\path[draw=drawColor,line width= 0.6pt,line join=round] (625.14,210.06) --
	(625.14,212.61);

\path[draw=drawColor,line width= 0.6pt,line join=round] (625.14,211.34) --
	(621.57,211.34);

\path[draw=drawColor,line width= 0.6pt,line join=round] (621.57,210.06) --
	(621.57,212.61);
\definecolor{drawColor}{RGB}{51,34,136}

\path[draw=drawColor,line width= 0.6pt,line join=round] (619.03,193.87) --
	(619.03,195.57);

\path[draw=drawColor,line width= 0.6pt,line join=round] (619.03,194.72) --
	(606.24,194.72);

\path[draw=drawColor,line width= 0.6pt,line join=round] (606.24,193.87) --
	(606.24,195.57);
\definecolor{drawColor}{RGB}{17,119,51}

\path[draw=drawColor,line width= 0.6pt,line join=round] (616.15,188.76) --
	(616.15,190.46);

\path[draw=drawColor,line width= 0.6pt,line join=round] (616.15,189.61) --
	(610.16,189.61);

\path[draw=drawColor,line width= 0.6pt,line join=round] (610.16,188.76) --
	(610.16,190.46);
\definecolor{drawColor}{RGB}{221,204,119}

\path[draw=drawColor,line width= 0.6pt,line join=round] (616.61,183.65) --
	(616.61,185.35);

\path[draw=drawColor,line width= 0.6pt,line join=round] (616.61,184.50) --
	(601.97,184.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] (601.97,183.65) --
	(601.97,185.35);
\definecolor{drawColor}{RGB}{51,34,136}

\path[draw=drawColor,line width= 0.6pt,line join=round] (591.94, 91.63) --
	(591.94, 93.33);

\path[draw=drawColor,line width= 0.6pt,line join=round] (591.94, 92.48) --
	(574.54, 92.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] (574.54, 91.63) --
	(574.54, 93.33);
\definecolor{drawColor}{RGB}{17,119,51}

\path[draw=drawColor,line width= 0.6pt,line join=round] (603.58, 86.52) --
	(603.58, 88.22);

\path[draw=drawColor,line width= 0.6pt,line join=round] (603.58, 87.37) --
	(594.82, 87.37);

\path[draw=drawColor,line width= 0.6pt,line join=round] (594.82, 86.52) --
	(594.82, 88.22);
\definecolor{drawColor}{RGB}{221,204,119}

\path[draw=drawColor,line width= 0.6pt,line join=round] (609.58, 81.41) --
	(609.58, 83.11);

\path[draw=drawColor,line width= 0.6pt,line join=round] (609.58, 82.26) --
	(599.09, 82.26);

\path[draw=drawColor,line width= 0.6pt,line join=round] (599.09, 81.41) --
	(599.09, 83.11);
\definecolor{drawColor}{RGB}{51,34,136}
\definecolor{fillColor}{RGB}{51,34,136}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (616.73,143.60) circle (  1.96);
\definecolor{drawColor}{RGB}{17,119,51}
\definecolor{fillColor}{RGB}{17,119,51}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (613.61,138.49) circle (  1.96);
\definecolor{drawColor}{RGB}{221,204,119}
\definecolor{fillColor}{RGB}{221,204,119}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (610.16,133.38) circle (  1.96);
\definecolor{drawColor}{RGB}{51,34,136}
\definecolor{fillColor}{RGB}{51,34,136}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (619.84,169.16) circle (  1.96);
\definecolor{drawColor}{RGB}{17,119,51}
\definecolor{fillColor}{RGB}{17,119,51}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (617.88,164.05) circle (  1.96);
\definecolor{drawColor}{RGB}{221,204,119}
\definecolor{fillColor}{RGB}{221,204,119}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (615.57,158.94) circle (  1.96);
\definecolor{drawColor}{RGB}{51,34,136}
\definecolor{fillColor}{RGB}{51,34,136}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (612.00,500.16) circle (  1.96);
\definecolor{drawColor}{RGB}{17,119,51}
\definecolor{fillColor}{RGB}{17,119,51}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (597.24,492.49) circle (  1.96);
\definecolor{drawColor}{RGB}{51,34,136}
\definecolor{fillColor}{RGB}{51,34,136}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (611.31,475.88) circle (  1.96);
\definecolor{drawColor}{RGB}{17,119,51}
\definecolor{fillColor}{RGB}{17,119,51}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (609.92,470.77) circle (  1.96);
\definecolor{drawColor}{RGB}{221,204,119}
\definecolor{fillColor}{RGB}{221,204,119}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (609.46,465.65) circle (  1.96);
\definecolor{drawColor}{RGB}{51,34,136}
\definecolor{fillColor}{RGB}{51,34,136}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (613.84,450.32) circle (  1.96);
\definecolor{drawColor}{RGB}{17,119,51}
\definecolor{fillColor}{RGB}{17,119,51}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (622.84,445.21) circle (  1.96);
\definecolor{drawColor}{RGB}{221,204,119}
\definecolor{fillColor}{RGB}{221,204,119}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (620.18,440.09) circle (  1.96);
\definecolor{drawColor}{RGB}{17,119,51}
\definecolor{fillColor}{RGB}{17,119,51}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (626.18,423.48) circle (  1.96);
\definecolor{drawColor}{RGB}{221,204,119}
\definecolor{fillColor}{RGB}{221,204,119}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (630.10,415.81) circle (  1.96);
\definecolor{drawColor}{RGB}{51,34,136}
\definecolor{fillColor}{RGB}{51,34,136}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (637.13,399.20) circle (  1.96);
\definecolor{drawColor}{RGB}{17,119,51}
\definecolor{fillColor}{RGB}{17,119,51}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (635.75,394.09) circle (  1.96);
\definecolor{drawColor}{RGB}{221,204,119}
\definecolor{fillColor}{RGB}{221,204,119}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (632.52,388.98) circle (  1.96);
\definecolor{drawColor}{RGB}{51,34,136}
\definecolor{fillColor}{RGB}{51,34,136}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (603.93,373.64) circle (  1.96);
\definecolor{drawColor}{RGB}{17,119,51}
\definecolor{fillColor}{RGB}{17,119,51}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (613.84,368.53) circle (  1.96);
\definecolor{drawColor}{RGB}{221,204,119}
\definecolor{fillColor}{RGB}{221,204,119}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (610.50,363.42) circle (  1.96);
\definecolor{drawColor}{RGB}{51,34,136}
\definecolor{fillColor}{RGB}{51,34,136}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (623.87,348.08) circle (  1.96);
\definecolor{drawColor}{RGB}{17,119,51}
\definecolor{fillColor}{RGB}{17,119,51}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (611.54,342.97) circle (  1.96);
\definecolor{drawColor}{RGB}{221,204,119}
\definecolor{fillColor}{RGB}{221,204,119}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (617.53,337.86) circle (  1.96);
\definecolor{drawColor}{RGB}{51,34,136}
\definecolor{fillColor}{RGB}{51,34,136}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (563.01,118.04) circle (  1.96);
\definecolor{drawColor}{RGB}{17,119,51}
\definecolor{fillColor}{RGB}{17,119,51}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (556.21,112.93) circle (  1.96);
\definecolor{drawColor}{RGB}{221,204,119}
\definecolor{fillColor}{RGB}{221,204,119}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (570.50,107.82) circle (  1.96);
\definecolor{drawColor}{RGB}{51,34,136}
\definecolor{fillColor}{RGB}{51,34,136}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (625.95,296.96) circle (  1.96);
\definecolor{drawColor}{RGB}{17,119,51}
\definecolor{fillColor}{RGB}{17,119,51}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (620.99,291.85) circle (  1.96);
\definecolor{drawColor}{RGB}{221,204,119}
\definecolor{fillColor}{RGB}{221,204,119}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (608.89,286.74) circle (  1.96);
\definecolor{drawColor}{RGB}{51,34,136}
\definecolor{fillColor}{RGB}{51,34,136}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (614.31,271.40) circle (  1.96);
\definecolor{drawColor}{RGB}{17,119,51}
\definecolor{fillColor}{RGB}{17,119,51}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (614.65,266.29) circle (  1.96);
\definecolor{drawColor}{RGB}{221,204,119}
\definecolor{fillColor}{RGB}{221,204,119}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (609.69,261.18) circle (  1.96);
\definecolor{drawColor}{RGB}{51,34,136}
\definecolor{fillColor}{RGB}{51,34,136}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (619.95,245.84) circle (  1.96);
\definecolor{drawColor}{RGB}{17,119,51}
\definecolor{fillColor}{RGB}{17,119,51}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (625.37,240.73) circle (  1.96);
\definecolor{drawColor}{RGB}{221,204,119}
\definecolor{fillColor}{RGB}{221,204,119}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (628.14,235.62) circle (  1.96);
\definecolor{drawColor}{RGB}{51,34,136}
\definecolor{fillColor}{RGB}{51,34,136}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (622.95,219.00) circle (  1.96);
\definecolor{drawColor}{RGB}{17,119,51}
\definecolor{fillColor}{RGB}{17,119,51}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (623.41,211.34) circle (  1.96);
\definecolor{drawColor}{RGB}{51,34,136}
\definecolor{fillColor}{RGB}{51,34,136}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (612.58,194.72) circle (  1.96);
\definecolor{drawColor}{RGB}{17,119,51}
\definecolor{fillColor}{RGB}{17,119,51}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (613.15,189.61) circle (  1.96);
\definecolor{drawColor}{RGB}{221,204,119}
\definecolor{fillColor}{RGB}{221,204,119}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (609.35,184.50) circle (  1.96);
\definecolor{drawColor}{RGB}{51,34,136}
\definecolor{fillColor}{RGB}{51,34,136}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (583.18, 92.48) circle (  1.96);
\definecolor{drawColor}{RGB}{17,119,51}
\definecolor{fillColor}{RGB}{17,119,51}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (599.20, 87.37) circle (  1.96);
\definecolor{drawColor}{RGB}{221,204,119}
\definecolor{fillColor}{RGB}{221,204,119}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (604.39, 82.26) circle (  1.96);
\definecolor{drawColor}{RGB}{169,169,169}

\path[draw=drawColor,line width= 0.6pt,line join=round] (518.63,176.83) -- (645.43,176.83);

\path[draw=drawColor,line width= 0.6pt,line join=round] (518.63,151.27) -- (645.43,151.27);

\path[draw=drawColor,line width= 0.6pt,line join=round] (518.63,125.71) -- (645.43,125.71);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.5pt,line join=round,line cap=round] (518.63, 72.04) rectangle (645.43,511.66);
\end{scope}
\begin{scope}
\path[clip] (123.21,511.66) rectangle (250.02,524.61);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.5pt,line join=round,line cap=round,fill=fillColor] (123.21,511.66) rectangle (250.02,524.61);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (186.62,516.21) {All};
\end{scope}
\begin{scope}
\path[clip] (255.02,511.66) rectangle (381.82,524.61);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.5pt,line join=round,line cap=round,fill=fillColor] (255.02,511.66) rectangle (381.82,524.61);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (318.42,516.21) {By gender};
\end{scope}
\begin{scope}
\path[clip] (386.82,511.66) rectangle (513.63,524.61);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.5pt,line join=round,line cap=round,fill=fillColor] (386.82,511.66) rectangle (513.63,524.61);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (450.22,516.21) {By education};
\end{scope}
\begin{scope}
\path[clip] (518.63,511.66) rectangle (645.43,524.61);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.5pt,line join=round,line cap=round,fill=fillColor] (518.63,511.66) rectangle (645.43,524.61);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (582.03,516.21) {By age};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.5pt,line join=round] (128.98, 69.54) --
	(128.98, 72.04);

\path[draw=drawColor,line width= 0.5pt,line join=round] (157.80, 69.54) --
	(157.80, 72.04);

\path[draw=drawColor,line width= 0.5pt,line join=round] (186.62, 69.54) --
	(186.62, 72.04);

\path[draw=drawColor,line width= 0.5pt,line join=round] (215.44, 69.54) --
	(215.44, 72.04);

\path[draw=drawColor,line width= 0.5pt,line join=round] (244.25, 69.54) --
	(244.25, 72.04);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (128.98, 63.68) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (157.80, 63.68) {25};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (186.62, 63.68) {50};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (215.44, 63.68) {75};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (244.25, 63.68) {100};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.5pt,line join=round] (260.78, 69.54) --
	(260.78, 72.04);

\path[draw=drawColor,line width= 0.5pt,line join=round] (289.60, 69.54) --
	(289.60, 72.04);

\path[draw=drawColor,line width= 0.5pt,line join=round] (318.42, 69.54) --
	(318.42, 72.04);

\path[draw=drawColor,line width= 0.5pt,line join=round] (347.24, 69.54) --
	(347.24, 72.04);

\path[draw=drawColor,line width= 0.5pt,line join=round] (376.06, 69.54) --
	(376.06, 72.04);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (260.78, 63.68) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (289.60, 63.68) {25};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (318.42, 63.68) {50};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (347.24, 63.68) {75};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (376.06, 63.68) {100};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.5pt,line join=round] (392.59, 69.54) --
	(392.59, 72.04);

\path[draw=drawColor,line width= 0.5pt,line join=round] (421.41, 69.54) --
	(421.41, 72.04);

\path[draw=drawColor,line width= 0.5pt,line join=round] (450.22, 69.54) --
	(450.22, 72.04);

\path[draw=drawColor,line width= 0.5pt,line join=round] (479.04, 69.54) --
	(479.04, 72.04);

\path[draw=drawColor,line width= 0.5pt,line join=round] (507.86, 69.54) --
	(507.86, 72.04);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (392.59, 63.68) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (421.41, 63.68) {25};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (450.22, 63.68) {50};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (479.04, 63.68) {75};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (507.86, 63.68) {100};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.5pt,line join=round] (524.39, 69.54) --
	(524.39, 72.04);

\path[draw=drawColor,line width= 0.5pt,line join=round] (553.21, 69.54) --
	(553.21, 72.04);

\path[draw=drawColor,line width= 0.5pt,line join=round] (582.03, 69.54) --
	(582.03, 72.04);

\path[draw=drawColor,line width= 0.5pt,line join=round] (610.85, 69.54) --
	(610.85, 72.04);

\path[draw=drawColor,line width= 0.5pt,line join=round] (639.67, 69.54) --
	(639.67, 72.04);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (524.39, 63.68) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (553.21, 63.68) {25};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (582.03, 63.68) {50};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (610.85, 63.68) {75};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (639.67, 63.68) {100};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 13.68, 89.76) {USA};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 13.68, 81.12) {(National, 1959)};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 13.68,115.32) {Russia};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 13.68,106.68) {(Subnational, 61 regions, 22125)};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 13.68,140.88) {All LMICs};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 13.68,132.24) {(National samples)};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 13.68,162.12) {All LMICs};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 13.68,192.00) {Uganda 2};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 13.68,183.36) {(Subnational, 1 district, 1366)};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 13.68,217.56) {Uganda 1};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 13.68,208.92) {(Subnational, 13 districts, 3362)};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 13.68,243.12) {Sierra Leone 2};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 13.68,234.48) {(National, 2110)};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 13.68,268.68) {Sierra Leone 1};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 13.68,260.04) {(National, 1070)};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 13.68,294.24) {Rwanda};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 13.68,285.60) {(National, 1355)};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 13.68,319.80) {Pakistan 2};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 13.68,311.16) {(Subnational, 1 province, 1492)};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 13.68,345.36) {Pakistan 1};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 13.68,336.72) {(Subnational, 2 districts, 1633)};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 13.68,370.92) {Nigeria};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 13.68,362.28) {(Subnational, 1 state, 1868)};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 13.68,396.48) {Nepal};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 13.68,387.84) {(Subnational, 2 districts, 1389)};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 13.68,422.04) {Mozambique};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 13.68,413.40) {(Subnational, 2 cities, 862)};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 13.68,447.60) {India};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 13.68,438.96) {(Subnational, Slums in 2 cities, 1680)};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 13.68,473.16) {Colombia};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 13.68,464.52) {(National, 1012)};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 13.68,498.72) {Burkina Faso};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 13.68,490.08) {(National, 977)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.5pt,line join=round] (120.71, 87.37) --
	(123.21, 87.37);

\path[draw=drawColor,line width= 0.5pt,line join=round] (120.71,112.93) --
	(123.21,112.93);

\path[draw=drawColor,line width= 0.5pt,line join=round] (120.71,138.49) --
	(123.21,138.49);

\path[draw=drawColor,line width= 0.5pt,line join=round] (120.71,164.05) --
	(123.21,164.05);

\path[draw=drawColor,line width= 0.5pt,line join=round] (120.71,189.61) --
	(123.21,189.61);

\path[draw=drawColor,line width= 0.5pt,line join=round] (120.71,215.17) --
	(123.21,215.17);

\path[draw=drawColor,line width= 0.5pt,line join=round] (120.71,240.73) --
	(123.21,240.73);

\path[draw=drawColor,line width= 0.5pt,line join=round] (120.71,266.29) --
	(123.21,266.29);

\path[draw=drawColor,line width= 0.5pt,line join=round] (120.71,291.85) --
	(123.21,291.85);

\path[draw=drawColor,line width= 0.5pt,line join=round] (120.71,317.41) --
	(123.21,317.41);

\path[draw=drawColor,line width= 0.5pt,line join=round] (120.71,342.97) --
	(123.21,342.97);

\path[draw=drawColor,line width= 0.5pt,line join=round] (120.71,368.53) --
	(123.21,368.53);

\path[draw=drawColor,line width= 0.5pt,line join=round] (120.71,394.09) --
	(123.21,394.09);

\path[draw=drawColor,line width= 0.5pt,line join=round] (120.71,419.65) --
	(123.21,419.65);

\path[draw=drawColor,line width= 0.5pt,line join=round] (120.71,445.21) --
	(123.21,445.21);

\path[draw=drawColor,line width= 0.5pt,line join=round] (120.71,470.77) --
	(123.21,470.77);

\path[draw=drawColor,line width= 0.5pt,line join=round] (120.71,496.33) --
	(123.21,496.33);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (384.32, 55.27) {estimate};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (256.56,  5.00) rectangle (512.09, 43.91);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (261.56, 22.04) {Subgroups};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (302.90, 24.45) rectangle (317.35, 38.91);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{drawColor}{RGB}{136,34,85}

\path[draw=drawColor,line width= 0.6pt,line join=round] (304.34, 31.68) -- (315.90, 31.68);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{drawColor}{RGB}{136,34,85}
\definecolor{fillColor}{RGB}{136,34,85}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (310.12, 31.68) circle (  1.96);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (302.90, 10.00) rectangle (317.35, 24.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{drawColor}{RGB}{153,153,51}

\path[draw=drawColor,line width= 0.6pt,line join=round] (304.34, 17.23) -- (315.90, 17.23);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{drawColor}{RGB}{153,153,51}
\definecolor{fillColor}{RGB}{153,153,51}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (310.12, 17.23) circle (  1.96);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (346.90, 24.45) rectangle (361.35, 38.91);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{drawColor}{RGB}{68,170,153}

\path[draw=drawColor,line width= 0.6pt,line join=round] (348.35, 31.68) -- (359.91, 31.68);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{drawColor}{RGB}{68,170,153}
\definecolor{fillColor}{RGB}{68,170,153}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (354.13, 31.68) circle (  1.96);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (346.90, 10.00) rectangle (361.35, 24.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{drawColor}{RGB}{170,68,153}

\path[draw=drawColor,line width= 0.6pt,line join=round] (348.35, 17.23) -- (359.91, 17.23);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{drawColor}{RGB}{170,68,153}
\definecolor{fillColor}{RGB}{170,68,153}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (354.13, 17.23) circle (  1.96);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (431.22, 24.45) rectangle (445.68, 38.91);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{drawColor}{RGB}{51,34,136}

\path[draw=drawColor,line width= 0.6pt,line join=round] (432.67, 31.68) -- (444.23, 31.68);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{drawColor}{RGB}{51,34,136}
\definecolor{fillColor}{RGB}{51,34,136}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (438.45, 31.68) circle (  1.96);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (431.22, 10.00) rectangle (445.68, 24.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{drawColor}{RGB}{17,119,51}

\path[draw=drawColor,line width= 0.6pt,line join=round] (432.67, 17.23) -- (444.23, 17.23);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{drawColor}{RGB}{17,119,51}
\definecolor{fillColor}{RGB}{17,119,51}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (438.45, 17.23) circle (  1.96);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (476.35, 24.45) rectangle (490.80, 38.91);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{drawColor}{RGB}{221,204,119}

\path[draw=drawColor,line width= 0.6pt,line join=round] (477.79, 31.68) -- (489.35, 31.68);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{drawColor}{RGB}{221,204,119}
\definecolor{fillColor}{RGB}{221,204,119}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (483.57, 31.68) circle (  1.96);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (476.35, 10.00) rectangle (490.80, 24.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{drawColor}{RGB}{204,102,119}

\path[draw=drawColor,line width= 0.6pt,line join=round] (477.79, 17.23) -- (489.35, 17.23);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{drawColor}{RGB}{204,102,119}
\definecolor{fillColor}{RGB}{204,102,119}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (483.57, 17.23) circle (  1.96);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.80] at (322.35, 29.75) {Female};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.80] at (322.35, 15.30) {Male};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.80] at (366.35, 29.75) {Up to Secondary};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.80] at (366.35, 15.30) {More than Secondary};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.80] at (450.68, 29.75) {$< 25$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.80] at (450.68, 15.30) {$25-54$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.80] at (495.80, 29.75) {$55 +$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.80] at (495.80, 15.30) {All};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (650.43,542.02);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.20] at (  5.00,531.24) {If a COVID-19 vaccine becomes available in [country], would you take it?};
\end{scope}
\end{tikzpicture}

\end{document}