### 2021-06-22 paper_files/figure-latex/fig3paper-1.tex

% Created by tikzDevice version 0.12.3.1 on 2021-06-22 11:48:27
% !TEX encoding = UTF-8 Unicode
% !TEX program = xelatex

\documentclass[fontsize=7pt,DIV=calc]{scrartcl}

\nonstopmode

\usepackage{tikz}

\usepackage[active,tightpage,xetex]{preview}

\usepackage{fontspec,xunicode}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}

\setmainfont{Arial}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (722.70,650.43);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.00) rectangle (722.70,650.43);
\end{scope}
\begin{scope}
\path[clip] ( 41.25,541.96) rectangle (369.36,631.27);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 41.25,541.96) rectangle (369.36,631.27);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 41.25,557.39) --
	(369.36,557.39);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 41.25,580.12) --
	(369.36,580.12);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 41.25,602.85) --
	(369.36,602.85);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 41.25,625.57) --
	(369.36,625.57);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 41.25,546.02) --
	(369.36,546.02);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 41.25,568.75) --
	(369.36,568.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 41.25,591.48) --
	(369.36,591.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 41.25,614.21) --
	(369.36,614.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 68.60,541.96) --
	( 68.60,631.27);

\path[draw=drawColor,line width= 0.6pt,line join=round] (114.17,541.96) --
	(114.17,631.27);

\path[draw=drawColor,line width= 0.6pt,line join=round] (159.74,541.96) --
	(159.74,631.27);

\path[draw=drawColor,line width= 0.6pt,line join=round] (205.31,541.96) --
	(205.31,631.27);

\path[draw=drawColor,line width= 0.6pt,line join=round] (250.88,541.96) --
	(250.88,631.27);

\path[draw=drawColor,line width= 0.6pt,line join=round] (296.45,541.96) --
	(296.45,631.27);

\path[draw=drawColor,line width= 0.6pt,line join=round] (342.02,541.96) --
	(342.02,631.27);
\definecolor{fillColor}{RGB}{221,204,119}

\path[fill=fillColor] (139.23,546.02) rectangle (180.24,562.75);

\path[fill=fillColor] (230.37,546.02) rectangle (271.38,547.75);

\path[fill=fillColor] (184.80,546.02) rectangle (225.81,547.66);

\path[fill=fillColor] (275.94,546.02) rectangle (316.95,556.57);

\path[fill=fillColor] (321.51,546.02) rectangle (362.52,547.57);

\path[fill=fillColor] ( 48.09,546.02) rectangle ( 89.10,592.94);

\path[fill=fillColor] ( 93.66,546.02) rectangle (134.67,557.75);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 41.25,541.96) rectangle (369.36,631.27);
\end{scope}
\begin{scope}
\path[clip] ( 41.25,447.16) rectangle (369.36,536.46);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 41.25,447.16) rectangle (369.36,536.46);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 41.25,462.58) --
	(369.36,462.58);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 41.25,485.31) --
	(369.36,485.31);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 41.25,508.04) --
	(369.36,508.04);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 41.25,530.77) --
	(369.36,530.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 41.25,451.22) --
	(369.36,451.22);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 41.25,473.95) --
	(369.36,473.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 41.25,496.67) --
	(369.36,496.67);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 41.25,519.40) --
	(369.36,519.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 68.60,447.16) --
	( 68.60,536.46);

\path[draw=drawColor,line width= 0.6pt,line join=round] (114.17,447.16) --
	(114.17,536.46);

\path[draw=drawColor,line width= 0.6pt,line join=round] (159.74,447.16) --
	(159.74,536.46);

\path[draw=drawColor,line width= 0.6pt,line join=round] (205.31,447.16) --
	(205.31,536.46);

\path[draw=drawColor,line width= 0.6pt,line join=round] (250.88,447.16) --
	(250.88,536.46);

\path[draw=drawColor,line width= 0.6pt,line join=round] (296.45,447.16) --
	(296.45,536.46);

\path[draw=drawColor,line width= 0.6pt,line join=round] (342.02,447.16) --
	(342.02,536.46);
\definecolor{fillColor}{RGB}{221,204,119}

\path[fill=fillColor] (139.23,451.22) rectangle (180.24,483.58);

\path[fill=fillColor] (230.37,451.22) rectangle (271.38,451.58);

\path[fill=fillColor] (184.80,451.22) rectangle (225.81,465.85);

\path[fill=fillColor] (275.94,451.22) rectangle (316.95,451.67);

\path[fill=fillColor] (321.51,451.22) rectangle (362.52,453.40);

\path[fill=fillColor] ( 48.09,451.22) rectangle ( 89.10,491.40);

\path[fill=fillColor] ( 93.66,451.22) rectangle (134.67,451.94);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 41.25,447.16) rectangle (369.36,536.46);
\end{scope}
\begin{scope}
\path[clip] ( 41.25,352.35) rectangle (369.36,441.66);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 41.25,352.35) rectangle (369.36,441.66);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 41.25,367.78) --
	(369.36,367.78);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 41.25,390.50) --
	(369.36,390.50);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 41.25,413.23) --
	(369.36,413.23);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 41.25,435.96) --
	(369.36,435.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 41.25,356.41) --
	(369.36,356.41);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 41.25,379.14) --
	(369.36,379.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 41.25,401.87) --
	(369.36,401.87);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 41.25,424.60) --
	(369.36,424.60);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 68.60,352.35) --
	( 68.60,441.66);

\path[draw=drawColor,line width= 0.6pt,line join=round] (114.17,352.35) --
	(114.17,441.66);

\path[draw=drawColor,line width= 0.6pt,line join=round] (159.74,352.35) --
	(159.74,441.66);

\path[draw=drawColor,line width= 0.6pt,line join=round] (205.31,352.35) --
	(205.31,441.66);

\path[draw=drawColor,line width= 0.6pt,line join=round] (250.88,352.35) --
	(250.88,441.66);

\path[draw=drawColor,line width= 0.6pt,line join=round] (296.45,352.35) --
	(296.45,441.66);

\path[draw=drawColor,line width= 0.6pt,line join=round] (342.02,352.35) --
	(342.02,441.66);
\definecolor{fillColor}{RGB}{221,204,119}

\path[fill=fillColor] (139.23,356.41) rectangle (180.24,370.05);

\path[fill=fillColor] (230.37,356.41) rectangle (271.38,356.96);

\path[fill=fillColor] (184.80,356.41) rectangle (225.81,357.41);

\path[fill=fillColor] (275.94,356.41) rectangle (316.95,386.59);

\path[fill=fillColor] (321.51,356.41) rectangle (362.52,356.77);

\path[fill=fillColor] ( 48.09,356.41) rectangle ( 89.10,377.32);

\path[fill=fillColor] ( 93.66,356.41) rectangle (134.67,380.69);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 41.25,352.35) rectangle (369.36,441.66);
\end{scope}
\begin{scope}
\path[clip] ( 41.25,257.55) rectangle (369.36,346.85);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 41.25,257.55) rectangle (369.36,346.85);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 41.25,272.97) --
	(369.36,272.97);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 41.25,295.70) --
	(369.36,295.70);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 41.25,318.43) --
	(369.36,318.43);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 41.25,341.16) --
	(369.36,341.16);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 41.25,261.60) --
	(369.36,261.60);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 41.25,284.33) --
	(369.36,284.33);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 41.25,307.06) --
	(369.36,307.06);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 41.25,329.79) --
	(369.36,329.79);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 68.60,257.55) --
	( 68.60,346.85);

\path[draw=drawColor,line width= 0.6pt,line join=round] (114.17,257.55) --
	(114.17,346.85);

\path[draw=drawColor,line width= 0.6pt,line join=round] (159.74,257.55) --
	(159.74,346.85);

\path[draw=drawColor,line width= 0.6pt,line join=round] (205.31,257.55) --
	(205.31,346.85);

\path[draw=drawColor,line width= 0.6pt,line join=round] (250.88,257.55) --
	(250.88,346.85);

\path[draw=drawColor,line width= 0.6pt,line join=round] (296.45,257.55) --
	(296.45,346.85);

\path[draw=drawColor,line width= 0.6pt,line join=round] (342.02,257.55) --
	(342.02,346.85);
\definecolor{fillColor}{RGB}{221,204,119}

\path[fill=fillColor] (139.23,261.60) rectangle (180.24,264.42);

\path[fill=fillColor] (230.37,261.60) rectangle (271.38,261.60);

\path[fill=fillColor] (184.80,261.60) rectangle (225.81,263.15);

\path[fill=fillColor] (275.94,261.60) rectangle (316.95,266.88);

\path[fill=fillColor] ( 48.09,261.60) rectangle ( 89.10,342.79);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 41.25,257.55) rectangle (369.36,346.85);
\end{scope}
\begin{scope}
\path[clip] ( 41.25,162.74) rectangle (369.36,252.05);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 41.25,162.74) rectangle (369.36,252.05);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 41.25,178.16) --
	(369.36,178.16);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 41.25,200.89) --
	(369.36,200.89);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 41.25,223.62) --
	(369.36,223.62);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 41.25,246.35) --
	(369.36,246.35);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 41.25,166.80) --
	(369.36,166.80);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 41.25,189.53) --
	(369.36,189.53);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 41.25,212.26) --
	(369.36,212.26);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 41.25,234.98) --
	(369.36,234.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 68.60,162.74) --
	( 68.60,252.05);

\path[draw=drawColor,line width= 0.6pt,line join=round] (114.17,162.74) --
	(114.17,252.05);

\path[draw=drawColor,line width= 0.6pt,line join=round] (159.74,162.74) --
	(159.74,252.05);

\path[draw=drawColor,line width= 0.6pt,line join=round] (205.31,162.74) --
	(205.31,252.05);

\path[draw=drawColor,line width= 0.6pt,line join=round] (250.88,162.74) --
	(250.88,252.05);

\path[draw=drawColor,line width= 0.6pt,line join=round] (296.45,162.74) --
	(296.45,252.05);

\path[draw=drawColor,line width= 0.6pt,line join=round] (342.02,162.74) --
	(342.02,252.05);
\definecolor{fillColor}{RGB}{221,204,119}

\path[fill=fillColor] (139.23,166.80) rectangle (180.24,182.07);

\path[fill=fillColor] (230.37,166.80) rectangle (271.38,168.16);

\path[fill=fillColor] (184.80,166.80) rectangle (225.81,171.62);

\path[fill=fillColor] (275.94,166.80) rectangle (316.95,176.44);

\path[fill=fillColor] (321.51,166.80) rectangle (362.52,168.98);

\path[fill=fillColor] ( 48.09,166.80) rectangle ( 89.10,210.53);

\path[fill=fillColor] ( 93.66,166.80) rectangle (134.67,184.34);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 41.25,162.74) rectangle (369.36,252.05);
\end{scope}
\begin{scope}
\path[clip] ( 41.25, 67.93) rectangle (369.36,157.24);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 41.25, 67.93) rectangle (369.36,157.24);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 41.25, 83.36) --
	(369.36, 83.36);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 41.25,106.09) --
	(369.36,106.09);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 41.25,128.81) --
	(369.36,128.81);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 41.25,151.54) --
	(369.36,151.54);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 41.25, 71.99) --
	(369.36, 71.99);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 41.25, 94.72) --
	(369.36, 94.72);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 41.25,117.45) --
	(369.36,117.45);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 41.25,140.18) --
	(369.36,140.18);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 68.60, 67.93) --
	( 68.60,157.24);

\path[draw=drawColor,line width= 0.6pt,line join=round] (114.17, 67.93) --
	(114.17,157.24);

\path[draw=drawColor,line width= 0.6pt,line join=round] (159.74, 67.93) --
	(159.74,157.24);

\path[draw=drawColor,line width= 0.6pt,line join=round] (205.31, 67.93) --
	(205.31,157.24);

\path[draw=drawColor,line width= 0.6pt,line join=round] (250.88, 67.93) --
	(250.88,157.24);

\path[draw=drawColor,line width= 0.6pt,line join=round] (296.45, 67.93) --
	(296.45,157.24);

\path[draw=drawColor,line width= 0.6pt,line join=round] (342.02, 67.93) --
	(342.02,157.24);
\definecolor{fillColor}{RGB}{221,204,119}

\path[fill=fillColor] (139.23, 71.99) rectangle (180.24, 82.45);

\path[fill=fillColor] (184.80, 71.99) rectangle (225.81, 74.17);

\path[fill=fillColor] (275.94, 71.99) rectangle (316.95, 91.90);

\path[fill=fillColor] (321.51, 71.99) rectangle (362.52, 72.08);

\path[fill=fillColor] ( 48.09, 71.99) rectangle ( 89.10,103.36);

\path[fill=fillColor] ( 93.66, 71.99) rectangle (134.67, 98.99);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 41.25, 67.93) rectangle (369.36,157.24);
\end{scope}
\begin{scope}
\path[clip] (389.10,541.96) rectangle (717.20,631.27);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (389.10,541.96) rectangle (717.20,631.27);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] (389.10,557.39) --
	(717.20,557.39);

\path[draw=drawColor,line width= 0.3pt,line join=round] (389.10,580.12) --
	(717.20,580.12);

\path[draw=drawColor,line width= 0.3pt,line join=round] (389.10,602.85) --
	(717.20,602.85);

\path[draw=drawColor,line width= 0.3pt,line join=round] (389.10,625.57) --
	(717.20,625.57);

\path[draw=drawColor,line width= 0.6pt,line join=round] (389.10,546.02) --
	(717.20,546.02);

\path[draw=drawColor,line width= 0.6pt,line join=round] (389.10,568.75) --
	(717.20,568.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] (389.10,591.48) --
	(717.20,591.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] (389.10,614.21) --
	(717.20,614.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] (416.44,541.96) --
	(416.44,631.27);

\path[draw=drawColor,line width= 0.6pt,line join=round] (462.01,541.96) --
	(462.01,631.27);

\path[draw=drawColor,line width= 0.6pt,line join=round] (507.58,541.96) --
	(507.58,631.27);

\path[draw=drawColor,line width= 0.6pt,line join=round] (553.15,541.96) --
	(553.15,631.27);

\path[draw=drawColor,line width= 0.6pt,line join=round] (598.72,541.96) --
	(598.72,631.27);

\path[draw=drawColor,line width= 0.6pt,line join=round] (644.29,541.96) --
	(644.29,631.27);

\path[draw=drawColor,line width= 0.6pt,line join=round] (689.86,541.96) --
	(689.86,631.27);
\definecolor{fillColor}{RGB}{221,204,119}

\path[fill=fillColor] (487.07,546.02) rectangle (528.09,579.30);

\path[fill=fillColor] (578.21,546.02) rectangle (619.22,547.57);

\path[fill=fillColor] (532.64,546.02) rectangle (573.65,547.66);

\path[fill=fillColor] (669.35,546.02) rectangle (710.36,554.11);

\path[fill=fillColor] (395.93,546.02) rectangle (436.95,581.75);

\path[fill=fillColor] (441.50,546.02) rectangle (482.52,556.57);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (389.10,541.96) rectangle (717.20,631.27);
\end{scope}
\begin{scope}
\path[clip] (389.10,447.16) rectangle (717.20,536.46);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (389.10,447.16) rectangle (717.20,536.46);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] (389.10,462.58) --
	(717.20,462.58);

\path[draw=drawColor,line width= 0.3pt,line join=round] (389.10,485.31) --
	(717.20,485.31);

\path[draw=drawColor,line width= 0.3pt,line join=round] (389.10,508.04) --
	(717.20,508.04);

\path[draw=drawColor,line width= 0.3pt,line join=round] (389.10,530.77) --
	(717.20,530.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] (389.10,451.22) --
	(717.20,451.22);

\path[draw=drawColor,line width= 0.6pt,line join=round] (389.10,473.95) --
	(717.20,473.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] (389.10,496.67) --
	(717.20,496.67);

\path[draw=drawColor,line width= 0.6pt,line join=round] (389.10,519.40) --
	(717.20,519.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] (416.44,447.16) --
	(416.44,536.46);

\path[draw=drawColor,line width= 0.6pt,line join=round] (462.01,447.16) --
	(462.01,536.46);

\path[draw=drawColor,line width= 0.6pt,line join=round] (507.58,447.16) --
	(507.58,536.46);

\path[draw=drawColor,line width= 0.6pt,line join=round] (553.15,447.16) --
	(553.15,536.46);

\path[draw=drawColor,line width= 0.6pt,line join=round] (598.72,447.16) --
	(598.72,536.46);

\path[draw=drawColor,line width= 0.6pt,line join=round] (644.29,447.16) --
	(644.29,536.46);

\path[draw=drawColor,line width= 0.6pt,line join=round] (689.86,447.16) --
	(689.86,536.46);
\definecolor{fillColor}{RGB}{221,204,119}

\path[fill=fillColor] (487.07,451.22) rectangle (528.09,458.49);

\path[fill=fillColor] (532.64,451.22) rectangle (573.65,458.49);

\path[fill=fillColor] (623.78,451.22) rectangle (664.79,454.76);

\path[fill=fillColor] (669.35,451.22) rectangle (710.36,455.04);

\path[fill=fillColor] (395.93,451.22) rectangle (436.95,503.95);

\path[fill=fillColor] (441.50,451.22) rectangle (482.52,467.58);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (389.10,447.16) rectangle (717.20,536.46);
\end{scope}
\begin{scope}
\path[clip] (389.10,352.35) rectangle (717.20,441.66);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (389.10,352.35) rectangle (717.20,441.66);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] (389.10,367.78) --
	(717.20,367.78);

\path[draw=drawColor,line width= 0.3pt,line join=round] (389.10,390.50) --
	(717.20,390.50);

\path[draw=drawColor,line width= 0.3pt,line join=round] (389.10,413.23) --
	(717.20,413.23);

\path[draw=drawColor,line width= 0.3pt,line join=round] (389.10,435.96) --
	(717.20,435.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] (389.10,356.41) --
	(717.20,356.41);

\path[draw=drawColor,line width= 0.6pt,line join=round] (389.10,379.14) --
	(717.20,379.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] (389.10,401.87) --
	(717.20,401.87);

\path[draw=drawColor,line width= 0.6pt,line join=round] (389.10,424.60) --
	(717.20,424.60);

\path[draw=drawColor,line width= 0.6pt,line join=round] (416.44,352.35) --
	(416.44,441.66);

\path[draw=drawColor,line width= 0.6pt,line join=round] (462.01,352.35) --
	(462.01,441.66);

\path[draw=drawColor,line width= 0.6pt,line join=round] (507.58,352.35) --
	(507.58,441.66);

\path[draw=drawColor,line width= 0.6pt,line join=round] (553.15,352.35) --
	(553.15,441.66);

\path[draw=drawColor,line width= 0.6pt,line join=round] (598.72,352.35) --
	(598.72,441.66);

\path[draw=drawColor,line width= 0.6pt,line join=round] (644.29,352.35) --
	(644.29,441.66);

\path[draw=drawColor,line width= 0.6pt,line join=round] (689.86,352.35) --
	(689.86,441.66);
\definecolor{fillColor}{RGB}{221,204,119}

\path[fill=fillColor] (487.07,356.41) rectangle (528.09,364.05);

\path[fill=fillColor] (578.21,356.41) rectangle (619.22,356.87);

\path[fill=fillColor] (532.64,356.41) rectangle (573.65,360.68);

\path[fill=fillColor] (623.78,356.41) rectangle (664.79,364.50);

\path[fill=fillColor] (669.35,356.41) rectangle (710.36,357.41);

\path[fill=fillColor] (395.93,356.41) rectangle (436.95,396.41);

\path[fill=fillColor] (441.50,356.41) rectangle (482.52,385.96);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (389.10,352.35) rectangle (717.20,441.66);
\end{scope}
\begin{scope}
\path[clip] (389.10,257.55) rectangle (717.20,346.85);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (389.10,257.55) rectangle (717.20,346.85);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] (389.10,272.97) --
	(717.20,272.97);

\path[draw=drawColor,line width= 0.3pt,line join=round] (389.10,295.70) --
	(717.20,295.70);

\path[draw=drawColor,line width= 0.3pt,line join=round] (389.10,318.43) --
	(717.20,318.43);

\path[draw=drawColor,line width= 0.3pt,line join=round] (389.10,341.16) --
	(717.20,341.16);

\path[draw=drawColor,line width= 0.6pt,line join=round] (389.10,261.60) --
	(717.20,261.60);

\path[draw=drawColor,line width= 0.6pt,line join=round] (389.10,284.33) --
	(717.20,284.33);

\path[draw=drawColor,line width= 0.6pt,line join=round] (389.10,307.06) --
	(717.20,307.06);

\path[draw=drawColor,line width= 0.6pt,line join=round] (389.10,329.79) --
	(717.20,329.79);

\path[draw=drawColor,line width= 0.6pt,line join=round] (416.44,257.55) --
	(416.44,346.85);

\path[draw=drawColor,line width= 0.6pt,line join=round] (462.01,257.55) --
	(462.01,346.85);

\path[draw=drawColor,line width= 0.6pt,line join=round] (507.58,257.55) --
	(507.58,346.85);

\path[draw=drawColor,line width= 0.6pt,line join=round] (553.15,257.55) --
	(553.15,346.85);

\path[draw=drawColor,line width= 0.6pt,line join=round] (598.72,257.55) --
	(598.72,346.85);

\path[draw=drawColor,line width= 0.6pt,line join=round] (644.29,257.55) --
	(644.29,346.85);

\path[draw=drawColor,line width= 0.6pt,line join=round] (689.86,257.55) --
	(689.86,346.85);
\definecolor{fillColor}{RGB}{221,204,119}

\path[fill=fillColor] (487.07,261.60) rectangle (528.09,270.24);

\path[fill=fillColor] (578.21,261.60) rectangle (619.22,266.70);

\path[fill=fillColor] (532.64,261.60) rectangle (573.65,268.15);

\path[fill=fillColor] (623.78,261.60) rectangle (664.79,270.88);

\path[fill=fillColor] (669.35,261.60) rectangle (710.36,261.70);

\path[fill=fillColor] (395.93,261.60) rectangle (436.95,293.42);

\path[fill=fillColor] (441.50,261.60) rectangle (482.52,291.06);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (389.10,257.55) rectangle (717.20,346.85);
\end{scope}
\begin{scope}
\path[clip] (389.10,162.74) rectangle (717.20,252.05);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (389.10,162.74) rectangle (717.20,252.05);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] (389.10,178.16) --
	(717.20,178.16);

\path[draw=drawColor,line width= 0.3pt,line join=round] (389.10,200.89) --
	(717.20,200.89);

\path[draw=drawColor,line width= 0.3pt,line join=round] (389.10,223.62) --
	(717.20,223.62);

\path[draw=drawColor,line width= 0.3pt,line join=round] (389.10,246.35) --
	(717.20,246.35);

\path[draw=drawColor,line width= 0.6pt,line join=round] (389.10,166.80) --
	(717.20,166.80);

\path[draw=drawColor,line width= 0.6pt,line join=round] (389.10,189.53) --
	(717.20,189.53);

\path[draw=drawColor,line width= 0.6pt,line join=round] (389.10,212.26) --
	(717.20,212.26);

\path[draw=drawColor,line width= 0.6pt,line join=round] (389.10,234.98) --
	(717.20,234.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] (416.44,162.74) --
	(416.44,252.05);

\path[draw=drawColor,line width= 0.6pt,line join=round] (462.01,162.74) --
	(462.01,252.05);

\path[draw=drawColor,line width= 0.6pt,line join=round] (507.58,162.74) --
	(507.58,252.05);

\path[draw=drawColor,line width= 0.6pt,line join=round] (553.15,162.74) --
	(553.15,252.05);

\path[draw=drawColor,line width= 0.6pt,line join=round] (598.72,162.74) --
	(598.72,252.05);

\path[draw=drawColor,line width= 0.6pt,line join=round] (644.29,162.74) --
	(644.29,252.05);

\path[draw=drawColor,line width= 0.6pt,line join=round] (689.86,162.74) --
	(689.86,252.05);
\definecolor{fillColor}{RGB}{221,204,119}

\path[fill=fillColor] (487.07,166.80) rectangle (528.09,192.35);

\path[fill=fillColor] (578.21,166.80) rectangle (619.22,171.25);

\path[fill=fillColor] (532.64,166.80) rectangle (573.65,168.80);

\path[fill=fillColor] (623.78,166.80) rectangle (664.79,181.89);

\path[fill=fillColor] (395.93,166.80) rectangle (436.95,199.53);

\path[fill=fillColor] (441.50,166.80) rectangle (482.52,177.98);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (389.10,162.74) rectangle (717.20,252.05);
\end{scope}
\begin{scope}
\path[clip] (374.86,541.96) rectangle (389.10,631.27);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (374.86,541.96) rectangle (389.10,631.27);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (384.10,586.62) {Colombia};
\end{scope}
\begin{scope}
\path[clip] (374.86,447.16) rectangle (389.10,536.46);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (374.86,447.16) rectangle (389.10,536.46);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (384.10,491.81) {Nigeria};
\end{scope}
\begin{scope}
\path[clip] (374.86,352.35) rectangle (389.10,441.66);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (374.86,352.35) rectangle (389.10,441.66);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (384.10,397.00) {Sierra Leone 1};
\end{scope}
\begin{scope}
\path[clip] (374.86,257.55) rectangle (389.10,346.85);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (374.86,257.55) rectangle (389.10,346.85);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (384.10,302.20) {Uganda 2};
\end{scope}
\begin{scope}
\path[clip] (374.86,162.74) rectangle (389.10,252.05);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (374.86,162.74) rectangle (389.10,252.05);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (384.10,207.39) {Russia};
\end{scope}
\begin{scope}
\path[clip] ( 27.01,541.96) rectangle ( 41.25,631.27);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 27.01,541.96) rectangle ( 41.25,631.27);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 36.26,586.62) {Burkina Faso};
\end{scope}
\begin{scope}
\path[clip] ( 27.01,447.16) rectangle ( 41.25,536.46);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 27.01,447.16) rectangle ( 41.25,536.46);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 36.26,491.81) {Nepal};
\end{scope}
\begin{scope}
\path[clip] ( 27.01,352.35) rectangle ( 41.25,441.66);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 27.01,352.35) rectangle ( 41.25,441.66);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 36.26,397.00) {Rwanda};
\end{scope}
\begin{scope}
\path[clip] ( 27.01,257.55) rectangle ( 41.25,346.85);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 27.01,257.55) rectangle ( 41.25,346.85);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 36.26,302.20) {Sierra Leone 2};
\end{scope}
\begin{scope}
\path[clip] ( 27.01,162.74) rectangle ( 41.25,252.05);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 27.01,162.74) rectangle ( 41.25,252.05);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 36.26,207.39) {All LMICs};
\end{scope}
\begin{scope}
\path[clip] ( 27.01, 67.93) rectangle ( 41.25,157.24);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 27.01, 67.93) rectangle ( 41.25,157.24);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 36.26,112.59) {USA};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 68.60, 65.18) --
	( 68.60, 67.93);

\path[draw=drawColor,line width= 0.6pt,line join=round] (114.17, 65.18) --
	(114.17, 67.93);

\path[draw=drawColor,line width= 0.6pt,line join=round] (159.74, 65.18) --
	(159.74, 67.93);

\path[draw=drawColor,line width= 0.6pt,line join=round] (205.31, 65.18) --
	(205.31, 67.93);

\path[draw=drawColor,line width= 0.6pt,line join=round] (250.88, 65.18) --
	(250.88, 67.93);

\path[draw=drawColor,line width= 0.6pt,line join=round] (296.45, 65.18) --
	(296.45, 67.93);

\path[draw=drawColor,line width= 0.6pt,line join=round] (342.02, 65.18) --
	(342.02, 67.93);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 70.72, 62.98) {Health workers};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (111.53, 62.98) {Government or};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (121.04, 62.98) {MoH};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (157.10, 62.98) {Family or};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (166.61, 62.98) {Friends};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (193.17, 62.98) {Famous person,};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (202.67, 62.98) {religious leader};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (212.18, 62.98) {or traditional};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (221.68, 62.98) {healers};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (243.49, 62.98) {Newspapers,};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (253.00, 62.98) {radio or online};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (262.50, 62.98) {groups};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (298.57, 62.98) {Other};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (339.38, 62.98) {Don't know or};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (348.89, 62.98) {Refuse};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (416.44,159.99) --
	(416.44,162.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] (462.01,159.99) --
	(462.01,162.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] (507.58,159.99) --
	(507.58,162.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] (553.15,159.99) --
	(553.15,162.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] (598.72,159.99) --
	(598.72,162.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] (644.29,159.99) --
	(644.29,162.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] (689.86,159.99) --
	(689.86,162.74);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (418.56,157.79) {Health workers};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (459.38,157.79) {Government or};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (468.88,157.79) {MoH};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (504.95,157.79) {Family or};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (514.45,157.79) {Friends};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (541.01,157.79) {Famous person,};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (550.52,157.79) {religious leader};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (560.02,157.79) {or traditional};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (569.53,157.79) {healers};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (591.34,157.79) {Newspapers,};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (600.84,157.79) {radio or online};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (610.34,157.79) {groups};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (646.41,157.79) {Other};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (687.23,157.79) {Don't know or};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (696.73,157.79) {Refuse};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,543.90) {0};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,566.63) {25};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,589.36) {50};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,612.09) {75};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.26,546.02) --
	( 27.01,546.02);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.26,568.75) --
	( 27.01,568.75);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.26,591.48) --
	( 27.01,591.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.26,614.21) --
	( 27.01,614.21);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,449.10) {0};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,471.82) {25};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,494.55) {50};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,517.28) {75};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.26,451.22) --
	( 27.01,451.22);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.26,473.95) --
	( 27.01,473.95);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.26,496.67) --
	( 27.01,496.67);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.26,519.40) --
	( 27.01,519.40);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,354.29) {0};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,377.02) {25};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,399.75) {50};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,422.48) {75};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.26,356.41) --
	( 27.01,356.41);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.26,379.14) --
	( 27.01,379.14);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.26,401.87) --
	( 27.01,401.87);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.26,424.60) --
	( 27.01,424.60);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,259.48) {0};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,282.21) {25};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,304.94) {50};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,327.67) {75};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.26,261.60) --
	( 27.01,261.60);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.26,284.33) --
	( 27.01,284.33);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.26,307.06) --
	( 27.01,307.06);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.26,329.79) --
	( 27.01,329.79);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,164.68) {0};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,187.41) {25};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,210.13) {50};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,232.86) {75};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.26,166.80) --
	( 27.01,166.80);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.26,189.53) --
	( 27.01,189.53);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.26,212.26) --
	( 27.01,212.26);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.26,234.98) --
	( 27.01,234.98);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05, 69.87) {0};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05, 92.60) {25};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,115.33) {50};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,138.06) {75};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.26, 71.99) --
	( 27.01, 71.99);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.26, 94.72) --
	( 27.01, 94.72);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.26,117.45) --
	( 27.01,117.45);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.26,140.18) --
	( 27.01,140.18);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 10.80,349.60) {estimate};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.32] at (  5.50,638.57) {Which of the following people would you trust MOST to help you decide whether you would get a COVID-19 vaccine?};
\end{scope}
\end{tikzpicture}

\end{document}