### 2021-06-22 paper_files/figure-latex/histcattrust-1.tex

% Created by tikzDevice version 0.12.3.1 on 2021-06-22 11:48:47
% !TEX encoding = UTF-8 Unicode
% !TEX program = xelatex

\documentclass[fontsize=7pt,DIV=calc]{scrartcl}

\nonstopmode

\usepackage{tikz}

\usepackage[active,tightpage,xetex]{preview}

\usepackage{fontspec,xunicode}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}

\setmainfont{Arial}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (722.70,650.43);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.00) rectangle (722.70,650.43);
\end{scope}
\begin{scope}
\path[clip] ( 41.25,548.04) rectangle (369.36,631.27);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 41.25,548.04) rectangle (369.36,631.27);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 41.25,561.87) --
	(369.36,561.87);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 41.25,581.98) --
	(369.36,581.98);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 41.25,602.08) --
	(369.36,602.08);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 41.25,622.18) --
	(369.36,622.18);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 41.25,551.82) --
	(369.36,551.82);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 41.25,571.92) --
	(369.36,571.92);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 41.25,592.03) --
	(369.36,592.03);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 41.25,612.13) --
	(369.36,612.13);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 68.60,548.04) --
	( 68.60,631.27);

\path[draw=drawColor,line width= 0.6pt,line join=round] (114.17,548.04) --
	(114.17,631.27);

\path[draw=drawColor,line width= 0.6pt,line join=round] (159.74,548.04) --
	(159.74,631.27);

\path[draw=drawColor,line width= 0.6pt,line join=round] (205.31,548.04) --
	(205.31,631.27);

\path[draw=drawColor,line width= 0.6pt,line join=round] (250.88,548.04) --
	(250.88,631.27);

\path[draw=drawColor,line width= 0.6pt,line join=round] (296.45,548.04) --
	(296.45,631.27);

\path[draw=drawColor,line width= 0.6pt,line join=round] (342.02,548.04) --
	(342.02,631.27);
\definecolor{fillColor}{RGB}{221,204,119}

\path[fill=fillColor] (166.57,551.82) rectangle (180.24,566.62);

\path[fill=fillColor] (257.71,551.82) rectangle (271.38,553.35);

\path[fill=fillColor] (212.14,551.82) rectangle (225.81,553.27);

\path[fill=fillColor] (303.28,551.82) rectangle (316.95,561.15);

\path[fill=fillColor] (348.85,551.82) rectangle (362.52,553.19);

\path[fill=fillColor] ( 75.43,551.82) rectangle ( 89.10,593.31);

\path[fill=fillColor] (121.00,551.82) rectangle (134.67,562.20);
\definecolor{fillColor}{RGB}{17,119,51}

\path[fill=fillColor] (152.90,551.82) rectangle (166.57,567.58);

\path[fill=fillColor] (244.04,551.82) rectangle (257.71,553.43);

\path[fill=fillColor] (198.47,551.82) rectangle (212.14,552.55);

\path[fill=fillColor] (289.61,551.82) rectangle (303.28,555.68);

\path[fill=fillColor] (335.18,551.82) rectangle (348.85,552.14);

\path[fill=fillColor] ( 61.76,551.82) rectangle ( 75.43,597.74);

\path[fill=fillColor] (107.33,551.82) rectangle (121.00,563.96);
\definecolor{fillColor}{RGB}{204,102,119}

\path[fill=fillColor] (139.23,551.82) rectangle (152.90,564.85);

\path[fill=fillColor] (230.37,551.82) rectangle (244.04,553.11);

\path[fill=fillColor] (184.80,551.82) rectangle (198.47,554.80);

\path[fill=fillColor] (275.94,551.82) rectangle (289.61,572.01);

\path[fill=fillColor] (321.51,551.82) rectangle (335.18,555.20);

\path[fill=fillColor] ( 48.09,551.82) rectangle ( 61.76,584.55);

\path[fill=fillColor] ( 93.66,551.82) rectangle (107.33,558.66);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 41.25,548.04) rectangle (369.36,631.27);
\end{scope}
\begin{scope}
\path[clip] ( 41.25,459.31) rectangle (369.36,542.54);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 41.25,459.31) rectangle (369.36,542.54);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 41.25,473.14) --
	(369.36,473.14);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 41.25,493.25) --
	(369.36,493.25);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 41.25,513.35) --
	(369.36,513.35);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 41.25,533.45) --
	(369.36,533.45);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 41.25,463.09) --
	(369.36,463.09);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 41.25,483.19) --
	(369.36,483.19);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 41.25,503.30) --
	(369.36,503.30);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 41.25,523.40) --
	(369.36,523.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 68.60,459.31) --
	( 68.60,542.54);

\path[draw=drawColor,line width= 0.6pt,line join=round] (114.17,459.31) --
	(114.17,542.54);

\path[draw=drawColor,line width= 0.6pt,line join=round] (159.74,459.31) --
	(159.74,542.54);

\path[draw=drawColor,line width= 0.6pt,line join=round] (205.31,459.31) --
	(205.31,542.54);

\path[draw=drawColor,line width= 0.6pt,line join=round] (250.88,459.31) --
	(250.88,542.54);

\path[draw=drawColor,line width= 0.6pt,line join=round] (296.45,459.31) --
	(296.45,542.54);

\path[draw=drawColor,line width= 0.6pt,line join=round] (342.02,459.31) --
	(342.02,542.54);
\definecolor{fillColor}{RGB}{221,204,119}

\path[fill=fillColor] (166.57,463.09) rectangle (180.24,491.72);

\path[fill=fillColor] (250.88,463.09) rectangle (271.38,463.41);

\path[fill=fillColor] (212.14,463.09) rectangle (225.81,476.04);

\path[fill=fillColor] (303.28,463.09) rectangle (316.95,463.49);

\path[fill=fillColor] (348.85,463.09) rectangle (362.52,465.02);

\path[fill=fillColor] ( 75.43,463.09) rectangle ( 89.10,498.63);

\path[fill=fillColor] (121.00,463.09) rectangle (134.67,463.74);
\definecolor{fillColor}{RGB}{17,119,51}

\path[fill=fillColor] (152.90,463.09) rectangle (166.57,492.20);

\path[fill=fillColor] (230.37,463.09) rectangle (250.88,463.41);

\path[fill=fillColor] (198.47,463.09) rectangle (212.14,476.04);

\path[fill=fillColor] (289.61,463.09) rectangle (303.28,463.49);

\path[fill=fillColor] (335.18,463.09) rectangle (348.85,464.14);

\path[fill=fillColor] ( 61.76,463.09) rectangle ( 75.43,499.03);

\path[fill=fillColor] (107.33,463.09) rectangle (121.00,463.65);
\definecolor{fillColor}{RGB}{204,102,119}

\path[fill=fillColor] (139.23,463.09) rectangle (152.90,478.13);

\path[fill=fillColor] (184.80,463.09) rectangle (198.47,476.60);

\path[fill=fillColor] (275.94,463.09) rectangle (289.61,463.90);

\path[fill=fillColor] (321.51,463.09) rectangle (335.18,488.18);

\path[fill=fillColor] ( 48.09,463.09) rectangle ( 61.76,487.38);

\path[fill=fillColor] ( 93.66,463.09) rectangle (107.33,464.78);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 41.25,459.31) rectangle (369.36,542.54);
\end{scope}
\begin{scope}
\path[clip] ( 41.25,370.58) rectangle (369.36,453.81);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 41.25,370.58) rectangle (369.36,453.81);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 41.25,384.41) --
	(369.36,384.41);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 41.25,404.51) --
	(369.36,404.51);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 41.25,424.62) --
	(369.36,424.62);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 41.25,444.72) --
	(369.36,444.72);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 41.25,374.36) --
	(369.36,374.36);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 41.25,394.46) --
	(369.36,394.46);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 41.25,414.57) --
	(369.36,414.57);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 41.25,434.67) --
	(369.36,434.67);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 68.60,370.58) --
	( 68.60,453.81);

\path[draw=drawColor,line width= 0.6pt,line join=round] (114.17,370.58) --
	(114.17,453.81);

\path[draw=drawColor,line width= 0.6pt,line join=round] (159.74,370.58) --
	(159.74,453.81);

\path[draw=drawColor,line width= 0.6pt,line join=round] (205.31,370.58) --
	(205.31,453.81);

\path[draw=drawColor,line width= 0.6pt,line join=round] (250.88,370.58) --
	(250.88,453.81);

\path[draw=drawColor,line width= 0.6pt,line join=round] (296.45,370.58) --
	(296.45,453.81);

\path[draw=drawColor,line width= 0.6pt,line join=round] (342.02,370.58) --
	(342.02,453.81);
\definecolor{fillColor}{RGB}{221,204,119}

\path[fill=fillColor] (166.57,374.36) rectangle (180.24,386.42);

\path[fill=fillColor] (250.88,374.36) rectangle (271.38,374.84);

\path[fill=fillColor] (212.14,374.36) rectangle (225.81,375.25);

\path[fill=fillColor] (303.28,374.36) rectangle (316.95,401.06);

\path[fill=fillColor] (348.85,374.36) rectangle (362.52,374.68);

\path[fill=fillColor] ( 75.43,374.36) rectangle ( 89.10,392.86);

\path[fill=fillColor] (121.00,374.36) rectangle (134.67,395.83);
\definecolor{fillColor}{RGB}{17,119,51}

\path[fill=fillColor] (152.90,374.36) rectangle (166.57,386.50);

\path[fill=fillColor] (230.37,374.36) rectangle (250.88,374.92);

\path[fill=fillColor] (198.47,374.36) rectangle (212.14,375.17);

\path[fill=fillColor] (289.61,374.36) rectangle (303.28,400.09);

\path[fill=fillColor] (335.18,374.36) rectangle (348.85,374.44);

\path[fill=fillColor] ( 61.76,374.36) rectangle ( 75.43,393.50);

\path[fill=fillColor] (107.33,374.36) rectangle (121.00,396.39);
\definecolor{fillColor}{RGB}{204,102,119}

\path[fill=fillColor] (139.23,374.36) rectangle (152.90,384.65);

\path[fill=fillColor] (184.80,374.36) rectangle (198.47,376.69);

\path[fill=fillColor] (275.94,374.36) rectangle (289.61,417.14);

\path[fill=fillColor] (321.51,374.36) rectangle (335.18,378.78);

\path[fill=fillColor] ( 48.09,374.36) rectangle ( 61.76,382.48);

\path[fill=fillColor] ( 93.66,374.36) rectangle (107.33,386.91);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 41.25,370.58) rectangle (369.36,453.81);
\end{scope}
\begin{scope}
\path[clip] ( 41.25,281.85) rectangle (369.36,365.08);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 41.25,281.85) rectangle (369.36,365.08);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 41.25,295.68) --
	(369.36,295.68);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 41.25,315.78) --
	(369.36,315.78);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 41.25,335.89) --
	(369.36,335.89);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 41.25,355.99) --
	(369.36,355.99);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 41.25,285.63) --
	(369.36,285.63);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 41.25,305.73) --
	(369.36,305.73);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 41.25,325.84) --
	(369.36,325.84);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 41.25,345.94) --
	(369.36,345.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 68.60,281.85) --
	( 68.60,365.08);

\path[draw=drawColor,line width= 0.6pt,line join=round] (114.17,281.85) --
	(114.17,365.08);

\path[draw=drawColor,line width= 0.6pt,line join=round] (159.74,281.85) --
	(159.74,365.08);

\path[draw=drawColor,line width= 0.6pt,line join=round] (205.31,281.85) --
	(205.31,365.08);

\path[draw=drawColor,line width= 0.6pt,line join=round] (250.88,281.85) --
	(250.88,365.08);

\path[draw=drawColor,line width= 0.6pt,line join=round] (296.45,281.85) --
	(296.45,365.08);

\path[draw=drawColor,line width= 0.6pt,line join=round] (342.02,281.85) --
	(342.02,365.08);
\definecolor{fillColor}{RGB}{221,204,119}

\path[fill=fillColor] (166.57,285.63) rectangle (180.24,288.12);

\path[fill=fillColor] (250.88,285.63) rectangle (271.38,285.63);

\path[fill=fillColor] (212.14,285.63) rectangle (225.81,287.00);

\path[fill=fillColor] (303.28,285.63) rectangle (316.95,290.29);

\path[fill=fillColor] ( 75.43,285.63) rectangle ( 89.10,357.44);
\definecolor{fillColor}{RGB}{17,119,51}

\path[fill=fillColor] (152.90,285.63) rectangle (166.57,288.04);

\path[fill=fillColor] (230.37,285.63) rectangle (250.88,285.71);

\path[fill=fillColor] (198.47,285.63) rectangle (212.14,286.35);

\path[fill=fillColor] (289.61,285.63) rectangle (303.28,287.16);

\path[fill=fillColor] ( 61.76,285.63) rectangle ( 75.43,361.30);
\definecolor{fillColor}{RGB}{204,102,119}

\path[fill=fillColor] (139.23,285.63) rectangle (152.90,288.77);

\path[fill=fillColor] (184.80,285.63) rectangle (198.47,291.66);

\path[fill=fillColor] (275.94,285.63) rectangle (289.61,312.89);

\path[fill=fillColor] ( 48.09,285.63) rectangle ( 61.76,329.61);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 41.25,281.85) rectangle (369.36,365.08);
\end{scope}
\begin{scope}
\path[clip] ( 41.25,193.12) rectangle (369.36,276.35);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 41.25,193.12) rectangle (369.36,276.35);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 41.25,206.95) --
	(369.36,206.95);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 41.25,227.05) --
	(369.36,227.05);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 41.25,247.16) --
	(369.36,247.16);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 41.25,267.26) --
	(369.36,267.26);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 41.25,196.90) --
	(369.36,196.90);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 41.25,217.00) --
	(369.36,217.00);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 41.25,237.10) --
	(369.36,237.10);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 41.25,257.21) --
	(369.36,257.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 68.60,193.12) --
	( 68.60,276.35);

\path[draw=drawColor,line width= 0.6pt,line join=round] (114.17,193.12) --
	(114.17,276.35);

\path[draw=drawColor,line width= 0.6pt,line join=round] (159.74,193.12) --
	(159.74,276.35);

\path[draw=drawColor,line width= 0.6pt,line join=round] (205.31,193.12) --
	(205.31,276.35);

\path[draw=drawColor,line width= 0.6pt,line join=round] (250.88,193.12) --
	(250.88,276.35);

\path[draw=drawColor,line width= 0.6pt,line join=round] (296.45,193.12) --
	(296.45,276.35);

\path[draw=drawColor,line width= 0.6pt,line join=round] (342.02,193.12) --
	(342.02,276.35);
\definecolor{fillColor}{RGB}{221,204,119}

\path[fill=fillColor] (166.57,196.90) rectangle (180.24,210.41);

\path[fill=fillColor] (257.71,196.90) rectangle (271.38,198.11);

\path[fill=fillColor] (212.14,196.90) rectangle (225.81,201.16);

\path[fill=fillColor] (303.28,196.90) rectangle (316.95,205.42);

\path[fill=fillColor] (348.85,196.90) rectangle (362.52,198.83);

\path[fill=fillColor] ( 75.43,196.90) rectangle ( 89.10,235.58);

\path[fill=fillColor] (121.00,196.90) rectangle (134.67,212.42);
\definecolor{fillColor}{RGB}{17,119,51}

\path[fill=fillColor] (152.90,196.90) rectangle (166.57,210.41);

\path[fill=fillColor] (244.04,196.90) rectangle (257.71,198.11);

\path[fill=fillColor] (198.47,196.90) rectangle (212.14,200.52);

\path[fill=fillColor] (289.61,196.90) rectangle (303.28,202.45);

\path[fill=fillColor] (335.18,196.90) rectangle (348.85,197.87);

\path[fill=fillColor] ( 61.76,196.90) rectangle ( 75.43,238.15);

\path[fill=fillColor] (107.33,196.90) rectangle (121.00,214.27);
\definecolor{fillColor}{RGB}{204,102,119}

\path[fill=fillColor] (139.23,196.90) rectangle (152.90,209.12);

\path[fill=fillColor] (230.37,196.90) rectangle (244.04,198.19);

\path[fill=fillColor] (184.80,196.90) rectangle (198.47,203.90);

\path[fill=fillColor] (275.94,196.90) rectangle (289.61,217.89);

\path[fill=fillColor] (321.51,196.90) rectangle (335.18,204.86);

\path[fill=fillColor] ( 48.09,196.90) rectangle ( 61.76,223.03);

\path[fill=fillColor] ( 93.66,196.90) rectangle (107.33,205.58);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 41.25,193.12) rectangle (369.36,276.35);
\end{scope}
\begin{scope}
\path[clip] ( 41.25,104.39) rectangle (369.36,187.62);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 41.25,104.39) rectangle (369.36,187.62);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 41.25,118.22) --
	(369.36,118.22);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 41.25,138.32) --
	(369.36,138.32);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 41.25,158.43) --
	(369.36,158.43);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 41.25,178.53) --
	(369.36,178.53);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 41.25,108.17) --
	(369.36,108.17);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 41.25,128.27) --
	(369.36,128.27);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 41.25,148.37) --
	(369.36,148.37);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 41.25,168.48) --
	(369.36,168.48);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 68.60,104.39) --
	( 68.60,187.62);

\path[draw=drawColor,line width= 0.6pt,line join=round] (114.17,104.39) --
	(114.17,187.62);

\path[draw=drawColor,line width= 0.6pt,line join=round] (159.74,104.39) --
	(159.74,187.62);

\path[draw=drawColor,line width= 0.6pt,line join=round] (205.31,104.39) --
	(205.31,187.62);

\path[draw=drawColor,line width= 0.6pt,line join=round] (250.88,104.39) --
	(250.88,187.62);

\path[draw=drawColor,line width= 0.6pt,line join=round] (296.45,104.39) --
	(296.45,187.62);

\path[draw=drawColor,line width= 0.6pt,line join=round] (342.02,104.39) --
	(342.02,187.62);
\definecolor{fillColor}{RGB}{221,204,119}

\path[fill=fillColor] (166.57,108.17) rectangle (180.24,117.42);

\path[fill=fillColor] (212.14,108.17) rectangle (225.81,110.10);

\path[fill=fillColor] (303.28,108.17) rectangle (316.95,125.78);

\path[fill=fillColor] (342.02,108.17) rectangle (362.52,108.25);

\path[fill=fillColor] ( 75.43,108.17) rectangle ( 89.10,135.91);

\path[fill=fillColor] (121.00,108.17) rectangle (134.67,132.05);
\definecolor{fillColor}{RGB}{17,119,51}

\path[fill=fillColor] (152.90,108.17) rectangle (166.57,115.17);

\path[fill=fillColor] (198.47,108.17) rectangle (212.14,109.54);

\path[fill=fillColor] (289.61,108.17) rectangle (303.28,123.13);

\path[fill=fillColor] ( 61.76,108.17) rectangle ( 75.43,138.81);

\path[fill=fillColor] (107.33,108.17) rectangle (121.00,134.70);
\definecolor{fillColor}{RGB}{204,102,119}

\path[fill=fillColor] (139.23,108.17) rectangle (152.90,123.21);

\path[fill=fillColor] (184.80,108.17) rectangle (198.47,111.55);

\path[fill=fillColor] (275.94,108.17) rectangle (289.61,132.53);

\path[fill=fillColor] (321.51,108.17) rectangle (342.02,108.33);

\path[fill=fillColor] ( 48.09,108.17) rectangle ( 61.76,128.51);

\path[fill=fillColor] ( 93.66,108.17) rectangle (107.33,125.30);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 41.25,104.39) rectangle (369.36,187.62);
\end{scope}
\begin{scope}
\path[clip] (389.10,548.04) rectangle (717.20,631.27);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (389.10,548.04) rectangle (717.20,631.27);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] (389.10,561.87) --
	(717.20,561.87);

\path[draw=drawColor,line width= 0.3pt,line join=round] (389.10,581.98) --
	(717.20,581.98);

\path[draw=drawColor,line width= 0.3pt,line join=round] (389.10,602.08) --
	(717.20,602.08);

\path[draw=drawColor,line width= 0.3pt,line join=round] (389.10,622.18) --
	(717.20,622.18);

\path[draw=drawColor,line width= 0.6pt,line join=round] (389.10,551.82) --
	(717.20,551.82);

\path[draw=drawColor,line width= 0.6pt,line join=round] (389.10,571.92) --
	(717.20,571.92);

\path[draw=drawColor,line width= 0.6pt,line join=round] (389.10,592.03) --
	(717.20,592.03);

\path[draw=drawColor,line width= 0.6pt,line join=round] (389.10,612.13) --
	(717.20,612.13);

\path[draw=drawColor,line width= 0.6pt,line join=round] (416.44,548.04) --
	(416.44,631.27);

\path[draw=drawColor,line width= 0.6pt,line join=round] (462.01,548.04) --
	(462.01,631.27);

\path[draw=drawColor,line width= 0.6pt,line join=round] (507.58,548.04) --
	(507.58,631.27);

\path[draw=drawColor,line width= 0.6pt,line join=round] (553.15,548.04) --
	(553.15,631.27);

\path[draw=drawColor,line width= 0.6pt,line join=round] (598.72,548.04) --
	(598.72,631.27);

\path[draw=drawColor,line width= 0.6pt,line join=round] (644.29,548.04) --
	(644.29,631.27);

\path[draw=drawColor,line width= 0.6pt,line join=round] (689.86,548.04) --
	(689.86,631.27);
\definecolor{fillColor}{RGB}{221,204,119}

\path[fill=fillColor] (514.41,551.82) rectangle (528.09,581.25);

\path[fill=fillColor] (605.55,551.82) rectangle (619.22,553.19);

\path[fill=fillColor] (559.98,551.82) rectangle (573.65,553.27);

\path[fill=fillColor] (696.69,551.82) rectangle (710.36,558.98);

\path[fill=fillColor] (423.27,551.82) rectangle (436.95,583.42);

\path[fill=fillColor] (468.84,551.82) rectangle (482.52,561.15);
\definecolor{fillColor}{RGB}{17,119,51}

\path[fill=fillColor] (500.74,551.82) rectangle (514.41,581.49);

\path[fill=fillColor] (591.88,551.82) rectangle (605.55,553.19);

\path[fill=fillColor] (546.31,551.82) rectangle (559.98,552.55);

\path[fill=fillColor] (683.02,551.82) rectangle (696.69,556.89);

\path[fill=fillColor] (409.60,551.82) rectangle (423.27,585.11);

\path[fill=fillColor] (455.17,551.82) rectangle (468.84,562.03);
\definecolor{fillColor}{RGB}{204,102,119}

\path[fill=fillColor] (487.07,551.82) rectangle (500.74,580.37);

\path[fill=fillColor] (578.21,551.82) rectangle (591.88,552.95);

\path[fill=fillColor] (532.64,551.82) rectangle (546.31,556.08);

\path[fill=fillColor] (669.35,551.82) rectangle (683.02,566.94);

\path[fill=fillColor] (395.93,551.82) rectangle (409.60,577.15);

\path[fill=fillColor] (441.50,551.82) rectangle (455.17,557.93);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (389.10,548.04) rectangle (717.20,631.27);
\end{scope}
\begin{scope}
\path[clip] (389.10,459.31) rectangle (717.20,542.54);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (389.10,459.31) rectangle (717.20,542.54);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] (389.10,473.14) --
	(717.20,473.14);

\path[draw=drawColor,line width= 0.3pt,line join=round] (389.10,493.25) --
	(717.20,493.25);

\path[draw=drawColor,line width= 0.3pt,line join=round] (389.10,513.35) --
	(717.20,513.35);

\path[draw=drawColor,line width= 0.3pt,line join=round] (389.10,533.45) --
	(717.20,533.45);

\path[draw=drawColor,line width= 0.6pt,line join=round] (389.10,463.09) --
	(717.20,463.09);

\path[draw=drawColor,line width= 0.6pt,line join=round] (389.10,483.19) --
	(717.20,483.19);

\path[draw=drawColor,line width= 0.6pt,line join=round] (389.10,503.30) --
	(717.20,503.30);

\path[draw=drawColor,line width= 0.6pt,line join=round] (389.10,523.40) --
	(717.20,523.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] (416.44,459.31) --
	(416.44,542.54);

\path[draw=drawColor,line width= 0.6pt,line join=round] (462.01,459.31) --
	(462.01,542.54);

\path[draw=drawColor,line width= 0.6pt,line join=round] (507.58,459.31) --
	(507.58,542.54);

\path[draw=drawColor,line width= 0.6pt,line join=round] (553.15,459.31) --
	(553.15,542.54);

\path[draw=drawColor,line width= 0.6pt,line join=round] (598.72,459.31) --
	(598.72,542.54);

\path[draw=drawColor,line width= 0.6pt,line join=round] (644.29,459.31) --
	(644.29,542.54);

\path[draw=drawColor,line width= 0.6pt,line join=round] (689.86,459.31) --
	(689.86,542.54);
\definecolor{fillColor}{RGB}{221,204,119}

\path[fill=fillColor] (514.41,463.09) rectangle (528.09,469.52);

\path[fill=fillColor] (559.98,463.09) rectangle (573.65,469.52);

\path[fill=fillColor] (651.12,463.09) rectangle (664.79,466.23);

\path[fill=fillColor] (696.69,463.09) rectangle (710.36,466.47);

\path[fill=fillColor] (423.27,463.09) rectangle (436.95,509.73);

\path[fill=fillColor] (468.84,463.09) rectangle (482.52,477.57);
\definecolor{fillColor}{RGB}{17,119,51}

\path[fill=fillColor] (500.74,463.09) rectangle (514.41,468.16);

\path[fill=fillColor] (546.31,463.09) rectangle (559.98,467.19);

\path[fill=fillColor] (637.45,463.09) rectangle (651.12,465.18);

\path[fill=fillColor] (683.02,463.09) rectangle (696.69,463.57);

\path[fill=fillColor] (409.60,463.09) rectangle (423.27,514.39);

\path[fill=fillColor] (455.17,463.09) rectangle (468.84,480.46);
\definecolor{fillColor}{RGB}{204,102,119}

\path[fill=fillColor] (487.07,463.09) rectangle (500.74,474.27);

\path[fill=fillColor] (532.64,463.09) rectangle (546.31,477.40);

\path[fill=fillColor] (623.78,463.09) rectangle (637.45,469.93);

\path[fill=fillColor] (669.35,463.09) rectangle (683.02,476.44);

\path[fill=fillColor] (395.93,463.09) rectangle (409.60,493.33);

\path[fill=fillColor] (441.50,463.09) rectangle (455.17,467.59);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (389.10,459.31) rectangle (717.20,542.54);
\end{scope}
\begin{scope}
\path[clip] (389.10,370.58) rectangle (717.20,453.81);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (389.10,370.58) rectangle (717.20,453.81);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] (389.10,384.41) --
	(717.20,384.41);

\path[draw=drawColor,line width= 0.3pt,line join=round] (389.10,404.51) --
	(717.20,404.51);

\path[draw=drawColor,line width= 0.3pt,line join=round] (389.10,424.62) --
	(717.20,424.62);

\path[draw=drawColor,line width= 0.3pt,line join=round] (389.10,444.72) --
	(717.20,444.72);

\path[draw=drawColor,line width= 0.6pt,line join=round] (389.10,374.36) --
	(717.20,374.36);

\path[draw=drawColor,line width= 0.6pt,line join=round] (389.10,394.46) --
	(717.20,394.46);

\path[draw=drawColor,line width= 0.6pt,line join=round] (389.10,414.57) --
	(717.20,414.57);

\path[draw=drawColor,line width= 0.6pt,line join=round] (389.10,434.67) --
	(717.20,434.67);

\path[draw=drawColor,line width= 0.6pt,line join=round] (416.44,370.58) --
	(416.44,453.81);

\path[draw=drawColor,line width= 0.6pt,line join=round] (462.01,370.58) --
	(462.01,453.81);

\path[draw=drawColor,line width= 0.6pt,line join=round] (507.58,370.58) --
	(507.58,453.81);

\path[draw=drawColor,line width= 0.6pt,line join=round] (553.15,370.58) --
	(553.15,453.81);

\path[draw=drawColor,line width= 0.6pt,line join=round] (598.72,370.58) --
	(598.72,453.81);

\path[draw=drawColor,line width= 0.6pt,line join=round] (644.29,370.58) --
	(644.29,453.81);

\path[draw=drawColor,line width= 0.6pt,line join=round] (689.86,370.58) --
	(689.86,453.81);
\definecolor{fillColor}{RGB}{221,204,119}

\path[fill=fillColor] (514.41,374.36) rectangle (528.09,381.12);

\path[fill=fillColor] (605.55,374.36) rectangle (619.22,374.76);

\path[fill=fillColor] (559.98,374.36) rectangle (573.65,378.14);

\path[fill=fillColor] (651.12,374.36) rectangle (664.79,381.52);

\path[fill=fillColor] (696.69,374.36) rectangle (710.36,375.25);

\path[fill=fillColor] (423.27,374.36) rectangle (436.95,409.74);

\path[fill=fillColor] (468.84,374.36) rectangle (482.52,400.49);
\definecolor{fillColor}{RGB}{17,119,51}

\path[fill=fillColor] (500.74,374.36) rectangle (514.41,380.23);

\path[fill=fillColor] (591.88,374.36) rectangle (605.55,374.76);

\path[fill=fillColor] (546.31,374.36) rectangle (559.98,377.42);

\path[fill=fillColor] (637.45,374.36) rectangle (651.12,376.85);

\path[fill=fillColor] (683.02,374.36) rectangle (696.69,375.00);

\path[fill=fillColor] (409.60,374.36) rectangle (423.27,412.64);

\path[fill=fillColor] (455.17,374.36) rectangle (468.84,404.03);
\definecolor{fillColor}{RGB}{204,102,119}

\path[fill=fillColor] (487.07,374.36) rectangle (500.74,384.09);

\path[fill=fillColor] (578.21,374.36) rectangle (591.88,374.76);

\path[fill=fillColor] (532.64,374.36) rectangle (546.31,380.55);

\path[fill=fillColor] (623.78,374.36) rectangle (637.45,398.00);

\path[fill=fillColor] (669.35,374.36) rectangle (683.02,376.13);

\path[fill=fillColor] (395.93,374.36) rectangle (409.60,399.37);

\path[fill=fillColor] (441.50,374.36) rectangle (455.17,388.11);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (389.10,370.58) rectangle (717.20,453.81);
\end{scope}
\begin{scope}
\path[clip] (389.10,281.85) rectangle (717.20,365.08);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (389.10,281.85) rectangle (717.20,365.08);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] (389.10,295.68) --
	(717.20,295.68);

\path[draw=drawColor,line width= 0.3pt,line join=round] (389.10,315.78) --
	(717.20,315.78);

\path[draw=drawColor,line width= 0.3pt,line join=round] (389.10,335.89) --
	(717.20,335.89);

\path[draw=drawColor,line width= 0.3pt,line join=round] (389.10,355.99) --
	(717.20,355.99);

\path[draw=drawColor,line width= 0.6pt,line join=round] (389.10,285.63) --
	(717.20,285.63);

\path[draw=drawColor,line width= 0.6pt,line join=round] (389.10,305.73) --
	(717.20,305.73);

\path[draw=drawColor,line width= 0.6pt,line join=round] (389.10,325.84) --
	(717.20,325.84);

\path[draw=drawColor,line width= 0.6pt,line join=round] (389.10,345.94) --
	(717.20,345.94);

\path[draw=drawColor,line width= 0.6pt,line join=round] (416.44,281.85) --
	(416.44,365.08);

\path[draw=drawColor,line width= 0.6pt,line join=round] (462.01,281.85) --
	(462.01,365.08);

\path[draw=drawColor,line width= 0.6pt,line join=round] (507.58,281.85) --
	(507.58,365.08);

\path[draw=drawColor,line width= 0.6pt,line join=round] (553.15,281.85) --
	(553.15,365.08);

\path[draw=drawColor,line width= 0.6pt,line join=round] (598.72,281.85) --
	(598.72,365.08);

\path[draw=drawColor,line width= 0.6pt,line join=round] (644.29,281.85) --
	(644.29,365.08);

\path[draw=drawColor,line width= 0.6pt,line join=round] (689.86,281.85) --
	(689.86,365.08);
\definecolor{fillColor}{RGB}{221,204,119}

\path[fill=fillColor] (514.41,285.63) rectangle (528.09,293.27);

\path[fill=fillColor] (605.55,285.63) rectangle (619.22,290.13);

\path[fill=fillColor] (559.98,285.63) rectangle (573.65,291.42);

\path[fill=fillColor] (651.12,285.63) rectangle (664.79,293.83);

\path[fill=fillColor] (689.86,285.63) rectangle (710.36,285.71);

\path[fill=fillColor] (423.27,285.63) rectangle (436.95,313.77);

\path[fill=fillColor] (468.84,285.63) rectangle (482.52,311.68);
\definecolor{fillColor}{RGB}{17,119,51}

\path[fill=fillColor] (500.74,285.63) rectangle (514.41,293.51);

\path[fill=fillColor] (591.88,285.63) rectangle (605.55,289.65);

\path[fill=fillColor] (546.31,285.63) rectangle (559.98,291.26);

\path[fill=fillColor] (637.45,285.63) rectangle (651.12,288.45);

\path[fill=fillColor] (409.60,285.63) rectangle (423.27,316.43);

\path[fill=fillColor] (455.17,285.63) rectangle (468.84,314.98);
\definecolor{fillColor}{RGB}{204,102,119}

\path[fill=fillColor] (487.07,285.63) rectangle (500.74,292.47);

\path[fill=fillColor] (578.21,285.63) rectangle (591.88,291.66);

\path[fill=fillColor] (532.64,285.63) rectangle (546.31,291.90);

\path[fill=fillColor] (623.78,285.63) rectangle (637.45,311.36);

\path[fill=fillColor] (669.35,285.63) rectangle (689.86,286.11);

\path[fill=fillColor] (395.93,285.63) rectangle (409.60,305.33);

\path[fill=fillColor] (441.50,285.63) rectangle (455.17,300.99);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (389.10,281.85) rectangle (717.20,365.08);
\end{scope}
\begin{scope}
\path[clip] (389.10,193.12) rectangle (717.20,276.35);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (389.10,193.12) rectangle (717.20,276.35);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] (389.10,206.95) --
	(717.20,206.95);

\path[draw=drawColor,line width= 0.3pt,line join=round] (389.10,227.05) --
	(717.20,227.05);

\path[draw=drawColor,line width= 0.3pt,line join=round] (389.10,247.16) --
	(717.20,247.16);

\path[draw=drawColor,line width= 0.3pt,line join=round] (389.10,267.26) --
	(717.20,267.26);

\path[draw=drawColor,line width= 0.6pt,line join=round] (389.10,196.90) --
	(717.20,196.90);

\path[draw=drawColor,line width= 0.6pt,line join=round] (389.10,217.00) --
	(717.20,217.00);

\path[draw=drawColor,line width= 0.6pt,line join=round] (389.10,237.10) --
	(717.20,237.10);

\path[draw=drawColor,line width= 0.6pt,line join=round] (389.10,257.21) --
	(717.20,257.21);

\path[draw=drawColor,line width= 0.6pt,line join=round] (416.44,193.12) --
	(416.44,276.35);

\path[draw=drawColor,line width= 0.6pt,line join=round] (462.01,193.12) --
	(462.01,276.35);

\path[draw=drawColor,line width= 0.6pt,line join=round] (507.58,193.12) --
	(507.58,276.35);

\path[draw=drawColor,line width= 0.6pt,line join=round] (553.15,193.12) --
	(553.15,276.35);

\path[draw=drawColor,line width= 0.6pt,line join=round] (598.72,193.12) --
	(598.72,276.35);

\path[draw=drawColor,line width= 0.6pt,line join=round] (644.29,193.12) --
	(644.29,276.35);

\path[draw=drawColor,line width= 0.6pt,line join=round] (689.86,193.12) --
	(689.86,276.35);
\definecolor{fillColor}{RGB}{221,204,119}

\path[fill=fillColor] (514.41,196.90) rectangle (528.09,219.50);

\path[fill=fillColor] (605.55,196.90) rectangle (619.22,200.84);

\path[fill=fillColor] (559.98,196.90) rectangle (573.65,198.67);

\path[fill=fillColor] (651.12,196.90) rectangle (664.79,210.25);

\path[fill=fillColor] (423.27,196.90) rectangle (436.95,225.85);

\path[fill=fillColor] (468.84,196.90) rectangle (482.52,206.79);
\definecolor{fillColor}{RGB}{17,119,51}

\path[fill=fillColor] (500.74,196.90) rectangle (514.41,210.17);

\path[fill=fillColor] (591.88,196.90) rectangle (605.55,200.20);

\path[fill=fillColor] (546.31,196.90) rectangle (559.98,198.51);

\path[fill=fillColor] (637.45,196.90) rectangle (651.12,201.56);

\path[fill=fillColor] (409.60,196.90) rectangle (423.27,234.77);

\path[fill=fillColor] (455.17,196.90) rectangle (468.84,216.52);
\definecolor{fillColor}{RGB}{204,102,119}

\path[fill=fillColor] (487.07,196.90) rectangle (500.74,223.52);

\path[fill=fillColor] (578.21,196.90) rectangle (591.88,201.16);

\path[fill=fillColor] (532.64,196.90) rectangle (546.31,198.67);

\path[fill=fillColor] (623.78,196.90) rectangle (637.45,214.03);

\path[fill=fillColor] (395.93,196.90) rectangle (409.60,221.91);

\path[fill=fillColor] (441.50,196.90) rectangle (455.17,202.45);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (389.10,193.12) rectangle (717.20,276.35);
\end{scope}
\begin{scope}
\path[clip] (374.86,548.04) rectangle (389.10,631.27);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (374.86,548.04) rectangle (389.10,631.27);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (384.10,589.65) {Colombia};
\end{scope}
\begin{scope}
\path[clip] (374.86,459.31) rectangle (389.10,542.54);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (374.86,459.31) rectangle (389.10,542.54);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (384.10,500.92) {Nigeria};
\end{scope}
\begin{scope}
\path[clip] (374.86,370.58) rectangle (389.10,453.81);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (374.86,370.58) rectangle (389.10,453.81);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (384.10,412.19) {Sierra Leone 1};
\end{scope}
\begin{scope}
\path[clip] (374.86,281.85) rectangle (389.10,365.08);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (374.86,281.85) rectangle (389.10,365.08);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (384.10,323.46) {Uganda 2};
\end{scope}
\begin{scope}
\path[clip] (374.86,193.12) rectangle (389.10,276.35);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (374.86,193.12) rectangle (389.10,276.35);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (384.10,234.73) {Russia};
\end{scope}
\begin{scope}
\path[clip] ( 27.01,548.04) rectangle ( 41.25,631.27);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 27.01,548.04) rectangle ( 41.25,631.27);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 36.26,589.65) {Burkina Faso};
\end{scope}
\begin{scope}
\path[clip] ( 27.01,459.31) rectangle ( 41.25,542.54);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 27.01,459.31) rectangle ( 41.25,542.54);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 36.26,500.92) {Nepal};
\end{scope}
\begin{scope}
\path[clip] ( 27.01,370.58) rectangle ( 41.25,453.81);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 27.01,370.58) rectangle ( 41.25,453.81);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 36.26,412.19) {Rwanda};
\end{scope}
\begin{scope}
\path[clip] ( 27.01,281.85) rectangle ( 41.25,365.08);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 27.01,281.85) rectangle ( 41.25,365.08);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 36.26,323.46) {Sierra Leone 2};
\end{scope}
\begin{scope}
\path[clip] ( 27.01,193.12) rectangle ( 41.25,276.35);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 27.01,193.12) rectangle ( 41.25,276.35);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 36.26,234.73) {All LMICs};
\end{scope}
\begin{scope}
\path[clip] ( 27.01,104.39) rectangle ( 41.25,187.62);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] ( 27.01,104.39) rectangle ( 41.25,187.62);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 36.26,146.00) {USA};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 68.60,101.64) --
	( 68.60,104.39);

\path[draw=drawColor,line width= 0.6pt,line join=round] (114.17,101.64) --
	(114.17,104.39);

\path[draw=drawColor,line width= 0.6pt,line join=round] (159.74,101.64) --
	(159.74,104.39);

\path[draw=drawColor,line width= 0.6pt,line join=round] (205.31,101.64) --
	(205.31,104.39);

\path[draw=drawColor,line width= 0.6pt,line join=round] (250.88,101.64) --
	(250.88,104.39);

\path[draw=drawColor,line width= 0.6pt,line join=round] (296.45,101.64) --
	(296.45,104.39);

\path[draw=drawColor,line width= 0.6pt,line join=round] (342.02,101.64) --
	(342.02,104.39);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 70.72, 99.44) {Health workers};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (111.53, 99.44) {Government or};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (121.04, 99.44) {MoH};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (157.10, 99.44) {Family or};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (166.61, 99.44) {Friends};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (193.17, 99.44) {Famous person,};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (202.67, 99.44) {religious leader};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (212.18, 99.44) {or traditional};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (221.68, 99.44) {healers};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (243.49, 99.44) {Newspapers,};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (253.00, 99.44) {radio or online};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (262.50, 99.44) {groups};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (298.57, 99.44) {Other};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (339.38, 99.44) {Don't know or};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (348.89, 99.44) {Refuse};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] (416.44,190.37) --
	(416.44,193.12);

\path[draw=drawColor,line width= 0.6pt,line join=round] (462.01,190.37) --
	(462.01,193.12);

\path[draw=drawColor,line width= 0.6pt,line join=round] (507.58,190.37) --
	(507.58,193.12);

\path[draw=drawColor,line width= 0.6pt,line join=round] (553.15,190.37) --
	(553.15,193.12);

\path[draw=drawColor,line width= 0.6pt,line join=round] (598.72,190.37) --
	(598.72,193.12);

\path[draw=drawColor,line width= 0.6pt,line join=round] (644.29,190.37) --
	(644.29,193.12);

\path[draw=drawColor,line width= 0.6pt,line join=round] (689.86,190.37) --
	(689.86,193.12);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (418.56,188.17) {Health workers};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (459.38,188.17) {Government or};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (468.88,188.17) {MoH};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (504.95,188.17) {Family or};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (514.45,188.17) {Friends};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (541.01,188.17) {Famous person,};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (550.52,188.17) {religious leader};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (560.02,188.17) {or traditional};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (569.53,188.17) {healers};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (591.34,188.17) {Newspapers,};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (600.84,188.17) {radio or online};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (610.34,188.17) {groups};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (646.41,188.17) {Other};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (687.23,188.17) {Don't know or};

\node[text=drawColor,rotate= 90.00,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at (696.73,188.17) {Refuse};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,549.70) {0};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,569.80) {25};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,589.91) {50};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,610.01) {75};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.26,551.82) --
	( 27.01,551.82);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.26,571.92) --
	( 27.01,571.92);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.26,592.03) --
	( 27.01,592.03);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.26,612.13) --
	( 27.01,612.13);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,460.97) {0};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,481.07) {25};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,501.17) {50};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,521.28) {75};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.26,463.09) --
	( 27.01,463.09);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.26,483.19) --
	( 27.01,483.19);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.26,503.30) --
	( 27.01,503.30);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.26,523.40) --
	( 27.01,523.40);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,372.24) {0};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,392.34) {25};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,412.44) {50};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,432.55) {75};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.26,374.36) --
	( 27.01,374.36);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.26,394.46) --
	( 27.01,394.46);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.26,414.57) --
	( 27.01,414.57);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.26,434.67) --
	( 27.01,434.67);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,283.51) {0};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,303.61) {25};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,323.71) {50};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,343.82) {75};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.26,285.63) --
	( 27.01,285.63);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.26,305.73) --
	( 27.01,305.73);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.26,325.84) --
	( 27.01,325.84);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.26,345.94) --
	( 27.01,345.94);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,194.78) {0};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,214.88) {25};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,234.98) {50};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,255.09) {75};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.26,196.90) --
	( 27.01,196.90);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.26,217.00) --
	( 27.01,217.00);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.26,237.10) --
	( 27.01,237.10);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.26,257.21) --
	( 27.01,257.21);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,106.05) {0};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,126.15) {25};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,146.25) {50};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 15.05,166.35) {75};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.26,108.17) --
	( 27.01,108.17);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.26,128.27) --
	( 27.01,128.27);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.26,148.37) --
	( 27.01,148.37);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 24.26,168.48) --
	( 27.01,168.48);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 10.80,367.83) {estimate};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (285.48,  5.50) rectangle (472.97, 30.95);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.10] at (290.98, 15.58) {Answer};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (324.70, 11.00) rectangle (339.15, 25.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{fillColor}{RGB}{204,102,119}

\path[fill=fillColor] (325.41, 11.71) rectangle (338.44, 24.74);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (398.94, 11.00) rectangle (413.39, 25.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{fillColor}{RGB}{17,119,51}

\path[fill=fillColor] (399.65, 11.71) rectangle (412.68, 24.74);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (434.93, 11.00) rectangle (449.39, 25.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{fillColor}{RGB}{221,204,119}

\path[fill=fillColor] (435.65, 11.71) rectangle (448.68, 24.74);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (344.65, 16.11) {No, Don't know};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (418.89, 16.11) {Yes};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (454.89, 16.11) {Any};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (722.70,650.43);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.32] at (  5.50,638.57) {Which of the following people would you trust MOST to help you decide whether you would get a COVID-19 vaccine?};
\end{scope}
\end{tikzpicture}

\end{document}