### 2021-07-17 paper.Rmd
clean up
1_cleaning.Rmd preserved but not called
3_main_text.Rmd called rather than call to googledrive
removal of legacy files

---
title: "COVID-19 Vaccine Acceptance and Hesitancy in Low and Middle Income Countries"
output: 
  bookdown::pdf_document2:
    keep_tex: true
    latex_engine: xelatex
    pandoc_args:
      - --template=4_files/mytemplate.latex
    toc: false
    extra_dependencies:
      caption: ["font={bf,normalsize}"]
      mathptmx: null
  bookdown::word_document2: default
bibliography: 4_files/bib_final.bib
csl: 4_files/nature.csl
header-includes:
  - \setmainfont{Arial}
  - \renewcommand{\familydefault}{\sfdefault}
  - \usepackage{titlesec}
  - \usepackage{hyperref}
  - \titleformat*{\subsection}{\normalsize\bfseries}
  - \titleformat*{\section}{\large\bfseries}
  - \usepackage{float} \floatplacement{figure}{H} 
  - \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
# linestretch: 1.5
urlcolor: blue
link-citations: yes
linkcolor: blue
fontsize: 12pt
---

```{r setup, include=FALSE}

library(knitr)

knitr::opts_chunk$set(echo = FALSE)

```

```{r cleaning, child = "1_cleaning.Rmd", include = FALSE, eval = FALSE}

```

```{r tables, include = FALSE, message = FALSE}

knitr::purl("2_tables.Rmd", documentation = 2)

source("2_tables.R")

file.remove("2_tables.R")

nice <- function(x) prettyNum(x, big.mark = ",", scientific = FALSE)

knitr::opts_chunk$set(echo = FALSE)

options( tikzDefaultEngine = 'xetex',
         tikzXelatexPackages = c(getOption("tikzXelatexPackages"),
                                 "\\setmainfont{Arial}\n"),
         tikzDocumentDeclaration = "% !TEX program = xelatex\n
\\documentclass[fontsize=7pt,DIV=calc]{scrartcl}\n")

```

```{r main_text, child = "3_main_text.Rmd"}

```